### 2025-04-11 sec/authoring-workflow.tex
Tweaks.

\section{AI-Assisted Authoring Workflow}
\label{sec:authoring-workflow}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{fig/authoring-assistant-architecture.png}
    \caption{AI-assisted authoring workflow}
\end{figure}

\subsection{Editor Loop}
Design of main loop that would be integrated into an IDE. The configuration (state) that is being maintained
is a \kw{Paragraph} the user is authoring in Fluid. Consists of a sequence of text fragments, some of which
are uninterpreted (plain literals), the remainder have underlying expressions linking the text to raw or derived
data.

Workflow:
\begin{enumerate}
\item User selects (a substring of) one of the literal text fragments, indicating they want to link it in this
way
\item Authoring tool generates expression which is either:
  \begin{enumerate}
  \item computes the selected text (e.g.~``greater than'' might be computed by comparing two numbers)
  \item becomes the formal meaning of the selected text (e.g. the text ``carbon intensity of methane
emissions'' can be understood in a specific context as referring to the numerical value ``34 gCO2eq/kWh''
  \end{enumerate}
\item Some kind of validation step, including both automated validation (e.g. checking for runtime errors) and
user validation; if validation fails, goto (2)
\item Partition code into additional definitions and expression; add definitions to main source file, and
incorporate expression into \kw{Paragraph}
\item Goto (1) with updated editor state
\end{enumerate}

\subsection{Turning validation errors into improved prompts}
E.g.:
\begin{enumerate}
\item Turn ``Definition not found'' into prompt to generate definition?
\end{enumerate}

\subsection{IDE integration}

Could then be integrated into desktop IDE like VSCode or online IDE like CodeMirror. Might make a good
internship project, but could also be out-of-scope for this paper.

\subsection{Other potential enhancements}

We could also think about using an LLM in a couple of other complementary ways:
\begin{itemize}
\item identifying text fragments which might be linked;
\item validating generated expressions (perhaps by proposing test cases)
\end{itemize}

### 2025-04-11 src/main/java/explorableviz/transparenttext/AuthoringAssistant.java
[fix] remove cast

@@ -18,7 +18,7 @@ public class AuthoringAssistant {
 
     public final Logger logger = Logger.getLogger(AuthoringAssistant.class.getName());
     private final PromptList prompts;
-    private final LLMEvaluatorAgent llm;
+    private final LLMEvaluatorAgent<Expression> llm;
     private final Program templateProgram;
 
     public AuthoringAssistant(InContextLearning inContextLearning, String agentClassName, Program templateProgram) throws Exception {
@@ -40,6 +40,7 @@ public class AuthoringAssistant {
             programEdit.replaceParagraph(programEdit.getParagraph().splice(result.response() == null ? individualEdit.component2() : result.response()));
             results.add(new Pair<>(programEdit, result));
             programEdits = programEdit.asIndividualEdits(templateProgram);
+            programEdit.toWebsite();
         }
         return results;
     }
@@ -56,7 +57,7 @@ public class AuthoringAssistant {
         for (attempts = 0; attempts <= limit; attempts++) {
             logger.info(STR."Attempt #\{attempts}");
             // Send the program to the LLM to be processed
-            Expression candidate = (Expression) llm.evaluate(sessionPrompts, "");
+            Expression candidate = llm.evaluate(sessionPrompts, "");
             //Check each generated expressions
             logger.info(STR."Received response: \{candidate.getExpr()}");
             //program.writeFluidFiles(candidate.getExpr());
@@ -73,11 +74,11 @@ public class AuthoringAssistant {
         return new QueryResult(null, expected, attempts, System.currentTimeMillis() - start);
     }
 
-    private LLMEvaluatorAgent initialiseAgent(String agentClassName) throws ClassNotFoundException, NoSuchMethodException, InvocationTargetException, InstantiationException, IllegalAccessException {
+    private LLMEvaluatorAgent<Expression> initialiseAgent(String agentClassName) throws ClassNotFoundException, NoSuchMethodException, InvocationTargetException, InstantiationException, IllegalAccessException {
         logger.info(STR."Initializing agent: \{agentClassName}");
-        LLMEvaluatorAgent llmAgent;
+        LLMEvaluatorAgent<Expression> llmAgent;
         Class<?> agentClass = Class.forName(agentClassName);
-        llmAgent = (LLMEvaluatorAgent) agentClass
+        llmAgent = (LLMEvaluatorAgent<Expression>) agentClass
                 .getDeclaredConstructor(JSONObject.class)
                 .newInstance(Settings.getSettings());
 

### 2025-04-11 src/main/java/explorableviz/transparenttext/Program.java
[new] website template moved in a folder

@@ -277,51 +277,9 @@ public class Program {
             e.printStackTrace();
         }
         /* html generation */
-        String html = "<!DOCTYPE html>\n" +
-                "<html lang=\"en\">\n" +
-                "<head>\n" +
-                "   <link rel=\"icon\" href=\"/favicon.ico\" type=\"image/x-icon\">\n" +
-                "   <meta charset=\"UTF-8\">\n" +
-                "   <meta name=\"viewport\" content=\"width=device-width, initial-scale=1.0\">\n" +
-                "   <meta http-equiv=\"X-UA-Compatible\" content=\"ie=edge\">\n" +
-                "   <title>Fluid: "+Path.of(this.testCaseFileName).getParent().getFileName()+"</title>\n" +
-                "   <link href=\"../css/styles.css\" rel=\"stylesheet\" type=\"text/css\">\n" +
-                "   <link href=\"../css/view-styles.css\" rel=\"stylesheet\" type=\"text/css\">\n" +
-                "   <script src=\"https://kit.fontawesome.com/20cf8b42c0.js\" crossorigin=\"anonymous\"></script>\n" +
-                "   <script src=\"../shared/util.js\"></script>\n" +
-                "   <style>\n" +
-                "      .data-pane {\n" +
-                "         max-width: 400px;\n" +
-                "      }\n" +
-                "   </style>\n" +
-                "</head>\n" +
-                "<body>\n" +
-                "<div id=\"grid\" class=\"grid-container data-pane-hidden\">\n" +
-                "   <div></div>\n" +
-                "   <div></div>\n" +
-                "   <div class=\"flex-left-align\">\n" +
-                "      <h3>" + Path.of(this.testCaseFileName).getFileName() + "</h3>\n" +
-                "      <p>---</p>\n" +
-                "   </div>\n" +
-                "\n" +
-                "   <div class=\"flex-right-align data-pane\">\n" +
-                "      <div id=\"fig-input\"></div>\n" +
-                "   </div>\n" +
-                "   <div onclick=\"toggleDataPane('grid')\">\n" +
-                "      <i class=\"data-pane-button toggle-button fa-solid fa-eye-slash\"></i>\n" +
-                "   </div>\n" +
-                "   <div id=\"fig\">\n" +
-                "      <div class=\"fig-loading\">loading figure(s)</div>\n" +
-                "   </div>\n" +
-                "</div>\n" +
-                "\n" +
-                "<script type=\"module\">\n" +
[…]

### 2025-04-11 src/main/java/explorableviz/transparenttext/Program.java
[fix] creation directory

@@ -182,6 +182,7 @@ public class Program {
 
     public static void writeFluidFiles(String basePath, String fluidFileName, String response, Map<String, String> datasets, Map<String, String> loadedDatasets, List<String> imports, List<String> loadedImports,  String code) throws IOException {
         Files.createDirectories(Paths.get(basePath));
+        Files.createDirectories(Paths.get(STR."\{basePath}/\{fluidFileName}"));
         //Write temp fluid file
         try (PrintWriter out = new PrintWriter(STR."\{basePath}/\{fluidFileName}.fld")) {
             out.println(code);

### 2025-04-11 src/main/java/explorableviz/transparenttext/paragraph/Paragraph.java
[fix] enable passing value

@@ -1,5 +1,6 @@
 package explorableviz.transparenttext.paragraph;
 
+import explorableviz.transparenttext.Settings;
 import kotlin.Pair;
 
 import java.util.ArrayList;
@@ -16,7 +17,7 @@ public class Paragraph extends ArrayList<TextFragment> {
         return STR."Paragraph([\{stream().map(e -> {
             if (e instanceof Literal l && l.getSelectedRegion() != null)
             {
-                return STR."\{e.getValue().substring(0, l.getSelectedRegion().start())} [REPLACE]\{e.getValue().substring(l.getSelectedRegion().end())}";
+                return STR."\{e.getValue().substring(0, l.getSelectedRegion().start())} [REPLACE \{Settings.isAddExpectedValueEnabled() ? STR."value=\"\{e.getValue()}\"" : ""}]\{e.getValue().substring(l.getSelectedRegion().end())}";
             }
             else if (e instanceof Literal) return STR."Text \"\{e.getValue()} \"";
             else if (e instanceof Expression) {

### 2025-04-11 src/main/java/explorableviz/transparenttext/AuthoringAssistant.java
[update] fix evaluate in authoring assistant

@@ -14,6 +14,8 @@ import java.util.logging.Logger;
 
 import explorableviz.transparenttext.Program.QueryResult;
 
+import static explorableviz.transparenttext.Program.writeFluidFiles;
+
 public class AuthoringAssistant {
 
     public final Logger logger = Logger.getLogger(AuthoringAssistant.class.getName());
@@ -60,7 +62,7 @@ public class AuthoringAssistant {
             Expression candidate = llm.evaluate(sessionPrompts, "");
             //Check each generated expressions
             logger.info(STR."Received response: \{candidate.getExpr()}");
-            //program.writeFluidFiles(candidate.getExpr());
+            writeFluidFiles(Settings.getFluidTempFolder(), Program.fluidFileName, candidate.getExpr(), subProgram.getDatasets(), subProgram.get_loadedDatasets(), subProgram.getImports(), subProgram.get_loadedImports(), subProgram.getCode());
             final FluidCLI fluidCLI = new FluidCLI(subProgram.getDatasets(), subProgram.getImports());
             Optional<String> error = Program.validate(fluidCLI.evaluate(subProgram.getFluidFileName()), expected);
             if (error.isPresent()) {

### 2025-04-11 src/main/java/explorableviz/transparenttext/Program.java

@@ -35,7 +35,7 @@ public class Program {
     private final Paragraph paragraph;
     private final Map<String, String> _loadedDatasets;
     private final String testCaseFileName;
-    private static final String fluidFileName = "llmTest";
+    public static final String fluidFileName = "llmTest";
 
     public Program(Paragraph paragraph, Map<String,String> datasets, List<String> imports, String code, Map<String, String> loadedDataset, String testCaseFileName) throws IOException {
         this.datasets = datasets;

### 2025-04-11 sec/related-work.tex

\section{Related Work}
\label{sec:related}

\todo{flesh out}

\begin{itemize}
\item Transparent visualisations \cite{perera22,bond24}
\end{itemize}

### 2025-04-11 src/main/java/explorableviz/transparenttext/Program.java
[new] link generation in index.html

@@ -8,7 +8,23 @@ import explorableviz.transparenttext.variable.Variables;
 import kotlin.Pair;
 import org.json.JSONArray;
 import org.json.JSONObject;
+import org.w3c.dom.Document;
+import org.w3c.dom.Element;
+import org.w3c.dom.Node;
+import org.xml.sax.SAXException;
 
+import javax.xml.parsers.DocumentBuilder;
+import javax.xml.parsers.DocumentBuilderFactory;
+import javax.xml.parsers.ParserConfigurationException;
+import javax.xml.transform.Transformer;
+import javax.xml.transform.TransformerException;
+import javax.xml.transform.TransformerFactory;
+import javax.xml.transform.dom.DOMSource;
+import javax.xml.transform.stream.StreamResult;
+import javax.xml.xpath.XPath;
+import javax.xml.xpath.XPathConstants;
+import javax.xml.xpath.XPathExpressionException;
+import javax.xml.xpath.XPathFactory;
 import java.io.File;
 import java.io.FileWriter;
 import java.io.IOException;
@@ -182,8 +198,8 @@ public class Program {
 
     public static void writeFluidFiles(String basePath, String fluidFileName, String response, Map<String, String> datasets, Map<String, String> loadedDatasets, List<String> imports, List<String> loadedImports,  String code) throws IOException {
         Files.createDirectories(Paths.get(basePath));
-        Files.createDirectories(Paths.get(STR."\{basePath}/\{fluidFileName}"));
-        //Write temp fluid file
+        Files.createDirectories(Paths.get(STR."\{basePath}/\{fluidFileName}").getParent());
+
         try (PrintWriter out = new PrintWriter(STR."\{basePath}/\{fluidFileName}.fld")) {
             out.println(code);
             out.println(response);
@@ -267,7 +283,7 @@ public class Program {
         });
 
         imports.forEach(_import -> {
-            spec.getJSONArray("imports").put(STR."lib/\{_import}");
+            spec.getJSONArray("imports").put(_import);
         });
         spec.put("file", Path.of(this.testCaseFileName).getFileName());
         spec.put("inputs", new JSONArray("[\"tableData\"]"));
@@ -288,6 +304,7 @@ public class Program {
             e.printStackTrace();
         }
         /* copy datasets  & lib */
[…]

### 2025-04-30 sec/authoring-workflow.tex
[update] add website and fluid code into the paper

@@
     \includegraphics[width=0.95\linewidth]{fig/authoring-assistant-architecture.png}
-    \caption{AI-assisted authoring workflow}
+    \caption{AI-assisted authoring workflow}\label{fig:architecture}
 \end{figure}
 
-\subsection{Editor Loop}
+\subsection{Editor Loop}\label{subsec:editor-loop}
 Design of main loop that would be integrated into an IDE. The configuration (state) that is being maintained
@@
 
-\subsection{Turning validation errors into improved prompts}
+\subsection{Paragraph with generated expression}
+Figure \ref{fig:fluid-example-paragraph} shows the fluid code of a Paragraph with fluid expression generated by the authoring assistant.
+Figure \ref{fig:fluid-scigen} shows the fluid code written for the scigen datasets and examples.
+\begin{figure}[h]
+    \small
+    {\lstinputlisting[language=Fluid]{website/authoring-assistant/fluid/1805.02474v1-10.fld}}
+    \vspace{-0.5em}
+    \caption{Fluid source code example Paragraph}
+    \label{fig:fluid-example-paragraph}
+\end{figure}
+
+\begin{figure}[h]
+    \tiny
+    {\lstinputlisting[language=Fluid]{fluid-common/scigen.fld}}
+    \vspace{-0.5em}
+    \caption{Fluid library written for scigen}
+    \label{fig:fluid-scigen}
+\end{figure}
+
+\subsection{Render Example}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.95\linewidth]{fig/scigen-mr-1906.02780-10.png}
+    \caption{Website example with the generated expression from authoring assistant}\label{fig:scigen-example-website}
+\end{figure}
+
+\subsection{Turning validation errors into improved prompts}\label{subsec:turning-validation-errors-into-improved-prompts}
 E.g.:
@@
 
-\subsection{IDE integration}
+\subsection{IDE integration}\label{subsec:ide-integration}
 
@@
 
-\subsection{Other potential enhancements}
+\subsection{Other potential enhancements}\label{subsec:other-potential-enhancements}
 
@@
 \end{itemize}
+
+\section{Results}
+\label{sec:results}
+
+\subsection{Success Rate by Testcase}
+\label{subsec:success_rate_by_test_case}
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.95\linewidth]{fig/success_rate_by_test_case}
+    \caption{Success rate by testcases}\label{fig:success_rate_by_test_case}
+\end{figure}

### 2025-04-30 sec/authoring-workflow.tex

@@
 
-\subsection{Paragraph with generated expression}
-Figure \ref{fig:fluid-example-paragraph} shows the fluid code of a Paragraph with fluid expression generated by the authoring assistant.
-Figure \ref{fig:fluid-scigen} shows the fluid code written for the scigen datasets and examples.
-\begin{figure}[h]
-    \small
-    {\lstinputlisting[language=Fluid]{website/authoring-assistant/fluid/1805.02474v1-10.fld}}
-    \vspace{-0.5em}
-    \caption{Fluid source code example Paragraph}
-    \label{fig:fluid-example-paragraph}
-\end{figure}
-
-\begin{figure}[h]
-    \tiny
-    {\lstinputlisting[language=Fluid]{fluid-common/scigen.fld}}
-    \vspace{-0.5em}
-    \caption{Fluid library written for scigen}
-    \label{fig:fluid-scigen}
-\end{figure}
-
-\subsection{Render Example}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.95\linewidth]{fig/scigen-mr-1906.02780-10.png}
-    \caption{Website example with the generated expression from authoring assistant}\label{fig:scigen-example-website}
-\end{figure}
-
 \subsection{Turning validation errors into improved prompts}\label{subsec:turning-validation-errors-into-improved-prompts}

### 2025-05-06 sec/introduction.tex
[draft] table

\section{Introduction}

\begin{itemize}
\item Making sense of data-driven claims is hard, even if the evidence base (code/data) is open
\item We see this in peer review, misinformation, retracted papers, …
\end{itemize}

\begin{figure}[h]
   \includegraphics[width=0.9\textwidth]{fig/ipcc-mockup.png}
   \caption{Mockup of end-user transparency features (numbered 1 to 8)}
   \label{fig:ipcc-mockup}
\end{figure}

\subsection{Self-Certifying Text}

\begin{itemize}
\item Introduce basic idea, as complementary to transparent visualisations
\end{itemize}

\subsubsection{Use cases}
Two potential scenarios for this sort of technology:

\paragraph{Authoring transparent text.} Someone authoring content for an online article, wants to create text
linked to raw data (and derivative data such as charts or tabular summaries), so that the evidence base for
the claims made in the text can be explored \emph{in situ}, by interacting with the text.

\paragraph{Interpreting text after the fact.} Someone reading textual claims derived from open data (e.g. a
scientific paper or climate report), wants to retroactively link the text to queries over the available data
and gradually ``rationally reconstruct'' the relationship between the claims in the paper and the evidence
base. Perhaps just to aid their own comprehension, or to provide some kind of justified peer review.

\vspace{2mm}
\noindent Here we focus on the first one because the second one requires a certain amount of additional setup.

\subsubsection{Target idioms of natural language}

NLP aspect of the problem is potentially a big problem space in itself. We will restrict interest to certain
idiomatic uses of natural language in making/justifying scientific claims. Examples:

\begin{itemize}
\item quantitative expressions
\begin{itemize}
   \item pure numerical values
   \item percentages
   \item rounded or normalised numbers
\end{itemize}
\item aggregation
\begin{itemize}
   \item average
   \item totals
   \item count
\end{itemize}
\item trends
\begin{itemize}
   \item comparisons
   \item ranks
\end{itemize}
\item graded Adjectives (e.g.~\emph{virtually certain}, \emph{exceptionally unlikely})
\item references to visual elements and their parts (mereology)
\end{itemize}

\begin{table}[!ht]
   \centering
   \tiny
   \begin{tabular}{p{2cm}p{1.3cm}p{6.7cm}p{2cm}}
      \toprule
      Type & Category & Example & Status \\
      \midrule
      pure numerical values & Quantitative expressions & (findWithKey' "model" "LSTM" tableData).time\_s & Done \\
      percentages & Quantitative expressions &  "The Energy Sector accounts for total methane emissions of", (record.emissions / sum(map (fun x -> x.emissions) (getByYear year tableData))) * 100 & Done \\
      rounded or normalised numbers & Quantitative expressions & ~ & ~ \\
      average & Aggregation & "The average methane emissions for the year 2024 is", (sumEmissions year tableData / length records) & Done \\
      totals & Aggregation & "The total methane emissions for the year 2024 is", (sumEmissions year tableData) & Done \\
      count & Aggregation & ~ & ~ \\
      comparisons & Trends & trendWord (findWithKey' "model" "BiLSTM" tableData).time\_s (findWithKey' "model" "LSTM" tableData).time\_s "growing from" "shrinking to" "equal to" & Done \\
      ranks & Trends & let pos = findIndex "model" "CNN" (insertionSort cmpTime tableData) in rankLabel "lowest" pos & Done \\
      Graded adjectives  & Adjectives & ~ & ~ \\
      References to visual elements and their parts (mereology) & ~ & ~ & ~ \\
      \bottomrule
   \end{tabular}
\end{table}

\subsection{Contributions}

\begin{itemize}
\item Design and proof-of-concept implementation of AI-assisted workflow for authoring transparent text
(\secref{authoring-workflow})
\item Empirical evaluation of how effective current LLMs are at providing the ``AI-assisted'' part
\end{itemize}

### 2025-05-07 sec/introduction.tex

@@
    \tiny
-   \begin{tabular}{p{2cm}p{1.3cm}p{6.7cm}p{2cm}}
+   \renewcommand{\arraystretch}{1.5}
+   \begin{tabular}{p{1.5cm}p{1.5cm}p{5cm}p{4cm}}
       \toprule
-      Type & Category & Example & Status \\
+      \textbf{Category} & \textbf{Type} & \textbf{Example} & \textbf{Result} \\
       \midrule
-      pure numerical values & Quantitative expressions & (findWithKey' "model" "LSTM" tableData).time\_s & Done \\
-      percentages & Quantitative expressions &  "The Energy Sector accounts for total methane emissions of", (record.emissions / sum(map (fun x -> x.emissions) (getByYear year tableData))) * 100 & Done \\
-      rounded or normalised numbers & Quantitative expressions & ~ & ~ \\
-      average & Aggregation & "The average methane emissions for the year 2024 is", (sumEmissions year tableData / length records) & Done \\
-      totals & Aggregation & "The total methane emissions for the year 2024 is", (sumEmissions year tableData) & Done \\
-      count & Aggregation & ~ & ~ \\
-      comparisons & Trends & trendWord (findWithKey' "model" "BiLSTM" tableData).time\_s (findWithKey' "model" "LSTM" tableData).time\_s "growing from" "shrinking to" "equal to" & Done \\
-      ranks & Trends & let pos = findIndex "model" "CNN" (insertionSort cmpTime tableData) in rankLabel "lowest" pos & Done \\
-      Graded adjectives  & Adjectives & ~ & ~ \\
+      \rowcolor{green!20}
+      Quantitative expressions & Pure numerical values &
+      \begin{lstlisting}[language=Fluid]
+            (findWithKey' "model" "LSTM" tableData).time\_s
+      \end{lstlisting}
+      & 67 \\
+      \rowcolor{green!20}
+      Quantitative expressions & Percentages &
+      \begin{lstlisting}[language=Fluid]
+         "The Energy Sector accounts for total methane emissions of", (record.emissions / sum(map (fun x -> x.emissions) (getByYear year tableData))) * 100, "in 2030"
+      \end{lstlisting} &
+      The Energy Sector accounts for total methane emissions of 52.80\% in 2030.\\
+      \rowcolor{green!20}
+      Quantitative expressions & Rounded or normalised numbers & ~ & ~ \\
+      \rowcolor{green!20}
+      Aggregation & Average &
+      \begin{lstlisting}[language=Fluid]
+         The average methane emissions for the year 2030 is, (sumEmissions year tableData / length records)
+      \end{lstlisting}
+      & The average methane emissions for the year 2030 is 13.51 \\
+      \rowcolor{green!20}
+      Aggregation & Totals &
+      \begin{lstlisting}[language=Fluid]
+         "The total methane emissions for the year 2030 is", (sumEmissions year tableData), "for Agriculture"
+      \end{lstlisting} &
+      The total methane emissions for the year 2030 is 37.74 for Agriculture \\
+      Aggregation & Count &  ~ & ~ \\
+      \rowcolor{green!20}
+      Trends & Comparisons &
+      \begin{lstlisting}[language=Fluid]
+         trendWord (findWithKey' "model" "BiLSTM" tableData).time\_s (findWithKey' "model" "LSTM" tableData).time\_s "growing from" "shrinking to" "equal to"
+      \end{lstlisting}
+      & growing from \\
+      \rowcolor{green!20}
+      Trends & Ranks &
+      \begin{lstlisting}[language=Fluid]
+         let pos = findIndex "model" "CNN" (insertionSort cmpTime tableData) in rankLabel "lowest" pos
+      \end{lstlisting}
+      & lowest \\
+      Adjectives & Graded adjectives  &  ~ & ~ \\
       References to visual elements and their parts (mereology) & ~ & ~ & ~ \\
@@
    \end{tabular}
+   \caption{Example expressions generated by the Authoring Assistant, classified by type. In green, the expression is already tested.}
+   \label{tab:fluid_examples}
+   \renewcommand{\arraystretch}{1.0}
 \end{table}

### 2025-05-09 src/main/java/explorableviz/authoringassistant/AuthoringAssistant.java
[refactory] rewriting function trendWord

@@ -36,10 +36,10 @@ public class AuthoringAssistant {
         while (!programEdits.isEmpty()) {
             Pair<Program, Expression> individualEdit = programEdits.get(i);
             //selection
-            Program programEdit = individualEdit.component1();
+            Program programEdit = individualEdit.getFirst();
             QueryResult result = execute(individualEdit);
 
-            programEdit.replaceParagraph(programEdit.getParagraph().splice(result.response() == null ? individualEdit.component2() : result.response()));
+            programEdit.replaceParagraph(programEdit.getParagraph().splice(result.response() == null ? individualEdit.getSecond() : result.response()));
             results.add(new Pair<>(programEdit, result));
             programEdits = programEdit.asIndividualEdits(templateProgram);
             programEdit.toWebsite();
@@ -52,8 +52,8 @@ public class AuthoringAssistant {
         // Add the input query to the KB that will be sent to the LLM
         int attempts;
         final long start = System.currentTimeMillis();
-        Program subProgram = test.component1();
-        Expression expected = test.component2();
+        Program subProgram = test.getFirst();
+        Expression expected = test.getSecond();
         final PromptList sessionPrompts = (PromptList) prompts.clone();
         sessionPrompts.addUserPrompt(subProgram.toUserPrompt());
         for (attempts = 0; attempts <= limit; attempts++) {

### 2025-05-09 src/main/java/explorableviz/authoringassistant/InContextLearning.java

@@ -35,7 +35,7 @@ public class InContextLearning {
         for (Program initialStatesFromTemplate : this.cases) {
             List<Pair<Program, Expression>> initialProgramStates = initialStatesFromTemplate.asIndividualEdits(initialStatesFromTemplate);
             for(Pair<Program, Expression> initialProgramState : initialProgramStates) {
-                inContextLearning.addPairPrompt(initialProgramState.component1().toUserPrompt(), initialProgramState.component2().getExpr());
+                inContextLearning.addPairPrompt(initialProgramState.getFirst().toUserPrompt(), initialProgramState.getSecond().getExpr());
             }
         }
         return inContextLearning;

### 2025-05-09 src/main/java/explorableviz/authoringassistant/Main.java

@@ -60,10 +60,10 @@ public class Main {
             out.println(String.join(";", headers));
             String content = results.stream()
                     .map(result -> {
-                        Program program = result.component1();
-                        QueryResult queryResult = result.component2();
+                        Program program = result.getFirst();
+                        QueryResult queryResult = result.getSecond();
                         String[] values = {
-                                result.component1().getTestCaseFileName(),
+                                result.getFirst().getTestCaseFileName(),
                                 agent,
                                 String.valueOf(Settings.getTemperature()),
                                 String.valueOf(Settings.getNumContextToken()),
@@ -85,7 +85,7 @@ public class Main {
 
     private static float computeExactMatch(List<Pair<Program, QueryResult>> results) {
         logger.info("Computing accuracy");
-        long count = IntStream.range(0, results.size()).filter(i -> results.get(i).component2().response() != null && results.get(i).component2().expected().getExpr().equals(results.get(i).component2().response().getExpr())).count();
+        long count = IntStream.range(0, results.size()).filter(i -> results.get(i).getSecond().response() != null && results.get(i).getSecond().expected().getExpr().equals(results.get(i).getSecond().response().getExpr())).count();
         return (float) count / results.size();
     }
 
@@ -98,7 +98,7 @@ public class Main {
         }
         logger.info("Printing generated expression");
         for (Pair<Program, QueryResult> result : results) {
-            logger.info(result.component2().response() != null ? result.component2().response().getExpr() : "NULL");
+            logger.info(result.getSecond().response() != null ? result.getSecond().response().getExpr() : "NULL");
         }
         return results;
     }

### 2025-05-09 sec/authoring-workflow.tex
[refactory] created util.fld and moved insertion sort into it
[update] first changes on the paper

@@
 
+\subsection{Paragraph with generated expression}
+Figure \ref{fig:fluid-example-paragraph} shows the Fluid code of a Paragraph with Fluid expression generated by the authoring assistant.
+Figure \ref{fig:fluid-scigen} shows the Fluid code written for the scigen datasets and examples.
+\begin{figure}[h]
+    \small
+    {\lstinputlisting[language=Fluid]{website/authoring-assistant/fluid/1805.02474v1-10.fld}}
+    \vspace{-0.5em}
+    \caption{Fluid source code example Paragraph}
+    \label{fig:fluid-example-paragraph}
+\end{figure}
+
+\begin{figure}[h]
+    \small
+    {\lstinputlisting[language=Fluid]{fluid-common/scigen.fld}}
+    \vspace{-0.5em}
+    \caption{Fluid library written for scigen}
+    \label{fig:fluid-scigen}
+\end{figure}
+
 \subsection{Turning validation errors into improved prompts}\label{subsec:turning-validation-errors-into-improved-prompts}
@@
 \end{itemize}
-
-\section{Results}
-\label{sec:results}
-
-\subsection{Success Rate by Testcase}
-\label{subsec:success_rate_by_test_case}
-
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.95\linewidth]{fig/success_rate_by_test_case}
-    \caption{Success rate by testcases}\label{fig:success_rate_by_test_case}
-\end{figure}

### 2025-05-09 sec/evaluation.tex
[update] first changes on the paper

\section{Empirical Evaluation}
\label{sec:evaluation}

\subsection{Research questions}

These seem quite pertinent to the overall problem of transparent science communication:
\begin{itemize}
\item How does performance degrade with partial/ambiguous information?
\item How is performance affected by ``adversarial'' (intentionally/unintentionally misleading) information?
\end{itemize}

Some ways we might explore these axes:
\begin{itemize}
\item meaningful vs.~meaningless vs.~misleading identifiers
\item leaving out key criteria (e.g.~year) that would resolve query to unique answer
\end{itemize}

We might want to distinguish ambiguity inherent in the problem domain (e.g.~uncertainty) vs.~ambiguity in the
specification of a particular code-generation query.

\subsection{Success Rate by Testcase}
\label{subsec:success_rate_by_test_case}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.95\linewidth]{fig/success_rate_by_test_case}
    \caption{Success rate by testcases}\label{fig:success_rate_by_test_case}
\end{figure}

### 2025-05-09 sec/introduction.tex
[update] first changes on the paper

--- 2025-05-06 sec/introduction.tex
@@
 
-\begin{figure}[h]
-   \includegraphics[width=0.9\textwidth]{fig/ipcc-mockup.png}
-   \caption{Mockup of end-user transparency features (numbered 1 to 8)}
-   \label{fig:ipcc-mockup}
-\end{figure}
-
 \subsection{Self-Certifying Text}
@@
 
-\subsubsection{Use cases}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/scigen-mr-1906.02780-10.png}
+    \caption{Website example with the generated expression from authoring assistant}\label{fig:scigen-example-website}
+\end{figure}
+
+\subsection{Use cases}
 Two potential scenarios for this sort of technology:
@@
 NLP aspect of the problem is potentially a big problem space in itself. We will restrict interest to certain
-idiomatic uses of natural language in making/justifying scientific claims. Examples:
-
-\begin{itemize}
-\item quantitative expressions
-\begin{itemize}
-   \item pure numerical values
-   \item percentages
-   \item rounded or normalised numbers
-\end{itemize}
-\item aggregation
-\begin{itemize}
-   \item average
-   \item totals
-   \item count
-\end{itemize}
-\item trends
-\begin{itemize}
-   \item comparisons
-   \item ranks
-\end{itemize}
-\item graded Adjectives (e.g.~\emph{virtually certain}, \emph{exceptionally unlikely})
-\item references to visual elements and their parts (mereology)
-\end{itemize}
+idiomatic uses of natural language in making/justifying scientific claims.
+Examples \ref{tab:fluid_examples}
 
@@
     \centering
-   \tiny
-   \begin{tabular}{p{2cm}p{1.3cm}p{6.7cm}p{2cm}}
+    \footnotesize
+    \renewcommand{\arraystretch}{1.5}
+    \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6cm}}
         \toprule
-      Type & Category & Example & Status \\
+        \textbf{Type}                & \textbf{Example} & \textbf{Generated Expression} \\
         \midrule
-      pure numerical values & Quantitative expressions & (findWithKey' "model" "LSTM" tableData).time\_s & Done \\
-      percentages & Quantitative expressions &  "The Energy Sector accounts for total methane emissions of", (record.emissions / sum(map (fun x -> x.emissions) (getByYear year tableData))) * 100 & Done \\
-      rounded or normalised numbers & Quantitative expressions & ~ & ~ \\
-      average & Aggregation & "The average methane emissions for the year 2024 is", (sumEmissions year tableData / length records) & Done \\
-      totals & Aggregation & "The total methane emissions for the year 2024 is", (sumEmissions year tableData) & Done \\
-      count & Aggregation & ~ & ~ \\
-      comparisons & Trends & trendWord (findWithKey' "model" "BiLSTM" tableData).time\_s (findWithKey' "model" "LSTM" tableData).time\_s "growing from" "shrinking to" "equal to" & Done \\
-      ranks & Trends & let pos = findIndex "model" "CNN" (insertionSort cmpTime tableData) in rankLabel "lowest" pos & Done \\
-      Graded adjectives  & Adjectives & ~ & ~ \\
-      References to visual elements and their parts (mereology) & ~ & ~ & ~ \\
+        \multicolumn{3}{>{\raggedright\arraybackslash}c}{\textbf{Quantitative expressions.}} \\
+        \midrule
+        Numerical value
+        & BiLSTM gives significantly better accuracies compared to uni-directional LSTM2, with the training time per epoch growing from \hl{67} seconds to 106 seconds. &
+        \begin{lstlisting}[language=Fluid,numbers=none]
+(findWithKey' "model" "LSTM" tableData).time\_s
+        \end{lstlisting} \\
+        Percentage &
+        The Energy Sector accounts for total methane emissions of \hl{52.80\%} in 2030. &
+        \begin{lstlisting}[language=Fluid,numbers=none]
+(record.emissions / sum(map (fun x -> x.emissions) (getByYear year tableData))) * 100
+        \end{lstlisting}  \\
+        Rounded or normalised number & ~                & ~                             \\
+        \multicolumn{3}{>{\raggedright\arraybackslash}c}{\textbf{Aggregation.}} \\
+        \midrule
+        Average
+        & The average methane emissions for the year 2030 is \hl{13.51} &
+        \begin{lstlisting}[language=Fluid,numbers=none]
+(sumEmissions year tableData / length records)
+        \end{lstlisting} \\
+        Min                          & ~                & ~                             \\
+        Max                          & ~                & ~                             \\\
+        Ranks &
+        On the other hand, a 3-layer stacked CNN gives an accuracy of 81.46\%, which is also the \hl{lowest} compared with BiLSTM, hierarchical attention and S-LSTM  &
+        \begin{lstlisting}[language=Fluid,numbers=none]
+let pos = findIndex "model" "CNN" (insertionSort cmpTime tableData) in rankLabel "lowest" pos
+        \end{lstlisting} \\
+        Total &
+        The total methane emissions for the year 2030 is \hl{37.74} for Agriculture &
+        \begin{lstlisting}[language=Fluid,numbers=none]
+sumEmissions year tableData
+        \end{lstlisting} \\
+        *Count                       & ~                & ~                             \\
+        \multicolumn{3}{>{\raggedright\arraybackslash}c}{\textbf{Trends \& Adjectives.}} \\
+        \midrule
+        Comparisons
+        & BiLSTM gives significantly better accuracies compared to uni-directional LSTM2, with the training time per epoch \hl{growing from} 67 seconds to 106 seconds. &
+        \begin{lstlisting}[language=Fluid,numbers=none]
+trendWord (findWithKey' "model" "BiLSTM" tableData).time\_s (findWithKey' "model" "LSTM" tableData).time\_s "growing from" "shrinking to" "equal to"
+        \end{lstlisting} \\
+        *Graded adjective            & ~                & ~                             \\
+        *Reference to visual element                            & ~                & ~                             \\
         \bottomrule
     \end{tabular}
+    \caption{Example expressions generated by the Authoring Assistant, classified by type. In green, the expression is already tested.}
+    \label{tab:fluid_examples}
+    \renewcommand{\arraystretch}{1.0}
 \end{table}

### 2025-05-14 sec/introduction.tex
[update] changes listing and figures

@@
 \begin{itemize}
+    \item Interactive visualization of data
+    \item Making the text transparent, with the possibility to highlight the data related to a specific sentence.
     \item Making sense of data-driven claims is hard, even if the evidence base (code/data) is open
@@
         \midrule
+        \rowcolor{gray!20}
         \multicolumn{3}{>{\raggedright\arraybackslash}c}{\textbf{Quantitative expressions.}} \\
@@
         \begin{lstlisting}[language=Fluid,numbers=none]
-(findWithKey' "model" "LSTM" tableData).time\_s
+(findWithKey' "model" "LSTM" tableData).time_s
         \end{lstlisting} \\
         Percentage &
-        The Energy Sector accounts for total methane emissions of \hl{52.80\%} in 2030. &
-        \begin{lstlisting}[language=Fluid,numbers=none]
-(record.emissions / sum(map (fun x -> x.emissions) (getByYear year tableData))) * 100
+        The Energy Sector accounts for total methane emissions of \hl{52.80\%} in 2030.
+        & \begin{lstlisting}[language=Fluid,numbers=none]
+(record.emissions /
+ sum (map (fun x -> x.emissions)
+          (getByYear year tableData))) * 100
+
         \end{lstlisting}  \\
@@
         Min                          & ~                & ~                             \\
-        Max                          & ~                & ~                             \\\
+        Max                          & ~                & ~                             \\
         Ranks &
@@
         \begin{lstlisting}[language=Fluid,numbers=none]
-let pos = findIndex "model" "CNN" (insertionSort cmpTime tableData) in rankLabel "lowest" pos
+let pos =
+  findIndex
+    "model"
+    "CNN"
+    (insertionSort cmpTime tableData)
+in rankLabel "lowest" pos
         \end{lstlisting} \\
@@
         \begin{lstlisting}[language=Fluid,numbers=none]
-trendWord (findWithKey' "model" "BiLSTM" tableData).time\_s (findWithKey' "model" "LSTM" tableData).time\_s "growing from" "shrinking to" "equal to"
+trendWord
+ (findWithKey' "model" "BiLSTM" tableData).time_s
+ (findWithKey' "model" "LSTM" tableData).time_s
+ growShrink
         \end{lstlisting} \\

### 2025-05-16 sec/introduction.tex
[update] fix on latex table and update figure

@@
         \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}c}{\textbf{Quantitative expressions.}} \\
-        \midrule
+        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Quantitative expressions.}} \\
+
         Numerical value
@@
         Rounded or normalised number & ~                & ~                             \\
-        \multicolumn{3}{>{\raggedright\arraybackslash}c}{\textbf{Aggregation.}} \\
-        \midrule
+        \rowcolor{gray!20}
+        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Aggregation.}} \\
         Average
@@
 let pos =
-  findIndex
-    "model"
-    "CNN"
+  findIndex "model" "CNN"
     (insertionSort cmpTime tableData)
@@
         *Count                       & ~                & ~                             \\
-        \multicolumn{3}{>{\raggedright\arraybackslash}c}{\textbf{Trends \& Adjectives.}} \\
-        \midrule
+        \rowcolor{gray!20}
+        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Trends \& Adjectives.}} \\
+
         Comparisons
-        & BiLSTM gives significantly better accuracies compared to uni-directional LSTM2, with the training time per epoch \hl{growing from} 67 seconds to 106 seconds. &
+        & BiLSTM gives significantly better accuracies compared to uni-directional LSTM2, with the training time per epoch \hl{growing} from 67 seconds to 106 seconds. &
         \begin{lstlisting}[language=Fluid,numbers=none]

### 2025-05-16 src/main/java/explorableviz/authoringassistant/Program.java
[update] generate always at least 1 variant

@@ -176,12 +176,10 @@ public class Program {
                     System.exit(1);
                 }
                 Variables tv = Variables.fromJSON(new JSONObject(jsonContent).getJSONObject("testing-variables"));
-                int maxVariants = 0;
+                int maxVariants = 1;
                 for(Map.Entry<String, ValueOptions> entry : tv.entrySet() ) {
                     if(entry.getValue() instanceof ValueOptions.List values) {
                         maxVariants = Math.max(maxVariants, values.get().size());
-                    } else {
-                        maxVariants = 1;
                     }
                 }
                 Variables.Flat testVariables = expandVariables(tv, new SplittableRandom(k));

### 2025-05-20 sec/introduction.tex
[update] box for system prompt and minor improvements to the table

@@
         \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Quantitative expressions.}} \\
+        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Quantitative expressions}} \\
 
         Numerical value
-        & BiLSTM gives significantly better accuracies compared to uni-directional LSTM2, with the training time per epoch growing from \hl{67} seconds to 106 seconds. &
+        & the training time per epoch growing from \hl{67} seconds to 106 seconds.
+        &
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         \end{lstlisting}  \\
-        Rounded or normalised number & ~                & ~                             \\
+        Rounding & ~                & ~                             \\
+        Cardinals \& Multiplicative & ~                & ~                             \\
         \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Aggregation.}} \\
+        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Aggregation}} \\
         Average
@@
         \end{lstlisting} \\
-        Min                          & ~                & ~                             \\
-        Max                          & ~                & ~                             \\
+        Min/Max                          & ~                & ~                             \\
         Ranks &
-        On the other hand, a 3-layer stacked CNN gives an accuracy of 81.46\%, which is also the \hl{lowest} compared with BiLSTM, hierarchical attention and S-LSTM  &
+        3-layer stacked CNN gives an accuracy of 81.46\%, which is the \hl{lowest} compared with BiLSTM, and S-LSTM  &
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Trends \& Adjectives.}} \\
+        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Trends}} \\
 
         Comparisons
-        & BiLSTM gives significantly better accuracies compared to uni-directional LSTM2, with the training time per epoch \hl{growing} from 67 seconds to 106 seconds. &
+        & The training time per epoch \hl{growing} from 67 seconds to 106 seconds. &
         \begin{lstlisting}[language=Fluid,numbers=none]

### 2025-05-21 src/main/java/explorableviz/authoringassistant/Program.java
[update] string also in LC

@@ -88,7 +88,7 @@ public class Program {
         if (outputLines.length < 2) {
             throw new RuntimeException("Output format is invalid");
         }
-        return outputLines[FluidCLI.isWindows() ? 2 : 1].replaceAll("^\"|\"$", "");
+        return outputLines[FluidCLI.isWindows() ? 2 : 1];
     }
 
     public static Optional<String> validate(String commandLineResponse, Expression expectedExpression) {
@@ -98,6 +98,10 @@ public class Program {
         if (commandLineResponse.contains("Error: ")) {
             logger.info("Validation failed because interpreter error");
             return Optional.of(value);
+        }
+        //String error
+        if(!value.matches("^\".*\"$")) {
+            return Optional.of(STR."String expected. \{value} seems not a string");
         }
         if (value.equals(expectedExpression.getValue()) || roundedEquals(value, expectedExpression.getValue())) {
             logger.info("Validation passed");

### 2025-05-21 sec/authoring-workflow.tex
[update] added system-prompt
[update] box for system prompt and minor improvements to the table
[update] fix newline in system prompt

@@
     \caption{AI-assisted authoring workflow}\label{fig:architecture}
+\end{figure}
+
+\subsection{System Prompt}
+\label{subsec:system-prompt}
+Fig. \ref{fig:system-prompt} shows the system prompt defined for the Authoring Assistant.
+
+\begin{figure}[h]
+    \centering
+    \begin{tcolorbox}[colback=gray!10, colframe=gray!50, boxrule=0.5pt, arc=2pt,
+        left=6pt, right=6pt, top=4pt, bottom=4pt]
+        \VerbatimInput[fontsize=\small, formatcom=\normalfont]{system-prompt/system-prompt.txt}
+    \end{tcolorbox}
+    \vspace{-0.5em}
+    \caption{System-Prompt defined for the Authoring Assistant}
+    \label{fig:system-prompt}
 \end{figure}

### 2025-05-22 sec/introduction.tex
[update] fix spaces table

@@
     \footnotesize
-    \renewcommand{\arraystretch}{1.5}
+    \renewcommand{\arraystretch}{1.1}
     \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6cm}}
@@
         & the training time per epoch growing from \hl{67} seconds to 106 seconds.
-        &
+        &%check with v-space
+        \vspace{-5pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
 (findWithKey' "model" "LSTM" tableData).time_s
-        \end{lstlisting} \\
+        \end{lstlisting}
+        \\
         Percentage &
         The Energy Sector accounts for total methane emissions of \hl{52.80\%} in 2030.
-        & \begin{lstlisting}[language=Fluid,numbers=none]
+        &
+        \vspace{-5pt}
+        \begin{lstlisting}[language=Fluid,numbers=none]
 (record.emissions /
@@
         & The average methane emissions for the year 2030 is \hl{13.51} &
+        \vspace{-5pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         3-layer stacked CNN gives an accuracy of 81.46\%, which is the \hl{lowest} compared with BiLSTM, and S-LSTM  &
+        \vspace{-5pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         The total methane emissions for the year 2030 is \hl{37.74} for Agriculture &
+        \vspace{-5pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         \end{lstlisting} \\
-        *Count                       & ~                & ~                             \\
+        Count*                       & ~                & ~                             \\
         \rowcolor{gray!20}
@@
         & The training time per epoch \hl{growing} from 67 seconds to 106 seconds. &
+        \vspace{-5pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         \end{lstlisting} \\
-        *Graded adjective            & ~                & ~                             \\
-        *Reference to visual element                            & ~                & ~                             \\
+        Graded adjective*            & ~                & ~                             \\
+        Reference to visual element*                            & ~                & ~                             \\
         \bottomrule
@@
 \end{table}
-
 \subsection{Contributions}

### 2025-05-22 sec/introduction.tex
[update] added expression in min/max

@@
         \end{lstlisting} \\
-        Min/Max                          & ~                & ~                             \\
+        Min/Max                          & The Energy Sector recorded its highest methane emissions in \hl{2030}             &
+        \vspace{-5pt}
+        \begin{lstlisting}[language=Fluid,numbers=none]
+let maxEntry =
+    maximumBy (fun x -> x.emissions)
+              (filter
+                (fun x -> x.type == "Energy Sector")
+                tableData)
+in maxEntry.year
+        \end{lstlisting} \\                             \\
         Ranks &

### 2025-05-23 sec/authoring-workflow.tex

@@
 
-\subsection{Editor Loop}\label{subsec:editor-loop}
+\subsection{Editor Loop}
+\label{subsec:editor-loop}
 Design of main loop that would be integrated into an IDE. The configuration (state) that is being maintained
@@
 
+\subsection{Loopback System}
+\label{subsec:loopback-system}
+
+Once the Authoring Assistant has generated the expression, it is executed by the Fluid interpreter.
+If execution fails, the error is sent back to the LLM for regeneration.
+This \textit{loopback} mechanism enables iterative refinement of expressions, improving accuracy and robustness.
+Three types of errors are handled:
+
+\begin{itemize}
+    \item \textbf{Interpreter Error}.
+    Syntax errors due to incorrect use of the language (e.g., undeclared functions, invalid syntax, etc.).
+    \item \textbf{String Error}.
+    The expression does not return a string, violating the requirements of the \textit{System Prompt} (Fig.~\ref{fig:system-prompt}).
+    \item \textbf{Value Mismatch}.
+    If an expected value is known, it is compared with the actual result.
+    The mismatch generates an error.
+\end{itemize}
+
 \subsection{Paragraph with generated expression}
+\label{subsec:paragraph-with-generated-expression}
 Figure \ref{fig:fluid-example-paragraph} shows the Fluid code of a Paragraph with Fluid expression generated by the authoring assistant.

### 2025-05-23 sec/introduction.tex
Tweaks to Fig.1.

@@
         &%check with v-space
-        \vspace{-5pt}
+        \vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         &
-        \vspace{-5pt}
+        \vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         Rounding & ~                & ~                             \\
-        Cardinals \& Multiplicative & ~                & ~                             \\
+        Cardinal & three                & ~                             \\
+        Multiplicative & three times                & ~                             \\
         \rowcolor{gray!20}
@@
         & The average methane emissions for the year 2030 is \hl{13.51} &
-        \vspace{-5pt}
+        \vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         Min/Max                          & The Energy Sector recorded its highest methane emissions in \hl{2030}             &
-        \vspace{-5pt}
+        \vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         \end{lstlisting} \\                             \\
-        Ranks &
+        Rank &
         3-layer stacked CNN gives an accuracy of 81.46\%, which is the \hl{lowest} compared with BiLSTM, and S-LSTM  &
-        \vspace{-5pt}
+        \vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         The total methane emissions for the year 2030 is \hl{37.74} for Agriculture &
-        \vspace{-5pt}
+        \vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
     \end{tabular}
-    \caption{Example expressions generated by the Authoring Assistant, classified by type. In green, the expression is already tested.}
+    \caption{Quantitative/semi-quantitative natural language forms considered in this paper}
     \label{tab:fluid_examples}

### 2025-05-23 sec/introduction.tex
Improvements to Table 1.

@@
         \end{lstlisting}  \\
-        Rounding & ~                & ~                             \\
-        Cardinal & three                & ~                             \\
-        Multiplicative & three times                & ~                             \\
+        Rounding* & about 2 degrees warmer                & ~                             \\
+        Cardinal* & three                & ~                             \\
+        Multiplicative* & three times                & ~                             \\
         \rowcolor{gray!20}
@@
         \end{lstlisting} \\
-        Graded adjective*            & ~                & ~                             \\
+        Graded adjective*            & \hl{extremely likely} to be exceeded in the 20th century               & ~                             \\
         Reference to visual element*                            & ~                & ~                             \\

### 2025-05-28 src/main/java/explorableviz/authoringassistant/AuthoringAssistant.java
[refactoring] added evaluateExpression function

@@ -6,6 +6,7 @@ import it.unisa.cluelab.lllm.llm.prompt.PromptList;
 import kotlin.Pair;
 import org.json.JSONObject;
 
+import java.io.IOException;
 import java.lang.reflect.InvocationTargetException;
 import java.util.ArrayList;
 import java.util.List;
@@ -64,15 +65,12 @@ public class AuthoringAssistant {
             Expression candidate = llm.evaluate(sessionPrompts, "");
             //Check each generated expressions
             for (Map<String, String> dataset : subProgram.getTest_datasets()) {
-                final FluidCLI fluidCLI = new FluidCLI(subProgram.getDatasets(), subProgram.getImports());
-                logger.info(STR."Received response: \{candidate.getExpr()}");
-                //Compute expected value with the expected expression
-                writeFluidFiles(Settings.getFluidTempFolder(), Program.fluidFileName, expected.getExpr(), subProgram.getDatasets(), dataset, subProgram.getImports(), subProgram.get_loadedImports(), subProgram.getCode());
-                String expectedValue = fluidCLI.evaluate(subProgram.getFluidFileName());
 
-                //Compute the value with the candidate expression
-                writeFluidFiles(Settings.getFluidTempFolder(), Program.fluidFileName, candidate.getExpr(), subProgram.getDatasets(), dataset, subProgram.getImports(), subProgram.get_loadedImports(), subProgram.getCode());
-                Optional<String> error = Program.validate(fluidCLI.evaluate(subProgram.getFluidFileName()), new Expression(expected.getExpr(), expectedValue));
+                logger.info(STR."Received response: \{candidate.getExpr()}");
+
+                Optional<String> error = Program.validate(
+                        evaluateExpression(subProgram, dataset, candidate),
+                        new Expression(expected.getExpr(), evaluateExpression(subProgram, dataset, expected), expected.getCategory()));
 
                 if (error.isPresent()) {
                     sessionPrompts.addAssistantPrompt(candidate.getExpr() == null ? "NULL" : candidate.getExpr());
@@ -88,6 +86,12 @@ public class AuthoringAssistant {
         }
         logger.warning(STR."Expression validation failed after \{limit} attempts");
         return new QueryResult(null, expected, attempts, System.currentTimeMillis() - start);
+    }
+
+    private static String evaluateExpression(Program p, Map<String, String> dataset, Expression expression) throws IOException {
[…]

### 2025-05-28 sec/evaluation.tex
[charts] created new chart and inserted into the paper

@@
 \subsection{Success Rate by Testcase}
+To evaluate the system we used a sample of the Scigen dataset \cite{scigen_dataset_2021}.
+Figure \ref{fig:success_rate_by_test_case} shows the results obtained for each test case.
+
 \label{subsec:success_rate_by_test_case}
@@
 \end{figure}
+
+We also aggregated the generated expression in three categories: aggregation, trends, quantitative (See Table \ref{tab:fluid_examples}).
+Figure \ref{fig:success_rate_by_category} shows the results obtained in generating expression by categories.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.95\linewidth]{fig/success_rate_by_category}
+    \caption{Success rate by testcases}\label{fig:success_rate_by_category}
+\end{figure}

### 2025-05-28 script/chart-generation.py

@@ -3,6 +3,7 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 import glob
 import os
+import json
 
 def get_latest_csv(folder_path):
     csv_files = glob.glob(os.path.join(folder_path, "*.csv"))
@@ -51,6 +52,68 @@ def generate_summary_test_case_plot(df, plot):
     plt.savefig("fig/per_row_checkmarks.png")
     plt.close()
 
+def generate_aggregated_plot(df, plot):
+    summary = (
+        df.groupby("expression-type")
+        .agg(
+            success_rate=("success", "mean"),
+            std_dev=("success", "std"),
+            avg_attempts=("attempts", "mean"),
+            count=("success", "size")
+        )
+        .reset_index()
+    )
+    with open("settings.json", "r") as f:
+        config = json.load(f)
+    limit = config.get("agent-limit", "N/A")  # Use "N/A" if limit not found
+    # Create custom x labels: "category (count)"
+    summary["label"] = summary["expression-type"] + " (" + summary["count"].astype(str) + ")"
+
+    # Sort by success_rate
+    summary = summary.sort_values("success_rate", ascending=False).reset_index(drop=True)
+
+    # Set style
+    sns.set(style="whitegrid")
+
+    # Plot
+    plt.figure(figsize=(12, 6))
+    ax = sns.barplot(
+        data=summary,
+        x="label",
+        y="success_rate",
+        palette="Blues_d",
+        width=0.3  # Reduce bar width
+    )
+
+    # Add error bars manually
+    ax.errorbar(
+        x=range(len(summary)),
+        y=summary["success_rate"],
+        yerr=summary["std_dev"],
+        fmt='none',
+        ecolor='orange',
+        capsize=5,
+        linewidth=1.3
+    )
+
+    # Annotate bars
+    for i, row in summary.iterrows():
+        ax.text(i, row.success_rate + 0.01, f"{row.success_rate:.2f}", ha='center', va='bottom', fontsize=9)
+
+    # Labels and formatting
+    plt.title(f"Success Rate by Linguistic Category (Max attempts = {limit})", fontsize=14)
+    plt.xlabel("Linguistic Category (Count)", fontsize=12)
+    plt.ylabel("Success Rate", fontsize=12)
+    plt.ylim(-1, 1.5)
+    plt.xticks(rotation=45, ha="right")
+
+    # Final layout
+    plt.tight_layout()
[…]

### 2025-05-28 src/main/java/explorableviz/authoringassistant/paragraph/Expression.java
[update] manage label for enum

@@ -3,18 +3,18 @@ package explorableviz.authoringassistant.paragraph;
 public class Expression extends TextFragment {
 
     private final String expr;
-    public Expression(String expr, String value) {
+    private final ExpressionCategory category;
+    public Expression(String expr, String value, ExpressionCategory category) {
         super(value);
         this.expr = expr;
+        this.category = category;
     }
 
     public String getExpr() {
         return expr;
     }
 
-    @Override
-    public Expression clone() {
-        return new Expression(expr, getValue());
+    public ExpressionCategory getCategory() {
+        return category;
     }
-
 }

### 2025-05-28 src/main/java/explorableviz/authoringassistant/Program.java
parse category from json

@@ -1,13 +1,13 @@
 package explorableviz.authoringassistant;
 
 import com.fasterxml.jackson.databind.ObjectMapper;
+import explorableviz.authoringassistant.paragraph.ExpressionCategory;
 import explorableviz.authoringassistant.paragraph.Expression;
 import explorableviz.authoringassistant.paragraph.Literal;
 import explorableviz.authoringassistant.paragraph.Paragraph;
 import explorableviz.authoringassistant.variable.ValueOptions;
 import explorableviz.authoringassistant.variable.Variables;
 import kotlin.Pair;
-import org.checkerframework.checker.units.qual.A;
 import org.json.JSONArray;
 import org.json.JSONObject;
 
@@ -217,7 +217,7 @@ public class Program {
                 String expression = json_paragraph.getJSONObject(i).getString("expression");
                 writeFluidFiles(Settings.getFluidTempFolder(), fluidFileName, expression, datasetMapping, loadDatasetsFiles(datasetMapping, testVariables), imports, loadImports(imports), code);
                 String commandLineResult = new FluidCLI(datasetMapping, imports).evaluate(fluidFileName);
-                Expression candidate = new Expression(expression, extractValue(commandLineResult));
+                Expression candidate = new Expression(expression, extractValue(commandLineResult), ExpressionCategory.of(json_paragraph.getJSONObject(i).getString("category")));
                 paragraph.add(candidate);
                 validate(commandLineResult, candidate).ifPresent(value -> {
                     throw new RuntimeException(STR."[testCaseFile=\{casePath}] Invalid test exception\{value}");
@@ -338,6 +338,7 @@ public class Program {
             ObjectMapper objectMapper = new ObjectMapper();
             Object jsonObject = objectMapper.readValue(spec.toString(), Object.class);
             file.write(objectMapper.writerWithDefaultPrettyPrinter().writeValueAsString(jsonObject));
+            file.write("\n");
             file.flush();
         } catch (IOException e) {
             e.printStackTrace();

### 2025-05-28 src/main/java/explorableviz/authoringassistant/Main.java
[update] category to the report

@@ -55,7 +55,7 @@ public class Main {
         try (PrintWriter out = new PrintWriter(new FileOutputStream(STR."\{Settings.getLogFolder()}/log_\{System.currentTimeMillis()}.csv"))) {
             String[] headers = {
                     "test-case", "llm-agent", "temperature", "num-token", "in-context-learning-size",
-                    "attempts", "result", "generated-expression", "expected-value", "duration(ms)"
+                    "attempts", "result", "expression-type", "generated-expression", "expected-value", "duration(ms)"
             };
             out.println(String.join(";", headers));
             String content = results.stream()
@@ -71,7 +71,8 @@ public class Main {
                                 //program.getParagraph().toFluidSyntax(),
                                 String.valueOf(queryResult.attempt()),
                                 queryResult.response() != null ? "OK" : "KO",
-                                String.valueOf(queryResult.response() != null ? "generated" : "NULL"),
+                                queryResult.expected().getCategory().label,
+                                queryResult.response() != null ? "generated" : "NULL",
                                 //queryResult.expected().getExpr(),
                                 queryResult.expected().getValue(),
                                 String.valueOf(queryResult.duration())

### 2025-05-30 src/main/java/explorableviz/authoringassistant/Main.java
[refactory] change property name

@@ -89,7 +89,7 @@ public class Main {
 
     private static ArrayList<Pair<Program, QueryResult>> execute(InContextLearning inContextLearning, String agent, List<Program> programs) throws Exception {
         final ArrayList<Pair<Program, QueryResult>> results = new ArrayList<>();
-        for(int k = 0; k < Settings.numTestExecution(); k++)
+        for(int k = 0; k < Settings.numTestRuns(); k++)
         {
             int programId = 0;
             for (Program program : programs) {

### 2025-05-30 src/main/java/explorableviz/authoringassistant/Settings.java

@@ -6,7 +6,6 @@ import java.io.File;
 import java.io.IOException;
 import java.nio.file.Files;
 import java.nio.file.Paths;
-import java.util.Optional;
 
 public class Settings {
 
@@ -96,12 +95,8 @@ public class Settings {
         return getSettings().getInt("num-learning-case-to-generate");
     }
 
-    public static int numTestExecution() {
-        return getSettings().getInt("num-test-execution");
-    }
-
-    public static Optional<Integer> getNumQueryToExecute() {
-        return (getSettings().has("num-query-to-execute") && !getSettings().isNull("num-query-to-execute")) ? Optional.of(getSettings().getInt("num-query-to-execute")) : Optional.empty();
+    public static int numTestRuns() {
+        return getSettings().getInt("num-test-runs");
     }
 
     public static boolean isEditorLoopEnabled() {

### 2025-05-30 sec/evaluation.tex

@@
 To evaluate the system we used a sample of the Scigen dataset \cite{scigen_dataset_2021}.
-Figure \ref{fig:success_rate_by_test_case} shows the results obtained for each test case.
 
-\label{subsec:success_rate_by_test_case}
-
-\begin{figure}[!h]
-    \centering
-    \includegraphics[width=0.95\linewidth]{fig/success_rate_by_test_case}
-    \caption{Success rate by testcases}\label{fig:success_rate_by_test_case}
-\end{figure}
-
-We also aggregated the generated expression in three categories: aggregation, trends, quantitative (See Table \ref{tab:fluid_examples}).
+We aggregated the generated expression in three categories: aggregation, trends, quantitative (See Table \ref{tab:fluid_examples}).
 Figure \ref{fig:success_rate_by_category} shows the results obtained in generating expression by categories.

### 2025-06-05 sec/introduction.tex
[update] update error definitions and margin table

--- 2025-05-22 sec/introduction.tex
@@
 \section{Introduction}
+
+\lstset{
+    aboveskip=-2.3pt,    % riduce lo spazio sopra
+    belowskip=-2pt,    % riduce lo spazio sotto
+    xleftmargin=0pt,   % riduce il margine sinistro
+    xrightmargin=0pt   % riduce il margine destro
+}
 
@@
     \footnotesize
-    \renewcommand{\arraystretch}{1.1}
+    \renewcommand{\arraystretch}{0.9}
     \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6cm}}
@@
         &%check with v-space
-        \vspace{-5pt}
+        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         &
-        \vspace{-5pt}
+        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         & The average methane emissions for the year 2030 is \hl{13.51} &
-        \vspace{-5pt}
+        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         Min/Max                          & The Energy Sector recorded its highest methane emissions in \hl{2030}             &
-        \vspace{-5pt}
+        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
 in maxEntry.year
-        \end{lstlisting} \\                             \\
+        \end{lstlisting} \\
         Ranks &
         3-layer stacked CNN gives an accuracy of 81.46\%, which is the \hl{lowest} compared with BiLSTM, and S-LSTM  &
-        \vspace{-5pt}
+        %\vspace{-10pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         The total methane emissions for the year 2030 is \hl{37.74} for Agriculture &
-        \vspace{-5pt}
+        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         & The training time per epoch \hl{growing} from 67 seconds to 106 seconds. &
-        \vspace{-5pt}
+        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
     \label{tab:fluid_examples}
-    \renewcommand{\arraystretch}{1.0}
 \end{table}

### 2025-06-06 sec/introduction.tex
[update] padding for listing

@@
 \lstset{
-    aboveskip=-2.3pt,    % riduce lo spazio sopra
-    belowskip=-2pt,    % riduce lo spazio sotto
-    xleftmargin=0pt,   % riduce il margine sinistro
-    xrightmargin=0pt   % riduce il margine destro
+    aboveskip=-7pt,
+    belowskip=-8.5pt,
+    xleftmargin=0pt,
+    xrightmargin=0pt
 }
@@
     \footnotesize
-    \renewcommand{\arraystretch}{0.9}
+    \renewcommand{\arraystretch}{1.0}
     \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6cm}}
@@
         \textbf{Type}                & \textbf{Example} & \textbf{Generated Expression} \\
-        \midrule
         \rowcolor{gray!20}
@@
         3-layer stacked CNN gives an accuracy of 81.46\%, which is the \hl{lowest} compared with BiLSTM, and S-LSTM  &
-        %\vspace{-10pt}
+%        \vspace{-15pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
     (insertionSort cmpTime tableData)
-in rankLabel "lowest" pos
-        \end{lstlisting} \\
+in rankLabel "lowest" pos \end{lstlisting} \\\hline
         Total &

### 2025-06-06 script/chart-generation.py
[update] manage categories in chart

@@ -54,19 +54,24 @@ def generate_summary_test_case_plot(df, plot):
     plt.close()
 
 def generate_aggregated_plot(df, plot):
+    df['expression-type'] = df['expression-type'].str.strip('[]').str.split(',')
+    df_exploded = df.explode('expression-type')
+    df_exploded['expression-type'] = df_exploded['expression-type'].str.strip()
+    
     summary = (
-        df.groupby("expression-type")
+        df_exploded.groupby("expression-type")
         .agg(
             success_rate=("success", "mean"),
             std_dev=("success", "std"),
-            avg_attempts=("attempts", "mean"),
             count=("success", "size")
         )
         .reset_index()
     )
+    
     with open("settings.json", "r") as f:
         config = json.load(f)
     limit = config.get("agent-limit", "N/A")  # Use "N/A" if limit not found
+    
     # Create custom x labels: "category (count)"
     summary["label"] = summary["expression-type"] + " (" + summary["count"].astype(str) + ")"
 
@@ -75,8 +80,12 @@ def generate_aggregated_plot(df, plot):
     p = summary["success_rate"]
     n = summary["count"]
     ci = z * np.sqrt((p * (1 - p)) / n)
-    summary["ci"] = ci
-
+    
+    # IC limit error-bar.
+    lower_bound = np.maximum(0, p - ci)
+    upper_bound = np.minimum(1, p + ci)
+    yerr = np.array([p - lower_bound, upper_bound - p])
+    
     # Sort by success_rate
     summary = summary.sort_values("success_rate", ascending=False).reset_index(drop=True)
 
@@ -93,11 +102,11 @@ def generate_aggregated_plot(df, plot):
         width=0.7  # Reduce bar width
     )
 
-    # Add error bars manually
+    # Add error bars manually with limited CI
     ax.errorbar(
         x=range(len(summary)),
         y=summary["success_rate"],
-        yerr=summary["ci"],
+        yerr=yerr,
         fmt='none',
         ecolor='orange',
         capsize=5,

### 2025-06-06 sec/introduction.tex
cleanup

@@
     \footnotesize
-    \renewcommand{\arraystretch}{1.0}
+    \renewcommand{\arraystretch}{1.2}
     \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6cm}}
@@
         & the training time per epoch growing from \hl{67} seconds to 106 seconds.
-        &%check with v-space
-        %\vspace{-8pt}
+        &
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         &
-        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         Rounding & ~                & ~                             \\
-        Cardinals \& Multiplicative & ~                & ~                             \\
+        Cardinal  & ~                & ~                             \\
+        Multiplicative  & ~                & ~                             \\
         \rowcolor{gray!20}
@@
         & The average methane emissions for the year 2030 is \hl{13.51} &
-        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         Min/Max                          & The Energy Sector recorded its highest methane emissions in \hl{2030}             &
-        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
         \end{lstlisting} \\
-        Ranks &
+        Rank &
         3-layer stacked CNN gives an accuracy of 81.46\%, which is the \hl{lowest} compared with BiLSTM, and S-LSTM  &
-%        \vspace{-15pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
     (insertionSort cmpTime tableData)
-in rankLabel "lowest" pos \end{lstlisting} \\\hline
+in rankLabel "lowest" pos \end{lstlisting} \\
         Total &
         The total methane emissions for the year 2030 is \hl{37.74} for Agriculture &
-        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]
@@
 
-        Comparisons
+        Comparison
         & The training time per epoch \hl{growing} from 67 seconds to 106 seconds. &
-        %\vspace{-8pt}
         \begin{lstlisting}[language=Fluid,numbers=none]

### 2025-06-06 script/chart-generation.py
[new] chart for num of categories

@@ -129,6 +129,71 @@ def generate_aggregated_plot(df, plot):
     plt.savefig("fig/success_rate_by_category.png")
     plt.close()
 
+def generate_success_rate_by_category_count(df, plot):
+    # Count the number of categories for each test
+    df['category_count'] = df['expression-type'].str.len()
+    
+    # Calculate summary for each category count
+    summary = (
+        df.groupby("category_count")
+        .agg(
+            success_rate=("success", "mean"),
+            count=("success", "size")
+        )
+        .reset_index()
+    )
+    
+    # Calculate confidence intervals
+    z = 1.96
+    p = summary["success_rate"]
+    n = summary["count"]
+    ci = z * np.sqrt((p * (1 - p)) / n)
+    
+    # Limit confidence intervals between 0 and 1
+    lower_bound = np.maximum(0, p - ci)
+    upper_bound = np.minimum(1, p + ci)
+    yerr = np.array([p - lower_bound, upper_bound - p])
+    
+    # Create custom x labels: "number_of_categories (count)"
+    summary["label"] = summary["category_count"].astype(str) + " (" + summary["count"].astype(str) + ")"
+    
+    # Plot
+    plt.figure(figsize=(6, 6))
+    ax = sns.barplot(
+        data=summary,
+        x="label",
+        y="success_rate",
+        palette="Blues_d",
+        width=0.7
+    )
+    
+    # Add error bars with limited confidence intervals
+    ax.errorbar(
+        x=range(1),
+        y=summary["success_rate"],
+        yerr=yerr,
+        fmt='none',
+        ecolor='orange',
+        capsize=5,
+        linewidth=1.3
+    )
+    
+    # Annotate bars with success rate values
+    for i, row in summary.iterrows():
+        ax.text(i, row.success_rate + 0.01, f"{row.success_rate:.2f}", ha='center', va='bottom', fontsize=9)
+    
+    # Labels and formatting
+    plt.title("Success Rate by Number of Categories", fontsize=14)
+    plt.xlabel("Number of Categories (Number of Examples)", fontsize=12)
+    plt.ylabel("Average Success Rate (CI)", fontsize=12)
+    plt.ylim(0, 1.2)
+    plt.xticks(rotation=45, ha="right")
+    
+    # Final layout
+    plt.tight_layout()
+    plt.savefig("fig/success_rate_by_category_count.png")
+    plt.close()
+
[…]

### 2025-06-06 script/chart-generation.py
[new] chart for num of categories

@@ -169,7 +169,7 @@ def generate_success_rate_by_category_count(df, plot):
     
     # Add error bars with limited confidence intervals
     ax.errorbar(
-        x=range(1),
+        x=range(len(summary)),
         y=summary["success_rate"],
         yerr=yerr,
         fmt='none',

### 2025-06-06 sec/introduction.tex

@@
         \end{lstlisting}  \\
-        Rounding & ~                & ~                             \\
-        Cardinal  & ~                & ~                             \\
-        Multiplicative  & ~                & ~                             \\
+        Rounding* & about 2 degrees warmer                & ~                             \\
+        Cardinal* & three                & ~                             \\
+        Multiplicative* & three times                & ~                             \\
         \rowcolor{gray!20}
@@
 in maxEntry.year
-        \end{lstlisting} \\
+        \end{lstlisting} \\                             \\
         Rank &

### 2025-06-30 sec/authoring-workflow.tex

@@
 \begin{itemize}
-    \item \textbf{Interpreter Error}.
-    Syntax errors due to incorrect use of the language (e.g., undeclared functions, invalid syntax, etc.).
-    \item \textbf{String Error}.
-    The expression does not return a string, violating the requirements of the \textit{System Prompt} (Fig.~\ref{fig:system-prompt}).
-    \item \textbf{Value Mismatch}.
-    If an expected value is known, it is compared with the actual result.
-    The mismatch generates an error.
+    \item \textbf{Invalid Program}.
+    The generated expression is executed by the Fluid interpreter.
+    If the execution fails due to issues such as undeclared functions, syntax errors, or other interpreter-level problems, it results in an \textit{Invalid Program} error.
+    \item \textbf{Invalid Output Type}.
+    According to the \textit{System Prompt} (Fig.~\ref{fig:system-prompt}), the expression is expected to return a string value.
+    If the returned value is of a different type, an \textit{Invalid Output Type} error is raised.
+    \item \textbf{Invalid Output Value}.
+    The output produced by the generated expression is compared against the highlighted target text.
+    If the values do not match, an \textit{Invalid Output Value} error is reported.
 \end{itemize}
@@
 \label{subsec:paragraph-with-generated-expression}
-Figure \ref{fig:fluid-example-paragraph} shows the Fluid code of a Paragraph with Fluid expression generated by the authoring assistant.
+Figure \ref{fig:fluid-example-paragraph} shows the Fluid code of a Paragraph with the expression generated by the authoring assistant.
 Figure \ref{fig:fluid-scigen} shows the Fluid code written for the scigen datasets and examples.

### 2025-06-30 sec/evaluation.tex

@@
 
-\subsection{Success Rate by Testcase}
+\subsection{Success Rate by Category}
 To evaluate the system we used a sample of the Scigen dataset \cite{scigen_dataset_2021}.
@@
     \centering
-    \includegraphics[width=0.95\linewidth]{fig/success_rate_by_category}
+    \includegraphics[width=0.55\linewidth]{fig/success_rate_by_category}
     \caption{Success rate by testcases}\label{fig:success_rate_by_category}
 \end{figure}
+
+\subsection{Success Rate Category Count}
+
+There are expressions which fall in more than one category.
+Figure \ref{fig:success_rate_by_category_count} shows the results obtained in generating expression by the number of categories.
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=0.55\linewidth]{fig/success_rate_by_category_count}
+    \caption{Success rate by testcases}\label{fig:success_rate_by_category_count}
+\end{figure}

### 2025-08-05 src/main/java/explorableviz/authoringassistant/llm/DeepSeekAgent.java
[new] add deepseek agent

@@ -1,0 +1,18 @@
+package explorableviz.authoringassistant.llm;
+
+import explorableviz.authoringassistant.paragraph.Expression;
+import it.unisa.cluelab.lllm.llm.agents.generic.OLLAMAEvaluatorAgent;
+import org.json.JSONObject;
+
+public class DeepSeekAgent extends OLLAMAEvaluatorAgent<Expression> {
+
+    public DeepSeekAgent(JSONObject setttings) {
+        super(setttings);
+        setModel("deepseek-coder:33b");
+    }
+
+    @Override
+    public Expression parse(String s) {
+        return new Expression(s, null, null);
+    }
+}

### 2025-08-05 sec/related-work.tex
update literature

@@
 \begin{itemize}
-\item Transparent visualisations \cite{perera22,bond24}
+    \item Transparent visualisations \cite{perera22,bond25,psallidas18smoke}
+    \item recent advances in interpretable language understanding, highlighting the shift from fine-tuning to prompt-based learning in NLP and presenting research on improving the transparency of neural models through interpretable classifiers, uncertainty analysis, and explainable systems.
+\cite{Yulan_2023}
+    \item Quantity extraction in scientific text \cite{bolucu-etal-2023-investigating}
+    \item Scientific fact checking \cite{abu-ahmad-etal-2025-climatecheck-shared}
+
 \end{itemize}

### 2025-08-21 sec/related-work.tex
Import related-work changes from negative-cases.

@@
 
-\todo{flesh out}
+\paragraph{Interpretable Natural Language Processing}
 
@@
     \item Transparent visualisations \cite{perera22,bond25,psallidas18smoke}
-    \item recent advances in interpretable language understanding, highlighting the shift from fine-tuning to prompt-based learning in NLP and presenting research on improving the transparency of neural models through interpretable classifiers, uncertainty analysis, and explainable systems.
+    \item Recent advances in interpretable language understanding, highlighting the shift from fine-tuning to prompt-based learning in NLP and presenting research on improving the transparency of neural models through interpretable classifiers, uncertainty analysis, and explainable systems.
 \cite{Yulan_2023}
@@
     \item Scientific fact checking \cite{abu-ahmad-etal-2025-climatecheck-shared}
-
 \end{itemize}

### 2025-08-22 sec/evaluation.tex
Figures side by side.

--- 2025-05-30 sec/evaluation.tex
@@
 
-\begin{figure}[!h]
+\begin{figure}
     \centering
-    \includegraphics[width=0.95\linewidth]{fig/success_rate_by_category}
-    \caption{Success rate by testcases}\label{fig:success_rate_by_category}
+    \begin{subfigure}{0.48\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{fig/success_rate_by_category}
+        \caption{Success rate by testcases}
+        \label{fig:success_rate_by_category}
+    \end{subfigure}\hfill
+    \begin{subfigure}{0.48\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{fig/success_rate_by_category_count}
+        \caption{Success rate by category count}
+        \label{fig:success_rate_by_category_count}
+    \end{subfigure}
+    \caption{Comparison of success rates}
+    \label{fig:success_rate_comparison}
 \end{figure}

### 2025-08-22 sec/related-work.tex
Beramono.
Start on related work categories.

--- 2025-04-11 sec/related-work.tex
@@
 
-\todo{flesh out}
+\paragraph{Interpretable NLP.}
 
-\begin{itemize}
-\item Transparent visualisations \cite{perera22,bond24}
-\end{itemize}
+Recent advances in interpretable language understanding, highlighting the shift from fine-tuning to
+prompt-based learning in NLP and presenting research on improving the transparency of neural models through
+interpretable classifiers, uncertainty analysis, and explainable systems. \cite{Yulan_2023}
+
+\paragraph{AI interpretation of scientific text.}
+
+For example quantity extraction~\cite{bolucu-etal-2023-investigating}, scientific fact-checking
+\cite{abu-ahmad-etal-2025-climatecheck-shared}.
+
+\paragraph{Linguistic competences of LLMs.}
+
+\paragraph{Transparent programming languages.}
+
+\cite{perera22,bond25,psallidas18smoke}

### 2025-08-22 sec/related-work.tex
Bib fixes.

@@
 prompt-based learning in NLP and presenting research on improving the transparency of neural models through
-interpretable classifiers, uncertainty analysis, and explainable systems. \cite{Yulan_2023}
+interpretable classifiers, uncertainty analysis, and explainable systems \cite{yulan23}.
 
@@
 
-For example quantity extraction~\cite{bolucu-etal-2023-investigating}, scientific fact-checking
-\cite{abu-ahmad-etal-2025-climatecheck-shared}.
+For example quantity extraction~\cite{bolucu23}, scientific fact-checking
+\cite{abu-ahmad25}.
 
@@
 
-\cite{perera22,bond25,psallidas18smoke}
+\cite{atkey25,perera22,bond25,psallidas18smoke}

### 2025-08-22 sec/related-work.tex
WIP.

--- 2025-08-22 sec/related-work.tex
@@
 
-Recent advances in interpretable language understanding, highlighting the shift from fine-tuning to
-prompt-based learning in NLP and presenting research on improving the transparency of neural models through
-interpretable classifiers, uncertainty analysis, and explainable systems. \cite{Yulan_2023}
+As we saw earlier \todo{ref}, scientific texts routinely make use of comparatives like ``faster'' while
+leaving one of the argument slots implicit, with the context determining the omitted referent. LLMs
+demonstrate considerable competence in resolving these and other more syntatic forms of anaphor such as
+pronouns~\citep{zhu25}, but the resolved referent itself -- concretely, what was being referred to -- remains
+implicit. Interpretable NLP is a recent research direction which aims to support comprehension (and
+production) of text in a more explicit and transparent way~\citep{yulan23}. By generating code that formalises
+the interpretation of a comparative like ``faster'', our approach also makes these implicit references
+explicit; a richer version of our approach would allow the user to explore the linguistic interpretation as
+well. \todo{WIP}
 
-\paragraph{AI interpretation of scientific text.}
+\paragraph{Use of LLMs to interpret scientific text.}
 
-For example quantity extraction~\cite{bolucu-etal-2023-investigating}, scientific fact-checking
-\cite{abu-ahmad-etal-2025-climatecheck-shared}.
-
-\paragraph{Linguistic competences of LLMs.}
+The ability of LLMs to interpret natural language in scientific contexts has been studied extensively, for example quantity extraction~\citep{bolucu23}, scientific fact-checking~\citep{abu-ahmad25}.
 

### 2025-08-22 src/main/java/explorableviz/authoringassistant/Program.java
Drop 'file' from generated .json; replace by argument to loadFigure.

@@ -335,7 +335,8 @@ public class Program {
 
         /* spec generation */
         JSONObject spec = new JSONObject();
-        spec.put("fluidSrcPath", new JSONArray("[\"../fluid\"]"));
+        String fluidSrcPath = "../fluid";
+        spec.put("fluidSrcPath", new JSONArray(STR."[\"\{fluidSrcPath}\"]"));
         spec.put("datasets", new JSONArray());
         spec.put("imports", new JSONArray());
 
@@ -349,7 +350,6 @@ public class Program {
         imports.forEach(_import -> {
             spec.getJSONArray("imports").put(_import);
         });
-        spec.put("file", Path.of(this.testCaseFileName).getFileName());
         spec.put("inputs", new JSONArray("[\"tableData\"]"));
         try (FileWriter file = new FileWriter(STR."\{sitePath}/spec.json")) {
             ObjectMapper objectMapper = new ObjectMapper();
@@ -364,6 +364,7 @@ public class Program {
         String html = new String(Files.readAllBytes(Paths.get(new File(STR."\{path}/template.html").toURI())));
         html = html.replaceAll("##TITLE##", String.valueOf(Path.of(this.testCaseFileName).getParent().getFileName()));
         html = html.replaceAll("##TEST_NAME##", String.valueOf(Path.of(this.testCaseFileName).getFileName()));
+        html = html.replaceAll("##FLUID_FILE##", STR."\"\{fluidSrcPath}/\{Path.of(this.testCaseFileName).getFileName()}.fld\"");
         try (FileWriter file = new FileWriter(STR."\{sitePath}/index.html")) {
             file.write(html);
             file.flush();
@@ -387,7 +388,7 @@ public class Program {
                         }
                     });
         } catch (IOException e) {
-            System.err.println(STR."Errore during cleanup of website folder: \{e.getMessage()}");
+            System.err.println(STR."Error during clean of website folder: \{e.getMessage()}");
         }
     }
 

### 2025-09-03 paper/sec/authoring-workflow.tex

--- 2025-06-30 sec/authoring-workflow.tex
@@
 
-\begin{figure}
+\begin{figure}[ht]
     \centering
-    \includegraphics[width=0.95\linewidth]{fig/authoring-assistant-architecture.png}
-    \caption{AI-assisted authoring workflow}\label{fig:architecture}
+    \begin{subfigure}{0.48\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{fig/entire-workflow}
+        \caption{Workflow}
+        \label{fig:architecture}
+    \end{subfigure}\hfill
+    \begin{subfigure}{0.48\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{fig/data-flow}
+        \caption{Data flow}
+        \label{fig:dataflow}
+    \end{subfigure}
+    \caption{AI-assisted authoring process: workflow (a) and data flow (b).}
+    \label{fig:overall}
 \end{figure}
@@
         left=6pt, right=6pt, top=4pt, bottom=4pt]
-        \VerbatimInput[fontsize=\small, formatcom=\normalfont]{system-prompt/system-prompt.txt}
+        \VerbatimInput[fontsize=\small, formatcom=\normalfont]{../system-prompt/system-prompt.txt}
     \end{tcolorbox}
@@
     \small
-    {\lstinputlisting[language=Fluid]{website/authoring-assistant/fluid/1805.02474v1-10.fld}}
+    {\lstinputlisting[language=Fluid]{../website/authoring-assistant/fluid/1805.02474v1-10.fld}}
     \vspace{-0.5em}
@@
     \small
-    {\lstinputlisting[language=Fluid]{fluid-common/scigen.fld}}
-    \vspace{-0.5em}
+    {\lstinputlisting[language=Fluid]{../fluid-common/scigen.fld}}
     \caption{Fluid library written for scigen}

### 2025-09-04 paper/sec/authoring-workflow.tex
description of the workflow and update flow image

@@
 \label{sec:authoring-workflow}
+
+Figure~\ref{fig:overall-architecture} shows the system architecture and an example flow.
+The system is composed of two agents:
+
+\begin{itemize}
+    \item \textbf{SuggestionAgent}.
+Based on an LLM, the SuggestionAgent receives as input the initial text and the reference data, and returns the elements in the text that can be replaced with expressions in the Fluid language.
+    \item \textbf{InterpretationAgent}.
+Also based on an LLM, the InterpretationAgent receives as input a selection (either obtained from the SuggestionAgent or manually defined by the author) and generates an interpretation of the text as a Fluid expression.
+\end{itemize}
+
+Figure~\ref{fig:architecture} shows the overall architecture of the system.
+The author initially provides the text and accompanying data.
+The SuggestionAgent analyses the input and identifies candidate fragments of natural language which could be replaced by Fluid expressions.
+The user highlightes a fragment of interest, which is then sent to the InterpretationAgent, which generates a corresponding candidate expression.
+Internally the InterpretationAgent uses a compiler-in-the-loop validation process to ensure the expression is well-formed.
+The user is then able to accept the expression (in which case it is incorporated into the program) or reject and leave the text uninterpreted.
 
@@
     \caption{AI-assisted authoring process: workflow (a) and data flow (b).}
-    \label{fig:overall}
+    \label{fig:overall-architecture}
 \end{figure}

### 2025-09-11 paper/sec/authoring-workflow.tex
[update] interpretation examples wf

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/entire-workflow}
+    \caption{Workflow}\label{fig:architecture}
+\end{figure}
+
 Figure~\ref{fig:architecture} shows the overall architecture of the system.
@@
     \centering
-    \begin{subfigure}{0.48\linewidth}
+    \begin{subfigure}{0.50\linewidth}
         \centering
-        \includegraphics[width=\linewidth]{fig/entire-workflow}
-        \caption{Workflow}
-        \label{fig:architecture}
+        \includegraphics[width=\linewidth]{fig/data-flow-correct}
+        \caption{Author Acceptance Dataflow}
+        \label{fig:data-flow-correct}
     \end{subfigure}\hfill
-    \begin{subfigure}{0.48\linewidth}
+    \begin{subfigure}{0.50\linewidth}
         \centering
-        \includegraphics[width=\linewidth]{fig/data-flow}
-        \caption{Data flow}
-        \label{fig:dataflow}
+        \includegraphics[width=\linewidth]{fig/data-flow-error}
+        \caption{Author Rejection Dataflow}
+        \label{fig:dataflow-error}
     \end{subfigure}
-    \caption{AI-assisted authoring process: workflow (a) and data flow (b).}
+    \caption{Two workflows of the Interpretation Agent.
+In Fig.
+\ref{fig:data-flow-correct}, the author accepts the generated expression after verifying its correctness online.
+In Fig.
+\ref{fig:dataflow-error}, the author identifies an error (the expression mentions accuracy, while the article highlights timing) and rejects it.}
     \label{fig:overall-architecture}
@@
 Figure \ref{fig:fluid-scigen} shows the Fluid code written for the scigen datasets and examples.
-\begin{figure}[h]
-    \small
-    {\lstinputlisting[language=Fluid]{../website/authoring-assistant/fluid/1805.02474v1-10.fld}}
-    \vspace{-0.5em}
-    \caption{Fluid source code example Paragraph}
-    \label{fig:fluid-example-paragraph}
-\end{figure}
+%\begin{figure}[h]
+%    \small
+%    {\lstinputlisting[language=Fluid]{../website/authoring-assistant/fluid/1805.02474v1-10.fld}}
+%    \vspace{-0.5em}
+%    \caption{Fluid source code example Paragraph}
+%    \label{fig:fluid-example-paragraph}
+%\end{figure}
 

### 2025-09-13 paper/tex/authors.tex
New authors.

\author{
   Alfonso Piscitelli \\
   University of Salerno \\
   Fisciano, Italy \\
   \texttt{apiscitelli@unisa.it} \\
   \And
   Joe Bond \\
   University of Bristol \\
   Bristol, UK \\
   \texttt{j.bond@bristol.ac.uk} \\
   \And
   Cristina David \\
   University of Bristol \\
   Bristol, UK \\
   \texttt{cristina.david@bristol.ac.uk} \\
   \And
   Ali Mohammed \\
   Manchester Metropolitan University \\
   Manchester, UK \\
   \texttt{22465707@stu.mmu.ac.uk} \\
   \And
   Jacob Pake \\
   University of Kent \\
   Canterbury, UK \\
   \texttt{jwp24@kent.ac.uk} \\
   \And
   Roly Perera \\
   University of Cambridge \\
   Cambridge, UK \\
   \texttt{roly.perera@cl.cam.ac.uk} \\
   \And
   Jessy Sodimu \\
   University of Bath \\
   Bath, UK \\
   \texttt{js3720@bath.ac.uk} \\
   \And
   Chenyiqiu Zheng \\
   University College London \\
   London, UK \\
   \texttt{chenyiqiu.zheng.23@ucl.ac.uk} \\
}

% ORCID IDs (not shown in document)
% Alfonso Piscitelli: 0000-0002-3567-1504
% Joe Bond: 0009-0008-5995-6994
% Cristina David: 0000-0002-9106-934X
% Roly Perera: 0000-0001-9249-9862

### 2025-09-13 paper/sec/evaluation.tex
[draft] added validation paragraph

--- 2025-08-22 sec/evaluation.tex
@@
 \label{sec:evaluation}
+
+\subsection{Validation}
+\label{subsec:validation-with-counterfactual-data}
+
+The validation process is structured in two phases
+
+\paragraph{Initial validation}
+In the initial phase, the generated expression is executed via Fluid CLI.
+If the execution is successful, the result obtained is compared with the expected result.
+In case of error, the output is forwarded to the interpretation agent, which proceeds to regenerate the expression.
+
+\paragraph{Validation via counterfactual testing}
+In the second phase, the data used for execution is deliberately modified in order to verify the consistency of the behaviours.
+Both the expected and generated expressions are executed again and the results are compared.
+If discrepancies emerge, the error is forwarded to the interpretation agent.
+If there are no differences, the expression is considered valid and is returned to the user.
 
@@
 \subsection{Success Rate by Category}
-To evaluate the system we used a sample of the Scigen dataset \cite{scigen_dataset_2021}.
+To evaluate the system we used a sample of the Scigen dataset~\citep{scigen_dataset_2021}.
 
@@
 \end{figure}
+
+\subsection{Negative Cases}
+\label{subsec:negative-cases}
+
+\begin{table}[t]
+    \centering
+    \small
+    \begin{tabular}{p{3cm} p{6cm} p{4cm}}
+        \toprule
+        \textbf{Ambiguity Type} & \textbf{Example} & \textbf{Explanation} \\
+        \midrule
+
+        ambiguous referent &
+        As shown in Table 3 \hl{BiLSTM} gives significantly better accuracies compared to uni-directional LSTM with the training time per epoch growing from 99 seconds to 106 seconds &
+        There are three models that have 99 seconds as their time; the assistant would not know which one to choose. \\
+
+        \midrule
+        missing-data &
+        The average methane emissions for the year 2015 is \hl{1.0} &
+        November is missing to calculate the average correctly. \\
+
+        \midrule
+        false statement &
+        We additionally make comparisons with stacked CNNs and hierarchical attention (Vaswani et al., 2017), shown in Table 3 (the CNN and Transformer rows), BiLSTM is \hl{the most efficient} among all models compared, with the highest model size &
+        BiLSTM is not the most efficient, nor does it have the largest size. \\
+
+        \midrule
+        incorrect numerical value &
+        LSTM is the fastest model with overall time taken being \hl{90} seconds &
+        It is not 90 but 106. \\
+
+        \midrule
+        ambiguous referent &
+        LSTM is the fastest model with overall time taken being \hl{90} seconds &
+        There are two type of time in the dataset (training\_time, execution\_time), both with a value of 90 seconds. \\
+
+        \bottomrule
+    \end{tabular}
+    \caption{Negative cases example without the \emph{Data} column.}
+    \label{tab:ambiguita}
+\end{table}

### 2025-09-13 src/main/java/explorableviz/authoringassistant/llm/FluidOpenAIGpt4oAgent.java
[fix] adapt evaluate method to the new signature

@@ -9,6 +9,7 @@ import com.theokanning.openai.service.OpenAiService;
 import explorableviz.authoringassistant.paragraph.Expression;
 import it.unisa.cluelab.lllm.llm.agents.generic.OpenAIEvaluatorAgent;
 import it.unisa.cluelab.lllm.llm.prompt.Prompt;
+import it.unisa.cluelab.lllm.llm.prompt.PromptList;
 import org.jetbrains.annotations.NotNull;
 import org.json.JSONObject;
 
@@ -23,7 +24,7 @@ public class FluidOpenAIGpt4oAgent extends OpenAIEvaluatorAgent<Expression> {
     }
 
 
-    public Expression evaluate(List<Prompt> prompts, String grid) {
+    public Expression evaluate(PromptList prompts, String grid) {
         OpenAiService service = new OpenAiService(getToken(), Duration.ofSeconds(90));
         ChatMessage responseMessage = service.createChatCompletion(getChatCompletionRequest(prompts)).getChoices().get(0).getMessage();
         ChatFunctionCall functionCall = responseMessage.getFunctionCall();

### 2025-09-13 paper/sec/introduction.tex
Extract figure. Delete some old figures.

--- 2025-06-06 sec/introduction.tex
@@
 
-\subsection{Self-Certifying Text}
+\subsection{Transparent Text}
+
+Recent work
 
@@
 \end{itemize}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{fig/scigen-mr-1906.02780-10.png}
-    \caption{Website example with the generated expression from authoring assistant}\label{fig:scigen-example-website}
-\end{figure}
 
@@
 
-\begin{table}[!ht]
-    \centering
-    \footnotesize
-    \renewcommand{\arraystretch}{1.2}
-    \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6cm}}
-        \toprule
-        \textbf{Type}                & \textbf{Example} & \textbf{Generated Expression} \\
-        \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Quantitative expressions}} \\
+\input{fig/natural-language-forms}
 
-        Numerical value
-        & the training time per epoch growing from \hl{67} seconds to 106 seconds.
-        &
-        \begin{lstlisting}[language=Fluid,numbers=none]
-(findWithKey' "model" "LSTM" tableData).time_s
-        \end{lstlisting}
-        \\
-        Percentage &
-        The Energy Sector accounts for total methane emissions of \hl{52.80\%} in 2030.
-        &
-        \begin{lstlisting}[language=Fluid,numbers=none]
-(record.emissions /
- sum (map (fun x -> x.emissions)
-          (getByYear year tableData))) * 100
-
-        \end{lstlisting}  \\
-        Rounding & ~                & ~                             \\
-        Cardinal  & ~                & ~                             \\
-        Multiplicative  & ~                & ~                             \\
-        \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Aggregation}} \\
-        Average
-        & The average methane emissions for the year 2030 is \hl{13.51} &
-        \begin{lstlisting}[language=Fluid,numbers=none]
-(sumEmissions year tableData / length records)
-        \end{lstlisting} \\
-        Min/Max                          & The Energy Sector recorded its highest methane emissions in \hl{2030}             &
-        \begin{lstlisting}[language=Fluid,numbers=none]
-let maxEntry =
-    maximumBy (fun x -> x.emissions)
-              (filter
-                (fun x -> x.type == "Energy Sector")
-                tableData)
-in maxEntry.year
-        \end{lstlisting} \\
-        Rank &
-        3-layer stacked CNN gives an accuracy of 81.46\%, which is the \hl{lowest} compared with BiLSTM, and S-LSTM  &
-        \begin{lstlisting}[language=Fluid,numbers=none]
-let pos =
-  findIndex "model" "CNN"
-    (insertionSort cmpTime tableData)
-in rankLabel "lowest" pos \end{lstlisting} \\
-        Total &
-        The total methane emissions for the year 2030 is \hl{37.74} for Agriculture &
-        \begin{lstlisting}[language=Fluid,numbers=none]
-sumEmissions year tableData
-        \end{lstlisting} \\
-        Count*                       & ~                & ~                             \\
-        \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Trends}} \\
-
-        Comparison
-        & The training time per epoch \hl{growing} from 67 seconds to 106 seconds. &
-        \begin{lstlisting}[language=Fluid,numbers=none]
-trendWord
- (findWithKey' "model" "BiLSTM" tableData).time_s
- (findWithKey' "model" "LSTM" tableData).time_s
- growShrink
-        \end{lstlisting} \\
-        Graded adjective*            & ~                & ~                             \\
-        Reference to visual element*                            & ~                & ~                             \\
-        \bottomrule
-    \end{tabular}
-    \caption{Example expressions generated by the Authoring Assistant, classified by type. In green, the expression is already tested.}
-    \label{tab:fluid_examples}
-\end{table}
 \subsection{Contributions}
@@
 \end{itemize}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/scigen-mr-1906.02780-10.png}
+    \caption{Sample text with interpretation generated by Authoring Assistant}\label{fig:scigen-example-website}
+\end{figure}

### 2025-09-13 paper/sec/introduction.tex
Purge more old figures.
Purge obsolete figures.

@@
 
-Recent work
+Recent work on data provenance and data visualisation has shown how interactive provenance queries can be
+supported automatically \cite{psallisas
 
@@
 
-\paragraph{Authoring transparent text.} Someone authoring content for an online article, wants to create text
+\begin{enumerate}
+\item \textbf{Authoring transparent text.} Someone authoring content for an online article, wants to create text
 linked to raw data (and derivative data such as charts or tabular summaries), so that the evidence base for
@@
 
-\paragraph{Interpreting text after the fact.} Someone reading textual claims derived from open data (e.g. a
+\item \textbf{Interpreting text after the fact.} Someone reading textual claims derived from open data (e.g. a
 scientific paper or climate report), wants to retroactively link the text to queries over the available data
@@
 base. Perhaps just to aid their own comprehension, or to provide some kind of justified peer review.
+\end{enumerate}
 
-\vspace{2mm}
 \noindent Here we focus on the first one because the second one requires a certain amount of additional setup.

### 2025-09-13 paper/sec/introduction.tex
More purging.

@@
 
-\subsection{Transparent Text}
+\subsection{Transparent Text via Interactive Provenance Queries}
 
-Recent work on data provenance and data visualisation has shown how interactive provenance queries can be
-supported automatically \cite{psallisas
+Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} developed
+software infrastructure for integrating interactive provenance queries into charts and figures. Visual outputs
+like this, which here we call \emph{transparent visualisations}, allow readers to interact with a visual
+elements and see the related data. In scientific papers, news articles and policy reports, charts and figures
+usually only play a supporting role, however. The main claims are usually presented in natural language,
+language which is rich with quantitative claims and allusions to data analyses which are not present in the
+final artifact.
 
-\begin{itemize}
-    \item Introduce basic idea, as complementary to transparent visualisations
-\end{itemize}
+\rpnote{Small example here}
 

### 2025-09-14 paper/sec/introduction.tex
High-resolution screen capture. Fix filename to match example.

--- 2025-09-13 paper/sec/introduction.tex
@@
 \begin{itemize}
-    \item Interactive visualization of data
     \item Making the text transparent, with the possibility to highlight the data related to a specific sentence.
@@
 
-\subsection{Transparent Text}
+\subsection{Transparent Text via Interactive Provenance Queries}
 
-Recent work on data provenance and data visualisation has shown how interactive provenance queries can be
-supported automatically \cite{psallisas
+Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} integrating
+interactive provenance queries into charts and figures using \emph{transparent programming languages} which
+track data provenance automatically as data flows through a computation. Visual outputs produced using these
+platforms allow readers to interact with visual elements to see the relevant data. In most article charts and
+figures usually only play a supporting role, however. The main claims are usually presented in natural
+language which also makes quantitative claims and refers to data analyses which are not present in the final
+artifact.
 
-\begin{itemize}
-    \item Introduce basic idea, as complementary to transparent visualisations
-\end{itemize}
+In principle transparent programming languages can also be used to create natural language which is bidirectly
+linked to data, by computing key fragments of natural language from data. For example, using the language
+presented in this paper,
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-with-pointer.png}
+    \caption{Transparent Text with Interactive Provenance Queries}\label{fig:scigen-example-website}
+\end{figure}
+
+\rpnote{Small example here}
 
@@
 \end{itemize}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{fig/scigen-mr-1906.02780-10.png}
-    \caption{Sample text with interpretation generated by Authoring Assistant}\label{fig:scigen-example-website}
-\end{figure}

### 2025-09-14 paper/fig/scigen-1805.02474v1-10-src.tex
Extract figure.

--- 2025-09-03 paper/sec/authoring-workflow.tex
@@
-\section{AI-Assisted Authoring Workflow}
-\label{sec:authoring-workflow}
-
-\begin{figure}[ht]
-    \centering
-    \begin{subfigure}{0.48\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{fig/entire-workflow}
-        \caption{Workflow}
-        \label{fig:architecture}
-    \end{subfigure}\hfill
-    \begin{subfigure}{0.48\linewidth}
-        \centering
-        \includegraphics[width=\linewidth]{fig/data-flow}
-        \caption{Data flow}
-        \label{fig:dataflow}
-    \end{subfigure}
-    \caption{AI-assisted authoring process: workflow (a) and data flow (b).}
-    \label{fig:overall}
-\end{figure}
-
-\subsection{System Prompt}
-\label{subsec:system-prompt}
-Fig. \ref{fig:system-prompt} shows the system prompt defined for the Authoring Assistant.
-
-\begin{figure}[h]
-    \centering
-    \begin{tcolorbox}[colback=gray!10, colframe=gray!50, boxrule=0.5pt, arc=2pt,
-        left=6pt, right=6pt, top=4pt, bottom=4pt]
-        \VerbatimInput[fontsize=\small, formatcom=\normalfont]{../system-prompt/system-prompt.txt}
-    \end{tcolorbox}
-    \vspace{-0.5em}
-    \caption{System-Prompt defined for the Authoring Assistant}
-    \label{fig:system-prompt}
-\end{figure}
-
-\subsection{Editor Loop}
-\label{subsec:editor-loop}
-Design of main loop that would be integrated into an IDE. The configuration (state) that is being maintained
-is a \kw{Paragraph} the user is authoring in Fluid. Consists of a sequence of text fragments, some of which
-are uninterpreted (plain literals), the remainder have underlying expressions linking the text to raw or derived
-data.
-
-Workflow:
-\begin{enumerate}
-\item User selects (a substring of) one of the literal text fragments, indicating they want to link it in this
-way
-\item Authoring tool generates expression which is either:
-  \begin{enumerate}
-  \item computes the selected text (e.g.~``greater than'' might be computed by comparing two numbers)
-  \item becomes the formal meaning of the selected text (e.g. the text ``carbon intensity of methane
-emissions'' can be understood in a specific context as referring to the numerical value ``34 gCO2eq/kWh''
-  \end{enumerate}
-\item Some kind of validation step, including both automated validation (e.g. checking for runtime errors) and
-user validation; if validation fails, goto (2)
-\item Partition code into additional definitions and expression; add definitions to main source file, and
-incorporate expression into \kw{Paragraph}
-\item Goto (1) with updated editor state
-\end{enumerate}
-
-\subsection{Loopback System}
-\label{subsec:loopback-system}
-
-Once the Authoring Assistant has generated the expression, it is executed by the Fluid interpreter.
-If execution fails, the error is sent back to the LLM for regeneration.
-This \textit{loopback} mechanism enables iterative refinement of expressions, improving accuracy and robustness.
-Three types of errors are handled:
-
-\begin{itemize}
-    \item \textbf{Invalid Program}.
-    The generated expression is executed by the Fluid interpreter.
-    If the execution fails due to issues such as undeclared functions, syntax errors, or other interpreter-level problems, it results in an \textit{Invalid Program} error.
-    \item \textbf{Invalid Output Type}.
-    According to the \textit{System Prompt} (Fig.~\ref{fig:system-prompt}), the expression is expected to return a string value.
-    If the returned value is of a different type, an \textit{Invalid Output Type} error is raised.
-    \item \textbf{Invalid Output Value}.
-    The output produced by the generated expression is compared against the highlighted target text.
-    If the values do not match, an \textit{Invalid Output Value} error is reported.
-\end{itemize}
-
-\subsection{Paragraph with generated expression}
-\label{subsec:paragraph-with-generated-expression}
-Figure \ref{fig:fluid-example-paragraph} shows the Fluid code of a Paragraph with the expression generated by the authoring assistant.
-Figure \ref{fig:fluid-scigen} shows the Fluid code written for the scigen datasets and examples.
 \begin{figure}[h]
     \small
-    {\lstinputlisting[language=Fluid]{../website/authoring-assistant/fluid/1805.02474v1-10.fld}}
+    {\lstinputlisting[language=Fluid,mathescape=false]{../website/authoring-assistant/fluid/1805.02474v1-10.fld}}
     \vspace{-0.5em}
-    \caption{Fluid source code example Paragraph}
+    \caption{Source code for \figref{scigen-example-website}}
     \label{fig:fluid-example-paragraph}
 \end{figure}
-
-\begin{figure}[h]
-    \small
-    {\lstinputlisting[language=Fluid]{../fluid-common/scigen.fld}}
-    \caption{Fluid library written for scigen}
-    \label{fig:fluid-scigen}
-\end{figure}
-
-\subsection{Turning validation errors into improved prompts}\label{subsec:turning-validation-errors-into-improved-prompts}
-E.g.:
-\begin{enumerate}
-\item Turn ``Definition not found'' into prompt to generate definition?
-\end{enumerate}
-
-\subsection{IDE integration}\label{subsec:ide-integration}
-
-Could then be integrated into desktop IDE like VSCode or online IDE like CodeMirror. Might make a good
-internship project, but could also be out-of-scope for this paper.
-
-\subsection{Other potential enhancements}\label{subsec:other-potential-enhancements}
-
-We could also think about using an LLM in a couple of other complementary ways:
-\begin{itemize}
-\item identifying text fragments which might be linked;
-\item validating generated expressions (perhaps by proposing test cases)
-\end{itemize}

### 2025-09-14 paper/fig/scigen-1805.02474v1-10-src.tex
Import source directly into figure.

@@
     \small
-    {\lstinputlisting[language=Fluid,mathescape=false]{../website/authoring-assistant/fluid/1805.02474v1-10.fld}}
+    \begin{lstlisting}[language=Fluid,mathescape=false]
+"""
+	As shown in Table 3, BiLSTM gives significantly ${trendWord (model "BiLSTM" tableData).acc (model "LSTM" tableData).acc betterWorse} accuracies compared to uni-directional LSTM2, with the training time per epoch ${trendWord (model "BiLSTM" tableData).time_s (model "LSTM" tableData).time_s growShrink} from ${numToStr (model "LSTM" tableData).time_s} seconds to ${numToStr (model "BiLSTM" tableData).time_s} seconds.
+Stacking 2 layers of BiLSTM gives ${trendWord (model "2 stacked BiLSTM" tableData).acc (model "BiLSTM" tableData).acc improve} to development results, with a ${trendWord (model "2 stacked BiLSTM" tableData).time_s (model "BiLSTM" tableData).time_s smallerHigher} time of ${numToStr (model "2 stacked BiLSTM" tableData).time_s} seconds.
+3 layers of stacked BiLSTM ${trendWord (model "3 stacked BiLSTM" tableData).acc (model "BiLSTM" tableData).acc improve} the results.
+In contrast, S-LSTM gives a development result of ${numToStr (model "S-LSTM" tableData).acc} %, which is significantly ${trendWord (model "S-LSTM" tableData).acc (model "2 stacked BiLSTM" tableData).acc betterWorse} compared to 2-layer stacked BiLSTM, with a ${trendWord (model "S-LSTM" tableData).param (model "2 stacked BiLSTM" tableData).param smallerHigher} number of model parameters and a ${trendWord (model "S-LSTM" tableData).time_s (model "2 stacked BiLSTM" tableData).time_s shorterLonger} time of ${numToStr (model "S-LSTM" tableData).time_s} seconds.
+We additionally make comparisons with stacked CNNs and hierarchical attention (Vaswani et al., 2017), shown in Table 3 (the CNN and Transformer rows), ${(findWithKey' "time_s" (minimum (map (fun y -> y.time_s) tableData)) tableData).model} is the ${let pos = findIndex "model" "CNN" (insertionSort (fun a b -> a.time_s < b.time_s) tableData) in rankLabel "most efficient" pos} among all models compared, with the ${let pos = findIndex "model" "CNN" (insertionSort (fun a b -> a.param < b.param) tableData) in rankLabel "smallest" pos} model size.
+On the other hand, a 3-layer stacked CNN gives an accuracy of ${numToStr (model "3 stacked CNN" tableData).acc} %, which is also the ${let pos = findIndex "model" "CNN" (insertionSort (fun a b -> a.time_s < b.time_s) tableData) in rankLabel "lowest" pos} compared with BiLSTM, hierarchical attention and S-LSTM.
+The ${let pos = findIndex "model" "S-LSTM+Attention" (insertionSort (fun a b -> b.acc < a.acc) tableData) in rankLabel "best" pos} performance of hierarchical attention is obtained by S-LSTM+Attention in terms of both accuracy and efficiency.
+S-LSTM gives significantly ${trendWord (model "S-LSTM" tableData).acc (model "CNN" tableData).acc betterWorse} accuracies compared with both CNN and hierarchical attention.
+Table 3 additionally shows the results of BiLSTM and S-LSTM when external attention is used  Attention leads to improved accuracies for both BiLSTM and S-LSTM in classification, with S-LSTM still ${trendWord (model "S-LSTM" tableData).acc (model "BiLSTM" tableData).acc underOverPerforming} BiLSTM significantly.
+"""
+    \end{lstlisting}
     \vspace{-0.5em}

### 2025-09-15 paper/fig/scigen-1805.02474v1-10-src.tex
Better helper functions.

--- 2025-09-14 paper/fig/scigen-1805.02474v1-10-src.tex
@@
     \small
-    {\lstinputlisting[language=Fluid,mathescape=false]{../website/authoring-assistant/fluid/1805.02474v1-10.fld}}
-    \vspace{-0.5em}
+    \begin{lstlisting}[language=Fluid,mathescape=false]
+let model_ name = model name tableData in """
+As shown in Table 3, BiLSTM gives significantly ${trendWord (model_ "BiLSTM").acc (model_ "LSTM").acc
+betterWorse} accuracies compared to uni-directional LSTM2, with the training time per epoch ${trendWord (model_
+"BiLSTM").time_s (model_ "LSTM").time_s growShrink} from ${numToStr (model_ "LSTM").time_s} seconds to ${numToStr
+(model_ "BiLSTM").time_s} seconds. Stacking 2 layers of BiLSTM gives ${trendWord (model_ "2 stacked BiLSTM").acc
+(model_ "BiLSTM").acc improve} to development results, with a ${trendWord (model_ "2 stacked BiLSTM").time_s
+(model_ "BiLSTM").time_s smallerHigher} time of ${numToStr (model_ "2 stacked BiLSTM").time_s} seconds. 3 layers
+of stacked BiLSTM ${trendWord (model_ "3 stacked BiLSTM").acc (model_ "BiLSTM").acc improve} the results. In
+contrast, S-LSTM gives a development result of ${numToStr (model_ "S-LSTM").acc} %, which is significantly
+${trendWord (model_ "S-LSTM").acc (model_ "2 stacked BiLSTM").acc betterWorse} compared to 2-layer stacked
+BiLSTM, with a ${trendWord (model_ "S-LSTM").param (model_ "2 stacked BiLSTM").param smallerHigher} number of
+model parameters and a ${trendWord (model_ "S-LSTM").time_s (model_ "2 stacked BiLSTM").time_s shorterLonger}
+time of ${numToStr (model_ "S-LSTM").time_s} seconds.  We additionally make comparisons with stacked CNNs and
+hierarchical attention (Vaswani et al., 2017), shown in Table 3 (the CNN and Transformer rows), ${(findWithKey'
+"time_s" (minimum (map (fun y -> y.time_s) tableData)) tableData).model} is the ${rankLabel "most efficient"
+(findIndex "model" "CNN" (insertionSort (fun a b -> a.time_s < b.time_s) tableData))} among all models compared,
+with the ${rankLabel "smallest" (findIndex "model" "CNN" (insertionSort (fun a b -> a.param < b.param) tableData)
+)} model size. On the other hand, a 3-layer stacked CNN gives an accuracy of ${numToStr (model "3 stacked CNN"
+tableData).acc} %, which is also the ${rankLabel "lowest" (findIndex "model" "CNN" (insertionSort (fun a b ->
+a.time_s < b.time_s) tableData))} compared with BiLSTM, hierarchical attention and S-LSTM. The ${rankLabel "best"
+(findIndex "model" "S-LSTM+Attention" (insertionSort (fun a b -> b.acc < a.acc) tableData))} performance of
+hierarchical attention is obtained by S-LSTM+Attention in terms of both accuracy and efficiency. S-LSTM gives
+significantly ${trendWord (model_ "S-LSTM").acc (model_ "CNN").acc betterWorse} accuracies compared with both CNN
+and hierarchical attention. Table 3 additionally shows the results of BiLSTM and S-LSTM when external attention
+is used. Attention leads to improved accuracies for both BiLSTM and S-LSTM in classification, with S-LSTM still
+${trendWord (model_ "S-LSTM").acc (model_ "BiLSTM").acc underOverPerforming} BiLSTM significantly."""
+    \end{lstlisting}
     \caption{Source code for \figref{scigen-example-website}}

### 2025-09-15 paper/sec/introduction.tex
Extract figure.
Import source directly into figure.
Better helper functions.

@@
 platforms allow readers to interact with visual elements to see the relevant data. In most article charts and
-figures usually only play a supporting role, however. The main claims are usually presented in natural
-language which also makes quantitative claims and refers to data analyses which are not present in the final
-artifact.
+figures usually only play a supporting role, however, with the main claims usually presented in natural
+language. This language makes quantitative claims and refers to ``behind the scenes'' data analyses.
 
@@
     \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-with-pointer.png}
-    \caption{Transparent Text with Interactive Provenance Queries}\label{fig:scigen-example-website}
+    \caption{\emph{Transparent text} links fragments of natural language to underlying data, via queries}\label{fig:scigen-example-website}
 \end{figure}
+
+\input{fig/scigen-1805.02474v1-10-src}
 

### 2025-09-15 paper/fig/natural-language-forms.tex
Prefer underscores to primes.

--- 2025-06-06 sec/introduction.tex
@@
-\section{Introduction}
-
-\lstset{
-    aboveskip=-7pt,
-    belowskip=-8.5pt,
-    xleftmargin=0pt,
-    xrightmargin=0pt
-}
-
-\begin{itemize}
-    \item Interactive visualization of data
-    \item Making the text transparent, with the possibility to highlight the data related to a specific sentence.
-    \item Making sense of data-driven claims is hard, even if the evidence base (code/data) is open
-    \item We see this in peer review, misinformation, retracted papers, …
-\end{itemize}
-
-\subsection{Self-Certifying Text}
-
-\begin{itemize}
-    \item Introduce basic idea, as complementary to transparent visualisations
-\end{itemize}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{fig/scigen-mr-1906.02780-10.png}
-    \caption{Website example with the generated expression from authoring assistant}\label{fig:scigen-example-website}
-\end{figure}
-
-\subsection{Use cases}
-Two potential scenarios for this sort of technology:
-
-\paragraph{Authoring transparent text.} Someone authoring content for an online article, wants to create text
-linked to raw data (and derivative data such as charts or tabular summaries), so that the evidence base for
-the claims made in the text can be explored \emph{in situ}, by interacting with the text.
-
-\paragraph{Interpreting text after the fact.} Someone reading textual claims derived from open data (e.g. a
-scientific paper or climate report), wants to retroactively link the text to queries over the available data
-and gradually ``rationally reconstruct'' the relationship between the claims in the paper and the evidence
-base. Perhaps just to aid their own comprehension, or to provide some kind of justified peer review.
-
-\vspace{2mm}
-\noindent Here we focus on the first one because the second one requires a certain amount of additional setup.
-
-\subsubsection{Target idioms of natural language}
-
-NLP aspect of the problem is potentially a big problem space in itself. We will restrict interest to certain
-idiomatic uses of natural language in making/justifying scientific claims.
-Examples \ref{tab:fluid_examples}
-
 \begin{table}[!ht]
@@
     \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6cm}}
-        \toprule
+        \hline
         \textbf{Type}                & \textbf{Example} & \textbf{Generated Expression} \\
@@
         \begin{lstlisting}[language=Fluid,numbers=none]
-(findWithKey' "model" "LSTM" tableData).time_s
+(findWithKey_ "model" "LSTM" tableData).time_s
         \end{lstlisting}
@@
 trendWord
- (findWithKey' "model" "BiLSTM" tableData).time_s
- (findWithKey' "model" "LSTM" tableData).time_s
+ (findWithKey_ "model" "BiLSTM" tableData).time_s
+ (findWithKey_ "model" "LSTM" tableData).time_s
  growShrink
@@
         Reference to visual element*                            & ~                & ~                             \\
-        \bottomrule
+        \hline
     \end{tabular}
@@
 \end{table}
-\subsection{Contributions}
-
-\begin{itemize}
-    \item Design and proof-of-concept implementation of AI-assisted workflow for authoring transparent text
-    (\secref{authoring-workflow})
-    \item Empirical evaluation of how effective current LLMs are at providing the ``AI-assisted'' part
-\end{itemize}

### 2025-09-15 paper/fig/natural-language-forms.tex

@@
         \begin{lstlisting}[language=Fluid,numbers=none]
-(findWithKey_ "model" "LSTM" tableData).time_s
+(findWithKey' "model" "LSTM" tableData).time_s
         \end{lstlisting}
@@
 trendWord
- (findWithKey_ "model" "BiLSTM" tableData).time_s
- (findWithKey_ "model" "LSTM" tableData).time_s
+ (findWithKey' "model" "BiLSTM" tableData).time_s
+ (findWithKey' "model" "LSTM" tableData).time_s
  growShrink

### 2025-09-15 paper/fig/scigen-1805.02474v1-10-src.tex

--- 2025-09-14 paper/fig/scigen-1805.02474v1-10-src.tex
@@
     \small
-    {\lstinputlisting[language=Fluid,mathescape=false]{../website/authoring-assistant/fluid/1805.02474v1-10.fld}}
-    \vspace{-0.5em}
-    \caption{Source code for \figref{scigen-example-website}}
+    % listings package seems to struggle, for now just use screenshot from VSCode
+    \includegraphics[scale=0.169]{fig/scigen-1805.02474v1-10-src-screenshot.png}
+    \caption{Reference solution for transparent text in \figref{scigen-example-website}}
     \label{fig:fluid-example-paragraph}

### 2025-09-15 paper/sec/evaluation.tex
[update] remove toprule

@@
     \begin{tabular}{p{3cm} p{6cm} p{4cm}}
-        \toprule
+        \hline
         \textbf{Ambiguity Type} & \textbf{Example} & \textbf{Explanation} \\
-        \midrule
+        \hline
 
@@
 
-        \midrule
         missing-data &
@@
 
-        \midrule
         false statement &
@@
 
-        \midrule
         incorrect numerical value &
@@
 
-        \midrule
         ambiguous referent &
@@
 
-        \bottomrule
+        \hline
     \end{tabular}

### 2025-09-17 paper/abstract.tex
Start on abstract.

\begin{abstract}
We present a proof-of-concept implementation and evaluation of an agent-based LLM framework for authoring
\emph{transparent text}, natural language articles which provide interactions allowing a reader to explore the
relationship to the underlying data. We evaluate our approach on a subset of SciGen, a dataset consisting of
tables from scientific articles and their corresponding descriptions.
\end{abstract}

### 2025-09-17 paper/sec/authoring-workflow.tex
[update] fix listing and add caption to flows
Extract figure.
More on abstract.

@@
         \includegraphics[width=\linewidth]{fig/data-flow-correct}
-        \caption{Author Acceptance Dataflow}
+        \caption{Author Acceptance workflow}
         \label{fig:data-flow-correct}
@@
         \includegraphics[width=\linewidth]{fig/data-flow-error}
-        \caption{Author Rejection Dataflow}
+        \caption{Author Rejection workflow}
         \label{fig:dataflow-error}
     \end{subfigure}
-    \caption{Two workflows of the Interpretation Agent.
-In Fig.
-\ref{fig:data-flow-correct}, the author accepts the generated expression after verifying its correctness online.
-In Fig.
-\ref{fig:dataflow-error}, the author identifies an error (the expression mentions accuracy, while the article highlights timing) and rejects it.}
+    \caption{Two possible paths through the basic editing loop. In Fig.~\ref{fig:data-flow-correct}, the
+    author accepts the generated expression after interactively verifying its correctness. In
+    Fig.~\ref{fig:dataflow-error}, the author identifies an error (the expression mentions accuracy, while the
+    article highlights timing) and rejects it.}
     \label{fig:overall-architecture}
@@
 Figure \ref{fig:fluid-scigen} shows the Fluid code written for the scigen datasets and examples.
-%\begin{figure}[h]
-%    \small
-%    {\lstinputlisting[language=Fluid]{../website/authoring-assistant/fluid/1805.02474v1-10.fld}}
-%    \vspace{-0.5em}
-%    \caption{Fluid source code example Paragraph}
-%    \label{fig:fluid-example-paragraph}
-%\end{figure}
 

### 2025-09-17 paper/sec/authoring-workflow.tex
Fig. 4 caption.

@@
     \end{subfigure}
-    \caption{Two possible paths through the basic editing loop. In Fig.~\ref{fig:data-flow-correct}, the
-    author accepts the generated expression after interactively verifying its correctness. In
-    Fig.~\ref{fig:dataflow-error}, the author identifies an error (the expression mentions accuracy, while the
-    article highlights timing) and rejects it.}
+    \caption{Two possible paths through the basic editing loop: (a) author accepts generated expression after
+    interactively verifying its correctness; (b) author rejects proposed expression, having identified an
+    error (natural language mentions accuracy, while interaction reveals use of timing data).}
     \label{fig:overall-architecture}

### 2025-09-17 paper/abstract.tex

@@
 \begin{abstract}
-We present a proof-of-concept implementation and evaluation of an agent-based LLM framework for authoring
-\emph{transparent text}, natural language articles which provide interactions allowing a reader to explore the
-relationship to the underlying data. We evaluate our approach on a subset of SciGen, a dataset consisting of
-tables from scientific articles and their corresponding descriptions.
+We present an agent-based LLM framework for authoring \emph{transparent text}, web-based scholarly articles
+which allow a reader to explore the relationship to the underlying data by hovering over fragments of text.
+Our approach builds on recent developments in data provenance techniques for general-purpose programming
+languages, and tasks LLM agents with identifying fragments of text which can be computed from data, including
+numerical values computed by aggregations like sum and mean, comparatives and superlatives like ``better
+than'' and ``largest'', and trend-adjectives like ``growing''. We evaluate our approach on a subset of SciGen,
+a dataset consisting of tables from scientific articles and their corresponding descriptions, which we extend
+with hand-generated counterfactual test cases for evaluating how machine-generated expressions generalise in
+the presence of changes to the underlying data. Our results show that some state-of-the-art models are able to
+synthesis compound expressions that generate data-driven text that agrees with gold solutions.
 \end{abstract}

### 2025-09-17 paper/sec/authoring-workflow.tex

@@
 \begin{itemize}
-    \item \textbf{SuggestionAgent}.
-Based on an LLM, the SuggestionAgent receives as input the initial text and the reference data, and returns the elements in the text that can be replaced with expressions in the Fluid language.
-    \item \textbf{InterpretationAgent}.
-Also based on an LLM, the InterpretationAgent receives as input a selection (either obtained from the SuggestionAgent or manually defined by the author) and generates an interpretation of the text as a Fluid expression.
+    \item \textbf{SuggestionAgent}. LLM-based agent which identifies text fragments potentially computable from data.
+    \item \textbf{InterpretationAgent}. LLM-based agent which, for a given text fragment (provided by
+    SuggestionAgent or by the author), attempts to synthesise a Fluid expression computing the target text.
 \end{itemize}
@@
     \vspace{-0.5em}
-    \caption{System-Prompt defined for the Authoring Assistant}
+    \caption{System Prompt defined for the Authoring Assistant}
     \label{fig:system-prompt}

### 2025-09-17 paper/sec/introduction.tex
Missing SciGen citation.
Start on abstract.

@@
 \section{Introduction}
-
-\lstset{
-    aboveskip=-7pt,
-    belowskip=-8.5pt,
-    xleftmargin=0pt,
-    xrightmargin=0pt
-}
 
@@
 \end{itemize}
+
+\rpnote{Small example here}
 
@@
 In principle transparent programming languages can also be used to create natural language which is bidirectly
-linked to data, by computing key fragments of natural language from data. For example, using the language
-presented in this paper,
+linked to data, by computing key fragments of natural language from data. The following example, adapted from
+\cite{moosavi21}, shows how the language presented in this paper can be used to achieve this:
 
@@
 
-\rpnote{Small example here}
-
-\subsection{Use cases}
-Two potential scenarios for this sort of technology:
+We envisage two possible use cases for transparent text:
 
 \begin{enumerate}
-\item \textbf{Authoring transparent text.} Someone authoring content for an online article, wants to create text
+\item \textbf{Authoring.} Someone authoring content for an online article, wants to create text
 linked to raw data (and derivative data such as charts or tabular summaries), so that the evidence base for
@@
 
-\item \textbf{Interpreting text after the fact.} Someone reading textual claims derived from open data (e.g. a
+\item \textbf{Reading or reviewing.} Someone reading textual claims derived from open data (e.g. a
 scientific paper or climate report), wants to retroactively link the text to queries over the available data
@@
 \end{enumerate}
-
-\noindent Here we focus on the first one because the second one requires a certain amount of additional setup.
 

### 2025-09-17 paper/abstract.tex
More on abstract.

--- 2025-09-17 paper/abstract.tex
@@
 \begin{abstract}
-We present a proof-of-concept implementation and evaluation of an agent-based LLM framework for authoring
-\emph{transparent text}, natural language articles which provide interactions allowing a reader to explore the
-relationship to the underlying data. We evaluate our approach on a subset of SciGen, a dataset consisting of
-tables from scientific articles and their corresponding descriptions.
+We introduce the idea of \emph{transparent text}, web-based data-driven scholarly articles which allow a
+reader to explore the relationship to the data set by hovering over fragments of text, and present an
+agent-based LLM framework for authoring transparent text. Our approach builds on recent developments in
+integrated data provenance for general-purpose programming languages. Our system uses the open source
+programming language Fluid, which uses provenance tracking to link visual elements to data, and show how Fluid
+can also be used as host platform for transparent text. Our architecture consists of two LLM agents which
+support the human author during the creation of a transparent document. A SuggestionAgent identifies fragments
+of text which could in principle be computed from data, including numerical values selected from records or
+computed by aggregations like sum and mean; comparatives and superlatives like ``better than'' and
+``largest''; trend-adjectives like ``growing''; and other idiomatic quantitative or semi-quantitative phrases.
+An InterpretationAgent synthesises suitable Fluid queries over the data. We evaluate our approach on a subset
+of SciGen, a dataset consisting of tables from scientific articles and their corresponding descriptions, which
+we extend with hand-generated counterfactual test cases for evaluating how machine-generated expressions
+generalise in the presence of changes to the underlying data. Our results show that some state-of-the-art
+models are able to synthesise compound expressions that generate data-driven text that agrees with gold
+solutions.
 \end{abstract}

### 2025-09-17 paper/abstract.tex
More on abstract.

@@
 agent-based LLM framework for authoring transparent text. Our approach builds on recent developments in
-integrated data provenance for general-purpose programming languages. Our system uses the open source
-programming language Fluid, which uses provenance tracking to link visual elements to data, and show how Fluid
-can also be used as host platform for transparent text. Our architecture consists of two LLM agents which
-support the human author during the creation of a transparent document. A SuggestionAgent identifies fragments
-of text which could in principle be computed from data, including numerical values selected from records or
-computed by aggregations like sum and mean; comparatives and superlatives like ``better than'' and
-``largest''; trend-adjectives like ``growing''; and other idiomatic quantitative or semi-quantitative phrases.
-An InterpretationAgent synthesises suitable Fluid queries over the data. We evaluate our approach on a subset
-of SciGen, a dataset consisting of tables from scientific articles and their corresponding descriptions, which
-we extend with hand-generated counterfactual test cases for evaluating how machine-generated expressions
-generalise in the presence of changes to the underlying data. Our results show that some state-of-the-art
-models are able to synthesise compound expressions that generate data-driven text that agrees with gold
-solutions.
+integrated data provenance for general-purpose programming languages. Our system uses Fluid, an open source
+programming language which is able to link visual elements to data using a provenance-tracking runtime, and
+uses it as host platform for transparent text. Our architecture consists of two LLM agents which support the
+human author during the creation of a transparent document. A SuggestionAgent identifies fragments of text
+which could in principle be computed from data, including numerical values selected from records or computed
+by aggregations like sum and mean, comparatives and superlatives like ``better than'' and ``largest'',
+trend-adjectives like ``growing'', and other idiomatic quantitative or semi-quantitative phrases. An
+InterpretationAgent synthesises suitable Fluid queries over the data, generating either a target string (when
+the desired text fragment is already known), or a candidate string (when the answer to the query is not
+known). The Fluid runtime infrastructure then interprets the resulting expression in the context of an
+interactive web page which the author can use to manually validate the generated code. We evaluate our
+approach on a subset of SciGen, a dataset consisting of tables from scientific articles and their
+corresponding descriptions, which we extend with hand-generated counterfactual test cases for evaluating how
+machine-generated expressions generalise in the presence of changes to the underlying data. Our results show
+that some state-of-the-art models are able to synthesise compound expressions that are extensionally
+compatible with gold solutions.
 \end{abstract}

### 2025-09-17 paper/sec/authoring-workflow.tex
Simplify loopback description.

@@
 
-Once the Authoring Assistant has generated the expression, it is executed by the Fluid interpreter.
-If execution fails, the error is sent back to the LLM for regeneration.
-This \textit{loopback} mechanism enables iterative refinement of expressions, improving accuracy and robustness.
-Three types of errors are handled:
+Expressions generated by the \InterpretationAgent are evaluated by the Fluid interpreter. If evaluation fails,
+the error is used to extend the prompt and the extended prompt is sent back to the LLM for regeneration. This
+loopback mechanism enables iterative refinement of expressions, improving accuracy and robustness. Three types
+of errors are handled:
 
 \begin{itemize}
-    \item \textbf{Invalid Program}.
-    The generated expression is executed by the Fluid interpreter.
-    If the execution fails due to issues such as undeclared functions, syntax errors, or other interpreter-level problems, it results in an \textit{Invalid Program} error.
+    \item \textbf{Invalid Program}. Evaluation fails due a syntax error, undeclared identifer, division by
+    zero, array index out of bounds or other runtime problem.
     \item \textbf{Invalid Output Type}.

### 2025-09-17 paper/sec/authoring-workflow.tex
Simplify loopback description.

@@
     zero, array index out of bounds or other runtime problem.
-    \item \textbf{Invalid Output Type}.
-    According to the \textit{System Prompt} (Fig.~\ref{fig:system-prompt}), the expression is expected to return a string value.
-    If the returned value is of a different type, an \textit{Invalid Output Type} error is raised.
-    \item \textbf{Invalid Output Value}.
-    The output produced by the generated expression is compared against the highlighted target text.
-    If the values do not match, an \textit{Invalid Output Value} error is reported.
+    \item \textbf{Invalid Expression Type}. The value of the expression is not a string.
+    \item \textbf{Mismatching String}. The value of the expression is a string, but is not equal to
+    highlighted target text.
 \end{itemize}

### 2025-09-17 paper/sec/authoring-workflow.tex
Simplify loopback description.

@@
 the error is used to extend the prompt and the extended prompt is sent back to the LLM for regeneration. This
-loopback mechanism enables iterative refinement of expressions, improving accuracy and robustness. Three types
-of errors are handled:
+loopback mechanism enables iterative refinement of expressions, improving accuracy and robustness. There are
+two kinds of loopback error:
 
 \begin{itemize}
-    \item \textbf{Invalid Program}. Evaluation fails due a syntax error, undeclared identifer, division by
+    \item \textbf{Invalid Program}. Evaluation fails due a syntax error, undeclared identifier, division by
     zero, array index out of bounds or other runtime problem.
-    \item \textbf{Invalid Expression Type}. The value of the expression is not a string.
-    \item \textbf{Mismatching String}. The value of the expression is a string, but is not equal to
-    highlighted target text.
+%    \item \textbf{Invalid Expression Type}. The value of the expression is not a string.
+    \item \textbf{Mismatching String}. The string value of the expression is not equal to the target text.
 \end{itemize}

### 2025-09-17 src/main/java/explorableviz/authoringassistant/Program.java
Add link example to scigen 1805.02474v1-10

@@ -97,10 +97,7 @@ public class Program {
             logger.info("Validation failed because interpreter error");
             return Optional.of(value);
         }
-        //String error
-        if(!value.matches("^\".*\"$")) {
-            return Optional.of(STR."String expected. \{value} seems not a string");
-        }
+
         if (value.equals(expectedExpression.getValue()) || roundedEquals(value, expectedExpression.getValue())) {
             logger.info("Validation passed");
             return Optional.empty();

### 2025-09-17 src/main/java/explorableviz/authoringassistant/paragraph/ExpressionCategory.java

@@ -6,7 +6,8 @@ public enum ExpressionCategory {
     PERCENTAGE("Percentage"),
     DATA_RETRIEVAL("Data Retrieval"),
     DIFFERENCE("Difference"),
-    TOTAL("Total");
+    TOTAL("Total"),
+    LINK("Link");
 
     public final String label;
     ExpressionCategory(String label) {

### 2025-09-18 paper/sec/introduction.tex
Tweaks to intro.

@@
 
-Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} integrating
-interactive provenance queries into charts and figures using \emph{transparent programming languages} which
-track data provenance automatically as data flows through a computation. Visual outputs produced using these
-platforms allow readers to interact with visual elements to see the relevant data. In most article charts and
-figures usually only play a supporting role, however, with the main claims usually presented in natural
-language. This language makes quantitative claims and refers to ``behind the scenes'' data analyses.
+Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} proposed
+integrating interactive ``provenance queries'' directly into charts and figures. These interactive queries
+allow readers to interact with visual elements to see the data relevant to those elements. The key feature of
+these approaches is that the interactions are provided automatically, via \emph{transparent programming
+languages} (either query languages or general-purpose languages) which track dependencies as data flows
+through a computation.
 
-In principle transparent programming languages can also be used to create natural language which is bidirectly
-linked to data, by computing key fragments of natural language from data. The following example, adapted from
-\cite{moosavi21}, shows how the language presented in this paper can be used to achieve this:
+In most articles, however, charts and figures usually only play a supporting role. The main claims are usually
+presented in natural language, language which also makes quantitative statements and refers to ``behind the
+scenes'' data analyses. In principle transparent programming languages can also be used to create natural
+language which is bidirectly linked to data, by computing key fragments of natural language from data. The
+following example, adapted from \cite{moosavi21}, shows how the language presented in this paper can be used
+to achieve this:
 

### 2025-09-18 paper/abstract.tex
Some rework of abstract.

--- 2025-09-17 paper/abstract.tex
@@
 \begin{abstract}
-We present an agent-based LLM framework for authoring \emph{transparent text}, web-based scholarly articles
-which allow a reader to explore the relationship to the underlying data by hovering over fragments of text.
-Our approach builds on recent developments in data provenance techniques for general-purpose programming
-languages, and tasks LLM agents with identifying fragments of text which can be computed from data, including
-numerical values computed by aggregations like sum and mean, comparatives and superlatives like ``better
-than'' and ``largest'', and trend-adjectives like ``growing''. We evaluate our approach on a subset of SciGen,
-a dataset consisting of tables from scientific articles and their corresponding descriptions, which we extend
-with hand-generated counterfactual test cases for evaluating how machine-generated expressions generalise in
-the presence of changes to the underlying data. Our results show that some state-of-the-art models are able to
-synthesis compound expressions that generate data-driven text that agrees with gold solutions.
+We introduce the idea of \emph{transparent documents}, web-based data-driven scholarly articles which allow
+readers to explore the relationship to the underlying data by hovering over fragments of text. We present an
+agent-based LLM framework for authoring transparent documents, building on recent developments in data
+provenance for general-purpose programming languages. In particular we build on Fluid, an open source
+functional programming language with a provenance-tracking runtime, which we use as a target platform. Our
+tool consists of two LLM agents which support a human author during the creation of a transparent document. A
+\SuggestionAgent helps identify fragments of text which could plausibly be computed from data, including
+numerical values selected from records or computed by aggregations like sum and mean, comparatives and
+superlatives like ``better than'' and ``largest'', trend-adjectives like ``growing'', and other idiomatic
+quantitative or semi-quantitative phrases. An \InterpretationAgent, given such a fragment, then attempts to
+synthesise a suitable Fluid query over the data which will generate the target string. Our tool then
+interprets the resulting expression in the context of an interactive web page, which turns the generated text
+into an interactable element able to reveal the relevant data. We evaluate our approach on a subset of SciGen,
+an open source dataset consisting of tables from scientific articles and their corresponding descriptions,
+which we extend with hand-generated counterfactual test cases to evaluate how well machine-generated
+expressions generalise in the presence of changes to the underlying data. Our results show that some
+state-of-the-art models are often able to synthesise compound expressions that are extensionally compatible
+with our gold solutions.
 \end{abstract}

### 2025-09-18 paper/abstract.tex
Some rework of abstract.

@@
 agent-based LLM framework for authoring transparent documents, building on recent developments in data
-provenance for general-purpose programming languages. In particular we build on Fluid, an open source
-functional programming language with a provenance-tracking runtime, which we use as a target platform. Our
-tool consists of two LLM agents which support a human author during the creation of a transparent document. A
-\SuggestionAgent helps identify fragments of text which could plausibly be computed from data, including
-numerical values selected from records or computed by aggregations like sum and mean, comparatives and
-superlatives like ``better than'' and ``largest'', trend-adjectives like ``growing'', and other idiomatic
-quantitative or semi-quantitative phrases. An \InterpretationAgent, given such a fragment, then attempts to
-synthesise a suitable Fluid query over the data which will generate the target string. Our tool then
-interprets the resulting expression in the context of an interactive web page, which turns the generated text
-into an interactable element able to reveal the relevant data. We evaluate our approach on a subset of SciGen,
-an open source dataset consisting of tables from scientific articles and their corresponding descriptions,
-which we extend with hand-generated counterfactual test cases to evaluate how well machine-generated
-expressions generalise in the presence of changes to the underlying data. Our results show that some
-state-of-the-art models are often able to synthesise compound expressions that are extensionally compatible
-with our gold solutions.
+provenance for general-purpose programming languages. Our implementation uses Fluid, an open source functional
+programming language with a provenance-tracking runtime, as a target platform. Our tool consists of two LLM
+agents which support a human author during the creation of a transparent document. A \SuggestionAgent helps
+identify fragments of text which could plausibly be computed from data, including numerical values selected
+from records or computed by aggregations like sum and mean, comparatives and superlatives like ``better than''
+and ``largest'', trend-adjectives like ``growing'', and other idiomatic quantitative or semi-quantitative
+phrases. An \InterpretationAgent, given such a fragment, then attempts to synthesise a suitable Fluid query
+over the data which will generate the target string. Our tool then interprets the resulting expression in the
+context of an interactive web page, which turns the generated text into an interactable element able to reveal
+the relevant data. We evaluate our approach on a subset of SciGen, an open source dataset consisting of tables
+from scientific articles and their corresponding descriptions, which we extend with hand-generated
+counterfactual test cases to evaluate how well machine-generated expressions generalise in the presence of
+changes to the underlying data. Our results show that some state-of-the-art models are often able to
+synthesise compound expressions that are extensionally compatible with our gold solutions.
 \end{abstract}

### 2025-09-18 paper/abstract.tex
Abstract tweaks.

@@
 phrases. An \InterpretationAgent, given such a fragment, then attempts to synthesise a suitable Fluid query
-over the data which will generate the target string. Our tool then interprets the resulting expression in the
-context of an interactive web page, which turns the generated text into an interactable element able to reveal
-the relevant data. We evaluate our approach on a subset of SciGen, an open source dataset consisting of tables
-from scientific articles and their corresponding descriptions, which we extend with hand-generated
-counterfactual test cases to evaluate how well machine-generated expressions generalise in the presence of
-changes to the underlying data. Our results show that some state-of-the-art models are often able to
-synthesise compound expressions that are extensionally compatible with our gold solutions.
+over the data which will generate the target string. The resulting expression is spliced into the source code
+for an interactive web page, turning the static text fragment into an interactable data-driven element able to
+reveal the data that underwrites the natural language claim. We evaluate our approach on a subset of SciGen,
+an open source dataset consisting of tables from scientific articles and their corresponding descriptions,
+which we extend with hand-generated counterfactual test cases to evaluate how well machine-generated
+expressions generalise in the presence of changes to the data. Our results show that some state-of-the-art
+models are often able to synthesise compound expressions that are extensionally compatible with our gold
+solutions.
 \end{abstract}

### 2025-09-18 paper/sec/introduction.tex
Merge 1.1 into 1.

@@
-\section{Introduction}
+\section{Introduction: Transparent Documents}
+
+Making sense of data-driven claims is hard, even if the evidence base (data sets and related analyses) is
+open. We see this in peer review, misinformation, retracted papers, \rpnote{supporting citations}.
+\rpnote{Small example here?}
+
+Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} proposed more
+transparent charts and figures by integrating interactive ``provenance queries'' directly into visual outputs.
+These interactive queries allow readers to interact with visual elements to see the data relevant to those
+elements. The key feature of these approaches is that the interactions are provided automatically, via
+\emph{transparent programming languages} (either query languages or general-purpose languages) which track
+dependencies as data flows through a computation.
 
@@
     \item Making the text transparent, with the possibility to highlight the data related to a specific sentence.
-    \item Making sense of data-driven claims is hard, even if the evidence base (code/data) is open
-    \item We see this in peer review, misinformation, retracted papers, …
 \end{itemize}
 
-\rpnote{Small example here}
-
 \subsection{Transparent Text via Interactive Provenance Queries}
-
-Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} proposed
-integrating interactive ``provenance queries'' directly into charts and figures. These interactive queries
-allow readers to interact with visual elements to see the data relevant to those elements. The key feature of
-these approaches is that the interactions are provided automatically, via \emph{transparent programming
-languages} (either query languages or general-purpose languages) which track dependencies as data flows
-through a computation.
 

### 2025-09-18 paper/sec/introduction.tex
Merge some related work.

@@
 
-Making sense of data-driven claims is hard, even if the evidence base (data sets and related analyses) is
-open. We see this in peer review, misinformation, retracted papers, \rpnote{supporting citations}.
-\rpnote{Small example here?}
+Making sense of and verifying data-driven claims is hard, even when the evidence base (data sets and related
+analyses) is open. We see this in peer review, misinformation, retracted papers, \rpnote{supporting
+citations}. The ability of LLMs to assist with these challenges has received considerable attention recently,
+for example in quantity extraction~\citep{bolucu23}, scientific fact-checking~\citep{abu-ahmad25}, and
+scientific figure interpretation~\citep{roberts24}.
 
-Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} proposed more
-transparent charts and figures by integrating interactive ``provenance queries'' directly into visual outputs.
-These interactive queries allow readers to interact with visual elements to see the data relevant to those
-elements. The key feature of these approaches is that the interactions are provided automatically, via
-\emph{transparent programming languages} (either query languages or general-purpose languages) which track
-dependencies as data flows through a computation.
+Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} attempts to
+address this issue through software infrastructure for ``transparent'' charts and figures. These are visual
+outputs that provide interactive \emph{provenance queries}, user interactions (e.g. mousing over a visual
+element) that allow readers to explore how visual features relate to data. In these approaches the
+interactions are provided automatically via an underlying programming language (either a query language or
+general-purpose language) which tracks dependencies as data flows through a computation.
 
-\begin{itemize}
-    \item Making the text transparent, with the possibility to highlight the data related to a specific sentence.
-\end{itemize}
+In this paper, we extend interactive provenance queries to the natural language that accompanies the charts
+and figures. We believe this is an important step because in most articles, charts and figures usually only
+play a supporting role. The main claims are usually presented in natural language, language which also makes
+quantitative statements and contains substantial implicit analytic content. In principle transparent
+programming languages can also be used to create natural language which is bidirectly linked to data, by
+computing key fragments of natural language from data. The following example, adapted from \cite{moosavi21},
+shows how the language presented in this paper can be used to achieve this:
 
-\subsection{Transparent Text via Interactive Provenance Queries}
-
-In most articles, however, charts and figures usually only play a supporting role. The main claims are usually
-presented in natural language, language which also makes quantitative statements and refers to ``behind the
-scenes'' data analyses. In principle transparent programming languages can also be used to create natural
-language which is bidirectly linked to data, by computing key fragments of natural language from data. The
-following example, adapted from \cite{moosavi21}, shows how the language presented in this paper can be used
-to achieve this:
-
-\begin{figure}[h]
+\begin{figure}%[h]
     \centering
@@
 \end{figure}
-
-\input{fig/scigen-1805.02474v1-10-src}
 
@@
 \end{enumerate}
+
+\input{fig/scigen-1805.02474v1-10-src}
 

### 2025-09-18 src/main/java/explorableviz/authoringassistant/Program.java
Drop numToStr from Fig 2.

@@ -333,14 +333,7 @@ public class Program {
         String sitePath = STR."\{path}\{Path.of(this.testCaseFileName).getParent().getFileName()}-\{Path.of(this.testCaseFileName).getFileName()}";
         Files.createDirectories(Path.of(sitePath));
 
-        /* spec generation */
-//        JSONObject spec = new JSONObject();
         String fluidSrcPath = "../fluid";
-//        spec.put("fluidSrcPath", new JSONArray(STR."[\"\{fluidSrcPath}\"]"));
-//        spec.put("linking", true);
-//        spec.put("query", false);
-//        spec.put("inputs", new JSONArray("[\"tableData\"]"));
-
         final String jsonSpec = STR."""
         const jsonSpec = {
                \"fluidSrcPath\": [\"\{fluidSrcPath}\"],
@@ -350,15 +343,6 @@ public class Program {
            }
         """;
 
-  //      try (FileWriter file = new FileWriter(STR."\{sitePath}/spec.json")) {
-  //          ObjectMapper objectMapper = new ObjectMapper();
-  //          Object jsonObject = objectMapper.readValue(spec.toString(), Object.class);
-  //          file.write(objectMapper.writerWithDefaultPrettyPrinter().writeValueAsString(jsonObject));
-  //          file.write("\n");
-  //          file.flush();
-  //      } catch (IOException e) {
-  //          e.printStackTrace();
-  //      }
         /* html generation */
         String html = new String(Files.readAllBytes(Paths.get(new File(STR."\{path}/template.html").toURI())));
         html = html.replaceAll("##TITLE##", String.valueOf(Path.of(this.testCaseFileName).getParent().getFileName()));

### 2025-09-18 src/main/java/explorableviz/authoringassistant/Program.java
Drop numToStr from Fig 2.

@@ -336,14 +336,7 @@ public class Program {
         String sitePath = STR."\{path}\{Path.of(this.testCaseFileName).getParent().getFileName()}-\{Path.of(this.testCaseFileName).getFileName()}";
         Files.createDirectories(Path.of(sitePath));
 
-        /* spec generation */
-//        JSONObject spec = new JSONObject();
         String fluidSrcPath = "../fluid";
-//        spec.put("fluidSrcPath", new JSONArray(STR."[\"\{fluidSrcPath}\"]"));
-//        spec.put("linking", true);
-//        spec.put("query", false);
-//        spec.put("inputs", new JSONArray("[\"tableData\"]"));
-
         final String jsonSpec = STR."""
         const jsonSpec = {
                \"fluidSrcPath\": [\"\{fluidSrcPath}\"],
@@ -353,15 +346,6 @@ public class Program {
            }
         """;
 
-  //      try (FileWriter file = new FileWriter(STR."\{sitePath}/spec.json")) {
-  //          ObjectMapper objectMapper = new ObjectMapper();
-  //          Object jsonObject = objectMapper.readValue(spec.toString(), Object.class);
-  //          file.write(objectMapper.writerWithDefaultPrettyPrinter().writeValueAsString(jsonObject));
-  //          file.write("\n");
-  //          file.flush();
-  //      } catch (IOException e) {
-  //          e.printStackTrace();
-  //      }
         /* html generation */
         String html = new String(Files.readAllBytes(Paths.get(new File(STR."\{path}/template.html").toURI())));
         html = html.replaceAll("##TITLE##", String.valueOf(Path.of(this.testCaseFileName).getParent().getFileName()));

### 2025-09-18 src/main/java/explorableviz/authoringassistant/llm/recognition/CodeLLamaAgent.java
[new] LLM agents for suggestions

@@ -1,0 +1,18 @@
+package explorableviz.authoringassistant.llm.recognition;
+
+import explorableviz.authoringassistant.paragraph.Expression;
+import it.unisa.cluelab.lllm.llm.agents.generic.OLLAMAEvaluatorAgent;
+import org.json.JSONObject;
+
+public class CodeLLamaAgent extends OLLAMAEvaluatorAgent<String> {
+
+    public CodeLLamaAgent(JSONObject setttings) {
+        super(setttings);
+        setModel("codellama:34b");
+    }
+
+    @Override
+    public String parse(String s) {
+        return s;
+    }
+}

### 2025-09-18 src/main/java/explorableviz/authoringassistant/llm/recognition/QwenCoderAgent.java

@@ -1,0 +1,17 @@
+package explorableviz.authoringassistant.llm.recognition;
+
+import it.unisa.cluelab.lllm.llm.agents.generic.OLLAMAEvaluatorAgent;
+import org.json.JSONObject;
+
+public class QwenCoderAgent extends OLLAMAEvaluatorAgent<String> {
+
+    public QwenCoderAgent(JSONObject setttings) {
+        super(setttings);
+        setModel("qwen2.5-coder:32b");
+    }
+
+    @Override
+    public String parse(String s) {
+        return s;
+    }
+}

### 2025-09-18 src/main/java/explorableviz/authoringassistant/SuggestionAgent.java
[new] class for suggestion agent

@@ -1,0 +1,156 @@
+package explorableviz.authoringassistant;
+
+import explorableviz.authoringassistant.paragraph.Expression;
+import explorableviz.authoringassistant.paragraph.Literal;
+import explorableviz.authoringassistant.paragraph.Paragraph;
+import it.unisa.cluelab.lllm.llm.LLMEvaluatorAgent;
+import it.unisa.cluelab.lllm.llm.prompt.PromptList;
+import kotlin.Pair;
+import org.json.JSONObject;
+
+import java.io.IOException;
+import java.lang.reflect.InvocationTargetException;
+import java.nio.charset.StandardCharsets;
+import java.nio.file.Files;
+import java.nio.file.Path;
+import java.util.ArrayList;
+import java.util.HashSet;
+import java.util.List;
+import java.util.regex.Matcher;
+import java.util.regex.Pattern;
+
+public class SuggestionAgent {
+
+    private static final Pattern REPLACE = Pattern.compile("\\[REPLACE value=\"?(.*?)\"?]");
+    private static final Path SYSTEM_PROMPT_PATH = Path.of("suggestion-agent", "system-prompt.txt");
+    private final LLMEvaluatorAgent<String> llm;
+
+    public SuggestionAgent(String agent) throws ClassNotFoundException, InvocationTargetException, NoSuchMethodException, InstantiationException, IllegalAccessException {
+        llm = initialiseAgent(agent);
+    }
+
+    public Program generateTemplateProgram(Program p) throws IOException {
+        String text;
+        if(p.getTestCaseFileName().equals("web-case")) {
+            text = p.getParagraph().getFirst().getValue();
+        } else
+        {
+            text = extractText(p);
+        }
+        PromptList prompts = buildPrompts(text);
+        String result = llm.evaluate(prompts, null);
+        Paragraph paragraph = parseParagraph(result);
+        return new Program(paragraph, p.getDatasets(),p.getImports(),p.getCode(),p.get_loadedDatasets(),p.getTestCaseFileName(),p.getTest_datasets());
+    }
+
+    private Paragraph parseParagraph(String text) {
+        Paragraph paragraph = new Paragraph();
+        Matcher matcher = REPLACE.matcher(text);
+
+        int lastIndex = 0;
+        while (matcher.find()) {
+            // Literal prima del replace
+            if (matcher.start() > lastIndex) {
+                String literalText = text.substring(lastIndex, matcher.start());
+                paragraph.add(new Literal(literalText, null));
[…]

### 2025-09-18 src/main/java/explorableviz/authoringassistant/AuthoringAssistant.java
[update] instegrate suggestion agent in authoring assistant

@@ -1,5 +1,6 @@
 package explorableviz.authoringassistant;
 
+import explorableviz.authoringassistant.llm.LLMDummyAgent;
 import explorableviz.authoringassistant.paragraph.Expression;
 import it.unisa.cluelab.lllm.llm.LLMEvaluatorAgent;
 import it.unisa.cluelab.lllm.llm.prompt.PromptList;
@@ -25,19 +26,25 @@ public class AuthoringAssistant {
     public final Logger logger = Logger.getLogger(AuthoringAssistant.class.getName());
     private final PromptList prompts;
     private final LLMEvaluatorAgent<Expression> llm;
-    private final Program templateProgram;
+    private Program templateProgram;
+    private final SuggestionAgent recogitionAgent;
     private final int runId;
-    public AuthoringAssistant(InContextLearning inContextLearning, String agentClassName, Program templateProgram, int runId) throws Exception {
+    public AuthoringAssistant(InContextLearning inContextLearning, String agentClassName, Program templateProgram, String recognitionAgentClassName, int runId) throws Exception {
         this.prompts = inContextLearning.toPromptList();
         llm = initialiseAgent(agentClassName);
+        this.recogitionAgent = new SuggestionAgent(recognitionAgentClassName);
         this.templateProgram = templateProgram;
         this.runId = runId;
     }
 
     public List<Pair<Program, QueryResult>> executePrograms() throws Exception {
         List<Pair<Program, QueryResult>> results = new ArrayList<>();
-        List<Pair<Program, Expression>> programEdits = templateProgram.asIndividualEdits(templateProgram);
+        List<Pair<Program, Expression>> programEdits;
         int i = 0;
+        if (Settings.isRecognitionAgentEnabled()) {
+            templateProgram = recogitionAgent.generateTemplateProgram(templateProgram);
+        }
+        programEdits = templateProgram.asIndividualEdits(templateProgram);
         while (!programEdits.isEmpty()) {
             Pair<Program, Expression> individualEdit = programEdits.get(i);
             //selection
@@ -53,7 +60,7 @@ public class AuthoringAssistant {
     }
 
     public QueryResult execute(Pair<Program, Expression> test) throws Exception {
-        final int limit = Settings.getLimit();
+        final int limit = llm instanceof LLMDummyAgent ? 1 : Settings.getLimit();
         // Add the input query to the KB that will be sent to the LLM
         int attempts;
         final long start = System.currentTimeMillis();
[…]

### 2025-09-18 src/main/java/explorableviz/authoringassistant/Main.java

@@ -25,15 +25,17 @@ public class Main {
         logger.info(arguments.toString().replace(",", "\n"));
         final ArrayList<Program> programs;
         final InContextLearning inContextLearning;
-        final String agent = arguments.get("agent");
+
         try {
             Settings.init("settings.json");
+            final String agent = Settings.getAuthoringAgentName();
+            final String recognitionAgent = Settings.getRecognitionAgentName();
             //Create directory for logs and json
             Files.createDirectories(Paths.get(STR."\{Settings.getLogFolder()}/json"));
             cleanWebsiteFolders("website/authoring-assistant/");
             inContextLearning = InContextLearning.loadLearningCases(Settings.getSystemPromptPath(), Settings.getNumLearningCaseToGenerate());
             programs = Program.loadPrograms(Settings.getTestCaseFolder(), Settings.maxProgramVariants());
-            final ArrayList<Pair<Program, QueryResult>> results = execute(inContextLearning, agent, programs);
+            final ArrayList<Pair<Program, QueryResult>> results = execute(inContextLearning, agent, recognitionAgent, programs);
             float accuracy = computeAccuracy(results);
             generateLinks();
             writeLog(results, agent, inContextLearning.size());
@@ -95,13 +97,13 @@ public class Main {
         return (float) count / results.size();
     }
 
-    private static ArrayList<Pair<Program, QueryResult>> execute(InContextLearning inContextLearning, String agent, List<Program> programs) throws Exception {
+    private static ArrayList<Pair<Program, QueryResult>> execute(InContextLearning inContextLearning, String agent, String recognitionAgent, List<Program> programs) throws Exception {
         final ArrayList<Pair<Program, QueryResult>> results = new ArrayList<>();
         for(int k = 0; k < Settings.numTestRuns(); k++)
         {
             int programId = 0;
             for (Program program : programs) {
-                AuthoringAssistant workflow = new AuthoringAssistant(inContextLearning, agent, program, k);
+                AuthoringAssistant workflow = new AuthoringAssistant(inContextLearning, agent, program, recognitionAgent, k);
                 logger.info(STR."Analysing program id=\{(programId++)}");
                 results.addAll(workflow.executePrograms());
             }
@@ -135,7 +137,7 @@ public class Main {
                     .filter(Files::isDirectory)
                     .map(Path::getFileName)
                     .map(Path::toString)
[…]

### 2025-09-18 src/main/java/explorableviz/authoringassistant/llm/suggestion/CodeLLamaAgent.java
[update] align package name

@@ -1,6 +1,5 @@
-package explorableviz.authoringassistant.llm.recognition;
+package explorableviz.authoringassistant.llm.suggestion;
 
-import explorableviz.authoringassistant.paragraph.Expression;
 import it.unisa.cluelab.lllm.llm.agents.generic.OLLAMAEvaluatorAgent;
 import org.json.JSONObject;
 

### 2025-09-18 src/main/java/explorableviz/authoringassistant/llm/suggestion/QwenCoderAgent.java

@@ -1,4 +1,4 @@
-package explorableviz.authoringassistant.llm.recognition;
+package explorableviz.authoringassistant.llm.suggestion;
 
 import it.unisa.cluelab.lllm.llm.agents.generic.OLLAMAEvaluatorAgent;
 import org.json.JSONObject;

### 2025-09-18 src/main/java/explorableviz/authoringassistant/llm/suggestion/RecognitionDummyAgent.java

@@ -1,6 +1,5 @@
-package explorableviz.authoringassistant.llm.recognition;
+package explorableviz.authoringassistant.llm.suggestion;
 
-import explorableviz.authoringassistant.paragraph.Expression;
 import it.unisa.cluelab.lllm.llm.LLMEvaluatorAgent;
 import it.unisa.cluelab.lllm.llm.prompt.PromptList;
 import org.json.JSONObject;

### 2025-09-18 src/main/java/explorableviz/authoringassistant/Settings.java
[update] add properties to settings

@@ -102,4 +102,14 @@ public class Settings {
     public static boolean isEditorLoopEnabled() {
         return getSettings().getBoolean("enable-editor-loop");
     }
+    public static boolean isRecognitionAgentEnabled() {
+        return getSettings().getBoolean("enable-suggestion-agent");
+    }
+
+    public static String getRecognitionAgentName() {
+        return getSettings().getString("suggestion-agent-class");
+    }
+    public static String getAuthoringAgentName() {
+        return getSettings().getString("authoring-agent-class");
+    }
 }

### 2025-09-18 paper/sec/introduction.tex
Drop numToStr from Fig 2.
Track tex-common on master.

@@
 Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} attempts to
-address this issue through software infrastructure for ``transparent'' charts and figures. These are visual
-outputs that provide interactive \emph{provenance queries}, user interactions (e.g. mousing over a visual
-element) that allow readers to explore how visual features relate to data. In these approaches the
-interactions are provided automatically via an underlying programming language (either a query language or
+address this issue through software infrastructure which transparently connections charts and figures to their
+data sources. \rpnote{Contrast with LLM-based approach -- trusted infrastructure.} This enables visual outputs
+to provide interactive \emph{provenance queries}, user interactions (e.g. mousing over a visual element) that
+allow readers to explore how visual features relate to data. The key value proposition of these approaches is
+that the interactions are provided automatically via an underlying programming language (query language or
 general-purpose language) which tracks dependencies as data flows through a computation.
@@
 programming languages can also be used to create natural language which is bidirectly linked to data, by
-computing key fragments of natural language from data. The following example, adapted from \cite{moosavi21},
-shows how the language presented in this paper can be used to achieve this:
+computing key fragments of natural language from data. The following example, adapted from \cite{zhang18} via
+\cite{moosavi21}, shows how the language presented in this paper can be used to achieve this \rpnote{Need to
+illustrate counterfactual/generalisability}:
 
@@
     \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-with-pointer.png}
-    \caption{\emph{Transparent text} links fragments of natural language to underlying data, via queries}\label{fig:scigen-example-website}
+    \vspace{1mm}
+    \hrule
+    \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-counterfactual-with-pointer.png}
+    \caption{\emph{Transparent document} example, showing counterfactual behaviour}\label{fig:scigen-example-website}
 \end{figure}

### 2025-09-18 paper/sec/appendix-prompts.tex

\section*{System Prompt Interpretation Agent}

\VerbatimInput[fontsize=\small]{../system-prompt/system-prompt.txt}

\section*{System Prompt Recognition Agent}

\VerbatimInput[fontsize=\small]{../recognition-agent/system-prompt.txt}

### 2025-09-18 paper/sec/authoring-workflow.tex
Tweak.
Roll out 'agent' macros.

@@
 \begin{itemize}
-    \item \textbf{SuggestionAgent}. LLM-based agent which identifies text fragments potentially computable from data.
-    \item \textbf{InterpretationAgent}. LLM-based agent which, for a given text fragment (provided by
+    \item \textbf{\SuggestionAgent}. LLM-based agent which identifies text fragments potentially computable from data.
+    \item \textbf{\InterpretationAgent}. LLM-based agent which, for a given text fragment (provided by
     SuggestionAgent or by the author), attempts to synthesise a Fluid expression computing the target text.
@@
 
-Figure~\ref{fig:architecture} shows the overall architecture of the system.
-The author initially provides the text and accompanying data.
-The SuggestionAgent analyses the input and identifies candidate fragments of natural language which could be replaced by Fluid expressions.
-The user highlightes a fragment of interest, which is then sent to the InterpretationAgent, which generates a corresponding candidate expression.
-Internally the InterpretationAgent uses a compiler-in-the-loop validation process to ensure the expression is well-formed.
-The user is then able to accept the expression (in which case it is incorporated into the program) or reject and leave the text uninterpreted.
+Figure~\ref{fig:architecture} shows the overall architecture of the system. The author initially provides the
+text and accompanying data. The \SuggestionAgent analyses the input and identifies candidate fragments of
+natural language which could be replaced by Fluid expressions. The user highlightes a fragment of interest,
+which is then sent to the \InterpretationAgent, which generates a corresponding candidate expression.
+Internally the \InterpretationAgent uses a compiler-in-the-loop validation process to ensure the expression is
+well-formed. The user is then able to accept the expression (in which case it is incorporated into the
+program) or reject and leave the text uninterpreted.
 
@@
 loopback mechanism enables iterative refinement of expressions, improving accuracy and robustness. There are
-two kinds of loopback error:
+three kinds of loopback error:
 
@@
     zero, array index out of bounds or other runtime problem.
-%    \item \textbf{Invalid Expression Type}. The value of the expression is not a string.
+    \item \textbf{Invalid Expression Type}. The value of the expression is not convertible to a string.
     \item \textbf{Mismatching String}. The string value of the expression is not equal to the target text.

### 2025-09-18 paper/sec/evaluation.tex
Roll out 'agent' macros.

@@
 \paragraph{Initial validation}
-In the initial phase, the generated expression is executed via Fluid CLI.
-If the execution is successful, the result obtained is compared with the expected result.
-In case of error, the output is forwarded to the interpretation agent, which proceeds to regenerate the expression.
+In the initial phase, the generated expression is executed via Fluid CLI. If the execution is successful, the
+result obtained is compared with the expected result. In case of error, the output is forwarded to the
+\InterpretationAgent, which proceeds to regenerate the expression.
 
 \paragraph{Validation via counterfactual testing}
-In the second phase, the data used for execution is deliberately modified in order to verify the consistency of the behaviours.
-Both the expected and generated expressions are executed again and the results are compared.
-If discrepancies emerge, the error is forwarded to the interpretation agent.
-If there are no differences, the expression is considered valid and is returned to the user.
+In the second phase, the data used for execution is deliberately modified in order to verify the consistency
+of the behaviours. Both the expected and generated expressions are executed again and the results are
+compared. If discrepancies emerge, the error is forwarded to the \InterpretationAgent. If there are no
+differences, the expression is considered valid and is returned to the user.
 

### 2025-09-18 paper/sec/evaluation.tex
Corrections to Table 2.
Minor corrections.

@@
     \small
-    \begin{tabular}{p{3cm} p{6cm} p{4cm}}
+    \begin{tabular}{l p{6cm} p{4cm}}
         \hline
-        \textbf{Ambiguity Type} & \textbf{Example} & \textbf{Explanation} \\
+        \textbf{Problem Type} & \textbf{Example} & \textbf{Explanation} \\
         \hline
@@
 
-        missing-data &
+        missing data &
         The average methane emissions for the year 2015 is \hl{1.0} &
@@
     \end{tabular}
-    \caption{Negative cases example without the \emph{Data} column.}
+    \caption{Categories of problematic example}
     \label{tab:ambiguita}

### 2025-09-18 webapp/script.js
Purge some obsolete stuff.

@@ -1,105 +1,0 @@
-let fileHandles = [];
-let jsonData = {};
-
-// Open modal to load JSON
-document.getElementById('openJson').addEventListener('click', () => {
-    const modal = new bootstrap.Modal(document.getElementById('jsonModal'));
-    modal.show();
-});
-
-// Load JSON and fetch files
-document.getElementById('loadJsonButton').addEventListener('click', () => {
-    try {
-        const input = document.getElementById('jsonInput').value;
-        jsonData = JSON.parse(input);
-
-        document.getElementById('systemPrompt').value = jsonData.system_prompt;
-        loadFilesFromPath(jsonData.learningCasesPath);
-        bootstrap.Modal.getInstance(document.getElementById('jsonModal')).hide();
-    } catch (error) {
-        alert('Invalid JSON format');
-    }
-});
-
-// Save system prompt
-document.getElementById('saveSystemPrompt').addEventListener('click', () => {
-    jsonData.system_prompt = document.getElementById('systemPrompt').value;
-    alert('System Prompt updated successfully');
-});
-
-// Load files from a given path
-async function loadFilesFromPath(path) {
-    try {
-        const directoryHandle = await window.showDirectoryPicker();
-        fileHandles = [];
-
-        const fileList = document.getElementById('fileItems');
-        fileList.innerHTML = '';
-
-        for await (const entry of directoryHandle.values()) {
-            if (entry.kind === 'file') {
-                fileHandles.push(entry);
-
-                const listItem = document.createElement('li');
-                listItem.textContent = entry.name;
-                listItem.className = 'list-group-item';
-                listItem.addEventListener('click', () => onFileClick(entry, listItem));
-                fileList.appendChild(listItem);
-            }
-        }
-    } catch (error) {
-        console.error('Error accessing directory:', error);
-    }
-}
-
-// Handle file click
-async function onFileClick(fileHandle, listItem) {
-    try {
-        // Highlight active file
-        document.querySelectorAll('.sidebar li').forEach(li => li.classList.remove('active'));
-        listItem.classList.add('active');
-
-        const file = await fileHandle.getFile();
-        const content = JSON.parse(await file.text());
-
-        displayEditor(content, fileHandle);
-    } catch (error) {
-        console.error('Error reading file:', error);
-    }
-}
-
-// Display editor with file content
-function displayEditor(content, fileHandle) {
-    document.getElementById('editor').classList.remove('hidden');
-
-    document.getElementById('variables').value = JSON.stringify(content.variables, null, 2);
[…]

### 2025-09-18 obsolete/webapp/script.js
Purge some obsolete stuff.

@@ -1,0 +1,105 @@
+let fileHandles = [];
+let jsonData = {};
+
+// Open modal to load JSON
+document.getElementById('openJson').addEventListener('click', () => {
+    const modal = new bootstrap.Modal(document.getElementById('jsonModal'));
+    modal.show();
+});
+
+// Load JSON and fetch files
+document.getElementById('loadJsonButton').addEventListener('click', () => {
+    try {
+        const input = document.getElementById('jsonInput').value;
+        jsonData = JSON.parse(input);
+
+        document.getElementById('systemPrompt').value = jsonData.system_prompt;
+        loadFilesFromPath(jsonData.learningCasesPath);
+        bootstrap.Modal.getInstance(document.getElementById('jsonModal')).hide();
+    } catch (error) {
+        alert('Invalid JSON format');
+    }
+});
+
+// Save system prompt
+document.getElementById('saveSystemPrompt').addEventListener('click', () => {
+    jsonData.system_prompt = document.getElementById('systemPrompt').value;
+    alert('System Prompt updated successfully');
+});
+
+// Load files from a given path
+async function loadFilesFromPath(path) {
+    try {
+        const directoryHandle = await window.showDirectoryPicker();
+        fileHandles = [];
+
+        const fileList = document.getElementById('fileItems');
+        fileList.innerHTML = '';
+
+        for await (const entry of directoryHandle.values()) {
+            if (entry.kind === 'file') {
+                fileHandles.push(entry);
+
+                const listItem = document.createElement('li');
+                listItem.textContent = entry.name;
+                listItem.className = 'list-group-item';
+                listItem.addEventListener('click', () => onFileClick(entry, listItem));
+                fileList.appendChild(listItem);
+            }
+        }
+    } catch (error) {
+        console.error('Error accessing directory:', error);
+    }
+}
+
+// Handle file click
+async function onFileClick(fileHandle, listItem) {
+    try {
+        // Highlight active file
+        document.querySelectorAll('.sidebar li').forEach(li => li.classList.remove('active'));
+        listItem.classList.add('active');
+
+        const file = await fileHandle.getFile();
+        const content = JSON.parse(await file.text());
+
+        displayEditor(content, fileHandle);
+    } catch (error) {
+        console.error('Error reading file:', error);
+    }
+}
+
+// Display editor with file content
+function displayEditor(content, fileHandle) {
+    document.getElementById('editor').classList.remove('hidden');
+
+    document.getElementById('variables').value = JSON.stringify(content.variables, null, 2);
[…]

### 2025-09-18 src/main/java/explorableviz/authoringassistant/SuggestionAgent.java
[refactoring] new path suggestion agent

@@ -22,7 +22,7 @@ import java.util.regex.Pattern;
 public class SuggestionAgent {
 
     private static final Pattern REPLACE = Pattern.compile("\\[REPLACE value=\"?(.*?)\"?]");
-    private static final Path SYSTEM_PROMPT_PATH = Path.of("suggestion-agent", "system-prompt.txt");
+    private static final Path SYSTEM_PROMPT_PATH = Path.of("system-prompt", "suggestion-agent", "system-prompt.txt");
     private final LLMEvaluatorAgent<String> llm;
 
     public SuggestionAgent(String agent) throws ClassNotFoundException, InvocationTargetException, NoSuchMethodException, InstantiationException, IllegalAccessException {

### 2025-09-18 paper/sec/appendix-prompts.tex
Roll out 'agent' macros.
Corrections to Table 2.
Minor corrections.
[update] ref
[update] change path of system prompts

@@
 
-\VerbatimInput[fontsize=\small]{../system-prompt/system-prompt.txt}
+\VerbatimInput[fontsize=\small]{../system-prompt/interpretation-agent/system-prompt.txt}
 
@@
 
-\VerbatimInput[fontsize=\small]{../recognition-agent/system-prompt.txt}
+\VerbatimInput[fontsize=\small]{../system-prompt/suggestion-agent/system-prompt.txt}

### 2025-09-18 paper/sec/authoring-workflow.tex
[update] change path of system prompts

--- 2025-09-17 paper/sec/authoring-workflow.tex
@@
         left=6pt, right=6pt, top=4pt, bottom=4pt]
-        \VerbatimInput[fontsize=\small, formatcom=\normalfont]{../system-prompt/system-prompt.txt}
+        \VerbatimInput[fontsize=\small, formatcom=\normalfont]{../system-prompt/interpretation-agent/system-prompt.txt}
     \end{tcolorbox}

### 2025-09-18 paper/sec/introduction.tex
Supporting citations for intro.

@@
 Making sense of and verifying data-driven claims is hard, even when the evidence base (data sets and related
-analyses) is open. We see this in peer review, misinformation, retracted papers, \rpnote{supporting
-citations}. The ability of LLMs to assist with these challenges has received considerable attention recently,
-for example in quantity extraction~\citep{bolucu23}, scientific fact-checking~\citep{abu-ahmad25}, and
-scientific figure interpretation~\citep{roberts24}.
+analyses) is in the public domain. Expert reviewers often struggle to evaluate empirical
+claims~\citep{weber20}; paper retractions are frequently attributable to (non-fraudulent but significant)
+errors in data management and analysis~\citep{hu25}. The ability of LLMs to assist with these challenges has
+received considerable attention recently, for example in scientific fact-checking~\citep{abu-ahmad25} and
+figure interpretation~\citep{roberts24}.
 
@@
     \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-counterfactual-with-pointer.png}
-    \caption{\emph{Transparent document} example, showing counterfactual behaviour}\label{fig:scigen-example-website}
+    \caption{\emph{Transparent document} example (top); counterfactual behaviour (bottom)}\label{fig:scigen-example-website}
 \end{figure}

### 2025-09-18 paper/sec/introduction.tex
Intro title.

@@
-\section{Introduction: Transparent Documents}
+\section{Introduction: Transparent, Data-Driven Documents}
 
 Making sense of and verifying data-driven claims is hard, even when the evidence base (data sets and related
-analyses) is in the public domain. Expert reviewers often struggle to evaluate empirical
-claims~\citep{weber20}; paper retractions are frequently attributable to (non-fraudulent but significant)
-errors in data management and analysis~\citep{hu25}. The ability of LLMs to assist with these challenges has
-received considerable attention recently, for example in scientific fact-checking~\citep{abu-ahmad25} and
-figure interpretation~\citep{roberts24}.
+analyses) is in the public domain. Expert reviewers often struggle to evaluate empirical claims without author
+-supplied annotations that provide access to underlying data~\citep{weber20}; paper retractions are frequently
+attributable to (non-fraudulent but significant) errors in data management and analysis~\citep{hu25}. The
+ability of LLMs to assist with these challenges has received considerable attention recently, for example in
+scientific fact-checking~\citep{abu-ahmad25} and figure interpretation~\citep{roberts24}.
 
@@
 
-\subsubsection{Target idioms of natural language}
+\subsection{Target idioms of natural language}
 
 NLP aspect of the problem is potentially a big problem space in itself. We will restrict interest to certain
-idiomatic uses of natural language in making/justifying scientific claims.
-Examples \ref{tab:fluid_examples}
+idiomatic uses of natural language in making/justifying scientific claims. Table~\ref{tab:fluid_examples}
 

### 2025-09-18 paper/sec/introduction.tex
Trusted infrastructure.

@@
 
-Making sense of and verifying data-driven claims is hard, even when the evidence base (data sets and related
-analyses) is in the public domain. Expert reviewers often struggle to evaluate empirical claims without author
--supplied annotations that provide access to underlying data~\citep{weber20}; paper retractions are frequently
-attributable to (non-fraudulent but significant) errors in data management and analysis~\citep{hu25}. The
-ability of LLMs to assist with these challenges has received considerable attention recently, for example in
-scientific fact-checking~\citep{abu-ahmad25} and figure interpretation~\citep{roberts24}.
+Making sense of and verifying data-driven claims is hard, even when the evidence base lives in the public
+domain, because of the difficulty in tracing specific claims back to the data. Peer review of empirical claims
+without author-supplied traceability to the underlying data can be challenging~\citep{weber20}; paper
+retractions are often attributable to (non-fraudulent but significant) errors in data management and
+analysis~\citep{hu25}. The use of LLMs to assist with these challenges has received considerable attention
+recently, on problems ranging from fact-checking~\citep{abu-ahmad25} to interpretation of charts and
+figures~\citep{roberts24}, LLM interfaces do not support direct interrogation of visual outputs.
 
-Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} attempts to
-address this issue through software infrastructure which transparently connections charts and figures to their
-data sources. \rpnote{Contrast with LLM-based approach -- trusted infrastructure.} This enables visual outputs
-to provide interactive \emph{provenance queries}, user interactions (e.g. mousing over a visual element) that
-allow readers to explore how visual features relate to data. The key value proposition of these approaches is
-that the interactions are provided automatically via an underlying programming language (query language or
-general-purpose language) which tracks dependencies as data flows through a computation.
+Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} approaches the
+transparency problem from this direction, through software infrastructure which transparently connections
+charts and figures to their data sources. \rpnote{Contrast with LLM-based approach -- trusted infrastructure.}
+This enables visual outputs to provide interactive \emph{provenance queries}, user interactions (e.g. mousing
+over a visual element) that allow readers to explore how visual features relate to data. The value proposition
+of these approaches is that the interactions are provided automatically via trusted infrastructure, namely an
+underlying programming language (say query language or general-purpose language) which tracks dependencies as
+data flows through a computation. However, these approaches are limited to outputs computed from data, such as
+visualisations.
 
@@
     \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-counterfactual-with-pointer.png}
-    \caption{\emph{Transparent document} example (top); counterfactual behaviour (bottom)}\label{fig:scigen-example-website}
+    \caption{\emph{Transparent document} example, showing data-driven natural language (top
+    vs.~bottom)}\label{fig:scigen-example-website}
 \end{figure}

### 2025-09-18 paper/sec/introduction.tex
Rehashing intro.

@@
 underlying programming language (say query language or general-purpose language) which tracks dependencies as
-data flows through a computation. However, these approaches are limited to outputs computed from data, such as
-visualisations.
+data flows through a computation. However, these approaches are limited to outputs computed from data, which
+usually precludes natural language, despite the fact that the main claims of a paper are usually made using
+natural language, with charts and figures only playing a supporting role.
 
-In this paper, we extend interactive provenance queries to the natural language that accompanies the charts
-and figures. We believe this is an important step because in most articles, charts and figures usually only
-play a supporting role. The main claims are usually presented in natural language, language which also makes
-quantitative statements and contains substantial implicit analytic content. In principle transparent
-programming languages can also be used to create natural language which is bidirectly linked to data, by
-computing key fragments of natural language from data. The following example, adapted from \cite{zhang18} via
-\cite{moosavi21}, shows how the language presented in this paper can be used to achieve this \rpnote{Need to
-illustrate counterfactual/generalisability}:
+In this paper, we make the case for combining these two approaches, leveraging the ability of LLMs to
+understand complex technical natural language and to synthesise programs that express the analyses that
+underwrite those claims, and extend interactive provenance queries to the natural language that accompanies
+the charts and figures.  ..language which also makes quantitative statements and contains substantial implicit
+analytic content. In principle transparent programming languages can also be used to create natural language
+which is bidirectly linked to data, by computing key fragments of natural language from data. The following
+example, adapted from \cite{zhang18} via \cite{moosavi21}, shows how the language presented in this paper can
+be used to achieve this \rpnote{Need to illustrate counterfactual/generalisability}:
 

### 2025-09-18 paper/sec/introduction.tex
Rehash para 1.

@@
 Making sense of and verifying data-driven claims is hard, even when the evidence base lives in the public
-domain, because of the difficulty in tracing specific claims back to the data. Peer review of empirical claims
-without author-supplied traceability to the underlying data can be challenging~\citep{weber20}; paper
-retractions are often attributable to (non-fraudulent but significant) errors in data management and
-analysis~\citep{hu25}. The use of LLMs to assist with these challenges has received considerable attention
-recently, on problems ranging from fact-checking~\citep{abu-ahmad25} to interpretation of charts and
-figures~\citep{roberts24}, LLM interfaces do not support direct interrogation of visual outputs.
+domain. A key challenge lies in tracing specific claims back to the underlying data. In peer review, for
+example, empirical claims typically lack author-supplied links to data, making it hard for reviewers to check
+them directly~\citep{weber20}. Paper retractions, meanwhile, are often attributable not to fraud, but to
+simple errors in data management or analysis~\citep{hu25}. Large language models (LLMs) can assist with
+fact-checking~\citep{abu-ahmad25} and interpretation of charts and figures~\citep{roberts24}, but currently
+LLM interfaces do not support direct interrogation of visual outputs.
 
-Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} approaches the
-transparency problem from this direction, through software infrastructure which transparently connections
-charts and figures to their data sources. \rpnote{Contrast with LLM-based approach -- trusted infrastructure.}
-This enables visual outputs to provide interactive \emph{provenance queries}, user interactions (e.g. mousing
-over a visual element) that allow readers to explore how visual features relate to data. The value proposition
-of these approaches is that the interactions are provided automatically via trusted infrastructure, namely an
-underlying programming language (say query language or general-purpose language) which tracks dependencies as
-data flows through a computation. However, these approaches are limited to outputs computed from data, which
-usually precludes natural language, despite the fact that the main claims of a paper are usually made using
-natural language, with charts and figures only playing a supporting role.
+Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25}, on the other
+hand, approaches the transparency problem from this direction, through software infrastructure which
+transparently connections charts and figures to their data sources. \rpnote{Contrast with LLM-based approach
+-- trusted infrastructure.} This enables visual outputs to provide interactive \emph{provenance queries}, user
+interactions (e.g. mousing over a visual element) that allow readers to explore how visual features relate to
+data. The value proposition of these approaches is that the interactions are provided automatically via
+trusted infrastructure, namely an underlying programming language (say query language or general-purpose
+language) which tracks dependencies as data flows through a computation. However, these approaches are limited
+to outputs computed from data, which usually precludes natural language, despite the fact that the main claims
+of a paper are usually made using natural language, with charts and figures only playing a supporting role.
 

### 2025-09-18 paper/sec/introduction.tex
Progress on intro.

@@
 simple errors in data management or analysis~\citep{hu25}. Large language models (LLMs) can assist with
-fact-checking~\citep{abu-ahmad25} and interpretation of charts and figures~\citep{roberts24}, but currently
-LLM interfaces do not support direct interrogation of visual outputs.
+fact-checking~\citep{abu-ahmad25} and interpretation of charts and figures~\citep{roberts24}, but current LLM
+interfaces do not support direct interrogation of visual or other outputs.
 
 Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25}, on the other
-hand, approaches the transparency problem from this direction, through software infrastructure which
-transparently connections charts and figures to their data sources. \rpnote{Contrast with LLM-based approach
--- trusted infrastructure.} This enables visual outputs to provide interactive \emph{provenance queries}, user
-interactions (e.g. mousing over a visual element) that allow readers to explore how visual features relate to
-data. The value proposition of these approaches is that the interactions are provided automatically via
-trusted infrastructure, namely an underlying programming language (say query language or general-purpose
-language) which tracks dependencies as data flows through a computation. However, these approaches are limited
-to outputs computed from data, which usually precludes natural language, despite the fact that the main claims
-of a paper are usually made using natural language, with charts and figures only playing a supporting role.
+hand, takes a more infrastructural approach to the transparency problem, by linking visualisations to their
+data sources directly. This allows visual outputs to support in situ \emph{provenance queries}, user
+interactions (e.g. mousing over a visual element) that reveal how visual features relate to data. The value
+proposition of these approaches is that the interactions are provided automatically via trusted
+infrastructure, namely a query language or general-purpose programming language which tracks dependencies as
+data flows through a computation. However, these approaches to transparency only support direct interrogation
+of outputs computed from data, such as visualisations. The accompanying natural language, which often plays a
+central role in making the claims of the paper, remains hard to interpret.
 

### 2025-09-18 paper/sec/introduction.tex

--- 2025-09-18 paper/sec/introduction.tex
@@
-\section{Introduction: Transparent, Data-Driven Documents}
+\section{Introduction: Transparent Documents}
 
-Making sense of and verifying data-driven claims is hard, even when the evidence base lives in the public
-domain, because of the difficulty in tracing specific claims back to the data. Peer review of empirical claims
-without author-supplied traceability to the underlying data can be challenging~\citep{weber20}; paper
-retractions are often attributable to (non-fraudulent but significant) errors in data management and
-analysis~\citep{hu25}. The use of LLMs to assist with these challenges has received considerable attention
-recently, on problems ranging from fact-checking~\citep{abu-ahmad25} to interpretation of charts and
-figures~\citep{roberts24}, LLM interfaces do not support direct interrogation of visual outputs.
+Making sense of and verifying data-driven claims is hard, even when the evidence base (data sets and related
+analyses) is open. We see this in peer review, misinformation, retracted papers, \rpnote{supporting
+citations}. The ability of LLMs to assist with these challenges has received considerable attention recently,
+for example in quantity extraction~\citep{bolucu23}, scientific fact-checking~\citep{abu-ahmad25}, and
+scientific figure interpretation~\citep{roberts24}.
 
-Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} approaches the
-transparency problem from this direction, through software infrastructure which transparently connections
-charts and figures to their data sources. \rpnote{Contrast with LLM-based approach -- trusted infrastructure.}
-This enables visual outputs to provide interactive \emph{provenance queries}, user interactions (e.g. mousing
-over a visual element) that allow readers to explore how visual features relate to data. The value proposition
-of these approaches is that the interactions are provided automatically via trusted infrastructure, namely an
-underlying programming language (say query language or general-purpose language) which tracks dependencies as
-data flows through a computation. However, these approaches are limited to outputs computed from data, which
-usually precludes natural language, despite the fact that the main claims of a paper are usually made using
-natural language, with charts and figures only playing a supporting role.
+Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} attempts to
+address this issue through software infrastructure which transparently connections charts and figures to their
+data sources. \rpnote{Contrast with LLM-based approach -- trusted infrastructure.} This enables visual outputs
+to provide interactive \emph{provenance queries}, user interactions (e.g. mousing over a visual element) that
+allow readers to explore how visual features relate to data. The key value proposition of these approaches is
+that the interactions are provided automatically via an underlying programming language (query language or
+general-purpose language) which tracks dependencies as data flows through a computation.
 
-In this paper, we make the case for combining these two approaches, leveraging the ability of LLMs to
-understand complex technical natural language and to synthesise programs that express the analyses that
-underwrite those claims, and extend interactive provenance queries to the natural language that accompanies
-the charts and figures.  ..language which also makes quantitative statements and contains substantial implicit
-analytic content. In principle transparent programming languages can also be used to create natural language
-which is bidirectly linked to data, by computing key fragments of natural language from data. The following
-example, adapted from \cite{zhang18} via \cite{moosavi21}, shows how the language presented in this paper can
-be used to achieve this \rpnote{Need to illustrate counterfactual/generalisability}:
+In this paper, we extend interactive provenance queries to the natural language that accompanies the charts
+and figures. We believe this is an important step because in most articles, charts and figures usually only
+play a supporting role. The main claims are usually presented in natural language, language which also makes
+quantitative statements and contains substantial implicit analytic content. In principle transparent
+programming languages can also be used to create natural language which is bidirectly linked to data, by
+computing key fragments of natural language from data. The following example, adapted from \cite{zhang18} via
+\cite{moosavi21}, shows how the language presented in this paper can be used to achieve this \rpnote{Need to
+illustrate counterfactual/generalisability}:
 
@@
     \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-counterfactual-with-pointer.png}
-    \caption{\emph{Transparent document} example, showing data-driven natural language (top
-    vs.~bottom)}\label{fig:scigen-example-website}
+    \caption{\emph{Transparent document} example (top); counterfactual behaviour (bottom)}\label{fig:scigen-example-website}
 \end{figure}
@@
 
-\subsection{Target idioms of natural language}
+\subsubsection{Target idioms of natural language}
 
 NLP aspect of the problem is potentially a big problem space in itself. We will restrict interest to certain
-idiomatic uses of natural language in making/justifying scientific claims. Table~\ref{tab:fluid_examples}
+idiomatic uses of natural language in making/justifying scientific claims.
+Examples \ref{tab:fluid_examples}
 

### 2025-09-18 paper/sec/introduction.tex
Rehashing para 3.

--- 2025-09-18 paper/sec/introduction.tex
@@
 domain. A key challenge lies in tracing specific claims back to the underlying data. In peer review, for
-example, empirical claims typically lack author-supplied links to data, making it hard for reviewers to check
-them directly~\citep{weber20}. Paper retractions, meanwhile, are often attributable not to fraud, but to
+example, empirical claims typically lack author-supplied links to data, making them hard for reviewers to
+check directly~\citep{weber20}. Paper retractions, meanwhile, are often attributable not to fraud, but to
 simple errors in data management or analysis~\citep{hu25}. Large language models (LLMs) can assist with
@@
 Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25}, on the other
-hand, takes a more infrastructural approach to the transparency problem, by linking visualisations to their
-data sources directly. This allows visual outputs to support in situ \emph{provenance queries}, user
-interactions (e.g. mousing over a visual element) that reveal how visual features relate to data. The value
-proposition of these approaches is that the interactions are provided automatically via trusted
-infrastructure, namely a query language or general-purpose programming language which tracks dependencies as
-data flows through a computation. However, these approaches to transparency only support direct interrogation
-of outputs computed from data, such as visualisations. The accompanying natural language, which often plays a
-central role in making the claims of the paper, remains hard to interpret.
+hand, takes a more infrastructural approach to the transparency problem, linking visualisations to their data
+sources directly. This allows visual outputs to support \emph{provenance queries}, user interactions (e.g.
+mousing over a visual element) that reveal how visual features relate to data. The advantage of this approach
+is that the interactions are provided automatically via trusted infrastructure, namely a query language or
+general-purpose programming language which tracks dependencies as data flows through a computation. However,
+these approaches are limited to outputs computed from data, such as visualisations. What is missing is a way
+to extend these ``direct interrogation'' features to natural language itself, where the main claims of most
+scholarly articles are actually made.
 
-In this paper, we make the case for combining these two approaches, leveraging the ability of LLMs to
-understand complex technical natural language and to synthesise programs that express the analyses that
-underwrite those claims, and extend interactive provenance queries to the natural language that accompanies
-the charts and figures.  ..language which also makes quantitative statements and contains substantial implicit
-analytic content. In principle transparent programming languages can also be used to create natural language
-which is bidirectly linked to data, by computing key fragments of natural language from data. The following
-example, adapted from \cite{zhang18} via \cite{moosavi21}, shows how the language presented in this paper can
-be used to achieve this \rpnote{Need to illustrate counterfactual/generalisability}:
+In this paper, we make the case that progress can be made on this problem by combining these two approaches.
+We introduce the idea of \emph{transparent documents}, web-based scholarly articles where key fragments of
+natural language are computed from data, like visualisations and other summaries, and that support interactive
+provenance queries that reveal how key quantitative statements in the document relate to the data. We report
+on a proof-of-concept LLM-based tool which supports a human-in-the-loop workflow for iteratively transforming
+a preexisting opaque document into a transparent, data-driven counterpart. Our approach leverages the ability
+of LLMs to understand technical language and synthesise code, and also the provenance-tracking runtime of an
+open source programming language called Fluid, which we use as our target platform.
+
+The following example, adapted from \cite{zhang18} via \cite{moosavi21}, shows how the language presented in
+this paper can be used to achieve this \rpnote{Need to illustrate counterfactual/generalisability}:
 

### 2025-09-18 src/main/java/explorableviz/authoringassistant/SuggestionAgent.java
[fix] dummy agent return null

@@ -1,5 +1,6 @@
 package explorableviz.authoringassistant;
 
+import explorableviz.authoringassistant.llm.suggestion.RecognitionDummyAgent;
 import explorableviz.authoringassistant.paragraph.Expression;
 import explorableviz.authoringassistant.paragraph.Literal;
 import explorableviz.authoringassistant.paragraph.Paragraph;
@@ -39,6 +40,9 @@ public class SuggestionAgent {
         }
         PromptList prompts = buildPrompts(text);
         String result = llm.evaluate(prompts, null);
+        if(result == null) {
+            return p;
+        }
         Paragraph paragraph = parseParagraph(result);
         return new Program(paragraph, p.getDatasets(),p.getImports(),p.getCode(),p.get_loadedDatasets(),p.getTestCaseFileName(),p.getTest_datasets());
     }

### 2025-09-18 src/main/java/explorableviz/authoringassistant/llm/suggestion/RecognitionDummyAgent.java

@@ -17,6 +17,6 @@ public class RecognitionDummyAgent extends LLMEvaluatorAgent<String> {
     public String evaluate(PromptList list, String s) throws IOException {
         logger.info("Execution of the DummyAgent");
         list.addSystemPrompt("dummy");
-        return "Test paragraph with one replace tag with value [REPLACE value=\"10\"].";
+        return null;
     }
 }

### 2025-09-18 src/main/java/explorableviz/authoringassistant/AuthoringAssistant.java
[refactoring] rename recognition->suggestion

@@ -29,10 +29,10 @@ public class AuthoringAssistant {
     private Program templateProgram;
     private final SuggestionAgent recogitionAgent;
     private final int runId;
-    public AuthoringAssistant(InContextLearning inContextLearning, String agentClassName, Program templateProgram, String recognitionAgentClassName, int runId) throws Exception {
+    public AuthoringAssistant(InContextLearning inContextLearning, String agentClassName, Program templateProgram, String suggestionAgentClassName, int runId) throws Exception {
         this.prompts = inContextLearning.toPromptList();
         llm = initialiseAgent(agentClassName);
-        this.recogitionAgent = new SuggestionAgent(recognitionAgentClassName);
+        this.recogitionAgent = new SuggestionAgent(suggestionAgentClassName);
         this.templateProgram = templateProgram;
         this.runId = runId;
     }
@@ -41,7 +41,7 @@ public class AuthoringAssistant {
         List<Pair<Program, QueryResult>> results = new ArrayList<>();
         List<Pair<Program, Expression>> programEdits;
         int i = 0;
-        if (Settings.isRecognitionAgentEnabled()) {
+        if (Settings.isSuggestionAgentEnabled()) {
             templateProgram = recogitionAgent.generateTemplateProgram(templateProgram);
         }
         programEdits = templateProgram.asIndividualEdits(templateProgram);

### 2025-09-18 src/main/java/explorableviz/authoringassistant/Main.java

@@ -29,13 +29,13 @@ public class Main {
         try {
             Settings.init("settings.json");
             final String agent = Settings.getAuthoringAgentName();
-            final String recognitionAgent = Settings.getRecognitionAgentName();
+            final String suggestionAgent = Settings.getSuggestionAgentName();
             //Create directory for logs and json
             Files.createDirectories(Paths.get(STR."\{Settings.getLogFolder()}/json"));
             cleanWebsiteFolders("website/authoring-assistant/");
             inContextLearning = InContextLearning.loadLearningCases(Settings.getSystemPromptPath(), Settings.getNumLearningCaseToGenerate());
             programs = Program.loadPrograms(Settings.getTestCaseFolder(), Settings.maxProgramVariants());
-            final ArrayList<Pair<Program, QueryResult>> results = execute(inContextLearning, agent, recognitionAgent, programs);
+            final ArrayList<Pair<Program, QueryResult>> results = execute(inContextLearning, agent, suggestionAgent, programs);
             float accuracy = computeAccuracy(results);
             generateLinks();
             writeLog(results, agent, inContextLearning.size());
@@ -97,13 +97,13 @@ public class Main {
         return (float) count / results.size();
     }
 
-    private static ArrayList<Pair<Program, QueryResult>> execute(InContextLearning inContextLearning, String agent, String recognitionAgent, List<Program> programs) throws Exception {
+    private static ArrayList<Pair<Program, QueryResult>> execute(InContextLearning inContextLearning, String agent, String suggestionAgent, List<Program> programs) throws Exception {
         final ArrayList<Pair<Program, QueryResult>> results = new ArrayList<>();
         for(int k = 0; k < Settings.numTestRuns(); k++)
         {
             int programId = 0;
             for (Program program : programs) {
-                AuthoringAssistant workflow = new AuthoringAssistant(inContextLearning, agent, program, recognitionAgent, k);
+                AuthoringAssistant workflow = new AuthoringAssistant(inContextLearning, agent, program, suggestionAgent, k);
                 logger.info(STR."Analysing program id=\{(programId++)}");
                 results.addAll(workflow.executePrograms());
             }

### 2025-09-18 src/main/java/explorableviz/authoringassistant/Settings.java

@@ -102,11 +102,11 @@ public class Settings {
     public static boolean isEditorLoopEnabled() {
         return getSettings().getBoolean("enable-editor-loop");
     }
-    public static boolean isRecognitionAgentEnabled() {
+    public static boolean isSuggestionAgentEnabled() {
         return getSettings().getBoolean("enable-suggestion-agent");
     }
 
-    public static String getRecognitionAgentName() {
+    public static String getSuggestionAgentName() {
         return getSettings().getString("suggestion-agent-class");
     }
     public static String getAuthoringAgentName() {

### 2025-09-18 src/main/java/explorableviz/authoringassistant/SuggestionAgent.java

@@ -1,6 +1,5 @@
 package explorableviz.authoringassistant;
 
-import explorableviz.authoringassistant.llm.suggestion.RecognitionDummyAgent;
 import explorableviz.authoringassistant.paragraph.Expression;
 import explorableviz.authoringassistant.paragraph.Literal;
 import explorableviz.authoringassistant.paragraph.Paragraph;

### 2025-09-18 src/main/java/explorableviz/authoringassistant/llm/suggestion/SuggestionDummyAgent.java

@@ -7,9 +7,9 @@ import org.json.JSONObject;
 import java.io.IOException;
 import java.util.logging.Logger;
 
-public class RecognitionDummyAgent extends LLMEvaluatorAgent<String> {
-    public static Logger logger = Logger.getLogger(RecognitionDummyAgent.class.getName());
-    public RecognitionDummyAgent(JSONObject settings) {
+public class SuggestionDummyAgent extends LLMEvaluatorAgent<String> {
+    public static Logger logger = Logger.getLogger(SuggestionDummyAgent.class.getName());
+    public SuggestionDummyAgent(JSONObject settings) {
         super(settings);
     }
 

### 2025-09-18 paper/sec/introduction.tex
Progress on intro.

@@
 
-Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25}, on the other
-hand, takes a more infrastructural approach to the transparency problem, linking visualisations to their data
+Recent work on data provenance and data visualisation~\citep{psallidas18smoke,bond25}, on the other hand,
+takes a more infrastructural approach to the transparency problem, linking computed outputs to their data
 sources directly. This allows visual outputs to support \emph{provenance queries}, user interactions (e.g.
@@
 
-In this paper, we make the case that progress can be made on this problem by combining these two approaches.
-We introduce the idea of \emph{transparent documents}, web-based scholarly articles where key fragments of
-natural language are computed from data, like visualisations and other summaries, and that support interactive
-provenance queries that reveal how key quantitative statements in the document relate to the data. We report
-on a proof-of-concept LLM-based tool which supports a human-in-the-loop workflow for iteratively transforming
-a preexisting opaque document into a transparent, data-driven counterpart. Our approach leverages the ability
-of LLMs to understand technical language and synthesise code, and also the provenance-tracking runtime of an
-open source programming language called Fluid, which we use as our target platform.
+In this paper, we this gap by combining these two approaches, leveraging the ability of LLMs to understand
+technical language and synthesise queries over data, and utilising the provenance-tracking infrastructure of
+an open source programming language called Fluid (\url{https://f.luid.org/}) designed originally for
+transparent visualisations~\citep{perera22,bond25}. We show how these two technologies can together support
+the creation and deployment of \emph{transparent documents}, web-based scholarly articles where key
+quantitative claims expressed in natural language --- such as the claim that one software system is faster
+than another at solving a problem --- are computed from the relevant data, and that support interactive
+provenance queries that reveal to a reader or reviewer the specific data elements that support the claim.
 
-The following example, adapted from \cite{zhang18} via \cite{moosavi21}, shows how the language presented in
-this paper can be used to achieve this \rpnote{Need to illustrate counterfactual/generalisability}:
+\figref{scigen-example-website}, adapted from \cite{zhang18}, shows how a transparent document appears to a
+user.
 
@@
     \caption{\emph{Transparent document} example, showing data-driven natural language (top
-    vs.~bottom)}\label{fig:scigen-example-website}
+    vs.~bottom)}
+    \label{fig:scigen-example-website}
 \end{figure}
+
+\rpnote{Need to illustrate counterfactual/generalisability}
+
+\paragraph{Contributions}
+
+Our specific contributions are:
+
+\begin{itemize}
+\item a proof-of-concept LLM-based tool for iteratively transforming a preexisting opaque document and
+associated data set into a transparent, data-driven counterpart
+\item an empirical evaluation of how well state-of-the-art models are able to solve the associated
+interpretation and code synthesis problems
+\end{itemize}
+
+We leave implementing a full IDE authoring plugin (based on VSCode or Cursor, for example) for future work.
+
+Our contribution is an agent-based tool which identifies text fragments which are candidates for being made
+data-driven, generates the corresponding Fluid code, and then incrementally incorporates the code into a
+data-driven version of the original document.
 

### 2025-09-18 paper/sec/introduction.tex
Progress on intro.
Progress on introduction.

@@
 Recent work on data provenance and data visualisation~\citep{psallidas18smoke,bond25}, on the other hand,
-takes a more infrastructural approach to the transparency problem, linking computed outputs to their data
-sources directly. This allows visual outputs to support \emph{provenance queries}, user interactions (e.g.
-mousing over a visual element) that reveal how visual features relate to data. The advantage of this approach
-is that the interactions are provided automatically via trusted infrastructure, namely a query language or
-general-purpose programming language which tracks dependencies as data flows through a computation. However,
-these approaches are limited to outputs computed from data, such as visualisations. What is missing is a way
-to extend these ``direct interrogation'' features to natural language itself, where the main claims of most
-scholarly articles are actually made.
+takes a more infrastructural approach to the transparency problem. These approaches link computed outputs to
+their data sources directly, by tracking dependency information. This allows visual outputs to support
+\emph{provenance queries}, user interactions (e.g. mousing over a visual element) that reveal how visual
+features relate to data. The advantage of this approach is that the interactions are provided automatically
+via trusted infrastructure, namely a query language or general-purpose programming language which tracks how
+data flows through a computation. However, these approaches are limited to outputs computed from data, such as
+visualisations. What is missing is a way to extend these ``direct interrogation'' features to natural language
+itself, where the main claims of most scholarly articles are actually made.
 
-In this paper, we this gap by combining these two approaches, leveraging the ability of LLMs to understand
-technical language and synthesise queries over data, and utilising the provenance-tracking infrastructure of
+In this paper, we address this gap by combining these two approaches, leveraging the ability of LLMs to
+understand technical language and synthesise queries over data, plus the provenance-tracking infrastructure of
 an open source programming language called Fluid (\url{https://f.luid.org/}) designed originally for
@@
 
-\paragraph{Contributions}
-
-Our specific contributions are:
+\paragraph{Contributions.} Our specific contributions are:
 

### 2025-09-18 src/main/java/explorableviz/authoringassistant/AuthoringAssistant.java
recogitionAgent -> suggestionAgent.

@@ -27,12 +27,12 @@ public class AuthoringAssistant {
     private final PromptList prompts;
     private final LLMEvaluatorAgent<Expression> llm;
     private Program templateProgram;
-    private final SuggestionAgent recogitionAgent;
+    private final SuggestionAgent suggestionAgent;
     private final int runId;
     public AuthoringAssistant(InContextLearning inContextLearning, String agentClassName, Program templateProgram, String suggestionAgentClassName, int runId) throws Exception {
         this.prompts = inContextLearning.toPromptList();
         llm = initialiseAgent(agentClassName);
-        this.recogitionAgent = new SuggestionAgent(suggestionAgentClassName);
+        this.suggestionAgent = new SuggestionAgent(suggestionAgentClassName);
         this.templateProgram = templateProgram;
         this.runId = runId;
     }
@@ -42,7 +42,7 @@ public class AuthoringAssistant {
         List<Pair<Program, Expression>> programEdits;
         int i = 0;
         if (Settings.isSuggestionAgentEnabled()) {
-            templateProgram = recogitionAgent.generateTemplateProgram(templateProgram);
+            templateProgram = suggestionAgent.generateTemplateProgram(templateProgram);
         }
         programEdits = templateProgram.asIndividualEdits(templateProgram);
         while (!programEdits.isEmpty()) {

### 2025-09-18 src/main/java/explorableviz/authoringassistant/Main.java
Special treatment of 'font' and 'image' folders no longer needed.

@@ -137,7 +137,7 @@ public class Main {
                     .filter(Files::isDirectory)
                     .map(Path::getFileName)
                     .map(Path::toString)
-                    .filter(name -> !name.equals("datasets") && !name.equals("fluid") && !name.equals("font") && !name.equals("css") && !name.equals("image") && !name.equals("shared"))
+                    .filter(name -> !name.equals("datasets") && !name.equals("fluid") && !name.equals("css") && !name.equals("shared"))
                     .sorted()
                     .map(name -> STR."<a href=\"\{name}\">\{name}</a> <br />")
                     .collect(Collectors.joining("\n"));

### 2025-09-18 paper/sec/introduction.tex
Progress on intro.

@@
 \figref{scigen-example-website}, adapted from \cite{zhang18}, shows how a transparent document appears to a
-user.
+user. \rpnote{Make important point about counterfactuals/generalisability}
 
@@
 
-\rpnote{Need to illustrate counterfactual/generalisability}
-
 \paragraph{Contributions.} Our specific contributions are:
@@
 
-We leave implementing a full IDE authoring plugin (based on VSCode or Cursor, for example) for future work.
-
-Our contribution is an agent-based tool which identifies text fragments which are candidates for being made
-data-driven, generates the corresponding Fluid code, and then incrementally incorporates the code into a
-data-driven version of the original document.
+We leave implementing a full authoring plugin for an IDE such as VSCode or Cursor for future work.
 

### 2025-09-19 paper/abstract.tex

@@
 \begin{abstract}
-We introduce the idea of \emph{transparent documents}, web-based data-driven scholarly articles which allow
-readers to explore the relationship to the underlying data by hovering over fragments of text. We present an
-agent-based LLM framework for authoring transparent documents, building on recent developments in data
-provenance for general-purpose programming languages. Our implementation uses Fluid, an open source functional
-programming language with a provenance-tracking runtime, as a target platform. Our tool consists of two LLM
-agents which support a human author during the creation of a transparent document. A \SuggestionAgent helps
-identify fragments of text which could plausibly be computed from data, including numerical values selected
-from records or computed by aggregations like sum and mean, comparatives and superlatives like ``better than''
-and ``largest'', trend-adjectives like ``growing'', and other idiomatic quantitative or semi-quantitative
-phrases. An \InterpretationAgent, given such a fragment, then attempts to synthesise a suitable Fluid query
-over the data which will generate the target string. The resulting expression is spliced into the source code
-for an interactive web page, turning the static text fragment into an interactable data-driven element able to
-reveal the data that underwrites the natural language claim. We evaluate our approach on a subset of SciGen,
-an open source dataset consisting of tables from scientific articles and their corresponding descriptions,
-which we extend with hand-generated counterfactual test cases to evaluate how well machine-generated
-expressions generalise in the presence of changes to the data. Our results show that some state-of-the-art
-models are often able to synthesise compound expressions that are extensionally compatible with our gold
-solutions.
+We introduce \emph{transparent documents}, interactive web-based scholarly articles which allow readers to
+explore the relationship to the underlying data by hovering over fragments of text. We present an agent-based
+LLM framework for authoring transparent documents, building on recent developments in data provenance for
+general-purpose programming languages. Our implementation uses Fluid, an open source functional programming
+language with a provenance-tracking runtime, as a target platform. Our tool consists of two LLM agents which
+support a human author during the creation of a transparent document. A \SuggestionAgent helps identify
+fragments of text which could plausibly be computed from data, including numerical values selected from
+records or computed by aggregations like sum and mean, comparatives and superlatives like ``better than'' and
+``largest'', trend-adjectives like ``growing'', and other idiomatic quantitative or semi-quantitative phrases.
+An \InterpretationAgent, given such a fragment, then attempts to synthesise a suitable Fluid query over the
+data which will generate the target string. The resulting expression is spliced into the source code for an
+interactive web page, turning the static text fragment into an interactable data-driven element able to reveal
+the data that underwrites the natural language claim. We evaluate our approach on a subset of SciGen, an open
+source dataset consisting of tables from scientific articles and their corresponding descriptions, which we
+extend with hand-generated counterfactual test cases to evaluate how well machine-generated expressions
+generalise in the presence of changes to the data. Our results show that some state-of-the-art models are
+often able to synthesise compound expressions extensionally compatible with our gold solutions.
 \end{abstract}

### 2025-09-19 paper/sec/introduction.tex

@@
 
-Making sense of and verifying data-driven claims is hard, even when the evidence base lives in the public
-domain. A key challenge lies in tracing specific claims back to the underlying data. In peer review, for
-example, empirical claims typically lack author-supplied links to data, making them hard for reviewers to
-check directly~\citep{weber20}. Paper retractions, meanwhile, are often attributable not to fraud, but to
-simple errors in data management or analysis~\citep{hu25}. Large language models (LLMs) can assist with
-fact-checking~\citep{abu-ahmad25} and interpretation of charts and figures~\citep{roberts24}, but current LLM
-interfaces do not support direct interrogation of visual or other outputs.
+Making sense of and verifying data-driven claims is hard, even when the evidence base is open and accessible.
+A key challenge lies in tracing specific claims back to the underlying data. In peer review, for example,
+empirical claims typically lack author-supplied links to data, making them hard for reviewers to check
+directly~\citep{weber20}. Paper retractions, meanwhile, are often attributable not to fraud, but to simple
+errors in data management or analysis~\citep{hu25}. The use of large language models (LLMs) to interpret
+scholarly documents has seen considerable attention recently, from fact-checking~\citep{abu-ahmad25} to
+interpretation of charts and figures~\citep{roberts24}, but current LLM interfaces do not support direct
+interrogation of visual or other outputs.
 
 Recent work on data provenance and data visualisation~\citep{psallidas18smoke,bond25}, on the other hand,
-takes a more infrastructural approach to the transparency problem. These approaches link computed outputs to
-their data sources directly, by tracking dependency information. This allows visual outputs to support
-\emph{provenance queries}, user interactions (e.g. mousing over a visual element) that reveal how visual
-features relate to data. The advantage of this approach is that the interactions are provided automatically
-via trusted infrastructure, namely a query language or general-purpose programming language which tracks how
-data flows through a computation. However, these approaches are limited to outputs computed from data, such as
-visualisations. What is missing is a way to extend these ``direct interrogation'' features to natural language
-itself, where the main claims of most scholarly articles are actually made.
+takes a more infrastructural approach to tracing outputs back to data. These approaches link computed outputs
+to their data sources directly, by tracking dependency information. This allows visual outputs to support
+\emph{provenance queries}, user interactions (e.g. mousing over visual elements) that reveal how output
+features relate to data. The advantage of this approach is that the relationships to data sources are exposed
+automatically via trusted infrastructure, typically a query language or general-purpose programming language
+which tracks how data flows through a computation. However, these approaches are limited to outputs computed
+from data, such as visualisations. What is missing is a way to extend these ``direct interrogation'' features
+to natural language itself, where the main claims of most scholarly articles are actually made.
 
-In this paper, we address this gap by combining these two approaches, leveraging the ability of LLMs to
+In this paper, we address this gap by combining two complementary approaches: the ability of LLMs to
 understand technical language and synthesise queries over data, plus the provenance-tracking infrastructure of
-an open source programming language called Fluid (\url{https://f.luid.org/}) designed originally for
-transparent visualisations~\citep{perera22,bond25}. We show how these two technologies can together support
-the creation and deployment of \emph{transparent documents}, web-based scholarly articles where key
-quantitative claims expressed in natural language --- such as the claim that one software system is faster
-than another at solving a problem --- are computed from the relevant data, and that support interactive
-provenance queries that reveal to a reader or reviewer the specific data elements that support the claim.
+an open source programming language called Fluid (\url{https://f.luid.org/})~\citep{perera22,bond25}.
+Together, these two technologies enable the creation of \emph{transparent documents}, web-based scholarly
+articles with two key transparency features:
+\begin{enumerate}
+\item \textbf{Data-driven:} Quantitative statements expressed in natural language --- e.g.~that system $X$ is
+faster than system $Y$ on some task --- are computed from the relevant data, rather than occurring
+merely as static fragments of text.
 
-\figref{scigen-example-website}, adapted from \cite{zhang18}, shows how a transparent document appears to a
-user. \rpnote{Make important point about counterfactuals/generalisability}
+\item \textbf{Data linking:} Readers and reviewers can interactively trace such claims back to the specific
+data elements that support them, through embedded provenance queries.
+\end{enumerate}
+
+\figref{scigen-example-website}, generated from our implementation, illustrates these two features. The upper
+section shows an excerpt from a research paper comparing text encoding techniques~\citep{zhang18}. When a
+reader hovers over the phrase ``does not further improve'', the corresponding data entries are highlighted in
+blue on the left. Other fragments (e.g. ``better than'', ``further improvements'') that refer to the same data
+are also marked, allowing readers to explore supporting and contrasting evidence. The lower section shows a
+counterfactual situation where we imagine the authors' experiments had produced different results: here the
+phrase ``does not further improve'' is replaced by the ``further improves''. Transparent documents thus turn
+static text into interactable, data-driven content able to expose the evidential basis of scholarly claims.
 

### 2025-09-19 paper/fig/natural-language-forms.tex
removed unimplemented cases and add pairwise comparisons

--- 2025-09-15 paper/fig/natural-language-forms.tex
@@
         \end{lstlisting}  \\
-        Rounding* & about 2 degrees warmer                & ~                             \\
-        Cardinal* & three                & ~                             \\
-        Multiplicative* & three times                & ~                             \\
         \rowcolor{gray!20}
@@
         \end{lstlisting} \\
-        Count*                       & ~                & ~                             \\
         \rowcolor{gray!20}
@@
         \end{lstlisting} \\
-        Graded adjective*            & \hl{extremely likely} to be exceeded in the 20th century               & ~                             \\
-        Reference to visual element*                            & ~                & ~                             \\
+        Pairwise Comparisons
+        & Overall the SVM with polynomial degree 1 kernel outperformed all other kernels with other kernels generally offering \hl{better} precision at a higher cost to recall. &
+        \begin{lstlisting}[language=Fluid,numbers=none]
+trendWord
+(mean (getTableColumn \"svm1_p\" tableData))
+(mean
+[ mean (getTableColumn \"naive_bayes_p\" tableData),
+  mean (getTableColumn \"svm2_p\" tableData),
+  mean (getTableColumn \"svm3_p\" tableData),
+  mean (getTableColumn \"svmr_p\" tableData)
+])
+betterWorse
+
+        \end{lstlisting} \\
         \hline

### 2025-09-19 src/main/java/explorableviz/authoringassistant/paragraph/ExpressionCategory.java
fix compiler error

@@ -6,8 +6,8 @@ public enum ExpressionCategory {
     PERCENTAGE("Percentage"),
     DATA_RETRIEVAL("Data Retrieval"),
     DIFFERENCE("Difference"),
-    TOTAL("Total");
-    MINMAX("Min Max")
+    TOTAL("Total"),
+    MINMAX("Min Max");
 
     public final String label;
     ExpressionCategory(String label) {

### 2025-09-19 src/main/java/explorableviz/authoringassistant/variable/ValueOptions.java
[fix] BigDecimal issue

@@ -25,6 +25,9 @@ public abstract class ValueOptions {
         if (value instanceof Float num) {
             return new ValueOptions.Number(num);
         }
+        if (value instanceof BigDecimal num) {
+            return new ValueOptions.Number(num.floatValue());
+        }
         throw new IllegalArgumentException("Unsupported value type: " + value.getClass().getSimpleName());
     }
     public abstract Variables expandVariable(SplittableRandom random, String varName);

### 2025-09-19 src/main/java/explorableviz/authoringassistant/paragraph/ExpressionCategory.java
fix minmax category

@@ -6,8 +6,8 @@ public enum ExpressionCategory {
     PERCENTAGE("Percentage"),
     DATA_RETRIEVAL("Data Retrieval"),
     DIFFERENCE("Difference"),
-    TOTAL("Total");
-    MINMAX("Min Max")
+    TOTAL("Total"),
+    MINMAX("Minmax");
 
     public final String label;
     ExpressionCategory(String label) {

### 2025-09-19 paper/sec/related-work.tex
Merge some related work.
Initial draft - tbc

--- 2025-08-22 sec/related-work.tex
@@
 \label{sec:related}
+
+\paragraph{Argument mining.}
+For over fifteen years the field of NLP has been exploring the problem of argument mining, which involves identifying argumentative structures in text, such as claims, premises, and conclusions, and
+mapping them to formal representations~\citep{}. Early work in this area focused on rule-based approaches, while more recent work has leveraged machine learning and deep learning techniques to improve
+performance~\citep{}. The field has also focused on defining annotation schemes for the task, such as the Argumentative Zoning framework~\citep{}.
+Argument mining remains a challenging task, particularly when it comes to handling implicit arguments and reasoning about the relationships between different argumentative
+components~\citep{}.
+While earlier work focused on specific domains such as legal texts or online resources such as Debatepedia, the community has also started engaging with work that explores the use of argument mining in scientific texts~\citep{liakata2012automatic,...} to better understand the structure of scientific arguments and the relationships between different claims and evidence.
+
+\paragraph{NLP and scientific writing.} The intersection of NLP and scientific writing has gained increasing attention in recent years, with a focus on improving the clarity, coherence, and overall quality of scientific texts.
+Various approaches have been proposed, including the use of automated writing assistants, text simplification techniques, and tools for enhancing the argumentative structure of scientific papers~\citep{..,}.
+These efforts aim to support researchers in effectively communicating their findings and engaging with the broader scientific community.
 
@@
 leaving one of the argument slots implicit, with the context determining the omitted referent. LLMs
-demonstrate considerable competence in resolving these and other more syntatic forms of anaphor such as
+demonstrate considerable competence in resolving these and other more syntactic forms of anaphora such as
 pronouns~\citep{zhu25}, but the resolved referent itself -- concretely, what was being referred to -- remains
@@
 well. \todo{WIP}
-
-\paragraph{Use of LLMs to interpret scientific text.}
-
-The ability of LLMs to interpret natural language in scientific contexts has been studied extensively, for example quantity extraction~\citep{bolucu23}, scientific fact-checking~\citep{abu-ahmad25}.
-
-\paragraph{Transparent programming languages.}
-
-\cite{perera22,bond25,psallidas18smoke}

### 2025-09-20 paper/sec/authoring-workflow.tex
New interpretation agent figure.
Figure.

--- 2025-09-18 paper/sec/authoring-workflow.tex
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/entire-workflow}
-    \caption{Workflow}\label{fig:architecture}
+    \includegraphics[width=0.8\linewidth]{fig/interpretation-agent}
+    \caption{Human-in-the-Loop \InterpretationAgent workflow (Author intervention states in grey)}
+    \label{fig:interpretation-agent}
 \end{figure}
 
-Figure~\ref{fig:architecture} shows the overall architecture of the system. The author initially provides the
-text and accompanying data. The \SuggestionAgent analyses the input and identifies candidate fragments of
-natural language which could be replaced by Fluid expressions. The user highlightes a fragment of interest,
-which is then sent to the \InterpretationAgent, which generates a corresponding candidate expression.
-Internally the \InterpretationAgent uses a compiler-in-the-loop validation process to ensure the expression is
-well-formed. The user is then able to accept the expression (in which case it is incorporated into the
-program) or reject and leave the text uninterpreted.
+Figure~\ref{fig:interpretation-agent} shows the overall architecture of the \InterpretationAgent. The author
+initially provides the text and accompanying data. The \SuggestionAgent analyses the input and identifies
+candidate fragments of natural language which could be replaced by Fluid expressions. The user highlightes a
+fragment of interest, which is then sent to the \InterpretationAgent, which generates a corresponding
+candidate expression. Internally the \InterpretationAgent uses a compiler-in-the-loop validation process to
+ensure the expression is well-formed. The user is then able to accept the expression (in which case it is
+incorporated into the program) or reject and leave the text uninterpreted.
 

### 2025-09-22 src/main/java/explorableviz/authoringassistant/AuthoringAssistant.java
[update] check for literal

@@ -90,6 +90,12 @@ public class AuthoringAssistant {
                 sessionPrompts.addUserPrompt("Null Expression");
                 continue;
             }
+            if(candidate.getExpr().equals(expected.getValue())) {
+                sessionPrompts.addAssistantPrompt(candidate.getExpr() == null ? "NULL" : candidate.getExpr());
+                sessionPrompts.addUserPrompt("ExpressionError: Received a static value instead of a dynamic expression. " +
+                        "Please provide a valid fluid expression that *evaluates to* the expected value, rather than the value itself.");
+                continue;
+            }
             for (Map<String, String> datasets : subProgram.getTest_datasets()) {
 
                 logger.info(STR."Received response: \{candidate.getExpr()}");

### 2025-09-22 paper/sec/related-work.tex
updated rel work

@@
 For over fifteen years the field of NLP has been exploring the problem of argument mining, which involves identifying argumentative structures in text, such as claims, premises, and conclusions, and
-mapping them to formal representations~\citep{}. Early work in this area focused on rule-based approaches, while more recent work has leveraged machine learning and deep learning techniques to improve
-performance~\citep{}. The field has also focused on defining annotation schemes for the task, such as the Argumentative Zoning framework~\citep{}.
+mapping them to formal representations~\citep{palau2009argumentation,lippi2015argument}  .
+Early work in this area focused on rule-based approaches, while more recent work has leveraged machine learning and deep learning techniques to improve
+performance~\citep{stab2014identifying,stab2017parsing}.
+The field has also focused on defining annotation schemes for the task, such as the Argumentative Zoning framework~\citep{teufel2002summarizing,teufel2009towards}.
 Argument mining remains a challenging task, particularly when it comes to handling implicit arguments and reasoning about the relationships between different argumentative
 components~\citep{}.
-While earlier work focused on specific domains such as legal texts or online resources such as Debatepedia, the community has also started engaging with work that explores the use of argument mining in scientific texts~\citep{liakata2012automatic,...} to better understand the structure of scientific arguments and the relationships between different claims and evidence.
+While earlier work focused on specific domains such as legal texts~\citep{toulmin2003uses} or online resources such as Debatepedia~\citep{cabrio2013natural}, the community has also rapidly engaged with work that explores the use of argument mining in scientific texts~\citep{liakata2012automatic,lauscher2018investigating} to better understand the structure of scientific arguments and the relationships between different claims and evidence.
 

### 2025-09-22 src/main/java/explorableviz/authoringassistant/paragraph/Paragraph.java
limit orderToStr

@@ -68,7 +68,8 @@ public class Paragraph extends ArrayList<TextFragment> {
                 literalBuilder.append(literal.getValue());
             } else if (element instanceof Expression expression) {
                 if (exprCount < numComputedExpr) {
-                    Expression computedExpr = (Expression) this.get(k);
+                    TextFragment fragment = this.get(k);
+                    Expression computedExpr = fragment instanceof Expression ? (Expression) fragment : null;
                     p.add(new Literal(literalBuilder.toString(), null));
                     literalBuilder.setLength(0);
                     p.add(computedExpr != null ? computedExpr : expression);

### 2025-09-22 paper/sec/evaluation.tex
first pass result

@@
 
+\subsection{Results}
+In this section, we present the experimental results.
+Overall, the system produced a valid fluid expression in 74.9\% of the cases (S.D. 3\%).
+Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category.
+
+The system achieved performance above 70\% in comparison, data retrieval, and min/max search tasks.
+In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (50\%).
+
+Figure~\ref{fig:success_rate_by_category_count} reports the success rate as a function of the number of categories assigned to each expression.
+Fluid instructions combining multiple operations were associated with multiple categories.
+Results show that the success rate was 75\% for simple expressions (one category), increased to 85\% for two categories,
+and dropped to 58\% when three categories were involved.
+
 \subsection{Negative Cases}

### 2025-09-22 paper/sec/evaluation.tex
first pass on negative cases

@@
 
+Table~\ref{tab:problematic_cases} presents examples of problematic cases in which the system was asked to generate a valid
+fluid expression.
+In these cases, producing a valid expression is challenging because the input data (or the NL query) contains ambiguities
+that prevent the construction of a well-formed expression.
+
+For instance, in the \textit{missing-data} problem, the \InterpretationAgent should generate a function to compute
+ the average of the emissions. However, the input data does not provide information for all months.
+
+In such cases, the system was unable to generate a general expression and instead produced the expected value directly.
+
 \begin{table}[t]
@@
     \caption{Categories of problematic example}
-    \label{tab:ambiguita}
+    \label{tab:problematic_cases}
 \end{table}

### 2025-09-22 paper/sec/conclusion.tex
[new] section conclusion-further works with table

\section{Conclusion}
\label{sec:conclusion}

\input{fig/nl-further-works}

### 2025-09-22 paper/sec/introduction.tex
Start writeup of table 1.

@@
 
-NLP aspect of the problem is potentially a big problem space in itself. We will restrict interest to certain
-idiomatic uses of natural language in making/justifying scientific claims. Table~\ref{tab:fluid_examples}
+Table~\ref{tab:fluid_examples} summarises the natural language idioms we have studied in this paper. In the
+realm of traditional quantitative expressions, and when using state-of-the-art models like gpt4o and gpt5, our
+system is able to recognise direct numerical values, percentages, averages, minima and maxima, ranks, and
+totals, each mapped to a corresponding computation over the source dataset. For example, phrases such as ``the
+Energy Sector accounts for 52.80\% of total emissions'' or ``the expected projected methane emissions for 2030
+is 13.5'' are interpreted as division and aggregation operations over data values. Similarly, ``recorded its
+highest emissions in 2030'' is interpreted as a `maximumBy` query, while rank-based statements such as ``CNN
+gives the lowest accuracy'' are interpreted as ranking computations.
+
+We also consider \emph{trend} expressions, which include relative or comparative natural language phrases
+describing how some property evolves over time. These include single comparisons (e.g.~``training time growing
+from 67 to 106 seconds'') and pairwise comparisons (e.g. “SVM outperformed all other kernels, offering better
+precision”). Such idioms are mapped to higher-level operators like trendWord, growShrink, and betterWorse,
+which combine statistical aggregation with semantic markers of relative performance. Taken together, these
+categories cover a substantial portion of the numerical reasoning idioms found in empirical scientific
+writing.
+
+Notably, the current coverage omits several common linguistic forms. We have yet to study the ability of LLMs
+to interpret approximate quantitative terms like ``around 50\%'' or ``roughly 100 instances'', nor
+interval-based descriptions such as ``between 30 and 40\%” or ``within 5–10 seconds''. We have no reason for
+thinking these will present specific difficulties. ``Graded'' modal adjectives like ``slightly better'',
+``significantly higher'' present more of a challenge linguistically~\cite{lassiter17}. Likewise, idioms
+expressing trends over time in nuanced or qualified ways (e.g. ``steady increase'', ``sharp decline'') or
+distributional descriptors like ``most models'', ``the majority of cases'', or ``usually performs better
+than'' present significant challenges to formal interpretation because of their reliance on non-standard
+quantifiers. Extending coverage to these idioms would substantially deepen our tool's ability to bridge
+natural language reporting with interpretation in terms of the underlying dataset. We discuss this further in
+\secref{future-work}.
 

### 2025-09-22 paper/sec/introduction.tex
Initial progress on target idioms.

--- 2025-09-19 paper/sec/introduction.tex
@@
 
-NLP aspect of the problem is potentially a big problem space in itself. We will restrict interest to certain
-idiomatic uses of natural language in making/justifying scientific claims. Table~\ref{tab:fluid_examples}
+Table~\ref{tab:fluid_examples} summarises the natural language idioms we study in this paper. When using
+state-of-the-art models like gpt4o and gpt5, our system is able to resolve basic table lookups of direct
+numerical values, as well as computations of percentages, averages, minima and maxima, and totals, each mapped
+to a corresponding aggregation over the source dataset. For example, phrases such as ``the Energy Sector
+accounts for 52.80\% of total emissions'' or ``the expected projected methane emissions for 2030 is 13.5'' are
+interpreted as division and averaging operations over the relevant data values. Similarly, ``recorded its
+highest emissions in 2030'' is interpreted as a \kw{maximumBy} query, while rank-based statements such as
+``CNN gives the lowest accuracy'' are interpreted as ranking computations.
+
+We also consider \emph{trend} expressions, which include relative or comparative natural language phrases
+describing how some property evolves over time. These include single comparisons (e.g.~``training time growing
+from 67 to 106 seconds'') and universally quantified comparisons (e.g.~``SVM outperformed all other
+kernels''). Such idioms are mapped to higher-level operators like \kw{trendWord} which are parameterised on
+helper functions such as \kw{growShrink} and \kw{betterWorse} which maps comparisons to appropriate natural
+language phrases. Taken together, these categories cover a substantial portion of the numerical reasoning
+idioms found in empirical scientific writing.
+
+Notably, the current coverage omits several common linguistic forms. We have yet to study the ability of LLMs
+to interpret approximate quantitative terms like \textbf{around 50\%} or \textbf{roughly 100 instances}, nor
+interval-based descriptions such as \textbf{between 30 and 40}\% or \textbf{within 5–10 seconds}. We have no
+reason for thinking these will present specific difficulties, however. So-called \emph{graded} modal
+adjectives which express comparatives and other adjectives in nuanced or qualified ways -- like
+\textbf{slightly better} and \textbf{significantly higher} -- are likely to be more
+challenging~\cite{lassiter17}, especially when combined with trends over time, as in \textbf{steady increase}
+or \textbf{sharp decline}. Distributional descriptors like ``most models'', ``the majority of cases'', or
+``usually performs better than'' present significant challenges to formal interpretation because of their
+reliance on non-standard quantifiers. Extending coverage to these idioms would substantially deepen our tool's
+ability to bridge natural language reporting with interpretation in terms of the underlying dataset. We
+discuss this further in \secref{future-work}.
 

### 2025-09-23 paper/sec/related-work.tex
updated rel work

@@
 Early work in this area focused on rule-based approaches, while more recent work has leveraged machine learning and deep learning techniques to improve
-performance~\citep{stab2014identifying,stab2017parsing}.
-The field has also focused on defining annotation schemes for the task, such as the Argumentative Zoning framework~\citep{teufel2002summarizing,teufel2009towards}.
-Argument mining remains a challenging task, particularly when it comes to handling implicit arguments and reasoning about the relationships between different argumentative
-components~\citep{}.
-While earlier work focused on specific domains such as legal texts~\citep{toulmin2003uses} or online resources such as Debatepedia~\citep{cabrio2013natural}, the community has also rapidly engaged with work that explores the use of argument mining in scientific texts~\citep{liakata2012automatic,lauscher2018investigating} to better understand the structure of scientific arguments and the relationships between different claims and evidence.
+performance~\citep{stab2014identifying,stab2017parsing,eger2017neural}.
+The field has also emphasized defining annotation schemes for the task, such as the Argumentative Zoning framework~\citep{teufel2002summarizing,teufel2009towards} }, as well as schemes more directly tailored to argument mining~\citep{stab2014annotating}.
+The field has focused on various domains, starting from legal texts~\citep{toulmin2003uses}, and has relied on online resources such as Debatepedia~\citep{cabrio2013natural}.
+The community has also rapidly engaged with work that explores the use of argument mining in scientific texts~\citep{liakata2012automatic,lauscher2018investigating} to better understand the structure of scientific arguments and the relationships between different claims and evidence.
+While the advent of LLMs has improved performance~\citep{gorur2025can,vrakatseli2025can}, argument mining remains a challenging task, particularly when it comes to identifying implicit argumentative relations between discourse units, and reasoning about relationships among different argumentative components — especially in cross-domain settings where models struggle to generalise~\citep{gemechu2024aries}.
+While in our work we do not directly perform traditional argument mining, our work similarly relies on the identification of claims in text, which we evaluate following established practices in the field.
 

### 2025-09-23 paper/sec/introduction.tex
Intro tweaks.

--- 2025-09-19 paper/sec/introduction.tex
@@
 
-Making sense of and verifying data-driven claims is hard, even when the evidence base is open and accessible.
-A key challenge lies in tracing specific claims back to the underlying data. In peer review, for example,
-empirical claims typically lack author-supplied links to data, making them hard for reviewers to check
+Making sense of and verifying data-driven claims is hard even with open and accessible data. A key challenge
+lies in tracing specific claims back to the relevant data. In peer review, for example, empirical claims
+typically lack author-supplied links to data, making them hard for reviewers to check
 directly~\citep{weber20}. Paper retractions, meanwhile, are often attributable not to fraud, but to simple
@@
 interpretation of charts and figures~\citep{roberts24}, but current LLM interfaces do not support direct
-interrogation of visual or other outputs.
+interrogation of visual or other outputs for traceability to inputs.
 
-Recent work on data provenance and data visualisation~\citep{psallidas18smoke,bond25}, on the other hand,
-takes a more infrastructural approach to tracing outputs back to data. These approaches link computed outputs
-to their data sources directly, by tracking dependency information. This allows visual outputs to support
+Recent advances in data provenance and data visualisation~\citep{psallidas18smoke,bond25}, on the other hand,
+have pushed in this direction using a more infrastructural approach. These approaches link computed outputs
+to their data sources directly by tracking dependency information. This allows visual outputs to support
 \emph{provenance queries}, user interactions (e.g. mousing over visual elements) that reveal how output

### 2025-09-23 paper/sec/related-work.tex
updated rel work

@@
 
-\paragraph{NLP and scientific writing.} The intersection of NLP and scientific writing has gained increasing attention in recent years, with a focus on improving the clarity, coherence, and overall quality of scientific texts.
-Various approaches have been proposed, including the use of automated writing assistants, text simplification techniques, and tools for enhancing the argumentative structure of scientific papers~\citep{..,}.
-These efforts aim to support researchers in effectively communicating their findings and engaging with the broader scientific community.
+\paragraph{NLP and scientific writing.} 
+The intersection of NLP and scientific writing has gained increasing attention in the last decade, with a focus on improving the clarity, coherence, and overall quality of scientific texts. 
+On the authoring side, tools such as automated writing assistants can support researchers in producing more fluent and accessible text, for instance through grammar correction, summarisation and text simplification~\citep{napoles2017jfleg,stiennon2020learning,takeshita2024cross,saggion2017automatic}.
+Other approaches specifically target the argumentative structure of scientific papers, helping writers to organise contributions and claims more effectively~\citep{lauscher2018arguminsci}.
+Nowadays, general purpose LLMs such as ChatGPT or tools tailored for the task such as Grammarly show the variety of support that NLP tools can provide to authors~\citep{wu2023chatgpt,ahn2024transformative,khalifa2024using}.
+
+At the same time, NLP methods are being developed to assist reviewers and editors in evaluating submissions.
+These include systems for detecting potential issues such as lack of clarity, weak argumentative support, or even factual inconsistencies and up to scientific fraud~\citep{thakkar2025can, fromm2021argument, freedman2024detecting}.
+Such tools can also facilitate meta-reviewing by providing summaries of peer reviews and identifying points of disagreement among reviewers~\citep{kumar2023reviewers}.
+While AI tools show promises in improving the peer-review process~\citep{tyser2024ai}, there are also various risks associated such as breaches of confidentiality, lack of transparency and biases~\citep{perlis2025artificial}.
+Our current work situates itself in the context of NLP tools for supporting the understanding of scientific writing; specifically, it addresses one of the major critiques toward the automation of such process by offering a transparent way of examining its workflow.
 

### 2025-09-23 paper/sec/introduction.tex
Extract figure for SciGen library.

--- 2025-09-23 paper/sec/introduction.tex
@@
 section shows an excerpt from a research paper comparing text encoding techniques~\citep{zhang18}. When a
-reader hovers over the phrase ``does not further improve'', the corresponding data entries are highlighted in
-blue on the left. Other fragments (e.g. ``better than'', ``further improvements'') that refer to the same data
-are also marked, allowing readers to explore supporting and contrasting evidence. The lower section shows a
-counterfactual situation where we imagine the authors' experiments had produced different results: here the
-phrase ``does not further improve'' is replaced by the ``further improves''. Transparent documents thus turn
-static text into interactable, data-driven content able to expose the evidential basis of scholarly claims.
+reader hovers over the phrase \textfrag{does not further improve}, the corresponding data entries are
+highlighted in blue on the left. Other fragments (e.g. \textfrag{better than}, \textfrag{further
+improvements}) that refer to the same data are also marked, allowing readers to explore supporting and
+contrasting evidence. The lower section shows a counterfactual situation where we imagine the authors'
+experiments had produced different results: here the phrase \textfrag{does not further improve} is replaced by
+\textfrag{further improves}. Transparent documents thus turn static text into interactable, data-driven
+content able to expose the evidential basis of scholarly claims.
 
@@
 
-NLP aspect of the problem is potentially a big problem space in itself. We will restrict interest to certain
-idiomatic uses of natural language in making/justifying scientific claims. Table~\ref{tab:fluid_examples}
+Table~\ref{tab:fluid_examples} summarises the natural language idioms studied in this paper. With
+state-of-the-art models like gpt4o and gpt5, our system is able to resolve basic table lookups of direct
+numerical values, as well as computations of percentages, averages, minima and maxima, and totals, each mapped
+to the corresponding aggregation over the source data. For example, phrases such as \textfrag{the Energy
+Sector accounts for 52.80\% of total emissions} or \textfrag{average methane emissions for 2030 is 13.51} are
+interpreted in terms of sum and mean operations respectively over the relevant data values. Similarly,
+\textfrag{recorded its highest emissions in 2030} is interpreted as a \kw{maximumBy} query, while a statement
+such as \textfrag{CNN gives the lowest accuracy} is interpreted as an explicit computation of rank.
+
+We also consider \emph{trend} expressions, which comparative natural language phrases describing how a data
+attribute evolves over time, such as \textfrag{training time growing from 67 to 106 seconds}. Such idioms are
+mapped to higher-order functions like \kw{trendWord} which are parameterised on additional helper functions
+such as \kw{growShrink} and \kw{betterWorse} (\figref{}) which map comparisons to appropriate natural language
+phrases.
+
+\input{fig/scigen-lib}
+
+Taken together, these categories cover a representative portion of the numerical reasoning idioms found in the
+SciGen benchmark. However, some common linguistic forms are not covered in our analysis. We have yet to study
+approximate quantitative terms like \textfrag{around 50\%} or \textfrag{roughly 100 instances}, nor
+interval-based descriptions such as \textfrag{between 30 and 40\%} or \textfrag{within 5–10 seconds}. We have
+no reason for thinking these will present specific difficulties, however. Other forms are likely to be more
+challenging. So-called \emph{graded} modal adverbs~\citep{lassiter17} which modify adjectival comparatives
+like ``better`` -- like \textfrag{slightly better} and \textfrag{significantly higher} -- especially when
+combined with trends over time, as in \textfrag{steadily increasing} or \textfrag{sharply declining}, because
+the interpretation of these qualifiers is quite subjective and contextual. Adverbial quantifiers like
+\textfrag{generally} and \textfrag{usually} also present challenges because colloquial use (``most'' meaning
+``greater than any other alternative proportion'' may differ from more formal uses (``most'' meaning a
+majority, i.e.~greater than 50\% of cases). Extending coverage to these idioms would substantially deepen our
+tool's ability to bridge natural language reporting with interpretation in terms of the underlying dataset. We
+discuss this further in \secref{future-work}.
 

### 2025-09-23 paper/sec/authoring-workflow.tex
Extract figure for SciGen library.
Just summarise system prompt; refer to appendix for full.

@@
         \includegraphics[width=\linewidth]{fig/data-flow-correct}
-        \caption{Author Acceptance workflow}
+        \caption{Author Acceptance pathway}
         \label{fig:data-flow-correct}
@@
         \includegraphics[width=\linewidth]{fig/data-flow-error}
-        \caption{Author Rejection workflow}
+        \caption{Author Rejection pathway}
         \label{fig:dataflow-error}
@@
 \label{subsec:system-prompt}
-Fig. \ref{fig:system-prompt} shows the system prompt defined for the Authoring Assistant.
+
+In the system prompt used to guide the Interpretation Assistant in replacing placeholder tags in user-authored
+paragraphs with executable Fluid expressions, we include the following information about the transparent
+document under construction, the exact task to perform, and constraints on the expected output. The full
+system prompt is given in \appref{system-prompt:interpretation-agent}.
+
+\begin{itemize}
+\item \textbf{Input structure:} summarises the imported datasets, modules, auxililiary definitions in scope in
+the current file, the current Fluid representation of the target paragraph with \kw{[REPLACE …]} tag
+indicating the text fragment to replace, plus the full text for the paragraph.
+\item \textbf{Task description:} description of the LLM task, namely to identify the placeholder, generate a
+Fluid expression that produces the target string using supplied dataset and definitions in scope.
+\item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
+with no extraneous content (such as comments).
+\end{itemize}
 
@@
 \label{subsec:paragraph-with-generated-expression}
-Figure \ref{fig:fluid-example-paragraph} shows the Fluid code of a Paragraph with the expression generated by the authoring assistant.
-Figure \ref{fig:fluid-scigen} shows the Fluid code written for the scigen datasets and examples.
-
-\begin{figure}[h]
-    \small
-    {\lstinputlisting[language=Fluid]{../fluid-common/scigen.fld}}
-    \caption{Fluid library written for scigen}
-    \label{fig:fluid-scigen}
-\end{figure}
+Figure \ref{fig:fluid-example-paragraph} shows the Fluid code of a Paragraph with the expression generated by
+the authoring assistant.
 

### 2025-09-23 paper/sec/authoring-workflow.tex
Just summarise system prompt; refer to appendix for full.

@@
 indicating the text fragment to replace, plus the full text for the paragraph.
-\item \textbf{Task description:} description of the LLM task, namely to identify the placeholder, generate a
-Fluid expression that produces the target string using supplied dataset and definitions in scope.
+\item \textbf{Task description:} description of the LLM task, namely to identify \kw{[REPLACE …]} placeholder
+and generate Fluid expression that produces target string using supplied dataset and definitions in scope.
 \item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
@@
 \end{itemize}
-
-\begin{figure}[h]
-    \centering
-    \begin{tcolorbox}[colback=gray!10, colframe=gray!50, boxrule=0.5pt, arc=2pt,
-        left=6pt, right=6pt, top=4pt, bottom=4pt]
-        \VerbatimInput[fontsize=\small, formatcom=\normalfont]{../system-prompt/interpretation-agent/system-prompt.txt}
-    \end{tcolorbox}
-    \vspace{-0.5em}
-    \caption{System Prompt defined for the Authoring Assistant}
-    \label{fig:system-prompt}
-\end{figure}
 

### 2025-09-23 paper/sec/introduction.tex
add conclusion.
Resolve figref.

@@
 mapped to higher-order functions like \kw{trendWord} which are parameterised on additional helper functions
-such as \kw{growShrink} and \kw{betterWorse} (\figref{}) which map comparisons to appropriate natural language
-phrases.
+such as \kw{growShrink} and \kw{betterWorse} (shown in \figref{fluid-scigen}) which map comparisons to
+appropriate natural language phrases.
 
@@
 tool's ability to bridge natural language reporting with interpretation in terms of the underlying dataset. We
-discuss this further in \secref{future-work}.
+discuss this further in \secref{conclusion:future-work}.
 

### 2025-09-23 paper/sec/evaluation.tex
Import as keyword.

--- 2025-09-18 paper/sec/evaluation.tex
@@
 
-We aggregated the generated expression in three categories: aggregation, trends, quantitative (See Table \ref{tab:fluid_examples}).
-Figure \ref{fig:success_rate_by_category} shows the results obtained in generating expression by categories.
+We aggregated the generated expression in three categories: aggregation, trends, quantitative (See
+Table~\ref{tab:natural-language-forms}). \figref{success_rate_by_category} shows the results obtained in
+generating expression by categories.
 

### 2025-09-23 paper/fig/natural-language-forms.tex
Import as keyword.
Restore layout of Table 1.

@@
         &
-        \begin{lstlisting}[language=Fluid,numbers=none]
+        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
 (findWithKey_ "model" "LSTM" tableData).time_s
@@
         &
-        \begin{lstlisting}[language=Fluid,numbers=none]
+        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
 (record.emissions /
@@
         & The average methane emissions for the year 2030 is \hl{13.51} &
-        \begin{lstlisting}[language=Fluid,numbers=none]
+        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
 (sumEmissions year tableData / length records)
@@
         Min/Max                          & The Energy Sector recorded its highest methane emissions in \hl{2030}             &
-        \begin{lstlisting}[language=Fluid,numbers=none]
+        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
 let maxEntry =
@@
         3-layer stacked CNN gives an accuracy of 81.46\%, which is the \hl{lowest} compared with BiLSTM, and S-LSTM  &
-        \begin{lstlisting}[language=Fluid,numbers=none]
+        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
 let pos =
@@
         The total methane emissions for the year 2030 is \hl{37.74} for Agriculture &
-        \begin{lstlisting}[language=Fluid,numbers=none]
+        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
 sumEmissions year tableData
@@
         & The training time per epoch \hl{growing} from 67 seconds to 106 seconds. &
-        \begin{lstlisting}[language=Fluid,numbers=none]
+        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
 trendWord
@@
         & Overall the SVM with polynomial degree 1 kernel outperformed all other kernels with other kernels generally offering \hl{better} precision at a higher cost to recall. &
-        \begin{lstlisting}[language=Fluid,numbers=none]
+        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
 trendWord
@@
     \caption{Quantitative/semi-quantitative natural language forms considered in this paper}
-    \label{tab:fluid_examples}
+    \label{tab:natural-language-forms}
 \end{table}

### 2025-09-23 paper/fig/natural-language-forms.tex
Restore layout of Table 1.
Improvements to Table 1.

@@
         \hline
-        \textbf{Type}                & \textbf{Example} & \textbf{Generated Expression} \\
+        \textbf{Type}                & \textbf{Example} & \textbf{Gold Solution} \\
         \rowcolor{gray!20}
@@
         Percentage &
-        The Energy Sector accounts for total methane emissions of \hl{52.80\%} in 2030.
+        The Energy Sector accounts for total methane emissions of \hl{52.80}\% in 2030.
         &
@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-let maxEntry =
-    maximumBy (fun x -> x.emissions)
-              (filter
-                (fun x -> x.type == "Energy Sector")
+let maxEntry = maximumBy
+   (fun x -> x.emissions)
+   (filter (fun x -> x.type == "Energy Sector")
       tableData)
@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-let pos =
-  findIndex "model" "CNN"
+let pos = findIndex "model" "CNN"
     (insertionSort cmpTime tableData)
@@
         \end{lstlisting} \\
-        Pairwise Comparisons
+        Pairwise comparisons
         & Overall the SVM with polynomial degree 1 kernel outperformed all other kernels with other kernels generally offering \hl{better} precision at a higher cost to recall. &

### 2025-09-23 paper/sec/introduction.tex

--- 2025-09-23 paper/sec/introduction.tex
@@
 
-Table~\ref{tab:fluid_examples} summarises the natural language idioms studied in this paper. With
+\input{fig/natural-language-forms}
+
+Table~\ref{tab:natural-language-forms} summarises the natural language idioms studied in this paper. With
 state-of-the-art models like gpt4o and gpt5, our system is able to resolve basic table lookups of direct
@@
 to the corresponding aggregation over the source data. For example, phrases such as \textfrag{the Energy
-Sector accounts for 52.80\% of total emissions} or \textfrag{average methane emissions for 2030 is 13.51} are
-interpreted in terms of sum and mean operations respectively over the relevant data values. Similarly,
-\textfrag{recorded its highest emissions in 2030} is interpreted as a \kw{maximumBy} query, while a statement
-such as \textfrag{CNN gives the lowest accuracy} is interpreted as an explicit computation of rank.
+Sector accounts for 52.80\% of total emissions} and \textfrag{average methane emissions for 2030 is 13.51} are
+interpreted in terms of sum and mean respectively over the relevant data values. Similarly, \textfrag{recorded
+its highest emissions in 2030} is interpreted as a \kw{maximumBy} query, while a statement such as
+\textfrag{CNN gives the lowest accuracy} is mapped to an explicit computation of rank.
 
@@
 attribute evolves over time, such as \textfrag{training time growing from 67 to 106 seconds}. Such idioms are
-mapped to higher-order functions like \kw{trendWord} which are parameterised on additional helper functions
-such as \kw{growShrink} and \kw{betterWorse} (\figref{}) which map comparisons to appropriate natural language
-phrases.
+mapped to higher-order functions like \kw{trendWord} parameterised on additional helper functions such as
+\kw{growShrink} and \kw{betterWorse} (shown in \figref{fluid-scigen}) which map comparisons to appropriate
+natural language phrases.
 
@@
 Taken together, these categories cover a representative portion of the numerical reasoning idioms found in the
-SciGen benchmark. However, some common linguistic forms are not covered in our analysis. We have yet to study
-approximate quantitative terms like \textfrag{around 50\%} or \textfrag{roughly 100 instances}, nor
-interval-based descriptions such as \textfrag{between 30 and 40\%} or \textfrag{within 5–10 seconds}. We have
-no reason for thinking these will present specific difficulties, however. Other forms are likely to be more
-challenging. So-called \emph{graded} modal adverbs~\citep{lassiter17} which modify adjectival comparatives
-like ``better`` -- like \textfrag{slightly better} and \textfrag{significantly higher} -- especially when
-combined with trends over time, as in \textfrag{steadily increasing} or \textfrag{sharply declining}, because
-the interpretation of these qualifiers is quite subjective and contextual. Adverbial quantifiers like
-\textfrag{generally} and \textfrag{usually} also present challenges because colloquial use (``most'' meaning
-``greater than any other alternative proportion'' may differ from more formal uses (``most'' meaning a
-majority, i.e.~greater than 50\% of cases). Extending coverage to these idioms would substantially deepen our
-tool's ability to bridge natural language reporting with interpretation in terms of the underlying dataset. We
-discuss this further in \secref{future-work}.
-
-\input{fig/natural-language-forms}
+SciGen benchmark. However, some linguistic forms that commonly arise in scholarly articles are not covered in
+our analysis. We have yet to study approximate quantitative terms like \textfrag{around 50\%} or
+\textfrag{roughly 100 instances}, nor interval-based descriptions such as \textfrag{between 30 and 40\%} or
+\textfrag{within 5–10 seconds}. While we have no reason for thinking these will present specific difficulties,
+other forms are likely to be more challenging. So-called \emph{graded} modal adverbs~\citep{lassiter17} which
+modify adjectival comparatives like \textfrag{better} -- as in \textfrag{slightly better} and
+\textfrag{significantly higher} -- especially when combined with trends over time, as in \textfrag{steadily
+increasing} or \textfrag{sharply declining} -- are likely to prove difficult because the interpretation of
+these qualifiers can be subjective and context-dependent. Adverbial quantifiers like \textfrag{generally} and
+\textfrag{usually} present similar challenges because colloquial use may differ from more formal uses (in some
+situations ``most'' might mean a majority, i.e.~greater than 50\% of cases, but in others may mean only
+``greater than any other alternative proportion''). On the other hand these difficulties also present
+themselves to human readers, so extending coverage to these idioms would substantially deepen our tool's
+ability to bridge natural language reporting with interpretation in terms of the underlying dataset, perhaps
+revealing inconsistent use of technical language on the part of the author. We discuss this further in
+\secref{conclusion:future-work}.
 

### 2025-09-23 paper/sec/introduction.tex
First pass over integrating use cases better into section 1,.

@@
 \figref{scigen-example-website}, generated from our implementation, illustrates these two features. The upper
-section shows an excerpt from a research paper comparing text encoding techniques~\citep{zhang18}. When a
-reader hovers over the phrase \textfrag{does not further improve}, the corresponding data entries are
-highlighted in blue on the left. Other fragments (e.g. \textfrag{better than}, \textfrag{further
-improvements}) that refer to the same data are also marked, allowing readers to explore supporting and
-contrasting evidence. The lower section shows a counterfactual situation where we imagine the authors'
-experiments had produced different results: here the phrase \textfrag{does not further improve} is replaced by
-\textfrag{further improves}. Transparent documents thus turn static text into interactable, data-driven
-content able to expose the evidential basis of scholarly claims.
+section shows an excerpt from \cite{zhang18}, a scholarly article comparing text encoding techniques. When a
+reader hovers over the phrase \textfrag{does not further improve}, the relevant data are highlighted on the
+left. Other fragments (e.g. \textfrag{better than}, \textfrag{further improvements}) that refer to the same
+data are also marked, allowing the reader to explore supporting and contrasting evidence. The lower section
+shows a counterfactual situation where the authors' experiments had produced different results: here the
+phrase \textfrag{does not further improve} is replaced by \textfrag{further improves}.
+
+Transparent documents thus turn static text into interactable, data-driven content able to expose the
+evidential basis of scholarly claims, for which we envisage two use cases. First, when \textbf{authoring}
+content for an online article, a journalist or scientific publisher may wish to link text to raw data so that
+the evidence base for the textual claims can be explored directly from the article. Second, when
+\textbf{reading} a document reporting on findings derived from open data (perhaps a scientific paper or
+climate report), the reader may want to retroactively interpret parts of the text as queries over the
+available data and gradually ``rationally reconstruct'' the relationship between claims in the paper and the
+evidence base. This might be just to aid their own comprehension, or part of a formal peer review process.
 
@@
 \begin{itemize}
-\item a proof-of-concept LLM-based tool for iteratively transforming a preexisting opaque document and
-associated data set into a transparent, data-driven counterpart
-\item an empirical evaluation of how well state-of-the-art models are able to solve the associated
-interpretation and code synthesis problems
+\item A proof-of-concept LLM-based tool for iteratively transforming a preexisting opaque document and
+associated data set into a transparent, data-driven counterpart;
+\item An empirical evaluation of how well state-of-the-art models are able to solve the associated
+interpretation and code synthesis problems.
 \end{itemize}
@@
 We leave implementing a full authoring plugin for an IDE such as VSCode or Cursor for future work.
-
-We envisage two possible use cases for transparent text:
-
-\begin{enumerate}
-\item \textbf{Authoring.} Someone authoring content for an online article, wants to create text
-linked to raw data (and derivative data such as charts or tabular summaries), so that the evidence base for
-the claims made in the text can be explored \emph{in situ}, by interacting with the text.
-
-\item \textbf{Reading or reviewing.} Someone reading textual claims derived from open data (e.g. a
-scientific paper or climate report), wants to retroactively link the text to queries over the available data
-and gradually ``rationally reconstruct'' the relationship between the claims in the paper and the evidence
-base. Perhaps just to aid their own comprehension, or to provide some kind of justified peer review.
-\end{enumerate}
 

### 2025-09-23 paper/sec/introduction.tex

@@
 evidential basis of scholarly claims, for which we envisage two use cases. First, when \textbf{authoring}
-content for an online article, a journalist or scientific publisher may wish to link text to raw data so that
-the evidence base for the textual claims can be explored directly from the article. Second, when
-\textbf{reading} a document reporting on findings derived from open data (perhaps a scientific paper or
-climate report), the reader may want to retroactively interpret parts of the text as queries over the
-available data and gradually ``rationally reconstruct'' the relationship between claims in the paper and the
-evidence base. This might be just to aid their own comprehension, or part of a formal peer review process.
+content for an online article, a journalist or scientific publisher may wish to provide text which is linked
+to the underlying data so that the evidence base for the textual claims can be explored directly from the
+article. Second, when \textbf{reading} a document reporting on findings derived from open data (perhaps a
+scientific paper or climate report), the reader may want to retroactively interpret parts of the text as
+queries over the available data and gradually ``rationally reconstruct'' the relationship between claims in
+the paper and the evidence base. This might be just to aid their own comprehension, or part of a formal peer
+review process.
 
@@
 
-We leave implementing a full authoring plugin for an IDE such as VSCode or Cursor for future work.
+We leave implementing a full Copilot-like authoring plugin for an IDE such as VSCode or Cursor for future
+work.
 

### 2025-09-23 src/main/java/explorableviz/authoringassistant/paragraph/Paragraph.java
1110.3094_gold_solution

@@ -68,8 +68,7 @@ public class Paragraph extends ArrayList<TextFragment> {
                 literalBuilder.append(literal.getValue());
             } else if (element instanceof Expression expression) {
                 if (exprCount < numComputedExpr) {
-                    TextFragment fragment = this.get(k);
-                    Expression computedExpr = fragment instanceof Expression ? (Expression) fragment : null;
+                    Expression computedExpr = (Expression) this.get(k);
                     p.add(new Literal(literalBuilder.toString(), null));
                     literalBuilder.setLength(0);
                     p.add(computedExpr != null ? computedExpr : expression);

### 2025-09-23 paper/sec/related-work.tex

@@
 performance~\citep{stab2014identifying,stab2017parsing,eger2017neural}.
-The field has also emphasized defining annotation schemes for the task, such as the Argumentative Zoning framework~\citep{teufel2002summarizing,teufel2009towards} }, as well as schemes more directly tailored to argument mining~\citep{stab2014annotating}.
+The field has also emphasized defining annotation schemes for the task, such as the Argumentative Zoning framework~\citep{teufel2002summarizing,teufel2009towards}, as well as schemes more directly tailored to argument mining~\citep{stab2014annotating}.
 The field has focused on various domains, starting from legal texts~\citep{toulmin2003uses}, and has relied on online resources such as Debatepedia~\citep{cabrio2013natural}.

### 2025-09-23 paper/sec/authoring-workflow.tex
Prep for IA summary.

@@
 Figure~\ref{fig:overall-architecture} shows the system architecture and an example flow.
-The system is composed of two agents:
+The system is composed of two LLM-based agents:
 
 \begin{itemize}
-    \item \textbf{\SuggestionAgent}. LLM-based agent which identifies text fragments potentially computable from data.
-    \item \textbf{\InterpretationAgent}. LLM-based agent which, for a given text fragment (provided by
-    SuggestionAgent or by the author), attempts to synthesise a Fluid expression computing the target text.
+    \item \textbf{\SuggestionAgent}. Identifies text fragments potentially computable from data.
+    \item \textbf{\InterpretationAgent}. For a given text fragment (provided by SuggestionAgent or author),
+    attempts to synthesise Fluid expression computing target fragment.
 \end{itemize}
@@
 
-Figure~\ref{fig:interpretation-agent} shows the overall architecture of the \InterpretationAgent. The author
-initially provides the text and accompanying data. The \SuggestionAgent analyses the input and identifies
-candidate fragments of natural language which could be replaced by Fluid expressions. The user highlightes a
-fragment of interest, which is then sent to the \InterpretationAgent, which generates a corresponding
-candidate expression. Internally the \InterpretationAgent uses a compiler-in-the-loop validation process to
-ensure the expression is well-formed. The user is then able to accept the expression (in which case it is
-incorporated into the program) or reject and leave the text uninterpreted.
+The author initially provides the text and accompanying data. The \SuggestionAgent analyses the inputs and
+identifies candidate fragments of natural language which could be replaced by Fluid expressions.
+Figure~\ref{fig:interpretation-agent} then shows the overall workflow of the \InterpretationAgent, with the
+human author in the loop. The user highlights a fragment of interest, which is then sent to the
+\InterpretationAgent, which generates a corresponding candidate expression. Internally the
+\InterpretationAgent uses a compiler-in-the-loop validation process to ensure the expression is well-formed.
+The user is then able to accept the expression (in which case it is incorporated into the program) or reject
+and leave the text uninterpreted.
 

### 2025-09-23 paper/sec/authoring-workflow.tex
Some progress on IA workflow.

@@
 identifies candidate fragments of natural language which could be replaced by Fluid expressions.
-Figure~\ref{fig:interpretation-agent} then shows the overall workflow of the \InterpretationAgent, with the
-human author in the loop. The user highlights a fragment of interest, which is then sent to the
-\InterpretationAgent, which generates a corresponding candidate expression. Internally the
-\InterpretationAgent uses a compiler-in-the-loop validation process to ensure the expression is well-formed.
-The user is then able to accept the expression (in which case it is incorporated into the program) or reject
-and leave the text uninterpreted.
+Figure~\ref{fig:interpretation-agent} then shows the human-in-the loop workflow of the \InterpretationAgent.
+The entry point is the selection by the author of a fragment of text $s$ to interpret, a fragment which may or
+may not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
+expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression is validated through a
+closed loop program synthesis step to check that it is well-formed, and, if so, that it evaluates to the
+target fragment of text $s$. If $e$ indeed evaluates to $s$, then the system replaces $s$ in the original
+top-level string with the interpolation expression \kw{\{$e$\}}. This is the primary successful path through
+the loop and returns the system to the ideal state where it is waiting for another top-level interaction from
+the author.
+
+Otherwise, the Fluid command-line either returned an error (either a syntax error or runtime error), or it
+succeeded but $e$ evaluated to some other string $s' \neq s$. Both cases trigger prompt augmentation with an
+appropriate error message, and the system retries generation. In cases where the expression remains invalid or
+mismatched, the user may intervene by resolving the mismatch manually or rejecting the candidate. Once a valid
+expression is approved, the updated text is spliced back into the document, and any downstream interactions
+(e.g. reloading the web page) are checked. This closed-loop design combines automated synthesis with
+validation and author oversight, ensuring both correctness and usability of the generated code.
+
+The user highlights a fragment of interest, which is then sent to the \InterpretationAgent, which generates a
+corresponding candidate expression. Internally the \InterpretationAgent uses a compiler-in-the-loop validation
+process to ensure the expression is well-formed. The user is then able to accept the expression (in which case
+it is incorporated into the program) or reject and leave the text uninterpreted.
 
@@
 
-\subsection{System Prompt}
-\label{subsec:system-prompt}
+\subsection{Prompt Design}
+\label{sec:prompt-design}
 

### 2025-09-23 paper/sec/authoring-workflow.tex
Progess on workflow.

@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\linewidth]{fig/interpretation-agent}
-    \caption{Human-in-the-Loop \InterpretationAgent workflow (Author intervention states in grey)}
-    \label{fig:interpretation-agent}
-\end{figure}
-
-The author initially provides the text and accompanying data. The \SuggestionAgent analyses the inputs and
-identifies candidate fragments of natural language which could be replaced by Fluid expressions.
-Figure~\ref{fig:interpretation-agent} then shows the human-in-the loop workflow of the \InterpretationAgent.
-The entry point is the selection by the author of a fragment of text $s$ to interpret, a fragment which may or
-may not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
-expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression is validated through a
-closed loop program synthesis step to check that it is well-formed, and, if so, that it evaluates to the
-target fragment of text $s$. If $e$ indeed evaluates to $s$, then the system replaces $s$ in the original
-top-level string with the interpolation expression \kw{\{$e$\}}. This is the primary successful path through
-the loop and returns the system to the ideal state where it is waiting for another top-level interaction from
-the author.
+The author initially imports the text and accompanying data into the system, which creates a Fluid
+representation of the target document. Initially that representation is just a string literal of the form
+\kw{"""..."""}, where the triple quotes are the Fluid syntax for an \emph{interpolated string}, i.e.~a string
+literal where expressions of the form \kw{\{$e$\}} are permitted within the string. The \SuggestionAgent
+analyses the inputs and identifies candidate fragments of natural language which could be replaced by Fluid
+expressions. Figure~\ref{fig:interpretation-agent} then shows the human-in-the loop workflow of the
+\InterpretationAgent. The entry point is the selection by the author of a fragment of text $s$ to interpret, a
+fragment which may or may not have been highlighted previously by the \SuggestionAgent. The system then
+generates a candidate Fluid expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression
+is validated through a closed loop program synthesis step to check that it is well-formed, and, if so, that it
+evaluates to the target fragment of text $s$. If $e$ indeed evaluates to $s$, then the system replaces $s$ in
+the original top-level string with the interpolation expression \kw{\{$e$\}}. This is the primary successful
+path through the loop and returns the system to the ideal state where it is waiting for another top-level
+interaction from the author.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\linewidth]{fig/interpretation-agent}
+    \caption{Human-in-the-Loop workflow (states requiring human intervention in grey)}
+    \label{fig:interpretation-agent}
+\end{figure}
+
+\subsection{Prompt Design}
+\label{sec:prompt-design}
+
+\paragraph{System Prompt.}
+For the Interpretation Assistant, the system prompt guides the LLM in replacing placeholder tags in the
+supplied paragraph with executable Fluid expressions. It includes the following information about the document
+under construction, the exact task to perform, and constraints on the expected output. The full system prompt
+is given in \appref{system-prompt:interpretation-agent}.
+
+\begin{itemize}
+\item \textbf{Input structure:} summarises the imported datasets, modules, auxililiary definitions in scope in
+the current file, the current Fluid representation of the target paragraph with \kw{[REPLACE …]} tag
+indicating the text fragment to replace, plus the full text for the paragraph.
+\item \textbf{Task description:} description of the LLM task, namely to identify \kw{[REPLACE …]} placeholder
+and generate Fluid expression that produces target string using supplied dataset and definitions in scope.
+\item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
+with no extraneous content (such as comments).
+\end{itemize}
+
+\rpnote{Now something about the rest of the prompt.}
+
+\begin{figure}[t]
     \centering
@@
 \end{figure}
-
-\subsection{Prompt Design}
-\label{sec:prompt-design}
-
-In the system prompt used to guide the Interpretation Assistant in replacing placeholder tags in user-authored
-paragraphs with executable Fluid expressions, we include the following information about the transparent
-document under construction, the exact task to perform, and constraints on the expected output. The full
-system prompt is given in \appref{system-prompt:interpretation-agent}.
-
-\begin{itemize}
-\item \textbf{Input structure:} summarises the imported datasets, modules, auxililiary definitions in scope in
-the current file, the current Fluid representation of the target paragraph with \kw{[REPLACE …]} tag
-indicating the text fragment to replace, plus the full text for the paragraph.
-\item \textbf{Task description:} description of the LLM task, namely to identify \kw{[REPLACE …]} placeholder
-and generate Fluid expression that produces target string using supplied dataset and definitions in scope.
-\item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
-with no extraneous content (such as comments).
-\end{itemize}
-
-\subsection{Editor Loop}
-\label{subsec:editor-loop}
-Design of main loop that would be integrated into an IDE. The configuration (state) that is being maintained
-is a \kw{Paragraph} the user is authoring in Fluid. Consists of a sequence of text fragments, some of which
-are uninterpreted (plain literals), the remainder have underlying expressions linking the text to raw or derived
-data.
-
-Workflow:
-\begin{enumerate}
-\item User selects (a substring of) one of the literal text fragments, indicating they want to link it in this
-way
-\item Authoring tool generates expression which is either:
-  \begin{enumerate}
-  \item computes the selected text (e.g.~``greater than'' might be computed by comparing two numbers)
-  \item becomes the formal meaning of the selected text (e.g. the text ``carbon intensity of methane
-emissions'' can be understood in a specific context as referring to the numerical value ``34 gCO2eq/kWh''
-  \end{enumerate}
-\item Some kind of validation step, including both automated validation (e.g. checking for runtime errors) and
-user validation; if validation fails, goto (2)
-\item Partition code into additional definitions and expression; add definitions to main source file, and
-incorporate expression into \kw{Paragraph}
-\item Goto (1) with updated editor state
-\end{enumerate}
 

### 2025-09-23 paper/sec/authoring-workflow.tex
Purge some obsolete stuff.
Rehash intro a bit.

@@
 the authoring assistant.
-
-\subsection{Turning validation errors into improved prompts}\label{subsec:turning-validation-errors-into-improved-prompts}
-E.g.:
-\begin{enumerate}
-\item Turn ``Definition not found'' into prompt to generate definition?
-\end{enumerate}
-
-\subsection{IDE integration}\label{subsec:ide-integration}
-
-Could then be integrated into desktop IDE like VSCode or online IDE like CodeMirror. Might make a good
-internship project, but could also be out-of-scope for this paper.
-
-\subsection{Other potential enhancements}\label{subsec:other-potential-enhancements}
-
-We could also think about using an LLM in a couple of other complementary ways:
-\begin{itemize}
-\item identifying text fragments which might be linked;
-\item validating generated expressions (perhaps by proposing test cases)
-\end{itemize}

### 2025-09-23 paper/sec/introduction.tex
Rehash intro a bit.

@@
 \figref{scigen-example-website}, generated from our implementation, illustrates these two features. The upper
-section shows an excerpt from \cite{zhang18}, a scholarly article comparing text encoding techniques. When a
-reader hovers over the phrase \textfrag{does not further improve}, the relevant data are highlighted on the
-left. Other fragments (e.g. \textfrag{better than}, \textfrag{further improvements}) that refer to the same
-data are also marked, allowing the reader to explore supporting and contrasting evidence. The lower section
-shows a counterfactual situation where the authors' experiments had produced different results: here the
-phrase \textfrag{does not further improve} is replaced by \textfrag{further improves}.
-
-Transparent documents thus turn static text into interactable, data-driven content able to expose the
-evidential basis of scholarly claims, for which we envisage two use cases. First, when \textbf{authoring}
-content for an online article, a journalist or scientific publisher may wish to provide text which is linked
-to the underlying data so that the evidence base for the textual claims can be explored directly from the
-article. Second, when \textbf{reading} a document reporting on findings derived from open data (perhaps a
-scientific paper or climate report), the reader may want to retroactively interpret parts of the text as
-queries over the available data and gradually ``rationally reconstruct'' the relationship between claims in
-the paper and the evidence base. This might be just to aid their own comprehension, or part of a formal peer
-review process.
+section shows a ``transparent'' excerpt from \cite{zhang18}, a scholarly article comparing text encoding
+techniques. When a reader hovers over the phrase \textfrag{does not further improve}, the relevant data are
+highlighted on the left. Other fragments (e.g. \textfrag{better than}, \textfrag{further improvements}) that
+refer to the same data are also marked, allowing the reader to explore supporting and contrasting evidence.
+The lower section shows a counterfactual situation where the authors' experiments had produced different
+results: here the phrase \textfrag{does not further improve} is replaced by \textfrag{further improves}.
 
@@
 \end{figure}
+
+This transparent version of the document was implemented in Fluid. The source code is shown in
+\figref{fluid-example-paragraph}, and makes use of several helper functions, a representative subset of which
+are shown in \figref{fluid-scigen}. Both the provenance-tracking runtime of Fluid \emph{and} the LLM-based
+authoring support contribute to the solution: the Fluid runtime provides the interactions, and the LLM-based
+authoring tool makes the whole process practical. Without tool support, authoring a scientific document as a
+(hand-crafted) \emph{program} is unlikely to be a feasible task.
+
+AI-assisted authoring of transparent documents thus support turning static text into interactable, data-driven
+content able to expose the evidential basis of scholarly claims, for which we envisage two use cases. First,
+when \textbf{authoring} content for an online article, a journalist or scientific publisher may wish to
+provide text which is linked to the underlying data so that the evidence base for the textual claims can be
+explored directly from the article. Second, when \textbf{reading} a document reporting on findings derived
+from open data (perhaps a scientific paper or climate report), the reader may want to retroactively interpret
+parts of the text as queries over the available data and gradually ``rationally reconstruct'' the relationship
+between claims in the paper and the evidence base. This might be just to aid their own comprehension, or part
+of a formal peer review process.
 
@@
 \secref{conclusion:future-work}.
-
-\subsection{Contributions}
-
-\begin{itemize}
-    \item Design and proof-of-concept implementation of AI-assisted workflow for authoring transparent text
-    (\secref{authoring-workflow})
-    \item Empirical evaluation of how effective current LLMs are at providing the ``AI-assisted'' part
-\end{itemize}

### 2025-09-23 paper/sec/authoring-workflow.tex
Some progress on workflow. Need a bit of decomposition though.

@@
 
-Figure~\ref{fig:overall-architecture} shows the system architecture and an example flow.
-The system is composed of two LLM-based agents:
+Our authoring tool is composed of two LLM-based agents:
 
@@
 analyses the inputs and identifies candidate fragments of natural language which could be replaced by Fluid
-expressions. Figure~\ref{fig:interpretation-agent} then shows the human-in-the loop workflow of the
-\InterpretationAgent. The entry point is the selection by the author of a fragment of text $s$ to interpret, a
-fragment which may or may not have been highlighted previously by the \SuggestionAgent. The system then
-generates a candidate Fluid expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression
-is validated through a closed loop program synthesis step to check that it is well-formed, and, if so, that it
-evaluates to the target fragment of text $s$. If $e$ indeed evaluates to $s$, then the system replaces $s$ in
-the original top-level string with the interpolation expression \kw{\{$e$\}}. This is the primary successful
-path through the loop and returns the system to the ideal state where it is waiting for another top-level
-interaction from the author.
+expressions. The system then enters the human-in-the loop workflow shown in
+Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The system is
+initially waiting for the author to select a fragment of text $s$ to interpret, a fragment which may or may
+not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
+expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression is validated through a
+closed loop program synthesis step which checks that it is well-formed and evaluates to the target fragment
+$s$. If this process succeeds, the system replaces the selected substring $s$ with the interpolation
+expression \kw{\{$e$\}}, creating a new document configuration. This is the primary successful path through
+the workflow and returns the system to the entry state where it is waiting for another top-level interaction
+from the author.
 
-Otherwise, the Fluid command-line either returned an error (either a syntax error or runtime error), or it
+Otherwise, the Fluid command-line either returned an error (either a syntax error or runtime error), or
 succeeded but $e$ evaluated to some other string $s' \neq s$. Both cases trigger prompt augmentation with an
-appropriate error message, and the system retries generation. In cases where the expression remains invalid or
-mismatched, the user may intervene by resolving the mismatch manually or rejecting the candidate. Once a valid
-expression is approved, the updated text is spliced back into the document, and any downstream interactions
-(e.g. reloading the web page) are checked. This closed-loop design combines automated synthesis with
-validation and author oversight, ensuring both correctness and usability of the generated code.
+appropriate error message, and the system retries generation. Failure of the code synthesis step means that,
+even after multiple retries with an augmented prompt, $e$ remained invalid or evaluated to $s' \neq s$. If $e$
+was invalid (resulting in an error), no remedial action is open to the author. This is considered an
+unsuccessful path through the workflow and returns the system to the start state. In the $s' \neq s$ case, the
+can choose to manually abort and return to the entry state, or optionally to \emph{revise the goal}, replacing
+$s$ with $s'$ in the target document and retaining $e$ as the candidate expression. This is intended to cover
+the situation where the author has made a claim which is \emph{incorrect}, but the data set and surrounding
+natural language has caused the LLM to synthesise an expression which generates a different value from the one
+specified by the user. Once a valid expression is approved, the updated text is spliced back into the
+document, and any downstream interactions (e.g. reloading the web page) are checked. This closed-loop design
+combines automated synthesis with validation and author oversight, ensuring both correctness and usability of
+the generated code.
 

### 2025-09-23 paper/sec/authoring-workflow.tex
Purge pointless paragraph.
Some reorg still needed.

@@
 
-The author initially imports the text and accompanying data into the system, which creates a Fluid
-representation of the target document. Initially that representation is just a string literal of the form
-\kw{"""..."""}, where the triple quotes are the Fluid syntax for an \emph{interpolated string}, i.e.~a string
-literal where expressions of the form \kw{\{$e$\}} are permitted within the string. The \SuggestionAgent
-analyses the inputs and identifies candidate fragments of natural language which could be replaced by Fluid
-expressions. The system then enters the human-in-the loop workflow shown in
-Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The system is
-initially waiting for the author to select a fragment of text $s$ to interpret, a fragment which may or may
-not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
+\paragraph{Initial configuration.} The author initially imports the text and accompanying data into the
+system, which creates a Fluid representation of the target document. Initially that representation is just a
+string literal of the form \kw{"""..."""}, where the triple quotes are the Fluid syntax for an
+\emph{interpolated string}, i.e.~a string literal where expressions of the form \kw{\{$e$\}} are permitted
+within the string. The \SuggestionAgent analyses the inputs and identifies candidate fragments of natural
+language which could be replaced by Fluid expressions.
+
+\paragraph{Authoring workflow.} The system then enters the human-in-the loop authoring workflow shown in
+Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The system
+initially waits for the author to select a fragment of text $s$ to interpret, a fragment which may or may not
+have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
 expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression is validated through a
@@
 natural language has caused the LLM to synthesise an expression which generates a different value from the one
-specified by the user. Once a valid expression is approved, the updated text is spliced back into the
-document, and any downstream interactions (e.g. reloading the web page) are checked. This closed-loop design
-combines automated synthesis with validation and author oversight, ensuring both correctness and usability of
-the generated code.
+specified by the user.
 
-The user highlights a fragment of interest, which is then sent to the \InterpretationAgent, which generates a
-corresponding candidate expression. Internally the \InterpretationAgent uses a compiler-in-the-loop validation
-process to ensure the expression is well-formed. The user is then able to accept the expression (in which case
-it is incorporated into the program) or reject and leave the text uninterpreted.
+\paragraph{Manual validation step.} Once a valid expression is approved, the updated text is spliced back into
+the document, and any downstream interactions (e.g. reloading the web page) are checked.
+
+This human-in-the-loop design combines automated synthesis with validation and author oversight, providing a
+substantial level of automation, but requiring the author to intervene at key steps to ensure correctness.
 
@@
 \end{itemize}
-
-\subsection{Paragraph with generated expression}
-\label{subsec:paragraph-with-generated-expression}
-Figure \ref{fig:fluid-example-paragraph} shows the Fluid code of a Paragraph with the expression generated by
-the authoring assistant.

### 2025-09-23 paper/sec/authoring-workflow.tex
50 mins.

@@
 
-\paragraph{Initial configuration.} The author initially imports the text and accompanying data into the
-system, which creates a Fluid representation of the target document. Initially that representation is just a
-string literal of the form \kw{"""..."""}, where the triple quotes are the Fluid syntax for an
+The high-level workflow is as follows:
+
+\begin{enumerate}
+
+\item \textbf{Initial configuration.} Author imports text and accompanying data into the system, creating a
+programmatic representation of the target document. Initially this representation is isomorphic to the target
+text, namely a string literal of the form \kw{"""..."""}, where the triple quotes are Fluid syntax for an
 \emph{interpolated string}, i.e.~a string literal where expressions of the form \kw{\{$e$\}} are permitted
-within the string. The \SuggestionAgent analyses the inputs and identifies candidate fragments of natural
-language which could be replaced by Fluid expressions.
+within the string. The \SuggestionAgent analyses these inputs and identifies any fragments of natural language
+which are candidates for being computed instead of remaining as literal substrings.
 
-\paragraph{Authoring workflow.} The system then enters the human-in-the loop authoring workflow shown in
+\item \textbf{Authoring workflow.} The system then enters the human-in-the loop authoring workflow shown in
 Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The system
@@
 from the author.
+
+\end{enumerate}
 

### 2025-09-23 paper/sec/authoring-workflow.tex
Tricky.

@@
 
-The high-level workflow is as follows:
+The main components of the workflow are as follows:
 
@@
 
-\item \textbf{Authoring workflow.} The system then enters the human-in-the loop authoring workflow shown in
-Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The system
-initially waits for the author to select a fragment of text $s$ to interpret, a fragment which may or may not
-have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
-expression $e$ by prompting the LLM (\secref{prompt-design} below). The expression is validated through a
-closed loop program synthesis step which checks that it is well-formed and evaluates to the target fragment
-$s$. If this process succeeds, the system replaces the selected substring $s$ with the interpolation
-expression \kw{\{$e$\}}, creating a new document configuration. This is the primary successful path through
-the workflow and returns the system to the entry state where it is waiting for another top-level interaction
-from the author.
+\item \textbf{High-level Authoring workflow.} The system then enters the human-in-the loop authoring workflow
+shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
+system initially waits for the author to select a fragment of text $s$ to interpret, a fragment which may or
+may not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
+expression $e$ using a closed-loop synthesis step ((3) below). If this process succeeds, the system replaces
+the selected substring $s$ with the interpolation expression \kw{\{$e$\}}, creating a new document
+configuration. This is the primary successful path through the workflow and returns the system to the entry
+state where it is waiting for another top-level interaction from the author.
+
+\item \textbf{Code synthesis step.} The expression synthesis step is itself an internal loop which starts with
+an initial prompt sent to the LLM (\secref{prompt-design} below) requesting the generation of an expression
+$e$. Using the Fluid command-line interface, the expression is validated by checking that it evaluates without
+error, produces a value coercible to a string $s'$, and finally that $s'$ is equal to the target fragment $s$.
+Any failure triggers prompt augmentation with the appropriate error message and the system retries generation.
+If code synthesis loop is able to yield an expression which computes $s$ within a maximum
+number of retries, the system proceeds to the manual validation step (4). Otherwise, the code synthesis step has failed and there are two possibilities:
+\begin{enumerate}
+\item If the last generated $e$ was invalid (resulting in an error), no remedial action is open to the author.
+This is considered an unsuccessful path through the workflow and returns the system to the start state.
+\item In the $s' \neq s$ case, the user can choose to manually abort and return to the entry state, or
+optionally to \emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the
+candidate expression. This is intended to cover the situation where the author has made a claim which is
+\emph{incorrect}, but the data set and surrounding natural language has caused the LLM to synthesise an
+expression which generates a different value from the one specified by the user.
+\end{enumerate}
+
+\item \textbf{Manual validation step.} Once a candidate expression has been generated, the author is able to
+republish the web page containing the Fluid article and interact with the fragment of text. As shown in
+\secref{}, this is an important validation step that can reveal errors in the generated expression.
 
 \end{enumerate}
-
-Otherwise, the Fluid command-line either returned an error (either a syntax error or runtime error), or
-succeeded but $e$ evaluated to some other string $s' \neq s$. Both cases trigger prompt augmentation with an
-appropriate error message, and the system retries generation. Failure of the code synthesis step means that,
-even after multiple retries with an augmented prompt, $e$ remained invalid or evaluated to $s' \neq s$. If $e$
-was invalid (resulting in an error), no remedial action is open to the author. This is considered an
-unsuccessful path through the workflow and returns the system to the start state. In the $s' \neq s$ case, the
-can choose to manually abort and return to the entry state, or optionally to \emph{revise the goal}, replacing
-$s$ with $s'$ in the target document and retaining $e$ as the candidate expression. This is intended to cover
-the situation where the author has made a claim which is \emph{incorrect}, but the data set and surrounding
-natural language has caused the LLM to synthesise an expression which generates a different value from the one
-specified by the user.
-
-\paragraph{Manual validation step.} Once a valid expression is approved, the updated text is spliced back into
-the document, and any downstream interactions (e.g. reloading the web page) are checked.
 
@@
 \end{figure}
-
-\subsection{Loopback System}
-\label{subsec:loopback-system}
-
-Expressions generated by the \InterpretationAgent are evaluated by the Fluid interpreter. If evaluation fails,
-the error is used to extend the prompt and the extended prompt is sent back to the LLM for regeneration. This
-loopback mechanism enables iterative refinement of expressions, improving accuracy and robustness. There are
-three kinds of loopback error:
-
-\begin{itemize}
-    \item \textbf{Invalid Program}. Evaluation fails due a syntax error, undeclared identifier, division by
-    zero, array index out of bounds or other runtime problem.
-    \item \textbf{Invalid Expression Type}. The value of the expression is not convertible to a string.
-    \item \textbf{Mismatching String}. The string value of the expression is not equal to the target text.
-\end{itemize}

### 2025-09-23 paper/sec/authoring-workflow.tex
wow this is hard.

@@
 
-\item \textbf{Initial configuration.} Author imports text and accompanying data into the system, creating a
-programmatic representation of the target document. Initially this representation is isomorphic to the target
-text, namely a string literal of the form \kw{"""..."""}, where the triple quotes are Fluid syntax for an
-\emph{interpolated string}, i.e.~a string literal where expressions of the form \kw{\{$e$\}} are permitted
-within the string. The \SuggestionAgent analyses these inputs and identifies any fragments of natural language
-which are candidates for being computed instead of remaining as literal substrings.
+\item \textbf{Initial configuration.} The author imports the target text and accompanying data into the
+system, creating a programmatic representation of the target document. Initially this is simply isomorphic to
+the target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax
+for a Python or JavaScript-style \emph{interpolated string}, i.e.~a literal where expressions of the form
+\kw{\{$e$\}} are permitted within the string. The \SuggestionAgent analyses these inputs and identifies any
+fragments of natural language which are candidates for being computed instead of remaining as literal
+substrings.
 
@@
 shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
-system initially waits for the author to select a fragment of text $s$ to interpret, a fragment which may or
-may not have been highlighted previously by the \SuggestionAgent. The system then generates a candidate Fluid
-expression $e$ using a closed-loop synthesis step ((3) below). If this process succeeds, the system replaces
-the selected substring $s$ with the interpolation expression \kw{\{$e$\}}, creating a new document
-configuration. This is the primary successful path through the workflow and returns the system to the entry
-state where it is waiting for another top-level interaction from the author.
+system initially waits for the author to select a fragment of text $s$ to interpret (perhaps but not
+necessarily highlighted by the \SuggestionAgent). The system then attempts to generates a candidate Fluid
+expression $e$ using the closed-loop synthesis step (3) below. If this process succeeds with an expression
+$e$, the system proceeds to the manual validation step (4) below. If the code generation step fails with no
+expression, no remedial action is open to the author; this is considered an unsuccessful path through the
+workflow and returns the system to the start state. Otherwise the code generation step fails with a mismatched
+string $s' \neq s$ outcome, and the user can choose to manually abort and return to the entry state, or
+optionally to \emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the
+candidate expression. This is intended to cover the situation where the author has made a claim which is
+\emph{incorrect}, but the data set and surrounding natural language has caused the LLM to synthesise an
+expression which generates a different value from the one specified by the user.
 
@@
 Any failure triggers prompt augmentation with the appropriate error message and the system retries generation.
-If code synthesis loop is able to yield an expression which computes $s$ within a maximum
-number of retries, the system proceeds to the manual validation step (4). Otherwise, the code synthesis step has failed and there are two possibilities:
-\begin{enumerate}
-\item If the last generated $e$ was invalid (resulting in an error), no remedial action is open to the author.
-This is considered an unsuccessful path through the workflow and returns the system to the start state.
-\item In the $s' \neq s$ case, the user can choose to manually abort and return to the entry state, or
-optionally to \emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the
-candidate expression. This is intended to cover the situation where the author has made a claim which is
-\emph{incorrect}, but the data set and surrounding natural language has caused the LLM to synthesise an
-expression which generates a different value from the one specified by the user.
-\end{enumerate}
+If code synthesis loop is able to yield an expression which computes $s$ within a maximum number of retries,
+the synthesis step succeeds with $e$. If the last generated $e$ was invalid (resulting in an error), the code
+synthesis step fails with no expression. Otherwise, the code synthesis fails with a mismatched string outcome.
 
@@
 republish the web page containing the Fluid article and interact with the fragment of text.  As shown in
-\secref{}, this is an important validation step that can reveal errors in the generated expression.
+\secref{}, this is an important validation step that can reveal errors in the generated expression. If the
+interactions look reasonable, the author can approve the replacement of the selected substring $s$ with the
+interpolation expression \kw{\{$e$\}}, creating a new document configuration. This is the primary successful
+path through the workflow and returns the system to the entry state where it is waiting for another top-level
+interaction from the author. Otherwise, the author can reject the proposed change and return to the start
+state.
 

### 2025-09-23 paper/sec/authoring-workflow.tex
Rehashing workflow description.

@@
 
-\item \textbf{Initial configuration.} The author imports the target text and accompanying data into the
-system, creating a programmatic representation of the target document. Initially this is simply isomorphic to
-the target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax
-for a Python or JavaScript-style \emph{interpolated string}, i.e.~a literal where expressions of the form
-\kw{\{$e$\}} are permitted within the string. The \SuggestionAgent analyses these inputs and identifies any
-fragments of natural language which are candidates for being computed instead of remaining as literal
-substrings.
+\item \textbf{Initial configuration.} The author imports the target text and accompanying data into the system
+to create a programmatic representation of the target document. Initially this is simply isomorphic to the
+target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax for
+a Python or JavaScript-style \emph{interpolated string}, i.e.~a literal where expressions of the form
+\kw{\{$e$\}} are permitted within the string. The \SuggestionAgent analyses the target text and identifies any
+fragments which are candidates for being computed instead of remaining as literal substrings.
 
@@
 shown in Figure~\ref{fig:interpretation-agent}, where the author interacts with the \InterpretationAgent. The
-system initially waits for the author to select a fragment of text $s$ to interpret (perhaps but not
-necessarily highlighted by the \SuggestionAgent). The system then attempts to generates a candidate Fluid
-expression $e$ using the closed-loop synthesis step (3) below. If this process succeeds with an expression
-$e$, the system proceeds to the manual validation step (4) below. If the code generation step fails with no
-expression, no remedial action is open to the author; this is considered an unsuccessful path through the
-workflow and returns the system to the start state. Otherwise the code generation step fails with a mismatched
-string $s' \neq s$ outcome, and the user can choose to manually abort and return to the entry state, or
-optionally to \emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the
-candidate expression. This is intended to cover the situation where the author has made a claim which is
-\emph{incorrect}, but the data set and surrounding natural language has caused the LLM to synthesise an
+system waits for the author to select a fragment of text $s$ to interpret (perhaps previously highlighted by
+the \SuggestionAgent). The system then attempts to generates a candidate Fluid expression $e$ using the
+closed-loop synthesis step (3) below. If code synthesis succeeds with an expression $e$, the system proceeds
+to the manual validation step (4) below. If the synthesis step fails with no expression, no remedial action is
+possible; this is considered an unsuccessful path through the workflow and returns the system to the entry
+state. Otherwise the synthesis step produces an expression $e$ which evaluates to a mismatched string $s' \neq
+s$ outcome, and the user can choose to manually abort and return to the entry state, or optionally to
+\emph{revise the goal}, replacing $s$ with $s'$ in the target document and retaining $e$ as the candidate
+expression. This is intended to cover the situation where the author has made a claim which is
+\emph{incorrect}, and the data set and surrounding natural language have led the LLM to synthesise an
 expression which generates a different value from the one specified by the user.
 
-\item \textbf{Code synthesis step.} The expression synthesis step is itself an internal loop which starts with
-an initial prompt sent to the LLM (\secref{prompt-design} below) requesting the generation of an expression
-$e$. Using the Fluid command-line interface, the expression is validated by checking that it evaluates without
-error, produces a value coercible to a string $s'$, and finally that $s'$ is equal to the target fragment $s$.
-Any failure triggers prompt augmentation with the appropriate error message and the system retries generation.
-If code synthesis loop is able to yield an expression which computes $s$ within a maximum number of retries,
-the synthesis step succeeds with $e$. If the last generated $e$ was invalid (resulting in an error), the code
-synthesis step fails with no expression. Otherwise, the code synthesis fails with a mismatched string outcome.
+\item \textbf{Code synthesis step.} The expression synthesis step is an internal loop which starts with an
+initial prompt sent to the LLM (\secref{prompt-design} below) requesting the generation of an expression $e$.
+Using the Fluid command-line interface, the expression is validated to check that it evaluates without error,
+produces a value coercible to a string $s'$, and finally that $s'$ is equal to the target fragment $s$. Any
+failure triggers prompt augmentation with the appropriate error message and the system retries generation. If
+code synthesis loop is able to yield an expression which computes $s$ within a maximum number of retries, the
+synthesis step succeeds with $e$. If the last generated $e$ was invalid (resulting in an error), the code
+synthesis step fails with no expression. Otherwise, code synthesis produces an expression $e$ but with a
+mismatched string outcome $s' \neq s$.
 

### 2025-09-23 paper/sec/authoring-workflow.tex
First cut of authoring workflow -- bit verbose.

@@
 
-\item \textbf{Manual validation step.} Once a candidate expression has been generated, the author is able to
-republish the web page containing the Fluid article and interact with the fragment of text.  As shown in
-\secref{}, this is an important validation step that can reveal errors in the generated expression. If the
-interactions look reasonable, the author can approve the replacement of the selected substring $s$ with the
-interpolation expression \kw{\{$e$\}}, creating a new document configuration. This is the primary successful
-path through the workflow and returns the system to the entry state where it is waiting for another top-level
-interaction from the author. Otherwise, the author can reject the proposed change and return to the start
-state.
+\item \textbf{Manual validation step.} Once a candidate expression has been generated, the system replaces the
+selected substring $s$ with the interpolation expression \kw{\{$e$\}}, creating a new (but only tentative)
+document configuration. The author can republish the web page hosting the document and interact with the
+proposed revision. As shown in \secref{}, this is an important validation step that can reveal errors in the
+generated expression. If the interactions look reasonable, the author can approve the new document state; this
+is the primary successful path through the workflow and returns the system to the entry state where it is
+waiting for another top-level interaction from the author. Otherwise, the author rejects the proposed change
+and returns to the entry state without any change to the document.
 

### 2025-09-23 paper/sec/introduction.tex

@@
 AI-assisted authoring of transparent documents thus support turning static text into interactable, data-driven
-content able to expose the evidential basis of scholarly claims, for which we envisage two use cases. First,
-when \textbf{authoring} content for an online article, a journalist or scientific publisher may wish to
-provide text which is linked to the underlying data so that the evidence base for the textual claims can be
-explored directly from the article. Second, when \textbf{reading} a document reporting on findings derived
-from open data (perhaps a scientific paper or climate report), the reader may want to retroactively interpret
-parts of the text as queries over the available data and gradually ``rationally reconstruct'' the relationship
-between claims in the paper and the evidence base. This might be just to aid their own comprehension, or part
-of a formal peer review process.
+content able to expose the evidential basis of scholarly claims. We envisage two use cases. First, when
+\textbf{authoring} content for an online article, a journalist or scientific publisher may wish to provide
+text which is linked to the underlying data so that the evidence base for the textual claims can be explored
+directly from the article. Second, when \textbf{reading} a document reporting on findings derived from open
+data (perhaps a scientific paper or climate report), the reader may want to retroactively interpret parts of
+the text as queries over the available data and gradually ``rationally reconstruct'' the relationship between
+claims in the paper and the evidence base. This might be just to aid their own comprehension, or part of a
+formal peer review process.
 

### 2025-09-23 script/chart-generation.py
[update] group bar chart

@@ -56,154 +56,107 @@ def generate_summary_test_case_plot(df, plot):
     plt.close()
 
 def generate_aggregated_plot(df, plot):
+    # esplodo le categorie
     df['expression-type'] = df['expression-type'].str.strip('[]').str.split(',')
     df_exploded = df.explode('expression-type')
     df_exploded['expression-type'] = df_exploded['expression-type'].str.strip()
-    
+
+    # calcolo success_rate SEPARATAMENTE per target-value
     summary = (
-        df_exploded.groupby("expression-type")
-        .agg(
-            success_rate=("success", "mean"),
-            std_dev=("success", "std"),
-            count=("success", "size")
-        )
+        df_exploded.groupby(["expression-type", "target-value"])
+        .agg(success_rate=("success", "mean"),
+             count=("success", "size"))
         .reset_index()
     )
-    
-    with open("settings.json", "r") as f:
-        config = json.load(f)
-    limit = config.get("agent-limit", "N/A")  # Use "N/A" if limit not found
-    
-    # Create custom x labels: "category (count)"
-    summary["label"] = summary["expression-type"] + " (" + summary["count"].astype(str) + ")"
 
-    # Calculate 95% CI
-    z = 1.96  # for 95% CI
-    p = summary["success_rate"]
-    n = summary["count"]
-    ci = z * np.sqrt((p * (1 - p)) / n)
-    
-    # IC limit error-bar.
-    lower_bound = np.maximum(0, p - ci)
-    upper_bound = np.minimum(1, p + ci)
-    yerr = np.array([p - lower_bound, upper_bound - p])
+    # etichette sull’asse x basate sul nome categoria
+    plt.figure(figsize=(6,6))
+    label_map = {
+        1: "Present",
+        0: "Absent"
+    }
+    summary["target_label"] = summary["target-value"].map(label_map)
 
-    # Set style
-    sns.set(style="whitegrid")
-
-    # Plot
-    plt.figure(figsize=(6, 6))
     ax = sns.barplot(
         data=summary,
-        x="label",
+        x="expression-type",   # oppure "label"
         y="success_rate",
-        palette="Blues_d",
-        width=0.7  # Reduce bar width
+        hue="target_label",
+        palette="Set2"
     )
+    ax.legend(title="Target value")
 
-    # Add error bars manually with limited CI
-    ax.errorbar(
-        x=range(len(summary)),
[…]

### 2025-09-23 src/main/java/explorableviz/authoringassistant/Main.java
[update] column target-value report

@@ -56,7 +56,7 @@ public class Main {
         try (PrintWriter out = new PrintWriter(new FileOutputStream(STR."\{Settings.getLogFolder()}/log_\{System.currentTimeMillis()}.csv"))) {
             String[] headers = {
                     "runId", "test-case", "llm-agent", "temperature", "num-token", "is-negative", "in-context-learning-size",
-                    "attempts", "result", "expression-type", "generated-expression", "expected-value", "expected-expression", "parseErrors", "counterfactualFails", "nullExpressions", "onlyLiteralExpressions", "duration(ms)"
+                    "attempts", "result", "target-value", "expression-type", "generated-expression", "expected-value", "expected-expression", "parseErrors", "counterfactualFails", "nullExpressions", "onlyLiteralExpressions", "duration(ms)"
             };
             out.println(String.join(";", headers));
             String content = results.stream()
@@ -74,6 +74,7 @@ public class Main {
                                 //program.getParagraph().toFluidSyntax(),
                                 String.valueOf(queryResult.attempt()),
                                 queryResult.correctResponse() != null ? "OK" : "KO",
+                                String.valueOf(Settings.isAddExpectedValueEnabled() ? 1 : 0),
                                 STR."[\{queryResult.expected().getCategories().stream().map(cat -> cat.label).collect(Collectors.joining(","))}]",
 //                                queryResult.correctResponse() != null ? "generated" : "NULL",
                                 queryResult.correctResponse() != null ? queryResult.correctResponse().getExpr() : "NULL",

### 2025-09-23 src/main/java/explorableviz/authoringassistant/Main.java
[update] fix report format

@@ -77,9 +77,9 @@ public class Main {
                                 String.valueOf(Settings.isAddExpectedValueEnabled() ? 1 : 0),
                                 STR."[\{queryResult.expected().getCategories().stream().map(cat -> cat.label).collect(Collectors.joining(","))}]",
 //                                queryResult.correctResponse() != null ? "generated" : "NULL",
-                                queryResult.correctResponse() != null ? queryResult.correctResponse().getExpr() : "NULL",
+                                queryResult.correctResponse() != null ? queryResult.correctResponse().getExpr().replaceAll("\n", "[NEWLINE]").replaceAll("\"", "\"\"") : "NULL",
                                 queryResult.expected().getValue(),
-                                queryResult.expected().getExpr().replaceAll("\n", "[NEWLINE]"),
+                                queryResult.expected().getExpr().replaceAll("\n", "[NEWLINE]").replaceAll("\"", "\"\""),
                                 String.valueOf(queryResult.parseErrors()),
                                 String.valueOf(queryResult.counterfactualFails()),
                                 String.valueOf(queryResult.nullExpressions()),

### 2025-09-23 paper/sec/evaluation.tex
[update] result section, add table with results

--- 2025-09-22 paper/sec/evaluation.tex
@@
 \subsection{Results}
-In this section, we present the experimental results.
+
+\begin{table}[h!]
+    \scriptsize
+
+    \centering
+    \begin{subtable}{0.45\linewidth}
+        \centering
+        \begin{tabular}{lcccc}
+            \hline
+            & \multicolumn{2}{c}{\textbf{gpt4o}} & \multicolumn{2}{c}{\textbf{gpt5}} \\
+            & \textbf{Success Rate} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
+            \hline
+            Comparison     & 74   & 3 & 77.4 & 2 \\
+            Data Retrieval & 76   & 3 & 77.4 & 2 \\
+            Difference     & 80   & 3 & 77.4 & 2 \\
+            Min/max        & 50   & 3 & 77.4 & 2 \\
+            Rank           & 44   & 3 & 77.4 & 2 \\
+            Total          & 33   & 3 & 77.4 & 2 \\
+            Overall        & 74.9 & 3 & 77.4 & 2 \\
+            \hline
+        \end{tabular}
+        \caption{Results with target-value = true}
+        \label{tab:results_true}
+    \end{subtable}%
+    \hfill
+    \begin{subtable}{0.45\linewidth}
+        \centering
+        \begin{tabular}{lcccc}
+            \hline
+            & \multicolumn{2}{c}{\textbf{gpt4o}} & \multicolumn{2}{c}{\textbf{gpt5}} \\
+            & \textbf{Success Rate} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
+            \hline
+            Comparison     & -- & -- & -- & -- \\
+            Data Retrieval & -- & -- & -- & -- \\
+            Difference     & -- & -- & -- & -- \\
+            Min/max        & -- & -- & -- & -- \\
+            Rank           & -- & -- & -- & -- \\
+            Total          & -- & -- & -- & -- \\
+            Overall        & -- & -- & -- & -- \\
+            \hline
+        \end{tabular}
+        \caption{Results without target-value}
+        \label{tab:results_no_target}
+    \end{subtable}
+    \caption{Comparison of success rates (with standard deviation) between \textbf{gpt4o} and \textbf{gpt5} across different task categories, reported with and without the use of target-values.}
+    \label{tab:results_combined}
+\end{table}
+
+In this section, we report the experimental results.
+Table \ref{tab:results_combined} summarizes the performance of \textit{gpt4o} and the most recent OpenAI model, \textit{gpt5}, under two different settings: with target-value sharing enabled and without target-value sharing.
 Overall, the system produced a valid fluid expression in 74.9\% of the cases (S.D. 3\%).
-Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category.
+
+Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and without target-value sharing.
+The results show that providing the target value is crucial for achieving a high success rate.
+Moreover, the \InterpretationAgent generates a correct expression in approximately XX\% of the cases.
 
@@
 
-\subsection{Negative Cases}
-\label{subsec:negative-cases}
+In addition to evaluating performance when the target expression is provided, we also assessed the model’s ability to produce a
+correct expression when no reference string was given.
+On the SciGen dataset, the model achieved a 74.9\% success rate when provided with the target string.
+Without the target string, performance decreased moderately to 39.4\%, indicating that the model retains substantial
+ability to reconstruct valid solutions even without explicit guidance
+
+\subsection{Problematic Cases}
+\label{subsec:problematic-cases}
 

### 2025-09-24 paper/sec/authoring-workflow.tex

@@
 
-\item \textbf{Code synthesis step.} The expression synthesis step is an internal loop which starts with an
-initial prompt sent to the LLM (\secref{prompt-design} below) requesting the generation of an expression $e$.
-Using the Fluid command-line interface, the expression is validated to check that it evaluates without error,
-produces a value coercible to a string $s'$, and finally that $s'$ is equal to the target fragment $s$. Any
-failure triggers prompt augmentation with the appropriate error message and the system retries generation. If
-code synthesis loop is able to yield an expression which computes $s$ within a maximum number of retries, the
-synthesis step succeeds with $e$. If the last generated $e$ was invalid (resulting in an error), the code
-synthesis step fails with no expression. Otherwise, code synthesis produces an expression $e$ but with a
-mismatched string outcome $s' \neq s$.
+\item \textbf{Code synthesis step.} The expression synthesis step is an error-guided iterative prompting
+loop~\citep{skreta23}, beginning with an initial prompt sent to the LLM (\secref{prompt-design} below)
+requesting the generation of an expression $e$. Using the Fluid command-line interface, the expression is
+validated to check that it evaluates without error, produces a value coercible to a string $s'$, and finally
+that $s'$ is equal to the target fragment $s$. Any failure triggers prompt augmentation with the appropriate
+error message and the system retries generation. If code synthesis loop is able to yield an expression which
+computes $s$ within a maximum number of retries, the synthesis step succeeds with $e$. If the last generated
+$e$ was invalid (resulting in an error), the code synthesis step fails with no expression. Otherwise, code
+synthesis produces an expression $e$ but with a mismatched string outcome $s' \neq s$.
 

### 2025-09-24 paper/sec/evaluation.tex
[update] fix on table and on captions

@@
         \includegraphics[width=\linewidth]{fig/success_rate_by_category}
-        \caption{Success rate by testcases}
+        \caption{Success rate by Linguistic Category}
         \label{fig:success_rate_by_category}
@@
         \centering
-        \includegraphics[width=\linewidth]{fig/success_rate_by_category_count}
-        \caption{Success rate by category count}
-        \label{fig:success_rate_by_category_count}
+        \includegraphics[width=\linewidth]{fig/success_rate_by_complexity}
+        \caption{Success rate by Complexity}
+        \label{fig:success_rate_by_complexity}
     \end{subfigure}
-    \caption{Comparison of success rates}
+    \caption{Success rate of the proposed system, measured over 5 runs with gpt-4o}
     \label{fig:success_rate_comparison}
@@
             & \multicolumn{2}{c}{\textbf{gpt4o}} & \multicolumn{2}{c}{\textbf{gpt5}} \\
-            & \textbf{Success Rate} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
+            & \textbf{Success Rate (\%)} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
             \hline
-            Comparison     & 74   & 3 & 77.4 & 2 \\
-            Data Retrieval & 76   & 3 & 77.4 & 2 \\
-            Difference     & 80   & 3 & 77.4 & 2 \\
-            Min/max        & 50   & 3 & 77.4 & 2 \\
-            Rank           & 44   & 3 & 77.4 & 2 \\
-            Total          & 33   & 3 & 77.4 & 2 \\
+            Comparison     & 74.7   & 3 & 77.4 & 2 \\
+            Data Retrieval & 76.4   & 3 & 77.4 & 2 \\
+            Difference     & 80.3   & 3 & 77.4 & 2 \\
+            Min/max        & 50.2   & 3 & 77.4 & 2 \\
+            Rank           & 44.1   & 3 & 77.4 & 2 \\
+            Total          & 33.2   & 3 & 77.4 & 2 \\
             Overall        & 74.9 & 3 & 77.4 & 2 \\
@@
             & \multicolumn{2}{c}{\textbf{gpt4o}} & \multicolumn{2}{c}{\textbf{gpt5}} \\
-            & \textbf{Success Rate} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
+            & \textbf{Success Rate (\%)} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
             \hline
@@
 
-Figure~\ref{fig:success_rate_by_category_count} reports the success rate as a function of the number of categories assigned to each expression.
+Figure~\ref{fig:success_rate_by_complexity} reports the success rate as a function of the number of categories assigned to each expression.
 Fluid instructions combining multiple operations were associated with multiple categories.

### 2025-09-24 script/chart-generation.py
[update] change labels

@@ -93,7 +93,7 @@ def generate_aggregated_plot(df, plot):
                     (p.get_x() + p.get_width()/2., h),
                     ha='center', va='bottom', fontsize=8)
 
-    plt.title("Success Rate by Linguistic Category (split by target-value)")
+    plt.title("Success Rate by Linguistic Category")
     plt.xlabel("Category")
     plt.ylabel("Average Success Rate")
     plt.xticks(rotation=45, ha="right")
@@ -115,7 +115,7 @@ def generate_success_rate_by_category_count(df, plot):
         .reset_index()
     )
 
-    summary["label"] = summary["category_count"].astype(str) + " (" + summary["count"].astype(str) + ")"
+    summary["label"] = summary["category_count"].astype(str)
 
     plt.figure(figsize=(6, 6))
     label_map = {
@@ -137,13 +137,13 @@ def generate_success_rate_by_category_count(df, plot):
                     (p.get_x() + p.get_width() / 2., p.get_height()),
                     ha='center', va='bottom', fontsize=8)
 
-    plt.title("Success Rate by Number of Categories (grouped by target-value)")
-    plt.xlabel("Number of Categories (Number of Examples)")
-    plt.ylabel("Average Success Rate")
+    plt.title("Success Rate by Complexity")
+    plt.xlabel("Complexity")
+    plt.ylabel("Average Success Rate over 5 runs")
     plt.xticks(rotation=45, ha="right")
     plt.ylim(0, 1.1)
     plt.tight_layout()
-    plt.savefig(f"{fig_dir}/success_rate_by_category_count.png")
+    plt.savefig(f"{fig_dir}/success_rate_by_complexity.png")
     plt.close()
 
 def generate_charts():

### 2025-09-24 script/chart-generation.py
[update] fix path log

@@ -147,7 +147,7 @@ def generate_success_rate_by_category_count(df, plot):
     plt.close()
 
 def generate_charts():
-    latest_csv = "logs/log_1758489858583.csv"
+    latest_csv = get_latest_csv("logs")
     df = pd.read_csv(latest_csv, delimiter=';', quotechar='"', encoding='utf-8')
     df["success"] = df["generated-expression"].notna().astype(int)
     df["target-value"] = df["target-value"].astype(int)

### 2025-09-24 paper/sec/conclusion.tex
Start on bulleted list for future work.

@@
-\section{Conclusion}
+\section{Conclusions and Future Work}
 \label{sec:conclusion}
 
-\input{fig/nl-further-works}
+\subsection{Future Work}
+\label{sec:conclusion:future-work}
+
+\begin{itemize}
+\item Till now we have assumed that helpers functions such as \kw{growShrink} and \kw{trendWord} are
+predefined. Add support for LLM query results that introduce definitions, perhaps adding a loopback condition
+which turns "Definition not found" errors into an augmented prompt to generate the appropriate definition.
+\item Extend prompt with values of variables in scope. We include statically known data sets into the prompt,
+but realistic data analyses make use of derived data resulting from cleaning, imputation, aggregation, etc.
+Natural language interpretation will likely proceed more accurately if we are able to include the values of
+some of these intermediate queries into the prompt as well, perhaps treating the prompt as a notebook-like
+script combining code fragments and their values.
+\item A related observation is that much of the natural language present in scholarly articles refer to
+\emph{visualisations}. Indeed, these also sometimes arise as intermediate artifacts in notebooks and similar
+workflows. These too should be made available to the LLM, perhaps as structured data rather than mere bitmaps,
+as one possible target for interpreting natural language claims (for example claims about trends often refer
+directly to charts).
+\item IDE integration. The implementation described in \secref{authoring-workflow} is a design prototype only,
+with no IDE integration. We plan to add integrate this into VSCode or Cursor, providing a smooth authoring
+experience for content creators.
+\end{itemize}
+
+LINK tags #109
+
+IDE integration
+
+Further validation of generated expressions (e.g. by generating counterfactual test cases)

### 2025-09-24 paper/sec/evaluation.tex
Tweaks to eval.

--- 2025-09-24 paper/sec/evaluation.tex
@@
 
-\subsection{Validation}
-\label{subsec:validation-with-counterfactual-data}
-
-The validation process is structured in two phases
-
-\paragraph{Initial validation}
-In the initial phase, the generated expression is executed via Fluid CLI. If the execution is successful, the
-result obtained is compared with the expected result. In case of error, the output is forwarded to the
-\InterpretationAgent, which proceeds to regenerate the expression.
-
-\paragraph{Validation via counterfactual testing}
-In the second phase, the data used for execution is deliberately modified in order to verify the consistency
-of the behaviours. Both the expected and generated expressions are executed again and the results are
-compared. If discrepancies emerge, the error is forwarded to the \InterpretationAgent. If there are no
-differences, the expression is considered valid and is returned to the user.
-
 \subsection{Research questions}
 
-These seem quite pertinent to the overall problem of transparent science communication:
 \begin{itemize}
@@
 \subsection{Success Rate by Category}
-To evaluate the system we used a sample of the Scigen dataset~\citep{scigen_dataset_2021}.
-
-We aggregated the generated expression in three categories: aggregation, trends, quantitative (See
+To evaluate the system we used a sample of the Scigen dataset~\citep{moosavi21}. We aggregated the generated
+expression in three categories: aggregation, trends, quantitative (See
 Table~\ref{tab:natural-language-forms}). \figref{success_rate_by_category} shows the results obtained in
@@
 
-\subsection{Results}
+\subsection{Counterfactual testing}
+\label{subsec:validation-with-counterfactual-data}
 
-\begin{table}[h!]
-    \scriptsize
+For the \emph{counterfactual testing} part of our analysis, the raw data execution is deliberately modified in
+order to verify the consistency of the behaviours. Both the expected and generated expressions are executed
+again and the results are compared. If discrepancies are found \rpnote{complete}
 
-    \centering
-    \begin{subtable}{0.45\linewidth}
-        \centering
-        \begin{tabular}{lcccc}
-            \hline
-            & \multicolumn{2}{c}{\textbf{gpt4o}} & \multicolumn{2}{c}{\textbf{gpt5}} \\
-            & \textbf{Success Rate} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
-            \hline
-            Comparison     & 74   & 3 & 77.4 & 2 \\
-            Data Retrieval & 76   & 3 & 77.4 & 2 \\
-            Difference     & 80   & 3 & 77.4 & 2 \\
-            Min/max        & 50   & 3 & 77.4 & 2 \\
-            Rank           & 44   & 3 & 77.4 & 2 \\
-            Total          & 33   & 3 & 77.4 & 2 \\
-            Overall        & 74.9 & 3 & 77.4 & 2 \\
-            \hline
-        \end{tabular}
-        \caption{Results with target-value = true}
-        \label{tab:results_true}
-    \end{subtable}%
-    \hfill
-    \begin{subtable}{0.45\linewidth}
-        \centering
-        \begin{tabular}{lcccc}
-            \hline
-            & \multicolumn{2}{c}{\textbf{gpt4o}} & \multicolumn{2}{c}{\textbf{gpt5}} \\
-            & \textbf{Success Rate} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
-            \hline
-            Comparison     & -- & -- & -- & -- \\
-            Data Retrieval & -- & -- & -- & -- \\
-            Difference     & -- & -- & -- & -- \\
-            Min/max        & -- & -- & -- & -- \\
-            Rank           & -- & -- & -- & -- \\
-            Total          & -- & -- & -- & -- \\
-            Overall        & -- & -- & -- & -- \\
-            \hline
-        \end{tabular}
-        \caption{Results without target-value}
-        \label{tab:results_no_target}
-    \end{subtable}
-    \caption{Comparison of success rates (with standard deviation) between \textbf{gpt4o} and \textbf{gpt5} across different task categories, reported with and without the use of target-values.}
-    \label{tab:results_combined}
-\end{table}
-
-In this section, we report the experimental results.
-Table \ref{tab:results_combined} summarizes the performance of \textit{gpt4o} and the most recent OpenAI model, \textit{gpt5}, under two different settings: with target-value sharing enabled and without target-value sharing.
-Overall, the system produced a valid fluid expression in 74.9\% of the cases (S.D. 3\%).
-
-Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and without target-value sharing.
-The results show that providing the target value is crucial for achieving a high success rate.
-Moreover, the \InterpretationAgent generates a correct expression in approximately XX\% of the cases.
-
-The system achieved performance above 70\% in comparison, data retrieval, and min/max search tasks.
-In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (50\%).
-
-Figure~\ref{fig:success_rate_by_category_count} reports the success rate as a function of the number of categories assigned to each expression.
-Fluid instructions combining multiple operations were associated with multiple categories.
-Results show that the success rate was 75\% for simple expressions (one category), increased to 85\% for two categories,
-and dropped to 58\% when three categories were involved.
-
-In addition to evaluating performance when the target expression is provided, we also assessed the model’s ability to produce a
-correct expression when no reference string was given.
-On the SciGen dataset, the model achieved a 74.9\% success rate when provided with the target string.
-Without the target string, performance decreased moderately to 39.4\%, indicating that the model retains substantial
-ability to reconstruct valid solutions even without explicit guidance
-
-\subsection{Problematic Cases}
-\label{subsec:problematic-cases}
-
-Table~\ref{tab:problematic_cases} presents examples of problematic cases in which the system was asked to generate a valid
-fluid expression.
-In these cases, producing a valid expression is challenging because the input data (or the NL query) contains ambiguities
-that prevent the construction of a well-formed expression.
-
-For instance, in the \textit{missing-data} problem, the \InterpretationAgent should generate a function to compute
- the average of the emissions. However, the input data does not provide information for all months.
-
-In such cases, the system was unable to generate a general expression and instead produced the expected value directly.
+\subsection{Negative Cases}
+\label{subsec:negative-cases}
 
@@
     \caption{Categories of problematic example}
-    \label{tab:problematic_cases}
+    \label{tab:ambiguita}
 \end{table}

### 2025-09-24 paper/sec/evaluation.tex

--- 2025-09-24 paper/sec/evaluation.tex
@@
 
+\subsection{Counterfactual testing}
+\label{subsec:validation-with-counterfactual-data}
+
+For the \emph{counterfactual testing} part of our analysis, the input data is deliberately modified in order
+to verify the consistency of the behaviours. Both the expected and generated expressions are executed again
+and the results are compared. If discrepancies are found \rpnote{complete}
+
 \subsection{Problematic Cases}

### 2025-09-24 paper/sec/evaluation.tex
add macros for gpt models and scigen

--- 2025-09-24 paper/sec/evaluation.tex
@@
 \subsection{Success Rate by Category}
-To evaluate the system we used a sample of the Scigen dataset~\citep{scigen_dataset_2021}.
+To evaluate the system we used a sample of the \SciGen dataset~\citep{scigen_dataset_2021}.
 
@@
     \end{subfigure}
-    \caption{Success rate of the proposed system, measured over 5 runs with gpt-4o}
+    \caption{Success rate of the proposed system, measured over 5 runs with \gptfour}
     \label{fig:success_rate_comparison}
@@
             \hline
-            & \multicolumn{2}{c}{\textbf{gpt4o}} & \multicolumn{2}{c}{\textbf{gpt5}} \\
+            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfour}} \\
             & \textbf{Success Rate (\%)} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
@@
             \hline
-            & \multicolumn{2}{c}{\textbf{gpt4o}} & \multicolumn{2}{c}{\textbf{gpt5}} \\
+            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
             & \textbf{Success Rate (\%)} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
@@
     \end{subtable}
-    \caption{Comparison of success rates (with standard deviation) between \textbf{gpt4o} and \textbf{gpt5} across different task categories, reported with and without the use of target-values.}
+    \caption{Comparison of success rates (with standard deviation) between \textbf{\gptfour} and \textbf{\gptfive} across different task categories, reported with and without the use of target-values.}
     \label{tab:results_combined}
@@
 In this section, we report the experimental results.
-Table \ref{tab:results_combined} summarizes the performance of \textit{gpt4o} and the most recent OpenAI model, \textit{gpt5}, under two different settings: with target-value sharing enabled and without target-value sharing.
-Overall, the system produced a valid fluid expression in 74.9\% of the cases (S.D. 3\%).
+Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model, \gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
+Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%).
 
@@
 correct expression when no reference string was given.
-On the SciGen dataset, the model achieved a 74.9\% success rate when provided with the target string.
+On the \SciGen dataset, the model achieved a 74.9\% success rate when provided with the target string.
 Without the target string, performance decreased moderately to 39.4\%, indicating that the model retains substantial
@@
 Table~\ref{tab:problematic_cases} presents examples of problematic cases in which the system was asked to generate a valid
-fluid expression.
+Fluid expression.
 In these cases, producing a valid expression is challenging because the input data (or the NL query) contains ambiguities

### 2025-09-24 paper/sec/introduction.tex
add macros for gpt models and scigen

@@
 Table~\ref{tab:natural-language-forms} summarises the natural language idioms studied in this paper. With
-state-of-the-art models like gpt4o and gpt5, our system is able to resolve basic table lookups of direct
+state-of-the-art models like \gptfour and \gptfive, our system is able to resolve basic table lookups of direct
 numerical values, as well as computations of percentages, averages, minima and maxima, and totals, each mapped
@@
 Taken together, these categories cover a representative portion of the numerical reasoning idioms found in the
-SciGen benchmark. However, some linguistic forms that commonly arise in scholarly articles are not covered in
+\SciGen benchmark. However, some linguistic forms that commonly arise in scholarly articles are not covered in
 our analysis. We have yet to study approximate quantitative terms like \textfrag{around 50\%} or

### 2025-09-24 paper/sec/evaluation.tex

@@
 \subsection{Success Rate by Category}
-To evaluate the system we used a sample of the \SciGen dataset~\citep{scigen_dataset_2021}.
-
-We aggregated the generated expression in three categories: aggregation, trends, quantitative (See
+To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
+expression in three categories: aggregation, trends, quantitative (See
 Table~\ref{tab:natural-language-forms}). \figref{success_rate_by_category} shows the results obtained in

### 2025-09-24 paper/sec/introduction.tex
rough pass over future work.

--- 2025-09-23 paper/sec/introduction.tex
@@
 increasing} or \textfrag{sharply declining} -- are likely to prove difficult because the interpretation of
-these qualifiers can be subjective and context-dependent. Adverbial quantifiers like \textfrag{generally} and
-\textfrag{usually} present similar challenges because colloquial use may differ from more formal uses (in some
-situations ``most'' might mean a majority, i.e.~greater than 50\% of cases, but in others may mean only
-``greater than any other alternative proportion''). On the other hand these difficulties also present
-themselves to human readers, so extending coverage to these idioms would substantially deepen our tool's
-ability to bridge natural language reporting with interpretation in terms of the underlying dataset, perhaps
-revealing inconsistent use of technical language on the part of the author. We discuss this further in
+these qualifiers can be subjective and context-dependent. Generalised quantifiers like \textfrag{generally}
+and \textfrag{usually}~\citep{barwise81} present similar challenges because colloquial use may differ from
+more formal uses (in some situations ``most'' might mean a majority, i.e.~greater than 50\% of cases, but in
+others may mean only ``greater than any other alternative proportion''). On the other hand these difficulties
+also present themselves to human readers, so extending coverage to these idioms would substantially deepen our
+tool's ability to bridge natural language reporting with interpretation in terms of the underlying dataset,
+perhaps revealing inconsistent use of technical language on the part of the author. We discuss this further in
 \secref{conclusion:future-work}.

### 2025-09-24 src/main/java/explorableviz/authoringassistant/paragraph/ExpressionCategory.java
mapping new categories

@@ -7,6 +7,8 @@ public enum ExpressionCategory {
     DATA_RETRIEVAL("Data Retrieval"),
     DIFFERENCE("Difference"),
     TOTAL("Total"),
+    AVERAGE("Average"),
+    SUM("Sum"),
     MIN_MAX("Min/Max");
 
     public final String label;

### 2025-09-24 src/main/java/explorableviz/authoringassistant/paragraph/ExpressionCategory.java
[update] remove old category

@@ -6,7 +6,6 @@ public enum ExpressionCategory {
     PERCENTAGE("Percentage"),
     DATA_RETRIEVAL("Data Retrieval"),
     DIFFERENCE("Difference"),
-    TOTAL("Total"),
     AVERAGE("Average"),
     SUM("Sum"),
     MIN_MAX("Min/Max");

### 2025-09-24 src/main/java/explorableviz/authoringassistant/paragraph/ExpressionCategory.java
[update] percentage to ratio

@@ -3,7 +3,7 @@ package explorableviz.authoringassistant.paragraph;
 public enum ExpressionCategory {
     COMPARISON("Comparison"),
     RANK("Rank"),
-    PERCENTAGE("Percentage"),
+    RATIO("Ratio"),
     DATA_RETRIEVAL("Data Retrieval"),
     DIFFERENCE("Difference"),
     AVERAGE("Average"),

### 2025-09-24 paper/sec/related-work.tex
updated relwork

@@
 The community has also rapidly engaged with work that explores the use of argument mining in scientific texts~\citep{liakata2012automatic,lauscher2018investigating} to better understand the structure of scientific arguments and the relationships between different claims and evidence.
-While the advent of LLMs has improved performance~\citep{gorur2025can,vrakatseli2025can}, argument mining remains a challenging task, particularly when it comes to identifying implicit argumentative relations between discourse units, and reasoning about relationships among different argumentative components — especially in cross-domain settings where models struggle to generalise~\citep{gemechu2024aries}.
-While in our work we do not directly perform traditional argument mining, our work similarly relies on the identification of claims in text, which we evaluate following established practices in the field.
+While the advent of LLMs has improved performance~\citep{gorur2025can,vrakatseli2025can}, argument mining remains a challenging task, particularly when it comes to identifying implicit argumentative relations between discourse units, and reasoning about relationships among different argumentative components, especially in cross-domain settings where models struggle to generalise~\citep{gemechu2024aries}.
+While in our work we do not directly perform traditional argument mining, this work similarly relies on the identification of claims in text, which we evaluate following established practices in the field.
 

### 2025-09-24 paper/abstract.tex
Add some notes

@@
 extend with hand-generated counterfactual test cases to evaluate how well machine-generated expressions
-generalise in the presence of changes to the data. Our results show that some state-of-the-art models are
-often able to synthesise compound expressions extensionally compatible with our gold solutions.
+generalise in the presence of changes to the data. Our results show that gpt4o is often able to synthesise
+compound expressions extensionally compatible with our gold solutions.
+
+\mdnote{The abstract feels a little longer than usual.
+Maybe it could be shortened slightly (for instance, the part on the two agents), though you may well consider every detail necessary.
+In any case, please ignore this if it’s too late in the process to revise.}
 \end{abstract}

### 2025-09-24 paper/fig/natural-language-forms.tex
Add some notes

@@
     \end{tabular}
-    \caption{Quantitative/semi-quantitative natural language forms considered in this paper}
+    \caption{Quantitative/semi-quantitative natural language forms considered in this paper \mdnote{In this table, the examples are written in Fluid.
+There are references to Fluid in the introduction~\citep{perera22,bond25}, but without reading them it may be difficult to understand.
+Would it be worth briefly explaining Fluid here as well, to make the paper more self-contained (assuming there is space for it)?}}
     \label{tab:natural-language-forms}

### 2025-09-24 paper/fig/scigen-1805.02474v1-10-src.tex
Add some notes

@@
-\begin{figure}[h]
+\begin{figure}%[h]
     \small
@@
     \includegraphics[scale=0.169]{fig/scigen-1805.02474v1-10-src-screenshot.png}
-    \caption{Reference solution for transparent text in \figref{scigen-example-website}}
+    \caption{Gold solution for transparent text in \figref{scigen-example-website} \mdnote{The figure is somewhat difficult to read (partly because there are no line breaks or indents).
+It also seems rather long, although I understand that probably nothing can be omitted.}}
     \label{fig:fluid-example-paragraph}

### 2025-09-24 paper/fig/natural-language-forms.tex
added generalised quntifier to table 1

--- 2025-09-23 paper/fig/natural-language-forms.tex
@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-trendWord
-(mean (getTableColumn \"svm1_p\" tableData))
-(mean
-[ mean (getTableColumn \"naive_bayes_p\" tableData),
-  mean (getTableColumn \"svm2_p\" tableData),
-  mean (getTableColumn \"svm3_p\" tableData),
-  mean (getTableColumn \"svmr_p\" tableData)
-])
-betterWorse
+let compareToSvmLinearP columnName = 
+overallComparison 
+  (map (compareCols \"svm1_p\" columnName) 
+    tableData) 
+in betterWorse (overallComparison 
+[compareToSvmLinearP c | c <- [
+  \"naive_bayes_p\", 
+  \"svm2_p\",
+  \"svm3_p\",
+  \"svmr_p\"]
+])",
 

### 2025-09-24 src/main/java/explorableviz/authoringassistant/paragraph/ExpressionCategory.java

@@ -7,7 +7,8 @@ public enum ExpressionCategory {
     DATA_RETRIEVAL("Data Retrieval"),
     DIFFERENCE("Difference"),
     TOTAL("Total"),
-    MIN_MAX("Min/Max");
+    MIN_MAX("Min/Max"),
+    GENERALISED_QUANTIFIER("Generalised Quantifier");
 
     public final String label;
     ExpressionCategory(String label) {

### 2025-09-24 paper/sec/authoring-workflow.tex

@@
 fragments which are candidates for being computed instead of remaining as literal substrings.
+\mdnote{Since this is the only occurrence of the term ``isomorphic'' in the paper, it might be worth adding a brief explanation, unless you feel it is clear enough as is.}
 
@@
 system waits for the author to select a fragment of text $s$ to interpret (perhaps previously highlighted by
-the \SuggestionAgent). The system then attempts to generates a candidate Fluid expression $e$ using the
+the \SuggestionAgent). The system then attempts to generate \mdnote{I changed ``generates'' to ``generate''} a candidate Fluid expression $e$ using the
 closed-loop synthesis step (3) below. If code synthesis succeeds with an expression $e$, the system proceeds

### 2025-09-24 paper/sec/conclusion.tex
Tweaks to eval.

--- 2025-09-22 paper/sec/conclusion.tex
@@
-\section{Conclusion}
+\section{Conclusions and Future Work}
 \label{sec:conclusion}
 
-\input{fig/nl-further-works}
+\subsection{Future Work}
+\label{sec:conclusion:future-work}
+
+Providing helper functions such as \kw{growShrink} and \kw{trendWord} as a predefined library is in some ways
+beneficial because they offer a uniform framework for interpreting a given scholarly document. We would also
+like make the system able to operate in the absence of these, perhaps adding a loopback condition to turn
+``Definition not found'' errors into an augmented prompt to generate the appropriate definition.
+
+Another important direction is to broaden the scope of the artifacts than can be referred to and the natural
+language forms that can be used to refer to them. For example, many claims in scholarly writing refer to
+visualisations and intermediate data sets derived via cleansing, imputation or aggregation. These should also
+be made available to the LLM (for visualisations, as structured data rather than as static bitmaps), enabling
+interpretation of natural language in terms of visual information and other intermediate analyses. Extending
+coverage to additional idioms beyond those considered in Table~\ref{tab:natural-language-forms}, such as
+cardinals and multiplicatives (\textfrag{three} and \textfrag{three times}), rounding (\textfrag{about 2
+degrees warmer}) and graded adjectives \textfrag{extremely likely}, would also expand the expressivity of the
+assistant.
+
+Consider the false statement \textfrag{The population of France is the largest in the EU}. There are at least
+two ways to repair this statement so that it comes out true: treat the subject (France) as fixed, but adjust
+the predicate so that it matches the facts (\textfrag{second largest}), or treat the predicate as fixed, but
+keep adjust the subject to match (\textfrag{Germany}). To make these sorts of (pragmatically important)
+distinctions, we may need to introduce the idea of \emph{referential terms} that have fixed denotations, in
+contrast to queries, whose values are contingent on data.
+
+Finally, integration into developer and authoring environments such as VSCode or Cursor would make the
+workflow more seamless for authors. Automatic generation of the counterfactual test cases discussed in
+\secref{validation-with-counterfactual-data} could further strengthen validation at authoring-time, improving
+the trustability of interpreted claims.

### 2025-09-24 paper/sec/introduction.tex

@@
     \caption{\emph{Transparent document} example, showing data-driven natural language (top
-    vs.~bottom)}
+    vs.~bottom) \mdnote{The caption could perhaps say a little more about the figure, although it is already described in the text.}}
     \label{fig:scigen-example-website}

### 2025-09-24 src/main/java/authoringassistant/Main.java
refactoring

@@ -1,8 +1,8 @@
-package explorableviz.authoringassistant;
+package authoringassistant;
 
 import kotlin.Pair;
 
-import explorableviz.authoringassistant.Program.QueryResult;
+import authoringassistant.Program.QueryResult;
 
 import java.io.*;
 import java.nio.file.Files;
@@ -14,7 +14,7 @@ import java.util.stream.Collectors;
 import java.util.stream.IntStream;
 import java.util.stream.Stream;
 
-import static explorableviz.authoringassistant.Program.cleanWebsiteFolders;
+import static authoringassistant.Program.cleanWebsiteFolders;
 
 public class Main {
     public static Logger logger = Logger.getLogger(Main.class.getName());

### 2025-09-24 src/main/java/authoringassistant/SuggestionAgent.java

@@ -1,8 +1,8 @@
-package explorableviz.authoringassistant;
+package authoringassistant;
 
-import explorableviz.authoringassistant.paragraph.Expression;
-import explorableviz.authoringassistant.paragraph.Literal;
-import explorableviz.authoringassistant.paragraph.Paragraph;
+import authoringassistant.paragraph.Expression;
+import authoringassistant.paragraph.Literal;
+import authoringassistant.paragraph.Paragraph;
 import it.unisa.cluelab.lllm.llm.LLMEvaluatorAgent;
 import it.unisa.cluelab.lllm.llm.prompt.PromptList;
 import kotlin.Pair;

### 2025-09-24 src/main/java/authoringassistant/llm/FluidOpenAIGpt4oAgent.java

@@ -1,4 +1,4 @@
-package explorableviz.authoringassistant.llm;
+package authoringassistant.llm;
 
 import com.fasterxml.jackson.annotation.JsonPropertyDescription;
 import com.theokanning.openai.completion.chat.ChatFunction;
@@ -6,16 +6,14 @@ import com.theokanning.openai.completion.chat.ChatFunctionCall;
 import com.theokanning.openai.completion.chat.ChatMessage;
 import com.theokanning.openai.service.FunctionExecutor;
 import com.theokanning.openai.service.OpenAiService;
-import explorableviz.authoringassistant.paragraph.Expression;
+import authoringassistant.paragraph.Expression;
 import it.unisa.cluelab.lllm.llm.agents.generic.OpenAIEvaluatorAgent;
-import it.unisa.cluelab.lllm.llm.prompt.Prompt;
 import it.unisa.cluelab.lllm.llm.prompt.PromptList;
 import org.jetbrains.annotations.NotNull;
 import org.json.JSONObject;
 
 import java.time.Duration;
 import java.util.Collections;
-import java.util.List;
 
 public class FluidOpenAIGpt4oAgent extends OpenAIEvaluatorAgent<Expression> {
     public FluidOpenAIGpt4oAgent(JSONObject setttings) {

### 2025-09-24 src/main/java/authoringassistant/llm/suggestion/CodeLLamaAgent.java

@@ -1,4 +1,4 @@
-package explorableviz.authoringassistant.llm.suggestion;
+package authoringassistant.llm.suggestion;
 
 import it.unisa.cluelab.lllm.llm.agents.generic.OLLAMAEvaluatorAgent;
 import org.json.JSONObject;

### 2025-09-24 src/main/java/authoringassistant/paragraph/Expression.java

@@ -1,4 +1,4 @@
-package explorableviz.authoringassistant.paragraph;
+package authoringassistant.paragraph;
 
 import java.util.Set;
 

### 2025-09-24 src/main/java/authoringassistant/paragraph/Literal.java

@@ -1,4 +1,4 @@
-package explorableviz.authoringassistant.paragraph;
+package authoringassistant.paragraph;
 
 public class Literal extends TextFragment {
 

### 2025-09-24 src/main/java/authoringassistant/variable/Variables.java

@@ -1,4 +1,4 @@
-package explorableviz.authoringassistant.variable;
+package authoringassistant.variable;
 
 import org.json.JSONArray;
 import org.json.JSONObject;
@@ -6,7 +6,6 @@ import org.json.JSONObject;
 import java.math.BigDecimal;
 import java.util.HashMap;
 import java.util.Map;
-import java.util.Random;
 import java.util.SplittableRandom;
 import java.util.stream.Collectors;
 import java.util.stream.IntStream;

### 2025-09-24 paper/fig/natural-language-forms.tex
fixed golden example

@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-let compareToSvmLinearP columnName = 
-overallComparison 
-  (map (compareCols \"svm1_p\" columnName) 
-    tableData) 
-in betterWorse (overallComparison 
-[compareToSvmLinearP c | c <- [
-  \"naive_bayes_p\", 
-  \"svm2_p\",
-  \"svm3_p\",
-  \"svmr_p\"]
-])",
+betterWorse (overallComparison 
+  [compareToSvmLinearP c 
+  | c <- ["naive_bayes_p", "svm2_p", "svm3_p", "svmr_p"]]
+)
 

### 2025-09-24 src/main/java/explorableviz/authoringassistant/paragraph/ExpressionCategory.java
fix expression categories

@@ -6,8 +6,6 @@ public enum ExpressionCategory {
     RATIO("Ratio"),
     DATA_RETRIEVAL("Data Retrieval"),
     DIFFERENCE("Difference"),
-    TOTAL("Total"),
-    MIN_MAX("Min/Max"),
     AVERAGE("Average"),
     SUM("Sum"),
     MIN_MAX("Min/Max"),

### 2025-09-24 paper/sec/authoring-workflow.tex
Drop 'isomorphic'.

@@
 \item \textbf{Initial configuration.} The author imports the target text and accompanying data into the system
-to create a programmatic representation of the target document. Initially this is simply isomorphic to the
+to create a programmatic representation of the target document. Initially this is simply equivalent to the
 target text, taking the form of a string literal \kw{"""..."""}, where the triple quotes are Fluid syntax for
@@
 fragments which are candidates for being computed instead of remaining as literal substrings.
-\mdnote{Since this is the only occurrence of the term ``isomorphic'' in the paper, it might be worth adding a brief explanation, unless you feel it is clear enough as is.}
 

### 2025-09-24 paper/abstract.tex

@@
 \begin{abstract}
-We introduce \emph{transparent documents}, interactive web-based scholarly articles which allow readers to
-explore the relationship to the underlying data by hovering over fragments of text. We present an agent-based
-LLM framework for authoring transparent documents, building on recent developments in data provenance for
-general-purpose programming languages. Our implementation uses Fluid, an open source functional programming
-language with a provenance-tracking runtime, as a target platform. Our tool consists of two LLM agents which
-support a human author during the creation of a transparent document. A \SuggestionAgent helps identify
-fragments of text which could plausibly be computed from data, including numerical values selected from
-records or computed by aggregations like sum and mean, comparatives and superlatives like ``better than'' and
-``largest'', trend-adjectives like ``growing'', and other idiomatic quantitative or semi-quantitative phrases.
-An \InterpretationAgent, given such a fragment, then attempts to synthesise a suitable Fluid query over the
-data which will generate the target string. The resulting expression is spliced into the source code for an
-interactive web page, turning the static text fragment into an interactable data-driven element able to reveal
-the data that underwrites the natural language claim. We evaluate our approach on a subset of SciGen, an open
-source dataset consisting of tables from scientific articles and their corresponding descriptions, which we
-extend with hand-generated counterfactual test cases to evaluate how well machine-generated expressions
-generalise in the presence of changes to the data. Our results show that gpt4o is often able to synthesise
-compound expressions extensionally compatible with our gold solutions.
-
-\mdnote{The abstract feels a little longer than usual.
-Maybe it could be shortened slightly (for instance, the part on the two agents), though you may well consider every detail necessary.
-In any case, please ignore this if it’s too late in the process to revise.}
+We introduce \emph{transparent documents}, interactive web-based scholarly articles which allow readers to explore the relationship to the underlying
+data by hovering over fragments of text, and present an LLM-based tool for authoring transparent documents, building on recent developments in data
+provenance for general-purpose programming languages. As a target platform, our implementation uses Fluid, an open source programming language with a
+provenance-tracking runtime. Our agent-based tool supports a human author during the creation of transparent documents, identifying fragments of text
+which can be computed from data, such as numerical values selected from records or computed by aggregations like sum and mean, comparatives and
+superlatives like ``better than'' and ``largest'', trend-adjectives like ``growing'', and similar quantitative or semi-quantitative phrases, and then
+attempts to synthesise a suitable Fluid query over the data which generates the target string. The resulting expression is inserted into the article's
+web page, turning the static text fragment into an interactable data-driven element able to reveal the data that underwrites the natural language
+claim. We evaluate our approach on a subset of SciGen, an open source dataset consisting of tables from scientific articles and their corresponding
+descriptions, which we extend with hand-generated counterfactual test cases to evaluate how well machine-generated expressions generalise. Our results
+show that gpt4o is often able to synthesise compound expressions extensionally compatible with our gold solutions.
 \end{abstract}

### 2025-09-24 paper/sec/evaluation.tex
rough pass over future work.

@@
 \subsection{Research questions}
+
+\cdnote{I've bulked up the RQs qlthough I'm not sure we're actually attempting to answer them}
+RQ1. How robust is the system when provided with incomplete or ambiguous information?
+In real-world authoring tasks, text fragments may omit key criteria (e.g. \textbf{?}) or contain vague identifiers (e.g. \textbf{?} without further context).
+We therefore ask how performance degrades when the specification is underspecified or when multiple interpretations are possible. Here, we distinguish between two forms of ambiguity:
+(1) domain-level ambiguity, which reflects genuine uncertainty in the data (e.g. two models achieving identical scores).
+(2) specification-level ambiguity, which arises from incomplete or imprecise descriptions in the query (e.g. missing fields that would otherwise resolve the expression uniquely).
+
+RQ2. How sensitive is the system to misleading or adversarial information?
+Authors may, intentionally or unintentionally, introduce misleading elements into the text. These include inconsistent identifiers, incorrect numerical values, or contradictory claims.
+We ask how the system responds in such cases: does it fail gracefully, does it default to literal reproduction of the misleading fragment, or does it attempt to generate a plausible, but incorrect, expression?
+
+Together, these questions probe the limits of the InterpretationAgent's ability to generate reliable code.
+By contrasting genuine uncertainty with misleading or ill-specified inputs, we aim to understand not only the model's raw success rate, but also the failure modes most likely to arise in practical use.
 
@@
 \subsection{Counterfactual testing}
-\label{subsec:validation-with-counterfactual-data}
+\label{sec:validation-with-counterfactual-data}
 

### 2025-09-24 paper/fig/natural-language-forms.tex

@@
     \end{tabular}
-    \caption{Quantitative/semi-quantitative natural language forms considered in this paper \mdnote{In this table, the examples are written in Fluid.
-There are references to Fluid in the introduction~\citep{perera22,bond25}, but without reading them it may be difficult to understand.
-Would it be worth briefly explaining Fluid here as well, to make the paper more self-contained (assuming there is space for it)?}}
+    \caption{Quantitative/semi-quantitative natural language forms considered in this paper}
     \label{tab:natural-language-forms}

### 2025-09-24 paper/fig/natural-language-forms.tex
add generalised quntifiers to table 1

--- 2025-09-24 paper/fig/natural-language-forms.tex
@@
         Pairwise comparisons
+        & In the case of one syndrome (Hemorrhagic) we noticed an \hl{unusually low} level of recall for SVM but not for NB.&
+        \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
+unusuallyHighLow (overallComparison 
+  [compareCols col \"naive_bayes_r\" 
+    (findWithKey_ \"synd\" \"Hem.\" tableData) 
+  | col <- [\"svm1_r\", \"svm2_r\", \"svm3_r\", \"svmr_r\"]]
+)
+        \end{lstlisting} \\
+        Generalised quantifiers
         & Overall the SVM with polynomial degree 1 kernel outperformed all other kernels with other kernels generally offering \hl{better} precision at a higher cost to recall. &
@@
 betterWorse (overallComparison 
-  [compareToSvmLinearP c 
-  | c <- ["naive_bayes_p", "svm2_p", "svm3_p", "svmr_p"]]
+  [overallComparison (map (compareCols \"svm1_p\" col) tableData)
+  | col <- ["naive_bayes_p", "svm2_p", "svm3_p", "svmr_p"]]
 )
-
         \end{lstlisting} \\

### 2025-09-24 script/serve.sh
revert settings change

@@ -1,3 +1,3 @@
 #!/bin/bash
 
-npx http-serve dist/$1 -c-1
+npx http-serve dist/$1

### 2025-09-25 paper/sec/authoring-workflow.tex

@@
 system waits for the author to select a fragment of text $s$ to interpret (perhaps previously highlighted by
-the \SuggestionAgent). The system then attempts to generate \mdnote{I changed ``generates'' to ``generate''} a candidate Fluid expression $e$ using the
+the \SuggestionAgent). The system then attempts to generate a candidate Fluid expression $e$ using the
 closed-loop synthesis step (3) below. If code synthesis succeeds with an expression $e$, the system proceeds
@@
 
-\subsection{Prompt Design}
-\label{sec:prompt-design}
+\subsection{The InterpertationAgent}
+\label{sec:interpretation-agent}
 
-\paragraph{System Prompt.}
-For the Interpretation Assistant, the system prompt guides the LLM in replacing placeholder tags in the
-supplied paragraph with executable Fluid expressions. It includes the following information about the document
-under construction, the exact task to perform, and constraints on the expected output. The full system prompt
-is given in \appref{system-prompt:interpretation-agent}.
+\cdnote{I integrated the prompt description in the description of the InterpretationAgentv (unsure if this is the best plece) -- the prompt description seemed to basic (it's commented if you want to revert)}
 
-\begin{itemize}
-\item \textbf{Input structure:} summarises the imported datasets, modules, auxililiary definitions in scope in
-the current file, the current Fluid representation of the target paragraph with \kw{[REPLACE …]} tag
-indicating the text fragment to replace, plus the full text for the paragraph.
-\item \textbf{Task description:} description of the LLM task, namely to identify \kw{[REPLACE …]} placeholder
-and generate Fluid expression that produces target string using supplied dataset and definitions in scope.
-\item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
-with no extraneous content (such as comments).
-\end{itemize}
+The InterpretationAgent is guided by a structured system prompt that
+frames code generation as a precise replacement task. The model
+receives the imported datasets, helper modules, and the current Fluid
+representation of the paragraph, in which a text fragment is marked
+with [REPLACE …]. Its task is to substitute this placeholder with a
+Fluid expression that evaluates exactly to the target string,
+reconstructing quantitative or comparative claims as dataset queries
+where appropriate. To ensure integration with the workflow, the output
+must consist solely of a syntactically valid Fluid expression, with no
+additional commentary. This focused design constrains the model while
+providing enough context to generate expressions that are both
+data-grounded and immediately executable; the full prompt is given in~\appref{system-prompt:interpretation-agent}.
 
-\rpnote{Now something about the rest of the prompt.}
+%\paragraph{System Prompt.}
+%For the Interpretation Assistant, the system prompt guides the LLM in replacing placeholder tags in the
+%supplied paragraph with executable Fluid expressions. It includes the following information about the document
+%under construction, the exact task to perform, and constraints on the expected output. The full system prompt
+%is given in \appref{system-prompt:interpretation-agent}.
+
+%\begin{itemize}
+%\item \textbf{Input structure:} summarises the imported datasets, modules, auxililiary definitions in scope in
+%the current file, the current Fluid representation of the target paragraph with \kw{[REPLACE …]} tag
+%indicating the text fragment to replace, plus the full text for the paragraph.
+%\item \textbf{Task description:} description of the LLM task, namely to identify \kw{[REPLACE …]} placeholder
+%and generate Fluid expression that produces target string using supplied dataset and definitions in scope.
+%\item \textbf{Output constraints:} specify that the LLM should return a syntactically valid Fluid expression,
+%with no extraneous content (such as comments).
+%\end{itemize}
+
+%\rpnote{Now something about the rest of the prompt.}
 

### 2025-09-25 paper/sec/introduction.tex
comment
rewrite of the RQs
Improve Fig 1 caption.
Tweak.

--- 2025-09-24 paper/sec/introduction.tex
@@
 
-Making sense of and verifying data-driven claims is hard even with open and accessible data. A key challenge
-lies in tracing specific claims back to the relevant data. In peer review, for example, empirical claims
-typically lack author-supplied links to data, making them hard for reviewers to check
-directly~\citep{weber20}. Paper retractions, meanwhile, are often attributable not to fraud, but to simple
-errors in data management or analysis~\citep{hu25}. The use of large language models (LLMs) to interpret
-scholarly documents has seen considerable attention recently, from fact-checking~\citep{abu-ahmad25} to
-interpretation of charts and figures~\citep{roberts24}, but current LLM interfaces do not support direct
-interrogation of visual or other outputs for traceability to inputs.
+When attempting to interpret or verify data-driven claims, a key challenge lies in tracing specific claims
+back to the relevant data. In peer review, for example, empirical claims typically lack author-supplied links
+to data, making them hard for reviewers to check directly~\citep{weber20}. Paper retractions, meanwhile, are
+often attributable not to fraud, but to simple errors in data management or analysis~\citep{hu25}. The use of
+large language models (LLMs) to interpret scholarly documents has seen considerable attention recently, from
+fact-checking~\citep{abu-ahmad25} to interpretation of charts and figures~\citep{roberts24}, but current LLM
+interfaces do not support direct interrogation of visual or other outputs for traceability to inputs.
 
@@
     \includegraphics[width=\linewidth]{fig/scigen-1805.02474v1-10-counterfactual-with-pointer.png}
-    \caption{\emph{Transparent document} example, showing data-driven natural language (top
-    vs.~bottom)}
+    \caption{Two versions of a transparent document, showing text fragments linked to data}
     \label{fig:scigen-example-website}
@@
 \subsection{Target idioms of natural language}
+\cdnote{Can we make this a section? intro is getting too long}
 

### 2025-09-25 paper/sec/introduction.tex
Move NL idioms later.

@@
 \input{fig/scigen-1805.02474v1-10-src}
-
-\subsection{Target idioms of natural language}
-\cdnote{Can we make this a section? intro is getting too long}
-
-\input{fig/natural-language-forms}
-
-Table~\ref{tab:natural-language-forms} summarises the natural language idioms studied in this paper. With
-state-of-the-art models like \gptfour and \gptfive, our system is able to resolve basic table lookups of direct
-numerical values, as well as computations of percentages, averages, minima and maxima, and totals, each mapped
-to the corresponding aggregation over the source data. For example, phrases such as \textfrag{the Energy
-Sector accounts for 52.80\% of total emissions} and \textfrag{average methane emissions for 2030 is 13.51} are
-interpreted in terms of sum and mean respectively over the relevant data values. Similarly, \textfrag{recorded
-its highest emissions in 2030} is interpreted as a \kw{maximumBy} query, while a statement such as
-\textfrag{CNN gives the lowest accuracy} is mapped to an explicit computation of rank.
-
-We also consider \emph{trend} expressions, which comparative natural language phrases describing how a data
-attribute evolves over time, such as \textfrag{training time growing from 67 to 106 seconds}. Such idioms are
-mapped to higher-order functions like \kw{trendWord} parameterised on additional helper functions such as
-\kw{growShrink} and \kw{betterWorse} (shown in \figref{fluid-scigen}) which map comparisons to appropriate
-natural language phrases.
-
-\input{fig/scigen-lib}
-
-Taken together, these categories cover a representative portion of the numerical reasoning idioms found in the
-\SciGen benchmark. However, some linguistic forms that commonly arise in scholarly articles are not covered in
-our analysis. We have yet to study approximate quantitative terms like \textfrag{around 50\%} or
-\textfrag{roughly 100 instances}, nor interval-based descriptions such as \textfrag{between 30 and 40\%} or
-\textfrag{within 5–10 seconds}. While we have no reason for thinking these will present specific difficulties,
-other forms are likely to be more challenging. So-called \emph{graded} modal adverbs~\citep{lassiter17} which
-modify adjectival comparatives like \textfrag{better} -- as in \textfrag{slightly better} and
-\textfrag{significantly higher} -- especially when combined with trends over time, as in \textfrag{steadily
-increasing} or \textfrag{sharply declining} -- are likely to prove difficult because the interpretation of
-these qualifiers can be subjective and context-dependent. Generalised quantifiers like \textfrag{generally}
-and \textfrag{usually}~\citep{barwise81} present similar challenges because colloquial use may differ from
-more formal uses (in some situations ``most'' might mean a majority, i.e.~greater than 50\% of cases, but in
-others may mean only ``greater than any other alternative proportion''). On the other hand these difficulties
-also present themselves to human readers, so extending coverage to these idioms would substantially deepen our
-tool's ability to bridge natural language reporting with interpretation in terms of the underlying dataset,
-perhaps revealing inconsistent use of technical language on the part of the author. We discuss this further in
-\secref{conclusion:future-work}.

### 2025-09-25 paper/sec/nl-idioms.tex
Move NL idioms later.

\section{Target idioms of natural language}

\input{fig/natural-language-forms}

Table~\ref{tab:natural-language-forms} summarises the natural language idioms studied in this paper. With
state-of-the-art models like \gptfour and \gptfive, our system is able to resolve basic table lookups of direct
numerical values, as well as computations of percentages, averages, minima and maxima, and totals, each mapped
to the corresponding aggregation over the source data. For example, phrases such as \textfrag{the Energy
Sector accounts for 52.80\% of total emissions} and \textfrag{average methane emissions for 2030 is 13.51} are
interpreted in terms of sum and mean respectively over the relevant data values. Similarly, \textfrag{recorded
its highest emissions in 2030} is interpreted as a \kw{maximumBy} query, while a statement such as
\textfrag{CNN gives the lowest accuracy} is mapped to an explicit computation of rank.

We also consider \emph{trend} expressions, which comparative natural language phrases describing how a data
attribute evolves over time, such as \textfrag{training time growing from 67 to 106 seconds}. Such idioms are
mapped to higher-order functions like \kw{trendWord} parameterised on additional helper functions such as
\kw{growShrink} and \kw{betterWorse} (shown in \figref{fluid-scigen}) which map comparisons to appropriate
natural language phrases.

\input{fig/scigen-lib}

Taken together, these categories cover a representative portion of the numerical reasoning idioms found in the
\SciGen benchmark. However, some linguistic forms that commonly arise in scholarly articles are not covered in
our analysis. We have yet to study approximate quantitative terms like \textfrag{around 50\%} or
\textfrag{roughly 100 instances}, nor interval-based descriptions such as \textfrag{between 30 and 40\%} or
\textfrag{within 5–10 seconds}. While we have no reason for thinking these will present specific difficulties,
other forms are likely to be more challenging. So-called \emph{graded} modal adverbs~\citep{lassiter17} which
modify adjectival comparatives like \textfrag{better} -- as in \textfrag{slightly better} and
\textfrag{significantly higher} -- especially when combined with trends over time, as in \textfrag{steadily
increasing} or \textfrag{sharply declining} -- are likely to prove difficult because the interpretation of
these qualifiers can be subjective and context-dependent. Generalised quantifiers like \textfrag{generally}
and \textfrag{usually}~\citep{barwise81} present similar challenges because colloquial use may differ from
more formal uses (in some situations ``most'' might mean a majority, i.e.~greater than 50\% of cases, but in
others may mean only ``greater than any other alternative proportion''). On the other hand these difficulties
also present themselves to human readers, so extending coverage to these idioms would substantially deepen our
tool's ability to bridge natural language reporting with interpretation in terms of the underlying dataset,
perhaps revealing inconsistent use of technical language on the part of the author. We discuss this further in
\secref{conclusion:future-work}.

### 2025-09-25 paper/fig/natural-language-forms.tex
Fixing up Table 1.

--- 2025-09-24 paper/fig/natural-language-forms.tex
@@
 
-        Numerical value
+        Data retrieval
         & the training time per epoch growing from \hl{67} seconds to 106 seconds.
@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-betterWorse (overallComparison 
-  [compareToSvmLinearP c 
-  | c <- ["naive_bayes_p", "svm2_p", "svm3_p", "svmr_p"]]
-)
+trendWord
+(mean (getTableColumn \"svm1_p\" tableData))
+(mean
+[ mean (getTableColumn \"naive_bayes_p\" tableData),
+  mean (getTableColumn \"svm2_p\" tableData),
+  mean (getTableColumn \"svm3_p\" tableData),
+  mean (getTableColumn \"svmr_p\" tableData)
+])
+betterWorse
 

### 2025-09-25 paper/fig/natural-language-forms.tex
Fixing up Table 1.

@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-(findWithKey_ "model" "LSTM" tableData).time_s
+(model_ "LSTM").time_s
         \end{lstlisting}
         \\
-        Percentage &
+        Ratio &
         The Energy Sector accounts for total methane emissions of \hl{52.80}\% in 2030.
@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-let maxEntry = maximumBy
-   (fun x -> x.emissions)
+let maxEntry = maximumBy (fun x -> x.emissions)
    (filter (fun x -> x.type == "Energy Sector")
@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-let pos = findIndex "model" "CNN"
-    (insertionSort cmpTime tableData)
-in rankLabel "lowest" pos \end{lstlisting} \\
-        Total &
+rankLabel "lowest"
+  (findIndex "model" "CNN" (sort cmpTime tableData)) \end{lstlisting} \\
+        Sum &
         The total methane emissions for the year 2030 is \hl{37.74} for Agriculture &

### 2025-09-25 paper/fig/natural-language-forms.tex
Fixing up Table 1.

@@
         \hline
-        \textbf{Type}                & \textbf{Example} & \textbf{Gold Solution} \\
+        \textbf{Label}                & \textbf{Example} & \textbf{Gold Solution for Example} \\
         \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Quantitative expressions}} \\
-
+%        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Quantitative expressions}} \\
         Data retrieval
@@
         \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Aggregation}} \\
+%        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Aggregation}} \\
         Average
@@
 in maxEntry.year
-        \end{lstlisting} \\                             \\
+        \end{lstlisting} \\
+        \rowcolor{gray!20}
         Rank &
@@
         \rowcolor{gray!20}
-        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Trends}} \\
-
+%        \multicolumn{3}{>{\raggedright\arraybackslash}l}{\textbf{Trends}} \\
         Comparison
@@
 trendWord
- (findWithKey_ "model" "BiLSTM" tableData).time_s
- (findWithKey_ "model" "LSTM" tableData).time_s
+  (model_ "BiLSTM" tableData).time_s
+  (model_ "LSTM" tableData).time_s
   growShrink

### 2025-09-25 paper/fig/natural-language-forms.tex
Fix table 1.

@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-(record.emissions /
+(getByCategory "Energy Sector" year).emissions /
  sum (map (fun x -> x.emissions)
-          (getByYear year tableData))) * 100
+   (getByYear year)) * 100
 
@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-(sumEmissions year tableData / length records)
+sum (map (fun x -> x.emissions)
+  (getByYear year)) / length records
         \end{lstlisting} \\
@@
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-sumEmissions year tableData
+sum (map (fun x -> x.emissions)
+   (getByYear year))
         \end{lstlisting} \\

### 2025-09-25 paper/sec/authoring-workflow.tex
Fix table 1.

@@
 
-Our authoring tool is composed of two LLM-based agents:
-
-\begin{itemize}
-    \item \textbf{\SuggestionAgent}. Identifies text fragments potentially computable from data.
-    \item \textbf{\InterpretationAgent}. For a given text fragment (provided by SuggestionAgent or author),
-    attempts to synthesise Fluid expression computing target fragment.
-\end{itemize}
-
-The main components of the workflow are as follows:
+Our authoring tool is composed of two LLM-based agents. A \textbf{\SuggestionAgent} identifies text fragments
+potentially computable from data, and an \textbf{\InterpretationAgent}, given a text fragment provided by the
+SuggestionAgent or by the author, attempts to synthesise a Fluid expression which computes the target
+fragment. The main components of the workflow are as follows:
 
@@
 
-\subsection{The InterpertationAgent}
+\subsection{\InterpretationAgent}
 \label{sec:interpretation-agent}

### 2025-09-25 paper/sec/introduction.tex
Contribs.

@@
 
-\paragraph{Contributions.} Our specific contributions are:
+\paragraph{Contributions.} Our specific contributions are as follows. (\secref{conclusion:future-work}). We
+leave implementing a full Copilot-like authoring plugin for an IDE such as VSCode or Cursor for future work.
 
@@
 \item A proof-of-concept LLM-based tool for iteratively transforming a preexisting opaque document and
-associated data set into a transparent, data-driven counterpart;
-\item An empirical evaluation of how well state-of-the-art models are able to solve the associated
-interpretation and code synthesis problems.
+associated data set into a transparent, data-driven counterpart (\secref{authoring-workflow});
+\item A summary of the natural language idioms we have studied (\secref{nl-idioms}) and an empirical
+evaluation of how well state-of-the-art models are able to solve the associated interpretation and code
+synthesis problems (\secref{evaluation}).
 \end{itemize}
 
-We leave implementing a full Copilot-like authoring plugin for an IDE such as VSCode or Cursor for future
-work.
-
 \input{fig/scigen-1805.02474v1-10-src}

### 2025-09-25 paper/sec/evaluation.tex
Contribs.

@@
             \hline
-            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfour}} \\
-            & \textbf{Success Rate (\%)} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
+            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
+            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
             \hline
-            Comparison     & 74.7   & 3 & 77.4 & 2 \\
-            Data Retrieval & 76.4   & 3 & 77.4 & 2 \\
-            Difference     & 80.3   & 3 & 77.4 & 2 \\
-            Min/max        & 50.2   & 3 & 77.4 & 2 \\
-            Rank           & 44.1   & 3 & 77.4 & 2 \\
-            Total          & 33.2   & 3 & 77.4 & 2 \\
+            Comparison     & 68.2   & 3.2 & 77.4 & 2 \\
+            Data Retrieval & 77.3   & 3.7 & 77.4 & 2 \\
+            Difference     & 20.0   & 44.7 & 77.4 & 2 \\
+            Min/max        & 97.0   & 5.6 & 77.4 & 2 \\
+            Rank           & 0.0   & 0.0 & 77.4 & 2 \\
+            Sum            & 100   & 0.0 & 77.4 & 2 \\
+            Ratio          & 100   & 0.0 & 77.4 & 2 \\
+            Agerage        & 100   & 0.0 & 77.4 & 2 \\
             Overall        & 74.9 & 3 & 77.4 & 2 \\
@@
             & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
-            & \textbf{Success Rate (\%)} & \textbf{SD} & \textbf{Success Rate} & \textbf{SD} \\
+            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
             \hline
-            Comparison     & -- & -- & -- & -- \\
-            Data Retrieval & -- & -- & -- & -- \\
-            Difference     & -- & -- & -- & -- \\
-            Min/max        & -- & -- & -- & -- \\
-            Rank           & -- & -- & -- & -- \\
-            Total          & -- & -- & -- & -- \\
-            Overall        & -- & -- & -- & -- \\
+            Comparison     & 33.3 & 3.1 & -- & -- \\
+            Data Retrieval & 62.0 & 1.6 & -- & -- \\
+            Difference     & 75.0 & 50.0 & -- & -- \\
+            Min/max        & 90.6 & 12.6 & -- & -- \\
+            Rank           & 0.0 & 0.0 & -- & -- \\
+            Sum            & 100.0 & 0.0 & -- & -- \\
+            Ratio          & 100.0 & 0.0 & -- & -- \\
+            Average        & 100.0 & 0.0 & -- & -- \\
+            Overall        & 57.1 & 0.8 & -- & -- \\
             \hline

### 2025-09-25 paper/sec/authoring-workflow.tex
Contribs.

@@
 
-\subsection{\InterpretationAgent}
-\label{sec:interpretation-agent}
+\paragraph{\InterpretationAgent prompt design.}
 
-\cdnote{I integrated the prompt description in the description of the InterpretationAgentv (unsure if this is the best plece) -- the prompt description seemed to basic (it's commented if you want to revert)}
-
-The InterpretationAgent is guided by a structured system prompt that
-frames code generation as a precise replacement task. The model
-receives the imported datasets, helper modules, and the current Fluid
-representation of the paragraph, in which a text fragment is marked
-with [REPLACE …]. Its task is to substitute this placeholder with a
-Fluid expression that evaluates exactly to the target string,
-reconstructing quantitative or comparative claims as dataset queries
-where appropriate. To ensure integration with the workflow, the output
-must consist solely of a syntactically valid Fluid expression, with no
-additional commentary. This focused design constrains the model while
-providing enough context to generate expressions that are both
-data-grounded and immediately executable; the full prompt is given in~\appref{system-prompt:interpretation-agent}.
+The \InterpretationAgent is guided by a structured system prompt that frames code generation as a precise
+replacement task. The model receives the imported datasets, helper modules, and the current Fluid
+representation of the paragraph, in which a text fragment is marked with the tag \kw{[REPLACE …]}. Its task is
+to substitute this placeholder with a Fluid expression that evaluates exactly to the target string,
+reconstructing quantitative or comparative claims as data queries. To ensure integration with the workflow,
+the output must consist solely of a syntactically valid Fluid expression, with no additional commentary. The
+full prompt is given in~\appref{system-prompt:interpretation-agent}.
 

### 2025-09-25 paper/sec/related-work.tex
Add some notes

@@
 \paragraph{Argument mining.}
-For over fifteen years the field of NLP has been exploring the problem of argument mining, which involves identifying argumentative structures in text, such as claims, premises, and conclusions, and
-mapping them to formal representations~\citep{palau2009argumentation,lippi2015argument} .
-Early work in this area focused on rule-based approaches, while more recent work has leveraged machine learning and deep learning techniques to improve
-performance~\citep{stab2014identifying,stab2017parsing,eger2017neural}.
-The field has also emphasized defining annotation schemes for the task, such as the Argumentative Zoning framework~\citep{teufel2002summarizing,teufel2009towards}, as well as schemes more directly tailored to argument mining~\citep{stab2014annotating}.
-The field has focused on various domains, starting from legal texts~\citep{toulmin2003uses}, and has relied on online resources such as Debatepedia~\citep{cabrio2013natural}.
-The community has also rapidly engaged with work that explores the use of argument mining in scientific texts~\citep{liakata2012automatic,lauscher2018investigating} to better understand the structure of scientific arguments and the relationships between different claims and evidence.
-While the advent of LLMs has improved performance~\citep{gorur2025can,vrakatseli2025can}, argument mining remains a challenging task, particularly when it comes to identifying implicit argumentative relations between discourse units, and reasoning about relationships among different argumentative components, especially in cross-domain settings where models struggle to generalise~\citep{gemechu2024aries}.
-While in our work we do not directly perform traditional argument mining, this work similarly relies on the identification of claims in text, which we evaluate following established practices in the field.
+For over fifteen years the field of NLP has been exploring the problem of argument mining, which involves
+identifying argumentative structures in text, such as claims, premises, and conclusions, and mapping them to
+formal representations~\citep{palau2009argumentation,lippi2015argument} . Early work in this area focused on
+rule-based approaches, while more recent work has leveraged machine learning and deep learning techniques to
+improve performance~\citep{stab2014identifying,stab2017parsing,eger2017neural}. The field has also emphasized
+defining annotation schemes for the task, such as the Argumentative Zoning
+framework~\citep{teufel2002summarizing,teufel2009towards}, as well as schemes more directly tailored to
+argument mining~\citep{stab2014annotating}. The field has focused on various domains, starting from legal
+texts~\citep{toulmin2003uses}, and has relied on online resources such as
+Debatepedia~\citep{cabrio2013natural}. The community has also rapidly engaged with work that explores the use
+of argument mining in scientific texts~\citep{liakata2012automatic,lauscher2018investigating} to better
+understand the structure of scientific arguments and the relationships between different claims and evidence.
+While the advent of LLMs has improved performance~\citep{gorur2025can,vrakatseli2025can}, argument mining
+remains a challenging task, particularly when it comes to identifying implicit argumentative relations between
+discourse units, and reasoning about relationships among different argumentative components, especially in
+cross-domain settings where models struggle to generalise~\citep{gemechu2024aries}. While in our work we do
+not directly perform traditional argument mining, this work similarly relies on the identification of claims
+in text, which we evaluate following established practices in the field.
 
 \paragraph{NLP and scientific writing.}
-The intersection of NLP and scientific writing has gained increasing attention in the last decade, with a focus on improving the clarity, coherence, and overall quality of scientific texts. 
-On the authoring side, tools such as automated writing assistants can support researchers in producing more fluent and accessible text, for instance through grammar correction, summarisation and text simplification~\citep{napoles2017jfleg,stiennon2020learning,takeshita2024cross,saggion2017automatic}.
-Other approaches specifically target the argumentative structure of scientific papers, helping writers to organise contributions and claims more effectively~\citep{lauscher2018arguminsci}.
-Nowadays, general purpose LLMs such as ChatGPT or tools tailored for the task such as Grammarly show the variety of support that NLP tools can provide to authors~\citep{wu2023chatgpt,ahn2024transformative,khalifa2024using}.
+The intersection of NLP and scientific writing has gained increasing attention in the last decade, with a
+focus on improving the clarity, coherence, and overall quality of scientific texts. On the authoring side,
+tools such as automated writing assistants can support researchers in producing more fluent and accessible
+text, for instance through grammar correction, summarisation and text
+simplification~\citep{napoles2017jfleg,stiennon2020learning,takeshita2024cross,saggion2017automatic}. Other
+approaches specifically target the argumentative structure of scientific papers, helping writers to organise
+contributions and claims more effectively~\citep{lauscher2018arguminsci}. Nowadays, general purpose LLMs such
+as ChatGPT or tools tailored for the task such as Grammarly show the variety of support that NLP tools can
+provide to authors~\citep{wu2023chatgpt,ahn2024transformative,khalifa2024using}.
 
 At the same time, NLP methods are being developed to assist reviewers and editors in evaluating submissions.
-These include systems for detecting potential issues such as lack of clarity, weak argumentative support, or even factual inconsistencies and up to scientific fraud~\citep{thakkar2025can, fromm2021argument, freedman2024detecting}.
-Such tools can also facilitate meta-reviewing by providing summaries of peer reviews and identifying points of disagreement among reviewers~\citep{kumar2023reviewers}.
-While AI tools show promises in improving the peer-review process~\citep{tyser2024ai}, there are also various risks associated such as breaches of confidentiality, lack of transparency and biases~\citep{perlis2025artificial}.
-Our current work situates itself in the context of NLP tools for supporting the understanding of scientific writing; specifically, it addresses one of the major critiques toward the automation of such process by offering a transparent way of examining its workflow.
+These include systems for detecting potential issues such as lack of clarity, weak argumentative support, or
+even factual inconsistencies and up to scientific fraud~\citep{thakkar2025can, fromm2021argument,
+freedman2024detecting}. Such tools can also facilitate meta-reviewing by providing summaries of peer reviews
+and identifying points of disagreement among reviewers~\citep{kumar2023reviewers}. While AI tools show
+promises in improving the peer-review process~\citep{tyser2024ai}, there are also various risks associated
+such as breaches of confidentiality, lack of transparency and biases~\citep{perlis2025artificial}. Our current
+work situates itself in the context of NLP tools for supporting the understanding of scientific writing;
+specifically, it addresses one of the major critiques toward the automation of such process by offering a
+transparent way of examining its workflow.
 
@@
 
-As we saw earlier \todo{ref}, scientific texts routinely make use of comparatives like ``faster'' while
-leaving one of the argument slots implicit, with the context determining the omitted referent. LLMs
-demonstrate considerable competence in resolving these and other more syntactic forms of anaphora such as
-pronouns~\citep{zhu25}, but the resolved referent itself -- concretely, what was being referred to -- remains
-implicit. Interpretable NLP is a recent research direction which aims to support comprehension (and
-production) of text in a more explicit and transparent way~\citep{yulan23}. By generating code that formalises
-the interpretation of a comparative like ``faster'', our approach also makes these implicit references
-explicit; a richer version of our approach would allow the user to explore the linguistic interpretation as
-well. \todo{WIP}
+As \figref{scigen-example-website} illustrates, scientific texts routinely make use of comparatives like
+``faster'' while leaving one of the argument slots implicit, with the context determining the omitted
+referent. LLMs demonstrate considerable competence in resolving these and other more syntactic forms of
+anaphora such as pronouns~\citep{zhu25}, but the resolved referent itself -- concretely, what was being
+referred to -- remains implicit. Interpretable NLP is a recent research direction which aims to support
+comprehension (and production) of text in a more explicit and transparent way~\citep{yulan23}. By generating
+code that formalises the interpretation of a comparative like ``faster'', our approach also makes these
+implicit references explicit; combining our system with interpretable NLP would allow the user to explore the
+linguistic interpretation as well.

### 2025-09-25 paper/sec/evaluation.tex
Purge obsolete content.

@@
-\section{Empirical Evaluation}
+\section{Experimental Evaluation}
 \label{sec:evaluation}
@@
 
-\cdnote{I've bulked up the RQs qlthough I'm not sure we're actually attempting to answer them}
 RQ1. How robust is the system when provided with incomplete or ambiguous information?
@@
 By contrasting genuine uncertainty with misleading or ill-specified inputs, we aim to understand not only the model's raw success rate, but also the failure modes most likely to arise in practical use.
-
-\begin{itemize}
-\item How does performance degrade with partial/ambiguous information?
-\item How is performance affected by ``adversarial'' (intentionally/unintentionally misleading) information?
-\end{itemize}
-
-Some ways we might explore these axes:
-\begin{itemize}
-\item meaningful vs.~meaningless vs.~misleading identifiers
-\item leaving out key criteria (e.g.~year) that would resolve query to unique answer
-\end{itemize}
-
-We might want to distinguish ambiguity inherent in the problem domain (e.g.~uncertainty) vs.~ambiguity in the
-specification of a particular code-generation query.
 
@@
     \end{subtable}
-    \caption{Comparison of success rates (with standard deviation) between \textbf{\gptfour} and \textbf{\gptfive} across different task categories, reported with and without the use of target-values.}
+    \caption{Success rates (with standard deviation) for \textbf{\gptfour} and \textbf{\gptfive} across different task categories}
     \label{tab:results_combined}

### 2025-09-25 paper/sec/evaluation.tex
liakata12.

@@
 
-RQ1. How robust is the system when provided with incomplete or ambiguous information?
-In real-world authoring tasks, text fragments may omit key criteria (e.g. \textbf{?}) or contain vague identifiers (e.g. \textbf{?} without further context).
-We therefore ask how performance degrades when the specification is underspecified or when multiple interpretations are possible. Here, we distinguish between two forms of ambiguity:
-(1) domain-level ambiguity, which reflects genuine uncertainty in the data (e.g. two models achieving identical scores).
-(2) specification-level ambiguity, which arises from incomplete or imprecise descriptions in the query (e.g. missing fields that would otherwise resolve the expression uniquely).
+We investigated the following research questions, which probe the limits of the \InterpretationAgent's ability
+to reliably provide a data-driven interpretation of natural language. By contrasting genuine uncertainty with
+misleading or ill-specified inputs, we aim to understand not only the model's raw success rate, but also the
+failure modes most likely to arise in practical use.
 
-RQ2. How sensitive is the system to misleading or adversarial information?
-Authors may, intentionally or unintentionally, introduce misleading elements into the text. These include inconsistent identifiers, incorrect numerical values, or contradictory claims.
-We ask how the system responds in such cases: does it fail gracefully, does it default to literal reproduction of the misleading fragment, or does it attempt to generate a plausible, but incorrect, expression?
+\paragraph{RQ1. How robust is the system when provided with incomplete or ambiguous information?}
 
-Together, these questions probe the limits of the InterpretationAgent's ability to generate reliable code.
-By contrasting genuine uncertainty with misleading or ill-specified inputs, we aim to understand not only the model's raw success rate, but also the failure modes most likely to arise in practical use.
+In real-world authoring tasks, text fragments may omit key criteria (e.g. \textbf{?}) or contain vague
+identifiers (e.g. \textbf{?} without further context). We therefore ask how performance degrades when the
+specification is underspecified or when multiple interpretations are possible. Here, we distinguish between
+two forms of ambiguity: (1) domain-level ambiguity, which reflects genuine uncertainty in the data (e.g. two
+models achieving identical scores). (2) specification-level ambiguity, which arises from incomplete or
+imprecise descriptions in the query (e.g. missing fields that would otherwise resolve the expression
+uniquely).
+
+\paragraph{RQ2. How sensitive is the system to misleading or adversarial information?}
+
+Authors may (intentionally or unintentionally) introduce misleading elements into the text. These include
+inconsistent use of language, incorrect references to numerical values, or contradictory claims. We ask how
+the system responds in such cases: does it fail gracefully, does it default to literal reproduction of the
+misleading fragment, or does it attempt to generate a plausible, but incorrect, expression?
+Table~\ref{tab:problematic_cases} shows some of cases which are ``problematic'' in this sense; in these cases,
+producing a valid expression is likely to be challenging because of ambiguities in the input data or
+accompanying natural language.
 
@@
 
-\subsection{Problematic Cases}
-\label{subsec:problematic-cases}
-
-Table~\ref{tab:problematic_cases} presents examples of problematic cases in which the system was asked to generate a valid
-Fluid expression.
-In these cases, producing a valid expression is challenging because the input data (or the NL query) contains ambiguities
-that prevent the construction of a well-formed expression.
-
-For instance, in the \textit{missing-data} problem, the \InterpretationAgent should generate a function to compute
- the average of the emissions. However, the input data does not provide information for all months.
-
-In such cases, the system was unable to generate a general expression and instead produced the expected value directly.
-
 \begin{table}[t]

### 2025-09-25 paper/sec/evaluation.tex
Rehash Eval intro.

@@
 
-\subsection{Success Rate by Category}
-To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
-expression in three categories: aggregation, trends, quantitative (See
-Table~\ref{tab:natural-language-forms}). \figref{success_rate_by_category} shows the results obtained in
-generating expression by categories.
-
 \begin{figure}
@@
 \subsection{Results}
+
+To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
+expression in the linguistic categories from Table~\ref{tab:natural-language-forms}.
+\figref{success_rate_by_category} shows the results obtained in generating expression by categories.
 
@@
 
-In this section, we report the experimental results.
-Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model, \gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
+Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model,
+\gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
 Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%).

### 2025-09-25 paper/sec/evaluation.tex
[update] counterfactual

--- 2025-09-25 paper/sec/evaluation.tex
@@
 to verify the consistency of the behaviours. Both the expected and generated expressions are executed again
-and the results are compared. If discrepancies are found \rpnote{complete}
+and the results are compared. If discrepancies are found the \InterpretationAgent ask for a new generation
+sending back that the failure is related to a counterfactual error.
+
+%\rpnote{complete}
+
+Across 300 test executions, 121 contained at least one counterfactual error (avg. 3.8), of which 42 ultimately succeeded.
+Such tests help detect expressions that may yield the expected output but not the correct solution.
+For example, the \texttt{\textbackslash InterpretationAgent} generated:
+
+\begin{verbatim}
+(findWithKey_ "model" "LSTM2" tableData).time_s
+\end{verbatim}
+
+intended to retrieve the execution time of the \texttt{LSTM} model, but incorrectly referred to \texttt{LSTM2}.
+Counterfactual testing exposed this error and prompted regeneration.
 

### 2025-09-25 paper/sec/evaluation.tex
removed gpt4 vs gpt5 comparison

--- 2025-09-25 paper/sec/evaluation.tex
@@
 
-\begin{table}[h!]
-    \scriptsize
+%\begin{table}[h!]
+%    \scriptsize
+%
+%    \centering
+%    \begin{subtable}{0.45\linewidth}
+%        \centering
+%        \begin{tabular}{lcccc}
+%            \hline
+%            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
+%            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
+%            \hline
+%            Comparison     & 68.2   & 3.2 & 77.4 & 2 \\
+%            Data Retrieval & 77.3   & 3.7 & 77.4 & 2 \\
+%            Difference     & 20.0   & 44.7 & 77.4 & 2 \\
+%            Min/max        & 97.0   & 5.6 & 77.4 & 2 \\
+%            Rank           & 0.0   & 0.0 & 77.4 & 2 \\
+%            Sum            & 100   & 0.0 & 77.4 & 2 \\
+%            Ratio          & 100   & 0.0 & 77.4 & 2 \\
+%            Agerage        & 100   & 0.0 & 77.4 & 2 \\
+%            Overall        & 74.9 & 3 & 77.4 & 2 \\
+%            \hline
+%        \end{tabular}
+%        \caption{Results with target-value = true}
+%        \label{tab:results_true}
+%    \end{subtable}%
+%    \hfill
+%    \begin{subtable}{0.45\linewidth}
+%        \centering
+%        \begin{tabular}{lcccc}
+%            \hline
+%            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
+%            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
+%            \hline
+%            Comparison     & 33.3 & 3.1 & -- & -- \\
+%            Data Retrieval & 62.0 & 1.6 & -- & -- \\
+%            Difference     & 75.0 & 50.0 & -- & -- \\
+%            Min/max        & 90.6 & 12.6 & -- & -- \\
+%            Rank           & 0.0 & 0.0 & -- & -- \\
+%            Sum            & 100.0 & 0.0 & -- & -- \\
+%            Ratio          & 100.0 & 0.0 & -- & -- \\
+%            Average        & 100.0 & 0.0 & -- & -- \\
+%            Overall        & 57.1 & 0.8 & -- & -- \\
+%            \hline
+%        \end{tabular}
+%        \caption{Results without target-value}
+%        \label{tab:results_no_target}
+%    \end{subtable}
+%    \caption{Success rates (with standard deviation) for \textbf{\gptfour} and \textbf{\gptfive} across different task categories}
+%    \label{tab:results_combined}
+%\end{table}
 
-    \centering
-    \begin{subtable}{0.45\linewidth}
-        \centering
-        \begin{tabular}{lcccc}
-            \hline
-            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
-            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
-            \hline
-            Comparison     & 68.2   & 3.2 & 77.4 & 2 \\
-            Data Retrieval & 77.3   & 3.7 & 77.4 & 2 \\
-            Difference     & 20.0   & 44.7 & 77.4 & 2 \\
-            Min/max        & 97.0   & 5.6 & 77.4 & 2 \\
-            Rank           & 0.0   & 0.0 & 77.4 & 2 \\
-            Sum            & 100   & 0.0 & 77.4 & 2 \\
-            Ratio          & 100   & 0.0 & 77.4 & 2 \\
-            Agerage        & 100   & 0.0 & 77.4 & 2 \\
-            Overall        & 74.9 & 3 & 77.4 & 2 \\
-            \hline
-        \end{tabular}
-        \caption{Results with target-value = true}
-        \label{tab:results_true}
-    \end{subtable}%
-    \hfill
-    \begin{subtable}{0.45\linewidth}
-        \centering
-        \begin{tabular}{lcccc}
-            \hline
-            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
-            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
-            \hline
-            Comparison     & 33.3 & 3.1 & -- & -- \\
-            Data Retrieval & 62.0 & 1.6 & -- & -- \\
-            Difference     & 75.0 & 50.0 & -- & -- \\
-            Min/max        & 90.6 & 12.6 & -- & -- \\
-            Rank           & 0.0 & 0.0 & -- & -- \\
-            Sum            & 100.0 & 0.0 & -- & -- \\
-            Ratio          & 100.0 & 0.0 & -- & -- \\
-            Average        & 100.0 & 0.0 & -- & -- \\
-            Overall        & 57.1 & 0.8 & -- & -- \\
-            \hline
-        \end{tabular}
-        \caption{Results without target-value}
-        \label{tab:results_no_target}
-    \end{subtable}
-    \caption{Success rates (with standard deviation) for \textbf{\gptfour} and \textbf{\gptfive} across different task categories}
-    \label{tab:results_combined}
-\end{table}
+%Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model, \gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
+%Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%).
 
 In this section, we report the experimental results.
-Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model, \gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
-Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%).
-
-Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and without target-value sharing.
+Figure~\ref{fig:success_rate_by_category} illustrates the success rate of \gptfour % under two different settings: with target-value sharing enabled and without target-value sharing.
+for each category, both with and without target-value sharing.
+%Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%) \cdnote{this needs to be updated for target sharing and not target sharing or removed}.
 The results show that providing the target value is crucial for achieving a high success rate.
@@
 to verify the consistency of the behaviours. Both the expected and generated expressions are executed again
-and the results are compared. If discrepancies are found the \InterpretationAgent ask for a new generation
-sending back that the failure is related to a counterfactual error.
-
-%\rpnote{complete}
-
-Across 300 test executions, 121 contained at least one counterfactual error (avg. 3.8), of which 42 ultimately succeeded.
-Such tests help detect expressions that may yield the expected output but not the correct solution.
-For example, the \texttt{\textbackslash InterpretationAgent} generated:
-
-\begin{verbatim}
-(findWithKey_ "model" "LSTM2" tableData).time_s
-\end{verbatim}
-
-intended to retrieve the execution time of the \texttt{LSTM} model, but incorrectly referred to \texttt{LSTM2}.
-Counterfactual testing exposed this error and prompted regeneration.
+and the results are compared. If discrepancies are found \rpnote{complete}
 

### 2025-09-25 paper/sec/related-work.tex
Related work tweaks.

@@
 \paragraph{Argument mining.}
-For over fifteen years the field of NLP has been exploring the problem of argument mining, which involves
-identifying argumentative structures in text, such as claims, premises, and conclusions, and mapping them to
-formal representations~\citep{palau2009argumentation,lippi2015argument} . Early work in this area focused on
-rule-based approaches, while more recent work has leveraged machine learning and deep learning techniques to
-improve performance~\citep{stab2014identifying,stab2017parsing,eger2017neural}. The field has also emphasized
-defining annotation schemes for the task, such as the Argumentative Zoning
+\emph{Argument mining} is an area of NLP which involves identifying argumentative structures in text, such as
+claims, premises, and conclusions, and mapping them to formal
+representations~\citep{palau2009argumentation,lippi2015argument}. Early work focused on rule-based approaches,
+while more recent work has leveraged machine learning and deep learning
+techniques~\citep{stab2014identifying,stab2017parsing,eger2017neural}. The field has also emphasized defining
+annotation schemes for the task, such as the Argumentative Zoning
 framework~\citep{teufel2002summarizing,teufel2009towards}, as well as schemes more directly tailored to
@@
 Debatepedia~\citep{cabrio2013natural}. The community has also rapidly engaged with work that explores the use
-of argument mining in scientific texts~\citep{liakata2012automatic,lauscher2018investigating} to better
-understand the structure of scientific arguments and the relationships between different claims and evidence.
-While the advent of LLMs has improved performance~\citep{gorur2025can,vrakatseli2025can}, argument mining
-remains a challenging task, particularly when it comes to identifying implicit argumentative relations between
-discourse units, and reasoning about relationships among different argumentative components, especially in
-cross-domain settings where models struggle to generalise~\citep{gemechu2024aries}. While in our work we do
-not directly perform traditional argument mining, this work similarly relies on the identification of claims
-in text, which we evaluate following established practices in the field.
+of argument mining in scientific texts~\citep{liakata12,lauscher2018investigating} to better understand the
+structure of scientific arguments and the relationships between different claims and evidence. While the
+advent of LLMs has improved performance~\citep{gorur2025can,vrakatseli2025can}, argument mining remains a
+challenging task, particularly when it comes to identifying implicit argumentative relations between discourse
+units, and reasoning about relationships among different argumentative components, especially in cross-domain
+settings where models struggle to generalise~\citep{gemechu2024aries}. While in our work we do not directly
+perform traditional argument mining, this work similarly relies on the identification of claims in text, which
+we evaluate following established practices in the field.
 

### 2025-09-25 paper/sec/evaluation.tex

--- 2025-09-25 paper/sec/evaluation.tex
@@
 
+%\begin{table}[h!]
+%    \scriptsize
+%
+%    \centering
+%    \begin{subtable}{0.45\linewidth}
+%        \centering
+%        \begin{tabular}{lcccc}
+%            \hline
+%            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
+%            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
+%            \hline
+%            Comparison     & 68.2   & 3.2 & 77.4 & 2 \\
+%            Data Retrieval & 77.3   & 3.7 & 77.4 & 2 \\
+%            Difference     & 20.0   & 44.7 & 77.4 & 2 \\
+%            Min/max        & 97.0   & 5.6 & 77.4 & 2 \\
+%            Rank           & 0.0   & 0.0 & 77.4 & 2 \\
+%            Sum            & 100   & 0.0 & 77.4 & 2 \\
+%            Ratio          & 100   & 0.0 & 77.4 & 2 \\
+%            Agerage        & 100   & 0.0 & 77.4 & 2 \\
+%            Overall        & 74.9 & 3 & 77.4 & 2 \\
+%            \hline
+%        \end{tabular}
+%        \caption{Results with target-value = true}
+%        \label{tab:results_true}
+%    \end{subtable}%
+%    \hfill
+%    \begin{subtable}{0.45\linewidth}
+%        \centering
+%        \begin{tabular}{lcccc}
+%            \hline
+%            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
+%            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
+%            \hline
+%            Comparison     & 33.3 & 3.1 & -- & -- \\
+%            Data Retrieval & 62.0 & 1.6 & -- & -- \\
+%            Difference     & 75.0 & 50.0 & -- & -- \\
+%            Min/max        & 90.6 & 12.6 & -- & -- \\
+%            Rank           & 0.0 & 0.0 & -- & -- \\
+%            Sum            & 100.0 & 0.0 & -- & -- \\
+%            Ratio          & 100.0 & 0.0 & -- & -- \\
+%            Average        & 100.0 & 0.0 & -- & -- \\
+%            Overall        & 57.1 & 0.8 & -- & -- \\
+%            \hline
+%        \end{tabular}
+%        \caption{Results without target-value}
+%        \label{tab:results_no_target}
+%    \end{subtable}
+%    \caption{Success rates (with standard deviation) for \textbf{\gptfour} and \textbf{\gptfive} across different task categories}
+%    \label{tab:results_combined}
+%\end{table}
+
 To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
@@
 
-\begin{table}[h!]
-    \scriptsize
+%Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model, \gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
+%Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%).
 
-    \centering
-    \begin{subtable}{0.45\linewidth}
-        \centering
-        \begin{tabular}{lcccc}
-            \hline
-            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
-            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
-            \hline
-            Comparison     & 68.2   & 3.2 & 77.4 & 2 \\
-            Data Retrieval & 77.3   & 3.7 & 77.4 & 2 \\
-            Difference     & 20.0   & 44.7 & 77.4 & 2 \\
-            Min/max        & 97.0   & 5.6 & 77.4 & 2 \\
-            Rank           & 0.0   & 0.0 & 77.4 & 2 \\
-            Sum            & 100   & 0.0 & 77.4 & 2 \\
-            Ratio          & 100   & 0.0 & 77.4 & 2 \\
-            Agerage        & 100   & 0.0 & 77.4 & 2 \\
-            Overall        & 74.9 & 3 & 77.4 & 2 \\
-            \hline
-        \end{tabular}
-        \caption{Results with target-value = true}
-        \label{tab:results_true}
-    \end{subtable}%
-    \hfill
-    \begin{subtable}{0.45\linewidth}
-        \centering
-        \begin{tabular}{lcccc}
-            \hline
-            & \multicolumn{2}{c}{\textbf{\gptfour}} & \multicolumn{2}{c}{\textbf{\gptfive}} \\
-            & \textbf{Success Rate (\%)} & \textbf{SD (\%)} & \textbf{Success Rate} & \textbf{SD (\%)} \\
-            \hline
-            Comparison     & 33.3 & 3.1 & -- & -- \\
-            Data Retrieval & 62.0 & 1.6 & -- & -- \\
-            Difference     & 75.0 & 50.0 & -- & -- \\
-            Min/max        & 90.6 & 12.6 & -- & -- \\
-            Rank           & 0.0 & 0.0 & -- & -- \\
-            Sum            & 100.0 & 0.0 & -- & -- \\
-            Ratio          & 100.0 & 0.0 & -- & -- \\
-            Average        & 100.0 & 0.0 & -- & -- \\
-            Overall        & 57.1 & 0.8 & -- & -- \\
-            \hline
-        \end{tabular}
-        \caption{Results without target-value}
-        \label{tab:results_no_target}
-    \end{subtable}
-    \caption{Success rates (with standard deviation) for \textbf{\gptfour} and \textbf{\gptfive} across different task categories}
-    \label{tab:results_combined}
-\end{table}
+%Figure~\ref{fig:success_rate_by_category} illustrates the success rate of \gptfour % under two different settings: with target-value sharing enabled and without target-value sharing.
+%for each category, both with and without target-value sharing.
+%Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%) \cdnote{this needs to be updated for target sharing and not target sharing or removed}.
 
-Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model,
-\gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
-Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%).
+%Table \ref{tab:results_combined} summarizes the performance of \gptfour and the most recent OpenAI model,
+%\gptfive, under two different settings: with target-value sharing enabled and without target-value sharing.
+%Overall, the system produced a valid Fluid expression in 74.9\% of the cases (S.D. 3\%).
 
+In this section, we report the experimental results.
 Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and without target-value sharing.

### 2025-09-25 paper/sec/evaluation.tex
fix conflict in evaluation
Remove redundant/confusing problemmatic examples. Condense paras in Eval.

@@
 
-To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
-expression in the linguistic categories from Table~\ref{tab:natural-language-forms}.
-\figref{success_rate_by_category} shows the results obtained in generating expression by categories.
+%\figref{success_rate_by_category} shows the results obtained in generating expression by categories.
 
@@
 
-In this section, we report the experimental results.
-Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and without target-value sharing.
-The results show that providing the target value is crucial for achieving a high success rate.
-Moreover, the \InterpretationAgent generates a correct expression in approximately XX\% of the cases.
+To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
+expression in the linguistic categories from Table~\ref{tab:natural-language-forms}.
+Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and
+without target-value sharing. The results show that providing the target value (text fragment or numerical
+value) is crucial for achieving a high success rate. Moreover, the \InterpretationAgent generates a correct
+expression in approximately XX\% of the cases.
 
@@
 
-Across 300 test executions, 121 contained at least one counterfactual error (avg. 3.8), of which 42 ultimately succeeded.
-Such tests help detect expressions that may yield the expected output but not the correct solution.
-For example, the \texttt{\textbackslash InterpretationAgent} generated:
+Across 300 test executions, 121 contained at least one counterfactual error (avg. 3.8), of which 42 ultimately
+succeeded. Such tests help detect expressions that may yield the expected output but not the correct solution.
+For example, the \InterpretationAgent generated:
 
@@
         \hline
-
-        ambiguous referent &
-        As shown in Table 3 \hl{BiLSTM} gives significantly better accuracies compared to uni-directional LSTM with the training time per epoch growing from 99 seconds to 106 seconds &
-        There are three models that have 99 seconds as their time; the assistant would not know which one to choose. \\
-
-        missing data &
-        The average methane emissions for the year 2015 is \hl{1.0} &
-        November is missing to calculate the average correctly. \\
-
-        false statement &
-        We additionally make comparisons with stacked CNNs and hierarchical attention (Vaswani et al., 2017), shown in Table 3 (the CNN and Transformer rows), BiLSTM is \hl{the most efficient} among all models compared, with the highest model size &
+        false comparison &
+        BiLSTM is \hl{the most efficient} among all models compared, with the highest model size &
         BiLSTM is not the most efficient, nor does it have the largest size. \\
 
-        incorrect numerical value &
+        wrong numerical value &
         LSTM is the fastest model with overall time taken being \hl{90} seconds &

### 2025-09-25 paper/sec/evaluation.tex
fix numbers

--- 2025-09-25 paper/sec/evaluation.tex
@@
 expression in the linguistic categories from Table~\ref{tab:natural-language-forms}.
-\figref{success_rate_by_category} shows the results obtained in generating expression by categories.
+%\figref{success_rate_by_category} shows the results obtained in generating expression by categories.
 
@@
 
-In this section, we report the experimental results.
 Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and without target-value sharing.
 The results show that providing the target value is crucial for achieving a high success rate.
-Moreover, the \InterpretationAgent generates a correct expression in approximately XX\% of the cases.
+Moreover, the \InterpretationAgent generates a correct expression in 74.9\% of the cases (S.D. 3.0\%).
 
-The system achieved performance above 70\% in comparison, data retrieval, and min/max search tasks.
-In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (50\%).
+The system achieved performance above 68\% in comparison, 77.3\% in data retrieval, and 97\% in min/max search tasks.
+In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (0\%).
 
@@
 Fluid instructions combining multiple operations were associated with multiple categories.
-Results show that the success rate was 75\% for simple expressions (one category), increased to 85\% for two categories,
-and dropped to 58\% when three categories were involved.
+Results show that the success rate was 62\% for simple expressions (one category), increased to 91\% for two categories,
+while no correct expression was generated for expression with 3 categories.
 
@@
 On the \SciGen dataset, the model achieved a 74.9\% success rate when provided with the target string.
-Without the target string, performance decreased moderately to 39.4\%, indicating that the model retains substantial
+Without the target string, performance decreased moderately to 57.1\%, indicating that the model retains substantial
 ability to reconstruct valid solutions even without explicit guidance

### 2025-09-25 paper/sec/evaluation.tex

--- 2025-09-25 paper/sec/evaluation.tex
@@
 value) is crucial for achieving a high success rate. Moreover, the \InterpretationAgent generates a correct
-expression in approximately XX\% of the cases.
+expression in approximately in 74.9\% of the cases (S.D. 3.0\%).
 
-The system achieved performance above 70\% in comparison, data retrieval, and min/max search tasks.
-In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (50\%).
+The system achieved performance above 68\% in comparison, 77.3\% in data retrieval, and 97\% in min/max search tasks.
+In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (0\%).
 

### 2025-09-25 paper/sec/conclusion.tex
conclusions and future work

--- 2025-09-22 paper/sec/conclusion.tex
@@
-\section{Conclusion}
+\section{Conclusions and Future Work}
 \label{sec:conclusion}
 
-\input{fig/nl-further-works}
+We introduced a proof-of-concept system for authoring transparent, data-driven documents by combining LLM-based code synthesis with Fluid's provenance-tracking runtime.
+Our evaluation on SciGen shows that the approach can reliably link natural language claims to their underlying data, while also revealing common failure modes such as ambiguity and misleading input.
+
+%\subsection{Future Work}
+%\label{sec:conclusion:future-work}
+
+Future work includes reducing reliance on predefined helper functions
+such as \kw{growShrink} and \kw{trendWord}. While these provide a uniform
+framework, we aim to enable the system to operate in their absence,
+for instance by turning ``definition not found'' errors into augmented
+prompts that trigger automatic generation of missing definitions. We
+also plan to broaden the scope of supported artifacts, extending
+interpretation to visualisations and intermediate datasets derived
+from cleansing or aggregation, and to cover additional idioms such as
+cardinals, multiplicatives, rounding, and graded adjectives.
+
+Another priority is improving integration and validation. Embedding
+the system into developer and authoring environments such as VSCode or
+Cursor would make the workflow more seamless, while automatic
+generation of counterfactual test cases could strengthen validation at
+authoring time. Finally, distinguishing between \emph{referential terms} with
+fixed denotations and queries with data-dependent values may help in
+repairing false or inconsistent statements, ensuring that generated
+expressions remain aligned with both the data and the author's intent.
+
+%Providing helper functions such as \kw{growShrink} and \kw{trendWord} as a predefined library is in some ways
+%beneficial because they offer a uniform framework for interpreting a given scholarly document. We would also
+%like make the system able to operate in the absence of these, perhaps adding a loopback condition to turn
+%``Definition not found'' errors into an augmented prompt to generate the appropriate definition.
+
+%Another important direction is to broaden the scope of the artifacts than can be referred to and the natural
+%language forms that can be used to refer to them. For example, many claims in scholarly writing refer to
+%visualisations and intermediate data sets derived via cleansing, imputation or aggregation. These should also
+%be made available to the LLM (for visualisations, as structured data rather than as static bitmaps), enabling
+%interpretation of natural language in terms of visual information and other intermediate analyses. Extending
+%coverage to additional idioms beyond those considered in Table~\ref{tab:natural-language-forms}, such as
+%cardinals and multiplicatives (\textfrag{three} and \textfrag{three times}), rounding (\textfrag{about 2
+%degrees warmer}) and graded adjectives \textfrag{extremely likely}, would also expand the expressivity of the
+%assistant.
+
+%Consider the false statement \textfrag{The population of France is the largest in the EU}. There are at least
+%two ways to repair this statement so that it comes out true: treat the subject (France) as fixed, but adjust
+%the predicate so that it matches the facts (\textfrag{second largest}), or treat the predicate as fixed, but
+%keep adjust the subject to match (\textfrag{Germany}). To make these sorts of (pragmatically important)
+%distinctions, we may need to introduce the idea of \emph{referential terms} that have fixed denotations, in
+%contrast to queries, whose values are contingent on data.
+
+%Finally, integration into developer and authoring environments such as VSCode or Cursor would make the
+%workflow more seamless for authors. Automatic generation of the counterfactual test cases discussed in
+%\secref{validation-with-counterfactual-data} could further strengthen validation at authoring-time, improving
+%the trustability of interpreted claims.

### 2025-09-25 paper/sec/evaluation.tex
conclusions and future work

--- 2025-09-25 paper/sec/evaluation.tex
@@
         \label{fig:success_rate_by_category}
-    \end{subfigure}\hfill
+    \end{subfigure} \hfill
     \begin{subfigure}{0.48\linewidth}
@@
 The results show that providing the target value is crucial for achieving a high success rate.
-Moreover, the \InterpretationAgent generates a correct expression in 74.9\% of the cases (S.D. 3.0\%).
+Moreover, the \InterpretationAgent generates a correct expression in approximately XX\% of the cases.
 
-The system achieved performance above 68\% in comparison, 77.3\% in data retrieval, and 97\% in min/max search tasks.
-In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (0\%).
+The system achieved performance above 70\% in comparison, data retrieval, and min/max search tasks.
+In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (50\%).
 
@@
 Fluid instructions combining multiple operations were associated with multiple categories.
-Results show that the success rate was 62\% for simple expressions (one category), increased to 91\% for two categories,
-while no correct expression was generated for expression with 3 categories.
+Results show that the success rate was 75\% for simple expressions (one category), increased to 85\% for two categories,
+and dropped to 58\% when three categories were involved.
 
@@
 On the \SciGen dataset, the model achieved a 74.9\% success rate when provided with the target string.
-Without the target string, performance decreased moderately to 57.1\%, indicating that the model retains substantial
+Without the target string, performance decreased moderately to 39.4\%, indicating that the model retains substantial
 ability to reconstruct valid solutions even without explicit guidance

### 2025-09-25 paper/sec/evaluation.tex

--- 2025-09-25 paper/sec/evaluation.tex
@@
 value) is crucial for achieving a high success rate. Moreover, the \InterpretationAgent generates a correct
-expression in approximately in 74.9\% of the cases (S.D. 3.0\%).
+expression in approximately XX\% of the cases.
 
-The system achieved performance above 68\% in comparison, 77.3\% in data retrieval, and 97\% in min/max search tasks.
-In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (0\%).
+The system achieved performance above 70\% in comparison, data retrieval, and min/max search tasks.
+In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (50\%).
 

### 2025-09-25 paper/sec/conclusion.tex
Tweak concl.

@@
 
-Future work includes reducing reliance on predefined helper functions
-such as \kw{growShrink} and \kw{trendWord}. While these provide a uniform
-framework, we aim to enable the system to operate in their absence,
-for instance by turning ``definition not found'' errors into augmented
-prompts that trigger automatic generation of missing definitions. We
-also plan to broaden the scope of supported artifacts, extending
-interpretation to visualisations and intermediate datasets derived
-from cleansing or aggregation, and to cover additional idioms such as
-cardinals, multiplicatives, rounding, and graded adjectives.
+Future work includes reducing reliance on predefined helper functions such as \kw{growShrink} and
+\kw{trendWord}. While there is an advantage in using a predefined set of helpers (in that they offer a uniform
+framework for interpreting a given scholarly document), we also aim to enable the system to operate in their
+absence, for instance by turning ``definition not found'' errors into augmented prompts that trigger automatic
+generation of missing definitions. We also plan to broaden the scope of supported artifacts, extending
+interpretation to visualisations and intermediate datasets derived from cleansing or aggregation, and to cover
+additional idioms such as cardinals, multiplicatives, rounding, and graded adjectives.
 
@@
 expressions remain aligned with both the data and the author's intent.
-
-%Providing helper functions such as \kw{growShrink} and \kw{trendWord} as a predefined library is in some ways
-%beneficial because they offer a uniform framework for interpreting a given scholarly document. We would also
-%like make the system able to operate in the absence of these, perhaps adding a loopback condition to turn
-%``Definition not found'' errors into an augmented prompt to generate the appropriate definition.
 

### 2025-09-25 paper/sec/introduction.tex

@@
 
-When attempting to interpret or verify data-driven claims, a key challenge lies in tracing specific claims
-back to the relevant data. In peer review, for example, empirical claims typically lack author-supplied links
-to data, making them hard for reviewers to check directly~\citep{weber20}. Paper retractions, meanwhile, are
-often attributable not to fraud, but to simple errors in data management or analysis~\citep{hu25}. The use of
-large language models (LLMs) to interpret scholarly documents has seen considerable attention recently, from
+When interpreting or verifying data-driven claims, a key challenge lies in tracing specific claims back to the
+relevant data. In peer review, for example, empirical claims typically lack author-supplied links to data,
+making them hard for reviewers to check directly~\citep{weber20}. Paper retractions, meanwhile, are often
+attributable not to fraud, but to simple errors in data management or analysis~\citep{hu25}. The use of large
+language models (LLMs) to interpret scholarly documents has seen considerable attention recently, from
 fact-checking~\citep{abu-ahmad25} to interpretation of charts and figures~\citep{roberts24}, but current LLM
@@
 \figref{fluid-example-paragraph}, and makes use of several helper functions, a representative subset of which
-are shown in \figref{fluid-scigen}. Both the provenance-tracking runtime of Fluid \emph{and} the LLM-based
-authoring support contribute to the solution: the Fluid runtime provides the interactions, and the LLM-based
-authoring tool makes the whole process practical. Without tool support, authoring a scientific document as a
-(hand-crafted) \emph{program} is unlikely to be a feasible task.
+are shown in \figref{fluid-scigen}. What makes our solution interesting is that the provenance-tracking
+runtime of Fluid \emph{and} the LLM-based authoring support are both essential components of the solution,
+with Fluid providing the interactions, and the LLM-based tool making the authoring process feasible.
+Generating code for a traditional language like Python would still result in a data-driven document, but
+crucially without the interactive provenance queries; and without AI-based tooling to support the authoring
+process, the author would be faced with creating the code in \figref{fluid-example-paragraph} by hand, which
+is unlikely to be feasible as part of the usual scientific writing process.
 

### 2025-09-25 paper/sec/authoring-workflow.tex

@@
         \includegraphics[width=\linewidth]{fig/data-flow-correct}
-        \caption{Author Acceptance pathway}
+        \caption{Author accepts expression}
         \label{fig:data-flow-correct}
@@
         \includegraphics[width=\linewidth]{fig/data-flow-error}
-        \caption{Author Rejection pathway}
+        \caption{Author identifies error and rejects}
         \label{fig:dataflow-error}
     \end{subfigure}
-    \caption{Two possible paths through the basic editing loop: (a) author accepts generated expression after
-    interactively verifying its correctness; (b) author rejects proposed expression, having identified an
-    error (natural language mentions accuracy, while interaction reveals use of timing data).}
+    \caption{Two possible paths through editing loop, with interactive verification of generated code}
+%  (natural language mentions accuracy, while interaction reveals use of timing data).
     \label{fig:overall-architecture}

### 2025-09-25 paper/sec/evaluation.tex
Better intro to 4.

--- 2025-09-25 paper/sec/evaluation.tex
@@
 
-We investigated the following research questions, which probe the limits of the \InterpretationAgent's ability
-to reliably provide a data-driven interpretation of natural language. By contrasting genuine uncertainty with
-misleading or ill-specified inputs, we aim to understand not only the model's raw success rate, but also the
-failure modes most likely to arise in practical use.
+Our evaluation tests the ability of the \InterpretationAgent to translate quantitative and semi-quantitative
+expressions from scholarly natural language into executable queries that operate on the underlying dataset.
+Beyond raw accuracy, we are also concerned with how performance varies with task complexity, and whether the
+generated expressions are robust under changes to data or in the presence of ambiguity or other low data
+quality issues. These are captured in two research questions:
 
@@
 without target-value sharing. The results show that providing the target value (text fragment or numerical
-value) is crucial for achieving a high success rate. Moreover, the \InterpretationAgent generates a correct
+value) is crucial for achieving a high success rate. Overall the \InterpretationAgent generates a correct
 expression in approximately in 74.9\% of the cases (S.D. 3.0\%).

### 2025-09-25 paper/sec/evaluation.tex
Better RQ1.

@@
 
-\paragraph{RQ1. How robust is the system when provided with incomplete or ambiguous information?}
+\paragraph{RQ1. Interpretation Accuracy across Linguistic Idioms and Complexity.}
 
-In real-world authoring tasks, text fragments may omit key criteria (e.g. \textbf{?}) or contain vague
-identifiers (e.g. \textbf{?} without further context). We therefore ask how performance degrades when the
-specification is underspecified or when multiple interpretations are possible. Here, we distinguish between
-two forms of ambiguity: (1) domain-level ambiguity, which reflects genuine uncertainty in the data (e.g. two
-models achieving identical scores). (2) specification-level ambiguity, which arises from incomplete or
-imprecise descriptions in the query (e.g. missing fields that would otherwise resolve the expression
-uniquely).
+To what extent can LLMs accurately interpret quantitative and semi-quantitative claims in scholarly text as
+data queries? We examine performance across a range of linguistic idioms (e.g. averages, percentages, min/max,
+ranks, as summarised in Table~\ref{tab:natural-language-forms}) and investigate how accuracy varies with task
+complexity, measured (somewhat crudely and approximately) by the number of query sub-expressions (e.g.
+retrieval, aggregation, or arithmetic) present in the gold solution.
 

### 2025-09-25 paper/sec/evaluation.tex
Better RQ2.

@@
 
-\paragraph{RQ2. How sensitive is the system to misleading or adversarial information?}
+\paragraph{RQ2. Generalisability and Robustness.}
 
-Authors may (intentionally or unintentionally) introduce misleading elements into the text. These include
-inconsistent use of language, incorrect references to numerical values, or contradictory claims. We ask how
-the system responds in such cases: does it fail gracefully, does it default to literal reproduction of the
-misleading fragment, or does it attempt to generate a plausible, but incorrect, expression?
-Table~\ref{tab:problematic_cases} shows some of cases which are ``problematic'' in this sense; in these cases,
-producing a valid expression is likely to be challenging because of ambiguities in the input data or
-accompanying natural language.
+How well do the generated expressions generalise when the underlying data changes, or when the input contains
+misleading or ill-specified information? We test whether generated queries continue to produce correct outputs
+under a set of hand-generated counterfactual modifications of the dataset, based on expected query results
+specific to each test case, and also how counterfactual performance is impacted by the presence of misleading
+or adversarial phrasing. able~\ref{tab:problematic_cases} shows some of cases which are ``problematic'' in
+this sense; in these cases, producing a valid expression is likely to be challenging because of ambiguities in
+the input data or accompanying natural language.
 

### 2025-09-25 paper/fig/natural-language-forms.tex

@@
     \renewcommand{\arraystretch}{1.2}
-    \begin{tabular}{>{\raggedright\arraybackslash}p{2cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6cm}}
+    \begin{tabular}{>{\raggedright\arraybackslash}p{1.4cm} >{\raggedright\arraybackslash}p{5cm} >{\raggedright\arraybackslash}p{6.3cm}}
         \hline
@@
 rankLabel "lowest"
-  (findIndex "model" "CNN" (sort cmpTime tableData)) \end{lstlisting} \\
+  (findIndex "model" "CNN"
+    (sort cmpTime tableData)) \end{lstlisting} \\
         Sum &
@@
         \end{lstlisting} \\
-        Pairwise comparisons
-        & Overall the SVM with polynomial degree 1 kernel outperformed all other kernels with other kernels generally offering \hl{better} precision at a higher cost to recall. &
+        Generalised quantifiers
+        & In the case of one syndrome (Hemorrhagic) we noticed an \hl{unusually low} level of recall for SVM but not for NB. &
         \begin{lstlisting}[language=Fluid,numbers=none,aboveskip=-7pt,belowskip=-8.5pt]
-trendWord
-(mean (getTableColumn \"svm1_p\" tableData))
-(mean
-[ mean (getTableColumn \"naive_bayes_p\" tableData),
-  mean (getTableColumn \"svm2_p\" tableData),
-  mean (getTableColumn \"svm3_p\" tableData),
-  mean (getTableColumn \"svmr_p\" tableData)
+unusuallyHighLow (overallComparison [
+  compareCols col "naive_bayes_r"
+    (findWithKey_ "synd" "Hem." tableData)
+  | col <- ["svm1_r", "svm2_r", "svm3_r", "svmr_r"]
 ])
-betterWorse
-
         \end{lstlisting} \\

### 2025-09-25 paper/sec/evaluation.tex
Unpack SciGen.
answer to RQ1

@@
 
-To evaluate the system we used a sample of the \SciGen dataset~\citep{moosavi21}. We aggregated the generated
+To evaluate RQ1, we used a sample of the \SciGen dataset~\citep{moosavi21}, an open source dataset consisting
+of tables from scientific articles and their corresponding descriptions. We aggregated the generated
 expression in the linguistic categories from Table~\ref{tab:natural-language-forms}.
 Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and
-without target-value sharing. The results show that providing the target value (text fragment or numerical
-value) is crucial for achieving a high success rate. Overall the \InterpretationAgent generates a correct
-expression in approximately in 74.9\% of the cases (S.D. 3.0\%).
+without target-value sharing.
 
-The system achieved performance above 68\% in comparison, 77.3\% in data retrieval, and 97\% in min/max search tasks.
-In contrast, accuracy decreased significantly in expressions requiring differences (20\%) and in ranking tasks (0\%).
+\emph{The results show that the system is robust when provided with sufficient guidance but degrades when underspecified.} With the target-value sharing, 
+the InterpretationAgent produced correct Fluid expressions in 74.9\%  (S.D. 3.0\%) of cases,
+but performance dropped to 57.1\% when the target was withheld. This highlights the system's reliance on explicit cues when resolving ambiguous fragments.
 
-Figure~\ref{fig:success_rate_by_complexity} reports the success rate as a function of the number of categories assigned to each expression.
-Fluid instructions combining multiple operations were associated with multiple categories.
-Results show that the success rate was 62\% for simple expressions (one category), increased to 91\% for two categories,
-while no correct expression was generated for expression with 3 categories.
+\emph{Performance also varied across linguistic categories.} Success rates exceeded 68\% for comparison, 77.3\% for data retrieval, and 97\% for min/max search tasks.
+In contrast, accuracy decreased significantly for expressions requiring differences (20\%) and for ranking tasks (0\%).
 
-In addition to evaluating performance when the target expression is provided, we also assessed the model’s ability to produce a
-correct expression when no reference string was given.
-On the \SciGen dataset, the model achieved a 74.9\% success rate when provided with the target string.
-Without the target string, performance decreased moderately to 57.1\%, indicating that the model retains substantial
-ability to reconstruct valid solutions even without explicit guidance
+The trend for compositional complexity is more nuanced as shown in Figure~\ref{fig:success_rate_by_complexity}, which reports the success rate as a function of the number of categories assigned to each expression: success rates are 62\% for single-category expressions, increase to 91\% when two categories are combined, but collapse to 0\% when three categories are involved.
+This suggests that \emph{moderate composition can actually aid performance, perhaps by giving the model clearer structural cues, but that complexity beyond a certain threshold overwhelms the synthesis process}.
+
+%The results show that providing the target value (text fragment or numerical
+%value) is crucial for achieving a high success rate. Overall the \InterpretationAgent generates a correct
+%expression in approximately in 74.9\% of the cases (S.D. 3.0\%).
+
+%Figure~\ref{fig:success_rate_by_complexity} reports the success rate as a function of the number of categories assigned to each expression.
+%Fluid instructions combining multiple operations were associated with multiple categories.
+%Results show that the success rate was 62\% for simple expressions (one category), increased to 91\% for two categories,
+%while no correct expression was generated for expression with 3 categories.
+
+%In addition to evaluating performance when the target expression is provided, we also assessed the model’s ability to produce a
+%correct expression when no reference string was given.
+%On the \SciGen dataset, the model achieved a 74.9\% success rate when provided with the target string.
+%Without the target string, performance decreased moderately to 57.1\%, indicating that the model retains substantial
+%ability to reconstruct valid solutions even without explicit guidance
 

### 2025-09-25 paper/sec/evaluation.tex
small tweak

@@
 
-\emph{The results show that the system is robust when provided with sufficient guidance but degrades when underspecified.} With the target-value sharing, 
+The results show that \emph{the system is robust when provided with sufficient guidance but degrades when underspecified.} With the target-value sharing, 
 the InterpretationAgent produced correct Fluid expressions in 74.9\%  (S.D. 3.0\%) of cases,

### 2025-09-25 shared/util.js
deploy: 541dd3bb5f7a7a7fd5f6027e46adae5d87fd4029

@@ -1,60 +1,3 @@
-function loadHeader () {
-   fetch('shared/header.html')
-      .then(response => response.text())
-      .then(data => {
-         const header = document.createElement('div')
-         header.innerHTML = data
-         activateCurrentLink(header)
-         const divElement = header.children[0]
-         const grid = document.getElementById('grid')
-         grid.parentNode.insertBefore(divElement, grid)
-      })
-      .catch(error => console.error('Error loading shared HTML:', error))
-
-   fetch('shared/footer.html')
-      .then(response => response.text())
-      .then(data => {
-         const footer = document.createElement('div')
-         footer.innerHTML = data
-         const divElement = footer.children[0]
-         const grid = document.getElementById('grid')
-         grid.parentNode.appendChild(divElement)
-      })
-      .catch(error => console.error('Error loading shared HTML:', error))
-}
-
-function eqPaths (path1, path2) {
-   const trailingSlashes = /\/+$/
-   return path1.replace(trailingSlashes, '') === path2.replace(trailingSlashes, '')
-}
-
-function activateCurrentLink (el) {
-   const listItems = el.querySelectorAll('li')
-   const n_ = Array.from(listItems).findIndex(li => {
-      const link = li.querySelector('a')
-      return link && eqPaths(link.getAttribute('href'), window.location.pathname)
-   })
-   if (n_ !== -1) {
-      listItems[n_].classList.add('active-page')
-   }
-}
-
-function loadSubHeader () {
-   fetch('shared/sub-header.html')
-      .then(response => response.text())
-      .then(data => {
-         const header = document.createElement('div')
-         header.innerHTML = data
-         activateCurrentLink(header)
-         const divElements = header.children
-         const grid = document.getElementById('grid')
-         for (let i = Math.min(3, divElements.length - 1); i >= 0; --i) {
-            grid.insertBefore(divElements[i], grid.firstChild)
-         }
-      })
-      .catch(error => console.error('Error loading shared HTML:', error))
-}
-
 function toggleDataPane(gridId) {
    const grid = document.getElementById(gridId)
    const hidden = grid.classList.contains('data-pane-hidden')

### 2025-09-25 paper/sec/authoring-workflow.tex
Tweaks.

@@
 \item \textbf{Code synthesis step.} The expression synthesis step is an error-guided iterative prompting
-loop~\citep{skreta23}, beginning with an initial prompt sent to the LLM (\secref{prompt-design} below)
+loop~\citep{skreta23}, beginning with an initial prompt sent to the LLM (see \emph{Prompt design} below)
 requesting the generation of an expression $e$. Using the Fluid command-line interface, the expression is
@@
 replacement task. The model receives the imported datasets, helper modules, and the current Fluid
-representation of the paragraph, in which a text fragment is marked with the tag \kw{[REPLACE …]}. Its task is
+representation of the paragraph, in which a text fragment is marked with the tag \kw{[REPLACE …]}. The task is
 to substitute this placeholder with a Fluid expression that evaluates exactly to the target string,

### 2025-09-25 paper/sec/evaluation.tex

@@
 specific to each test case, and also how counterfactual performance is impacted by the presence of misleading
-or adversarial phrasing. able~\ref{tab:problematic_cases} shows some of cases which are ``problematic'' in
+or adversarial phrasing. Table~\ref{tab:problematic_cases} shows some of cases which are ``problematic'' in
 this sense; in these cases, producing a valid expression is likely to be challenging because of ambiguities in

### 2025-09-25 paper/sec/evaluation.tex
Tweak.

--- 2025-09-25 paper/sec/evaluation.tex
@@
 ranks, as summarised in Table~\ref{tab:natural-language-forms}) and investigate how accuracy varies with task
-complexity, measured (somewhat crudely and approximately) by the number of query sub-expressions (e.g.
-retrieval, aggregation, or arithmetic) present in the gold solution.
+complexity, measured (somewhat crudely) by the number of query sub-expressions (e.g. retrieval, aggregation,
+or arithmetic) present in the gold solution.
 
@@
 specific to each test case, and also how counterfactual performance is impacted by the presence of misleading
-or adversarial phrasing. able~\ref{tab:problematic_cases} shows some of cases which are ``problematic'' in
-this sense; in these cases, producing a valid expression is likely to be challenging because of ambiguities in
-the input data or accompanying natural language.
+or adversarial phrasing. Table~\ref{tab:problematic_cases} shows some of cases we deem problematic in this
+sense; in these cases, producing a valid expression is likely to be challenging because of ambiguities in the
+input data or accompanying natural language.
 
@@
 
-To evaluate RQ1, we used a sample of the \SciGen dataset~\citep{moosavi21}, an open source dataset consisting
-of tables from scientific articles and their corresponding descriptions. We aggregated the generated
-expression in the linguistic categories from Table~\ref{tab:natural-language-forms}.
-Figure~\ref{fig:success_rate_by_category} illustrates the success rate for each category, both with and
-without target-value sharing.
+\paragraph{Interpretation Accuracy across Linguistic Idioms and Complexity.} To evaluate RQ1, we used a sample
+of the \SciGen dataset~\citep{moosavi21}, an open source dataset consisting of tables from scientific articles
+and their corresponding descriptions. We aggregated the results according to the linguistic categories from
+Table~\ref{tab:natural-language-forms}. Figure~\ref{fig:success_rate_by_category} illustrates the success rate
+for each category, both with and without target-value sharing.
 
-The results show that \emph{the system is robust when provided with sufficient guidance but degrades when underspecified.} With the target-value sharing, 
-the InterpretationAgent produced correct Fluid expressions in 74.9\%  (S.D. 3.0\%) of cases,
-but performance dropped to 57.1\% when the target was withheld. This highlights the system's reliance on explicit cues when resolving ambiguous fragments.
+The results show that \emph{the system is robust when provided with sufficient guidance but degrades when
+underspecified.} With the target-value sharing, the InterpretationAgent produced correct Fluid expressions in
+74.9\%  (S.D. 3.0\%) of cases, but performance dropped to 57.1\% when the target was withheld. This highlights
+the system's reliance on explicit cues when resolving ambiguous fragments.
 
-\emph{Performance also varied across linguistic categories.} Success rates exceeded 68\% for comparison, 77.3\% for data retrieval, and 97\% for min/max search tasks.
-In contrast, accuracy decreased significantly for expressions requiring differences (20\%) and for ranking tasks (0\%).
+\emph{Performance also varied across linguistic categories.} Success rates exceeded 68\% for comparison,
+77.3\% for data retrieval, and 97\% for min/max search tasks. In contrast, accuracy decreased significantly
+for expressions requiring differences (20\%) and for ranking tasks (0\%).
 
-The trend for compositional complexity is more nuanced as shown in Figure~\ref{fig:success_rate_by_complexity}, which reports the success rate as a function of the number of categories assigned to each expression: success rates are 62\% for single-category expressions, increase to 91\% when two categories are combined, but collapse to 0\% when three categories are involved.
-This suggests that \emph{moderate composition can actually aid performance, perhaps by giving the model clearer structural cues, but that complexity beyond a certain threshold overwhelms the synthesis process}.
+The trend for compositional complexity is more nuanced as shown in
+Figure~\ref{fig:success_rate_by_complexity}, which reports the success rate as a function of the number of
+categories assigned to each expression: success rates are 62\% for single-category expressions, increase to
+91\% when two categories are combined, but collapse to 0\% when three categories are involved. This suggests
+that \emph{moderate composition can actually aid performance, perhaps by giving the model clearer structural
+cues, but that complexity beyond a certain threshold overwhelms the synthesis process}.
 
@@
 
-\subsection{Counterfactual testing}
-\label{sec:validation-with-counterfactual-data}
+\paragraph{Generalisability and Robustness.} As a preliminary attempt to address RQ2, we carried out
+\emph{counterfactual testing} to evaluate the robustness of generated expressions under changes to the
+underlying data. In this setup, the input tables were modified according to hand-craft test specifications,
+and both the expected and generated expressions were re-executed to check whether the behaviours remained
+consistent. Across 300 test executions, 121 contained at least one counterfactual error (an average of 3.8 per
+case), of which 42 ultimately still succeeded. These tests highlight cases where an expression may
+coincidentally yield the correct output on the original data but fails to be extensionally equivalent more
+generally (i.e.~under perturbation). For example, in one test the system generated
 
-For the \emph{counterfactual testing} part of our analysis, the input data is deliberately modified in order
-to verify the consistency of the behaviours. Both the expected and generated expressions are executed again
-and the results are compared. If discrepancies are found the \InterpretationAgent ask for a new generation
-sending back that the failure is related to a counterfactual error.
-
-%\rpnote{complete}
-
-Across 300 test executions, 121 contained at least one counterfactual error (avg. 3.8), of which 42 ultimately
-succeeded. Such tests help detect expressions that may yield the expected output but not the correct solution.
-For example, the \InterpretationAgent generated:
-
-\begin{verbatim}
-(findWithKey_ "model" "LSTM2" tableData).time_s
-\end{verbatim}
+\hspace{3mm} \kw{(findWithKey\_ "model" "LSTM2" tableData).time\_s}
 
 intended to retrieve the execution time of the \texttt{LSTM} model, but incorrectly referred to \texttt{LSTM2}.
-Counterfactual testing exposed this error and prompted regeneration.
+Counterfactual testing exposed this mismatch, which would otherwise have gone undetected.
+
+At present, counterfactual tests are used only as an evaluation device, not as part of the authoring workflow
+itself. For future work (\secref{conclusion}), we plan to investigate automatic generation of counterfactual
+tests, allowing these additional robustness checks to be integrated into the document authoring workflow.
 

### 2025-09-25 paper/sec/evaluation.tex
Merge main.

@@
 ranks, as summarised in Table~\ref{tab:natural-language-forms}) and investigate how accuracy varies with task
-complexity, measured (somewhat crudely) by the number of query sub-expressions (e.g. retrieval, aggregation,
-or arithmetic) present in the gold solution.
+complexity, measured (somewhat crudely and approximately) by the number of query sub-expressions (e.g.
+retrieval, aggregation, or arithmetic) present in the gold solution.
 

### 2025-10-27 paper/build.sh
Start on script to build arXiv upload.

@@ -8,4 +8,4 @@ $PDFLATEX $TARGET
 bibtex $TARGET
 $PDFLATEX $TARGET
 $PDFLATEX $TARGET
-rm -f $TARGET.aux $TARGET.dvi $TARGET.log $TARGET.bbl $TARGET.blg $TARGET.out $TARGET.pag $TARGET.cb $TARGET.cb2 $TARGET.toc
+rm -f $TARGET.{aux,bbl,blg,,cb,cb2,dvi,logout,pag,toc}

### 2025-10-27 paper/arXiv.sh
First cut of arXiv script.

@@ -1,0 +1,10 @@
+#!/bin/bash
+set -xe
+
+NAME=arXiv.zip
+ARCHIVE="../$NAME"
+
+. build.sh
+
+zip -r - . > $ARCHIVE
+zip -d $ARCHIVE *.pdf \*.{DS_Store,gitignore}

### 2025-10-27 paper/build.sh

@@ -8,4 +8,4 @@ $PDFLATEX $TARGET
 bibtex $TARGET
 $PDFLATEX $TARGET
 $PDFLATEX $TARGET
-rm -f $TARGET.{aux,bbl,blg,,cb,cb2,dvi,logout,pag,toc}
+rm -f $TARGET.{aux,bbl,blg,,cb,cb2,dvi,log,out,pag,toc}

### 2025-10-27 paper/arXiv.sh
Pin tex-common to 2405cd5.

@@ -8,3 +8,4 @@ ARCHIVE="../$NAME"
 
 zip -r - . > $ARCHIVE
 zip -d $ARCHIVE *.pdf \*.{DS_Store,gitignore}
+zip -d $ARCHIVE arXiv.sh build.bat build.sh

### 2025-10-27 paper/build.sh

@@ -8,4 +8,4 @@ $PDFLATEX $TARGET
 bibtex $TARGET
 $PDFLATEX $TARGET
 $PDFLATEX $TARGET
-rm -f $TARGET.{aux,bbl,blg,,cb,cb2,dvi,log,out,pag,toc}
+rm -f $TARGET.{aux,blg,cb,cb2,dvi,log,out,pag,toc}

### 2025-10-27 script/chart-generation.py

@@ -34,27 +34,6 @@ def generate_success_rate_test_case_plot(df, plot):
     plt.savefig(f"{fig_dir}/success_rate_by_test_case.png")
     plt.close()
 
-def generate_summary_test_case_plot(df, plot):
-    df_plot = df[::-1].reset_index(drop=True)
-    row_height = 0.5
-    plt.figure(figsize=(10, len(df_plot) * row_height))
-
-    for idx, row in df_plot.iterrows():
-        is_success = not pd.isna(row["generated-expression"])
-        color = "green" if is_success else "red"
-        marker_text = f"{'OK' if is_success else 'KO'}"
-        plt.text(0.15, idx, marker_text, ha="left", va="center", fontsize=10.5, color=color)
-
-    plt.yticks(range(len(df_plot)), df_plot["test-case-short"], fontsize=8)
-    plt.xticks([])
-    plt.gca().spines[['top', 'right', 'left', 'bottom']].set_visible(False)
-    plt.grid(False)
-    plt.tick_params(left=False)
-    plt.title("Test Outcome per Test Case (OK = Success, KO = Failure). Max Attempts = 4", pad=20)
-    plt.tight_layout()
-    plt.savefig(f"{fig_dir}/per_row_checkmarks.png")
-    plt.close()
-
 def generate_aggregated_plot(df, plot):
     # esplodo le categorie
     df['expression-type'] = df['expression-type'].str.strip('[]').str.split(',')
@@ -159,7 +138,6 @@ def generate_charts():
 
     sns.set_style("whitegrid")
     generate_success_rate_test_case_plot(df, plt)
-    generate_summary_test_case_plot(df, plt)
     generate_aggregated_plot(df, plt)
     generate_success_rate_by_category_count(df, plt)
 

### 2025-10-27 paper/arXiv.sh
Getting there, but still some up-references from 'paper' folder.

@@ -7,5 +7,5 @@ ARCHIVE="../$NAME"
 . build.sh
 
 zip -r - . > $ARCHIVE
-zip -d $ARCHIVE *.pdf \*.{DS_Store,gitignore}
-zip -d $ARCHIVE arXiv.sh build.bat build.sh
+zip -d $ARCHIVE \*.{DS_Store,gitignore}
+zip -d $ARCHIVE *.pdf .git arXiv.sh build.bat build.sh

### 2025-10-27 paper/sec/appendix-prompts.tex
Getting there, but still some up-references from 'paper' folder.

@@
-\section*{System Prompt Interpretation Agent}
+\section{\InterpretationAgent System Prompt}
+\label{app:system-prompt:interpretation-agent}
 
@@
 
-\section*{System Prompt Recognition Agent}
+\section{\SuggestionAgent System Prompt}
 
-\VerbatimInput[fontsize=\small]{../system-prompt/suggestion-agent/system-prompt.txt}
+\begin{Verbatim}[fontsize=\small]
+You are an expression detector for Fluid language.
+Fluid is a functional programming language used to represent structured data queries and comparisons in a
+transparent way.
+
+TASK DESCRIPTION
+
+Given a natural language paragraph and a structured dataset, identify and annotate the parts of the
+paragraph that can be replaced by a Fluid expression.
+
+You must detect:
+- Explicit values (e.g., scores, names, numbers)
+- Comparative expressions (e.g., *better than*, *worse*, *higher*, *more than*)
+- Superlative or aggregated expressions (e.g., *the best*, *highest*, *maximum*, *top performer*)
+
+FORMAT
+
+Replace each detected expression with:
+
+[REPLACE value=...]
+
+Where `value` contains the **original text** of the expression (e.g., "91.57", "better", "the best") —
+not the rewritten logic or Fluid code.
+
+IMPORTANT RULE
+
+When replacing comparative or superlative expressions (like "better", "worse", "the best", "highest"),
+the `value` **must be the exact original word or phrase** from the paragraph.
+
+Correct:
+S-LSTM gives [REPLACE value="the best"] reported results.
+BiLSTM performs [REPLACE value="better"] than LSTM.
+
+Incorrect:
+S-LSTM gives [REPLACE value="getMaxBy f1 data"] results.
+BiLSTM performs [REPLACE value="BiLSTM.acc > LSTM.acc"] than LSTM.
+
+If needed, annotate separate values independently:
+
+Example:
+BiLSTM gives [REPLACE value="91.2"]% accuracy, which is [REPLACE value="better"] than LSTM.
+
+---
+
+EXAMPLES
+
+Example Fluid code:
+
+let bestModel = getMaxBy f1 data in bestModel.model
+
+---
+
+INPUT EXAMPLE
+
+Paragraph:
+For NER (Table 7), S-LSTM gives an F1-score of 91.57% on the CoNLL test set, which is significantly
+better compared with BiLSTMs. Stacking more layers of BiLSTMs leads to slightly better F1-scores
+compared with a single-layer BiLSTM. Our BiLSTM results are comparable to the results reported
+by Ma and Hovy (2016) and Lample et al. (2016).
+In contrast, S-LSTM gives the best reported results under the same settings.
+In the second section of Table 7,Yang et al. (2017) obtain an Fscore of 91.26%.
+
+Data:
+[
+  {model: "BiLSTM", f1: 90.96},
+  {model: "2 stacked BiLSTM", f1: 91.02},
+  {model: "3 stacked BiLSTM", f1: 91.06},
+  {model: "S-LSTM", f1: 91.57},
+  {model: "yang2017transfer", f1: 91.26}
+]
+
+---
+
+OUTPUT EXAMPLE
+
+For NER (Table 7), S-LSTM gives an F1-score of [REPLACE value=91.57]% on the CoNLL test set,
+which is [REPLACE value="better"] compared with BiLSTMs.
+Stacking more layers of BiLSTMs leads to [REPLACE value="better"] F1-scores compared with a single-layer BiLSTM.
+Our BiLSTM results are comparable to the results reported by Ma and Hovy (2016) and Lample et al. (2016).
+In contrast, S-LSTM gives [REPLACE value="the best"] reported results under the same settings.
+In the second section of Table 7, Yang et al. (2017) obtain an Fscore of [REPLACE value=91.26]%.
+\end{Verbatim}

### 2025-10-27 paper/sec/appendix-prompts.tex

@@
 
-\VerbatimInput[fontsize=\small]{../system-prompt/interpretation-agent/system-prompt.txt}
+\begin{Verbatim}[fontsize=\small]
+You are a specialized language model for the Fluid functional programming language.
+Your task is to analyze a JSON object that represents the user’s Fluid program and its context,
+and to generate the Fluid expression that must replace the [REPLACE value=] placeholder inside
+the paragraph.
+
+Input Structure
+The JSON input always contains:
+-datasets: one or more JSON-like arrays containing the data used by the program
+(scenario-related key–value pairs).
+-imports: Fluid helper libraries provided by the user’s program.
+-code: Additional Fluid functions and definitions from the user’s program.
+-paragraph: A description that includes exactly one [REPLACE ...] tag.
+-paragraphValue: The correct final version of the paragraph (ground truth).
+
+Note: imports, code, and datasets are part of the user’s Fluid program, not just supporting context.
+Your output must be consistent with these definitions.
+
+Task
+Identify the [REPLACE ...] tag in paragraph.
+If the tag has the value property, generate a Fluid expression that evaluates exactly to that value.
+If not, infer the correct value by comparing paragraph, paragraphValue, and (if needed) datasets.
+The result must always be a Fluid expression that evaluates to a string.
+
+Output Format
+Return only the Fluid expression, nothing else.
+
+Constraints
+-Output exactly one valid Fluid expression.
+-Ensure it is syntactically correct and consistent with the provided imports and code.
+\end{Verbatim}
 

### 2025-11-14 paper/reviews/iclr2026/response.md
Initial pass over response.

# Key points to respond to

## AyvP (rating 6, confidence 3)

### Limited to single papers

I guess not actually true, but will need to think a bit about how best to respond.

### Checks for more complex claims cannot be generated with LLMs

Need to think about how we envisage this improving -- perhaps some kind of iterative refinement where LLM and human collaboratively build interpretation.

### Human has to verify all generated links anyway

Again, probably need to think about how we envisage this improving -- automated testing/validation?

## Bezq (rating 4, confidence 4 -- substantial AI content in the review)

### Re. subset of SciGen, could the authors provide more details on the selection logic, sampling criteria, and the exact number of examples used in evaluation?

How realistic would it be to cover all of SciGen, perhaps using the SuggestionAgent to generate the REPLACE tags?

### No analysis of statistical variance beyond standard deviation

We've already discussed improving this.

### Assess generalization to unseen writing styles or datasets

Not sure this makes sense -- we didn't train our system on SciGen.

### Can authors report statistics about the manual validation process? For examples, the ratio of accepted to rejected edits, average validation time per fragment, or the most common sources of rejection?

We might have to concede this point and say that we plan to do a proper user study as future work. The point of this study is to evaluate LLM competence at the basic problem.

The criticism that we may not reduce author workload but simply shift it from writing to debugging is worth responding to as part of this. I guess this applies to AI-generated code in general. There is probably some evidence that net productivity increases even if manual validation of machine-generated code is required.

### Could the authors conduct an ablation study removing or varying these helper components to determine how much of the system’s success derives from the LLM’s own reasoning versus predefined components (e.g. helper routines)?

I think we can do this. (We can also argue that having a fixed/standardised set of definitions might also be helpful, but I think their point remains.)

### The large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning

Not sure I understand this point properly but one related thing we considered studying (and perhaps should in response to this) is what happens when the target value is incorrect. This is somewhat related to RQ2 and the ablation studies we considered for that.

## azjF (rating 3, confidence 3)

### More engineering-oriented, with relatively limited academic research value

Perhaps respond to this as part of respond to point below.

### Experimental design lacks comprehensive justification. Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered)

Maybe we have to concede this point and then propose more in-depth ablation studies.

## VRkd (rating 2, confidence 4)

### Isn't this supposed to be an HCI paper?

Yes, the next step of this work will go to CHI/UIST/IUI.

I think we need to argue that the technical contribution from an ML/NLP point of view is competency evaluation for LLMs.

### Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too.

Let's respond to this as part of the question above.

### The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some documented NLP tasks (like scientific literature?) to begin with.

I think this is more than we can plan to do for this version of the paper, but for a resubmission it might be worth looking at some of the NLP literature (with Fede's help). Unsurprisingly, there are well-known NLP problems that are highly relevant and could form part of the "capability set" required for a tool providing automated computational interpretation of a scientific paper. From a very quick glance, all of the following datasets (from the computational lingustics literature) are potentially relevant and might be of use for a future study:

SciERC (https://aclanthology.org/D18-1360/)
Qasper (https://aclanthology.org/2021.naacl-main.365/)
DROP (https://aclanthology.org/N19-1246/)
NUMGLUE (https://aclanthology.org/2022.acl-long.246.pdf)
SCIREX (https://aclanthology.org/2020.acl-main.670.pdf)
TAT-QA (https://aclanthology.org/2021.acl-long.254/)

(Fig. 1 in the TAT-QA paper looks very similar to some of ours.)

### Many of the niche details, like Table 1, are perhaps unnecessary for the main paper

Yes we wondered the same but in the end decided to be explicit about the kind of linguistic/quantitative competencies we wanted to evaluate because of the NLP component.

### 2025-11-17 paper/tex/authors.tex
Fix authors.
Fix spelling of Ali's name.

@@
    \And
-   Joe Bond \\
-   University of Bristol \\
-   Bristol, UK \\
-   \texttt{j.bond@bristol.ac.uk} \\
-   \And
    Cristina David \\
@@
    \And
-   Ali Mohammed \\
+   Mattia De Rosa \\
+   University of Salerno \\
+   Fisciano, Italy \\
+   \texttt{matderosa@unisa.it} \\
+   \And
+   Ali Muhammad \\
    Manchester Metropolitan University \\
@@
    \texttt{22465707@stu.mmu.ac.uk} \\
+   \And
+   Federico Nanni \\
+   The Alan Turing Institute \\
+   London, UK \\
+   \texttt{fnanni@turing.ac.uk} \\
    \And

### 2025-11-17 paper/arXiv.sh

@@ -1,7 +1,7 @@
 #!/bin/bash
 set -xe
 
-NAME=arXiv.zip
+NAME=preprint.zip
 ARCHIVE="../$NAME"
 
 . build.sh

### 2025-11-19 paper/reviews/iclr2026/response.md
Move old 'response' draft to notes, start on actual response.

@@
-# Key points to respond to
+# Response to reviewers
 
-## AyvP (rating 6, confidence 3)
+We thank our three reviewers for their useful feedback and suggestions for improving the paper. We address
+specific comments of each reviewer below, and then summarise our planned revisions to the paper.
 
-### Limited to single papers
-
-I guess not actually true, but will need to think a bit about how best to respond.
-
-### Checks for more complex claims cannot be generated with LLMs
-
-Need to think about how we envisage this improving -- perhaps some kind of iterative refinement where LLM and human collaboratively build interpretation.
-
-### Human has to verify all generated links anyway
-
-Again, probably need to think about how we envisage this improving -- automated testing/validation?
-
-## Bezq (rating 4, confidence 4 -- substantial AI content in the review)
-
-### Re. subset of SciGen, could the authors provide more details on the selection logic, sampling criteria, and the exact number of examples used in evaluation?
-
-How realistic would it be to cover all of SciGen, perhaps using the SuggestionAgent to generate the REPLACE tags?
-
-### No analysis of statistical variance beyond standard deviation
-
-We've already discussed improving this.
-
-### Assess generalization to unseen writing styles or datasets
-
-Not sure this makes sense -- we didn't train our system on SciGen.
-
-### Can authors report statistics about the manual validation process?
-For examples, the ratio of accepted to rejected edits, average validation time per fragment, or the most common sources of rejection?
-
-We might have to concede this point and say that we plan to do a proper user study as future work. The point of this study is to evaluate LLM competence at the basic problem.
-
-The criticism that we may not reduce author workload but simply shift it from writing to debugging is worth responding to as part of this.
-I guess this applies to AI-generated code in general.
-There is probably some evidence that net productivity increases even if manual validation of machine-generated code is required.
-
-### Could the authors conduct an ablation study removing or varying these helper components to determine how much of the system’s success derives from the LLM’s own reasoning versus predefined components (e.g.
-helper routines)?
-
-I think we can do this. (We can also argue that having a fixed/standardised set of definitions might also be helpful, but I think their point remains.)
-
-### The large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning
-
-Not sure I understand this point properly but one related thing we considered studying (and perhaps should in response to this) is what happens when the target value is incorrect.
-This is somewhat related to RQ2 and the ablation studies we considered for that.
-
-## azjF (rating 3, confidence 3)
-
-### More engineering-oriented, with relatively limited academic research value
-
-Perhaps respond to this as part of respond to point below.
-
-### Experimental design lacks comprehensive justification.
-Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered)
-
-Maybe we have to concede this point and then propose more in-depth ablation studies.
-
-## VRkd (rating 2, confidence 4)
-
-### Isn't this supposed to be an HCI paper?
-
-Yes, the next step of this work will go to CHI/UIST/IUI.
-
-I think we need to argue that the technical contribution from an ML/NLP point of view is competency evaluation for LLMs.
-
-### Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too.
-
-Let's respond to this as part of the question above.
-
-### The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some documented NLP tasks (like scientific literature?) to begin with.
-
-I think this is more than we can plan to do for this version of the paper, but for a resubmission it might be worth looking at some of the NLP literature (with Fede's help).
-Unsurprisingly, there are well-known NLP problems that are highly relevant and could form part of the "capability set" required for a tool providing automated computational interpretation of a scientific paper.
-From a very quick glance, all of the following datasets (from the computational lingustics literature) are potentially relevant and might be of use for a future study:
-
-SciERC (https://aclanthology.org/D18-1360/)
-Qasper (https://aclanthology.org/2021.naacl-main.365/)
-DROP (https://aclanthology.org/N19-1246/)
-NUMGLUE (https://aclanthology.org/2022.acl-long.246.pdf)
-SCIREX (https://aclanthology.org/2020.acl-main.670.pdf)
-TAT-QA (https://aclanthology.org/2021.acl-long.254/)
-
-(Fig. 1 in the TAT-QA paper looks very similar to some of ours.)
-
-### Many of the niche details, like Table 1, are perhaps unnecessary for the main paper
-
-Yes we wondered the same but in the end decided to be explicit about the kind of linguistic/quantitative competencies we wanted to evaluate because of the NLP component.
+## Planned revisions to paper

### 2025-11-19 paper/reviews/iclr2026/response.md
Reviewer AyvP.
Discussion on future work.

@@
 
+## Reviewer AyvP
+
+_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our
+evaluation we have only considered excerpts from single papers. A multi-paper setup is an interesting use case
+and one we will consider for future work.
+
+_Checks for more complex claims cannot be generated with LLMs. Even for the simple cases, a human has to
+verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
+trust point of view. However, there do seem to be some likely ways this could be improved: in particular for
+cases where an LLM is unable to generate a full solution in one step, we envisage a more interactive workflow
+similar to working with Copilot.
+
+The most important consideration here
+
 ## Planned revisions to paper

### 2025-11-19 paper/reviews/iclr2026/response.md
Some burble, but needs sharpening.

@@
 verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
-trust point of view. However, there do seem to be some likely ways this could be improved: in particular for
-cases where an LLM is unable to generate a full solution in one step, we envisage a more interactive workflow
-similar to working with Copilot.
+trust point of view. However, there do seem to be some ways this could be improved: in particular for cases
+where an LLM is unable to generate a full solution in one step, we envisage a more interactive workflow
+similar to working with Copilot. For the manual validation, automated testing should be able to generate
+counterexamples. But the key point is that for a relatively small amount of author effort, every _reader_
+benefits. So some kind of hybrid system where the "easy" cases are fully or semi-automated and the more
+complex cases are constructed by hand (with a Copilot-like assistant) seems the most plausible.
 

### 2025-11-19 paper/reviews/iclr2026/response.md
Some tweaks.

--- 2025-11-19 paper/reviews/iclr2026/response.md
@@
 verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
-trust point of view. However, there do seem to be some likely ways this could be improved: in particular for
-cases where an LLM is unable to generate a full solution in one step, we envisage a more interactive workflow
-similar to working with Copilot.
+trust point of view. For the simple cases, where it may seem unfortunate that author verification is still
+needed, the key point here is that a small amount of author effort translates into a significant _reader_
+benefit (since there are many more readers than authors). In future work we anticipate that automated testing
+can reduce the author burden in this cases by automatically generating counterexamples.
 
-The most important consideration here
+More the more complex cases where current LLMs are unable to generate a full solution in one step, we envisage
+a more interactive workflow similar to working with Copilot. some kind of hybrid system where the "easy" cases
+are fully or semi-automated and the more complex cases are constructed by hand (with a Copilot-like assistant)
+seems the most plausible.
 

### 2025-11-19 paper/reviews/iclr2026/response.md
Improvements to AyvP response.

--- 2025-11-19 paper/reviews/iclr2026/response.md
@@
 evaluation we have only considered excerpts from single papers. A multi-paper setup is an interesting use case
-and one we will consider for future work.
+and we will consider for a real-world tool.
 
@@
 verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
-trust point of view. However, there do seem to be some likely ways this could be improved: in particular for
-cases where an LLM is unable to generate a full solution in one step, we envisage a more interactive workflow
-similar to working with Copilot.
+trust point of view. For the simple cases, where it may seem unfortunate that author verification is still
+needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since
+there are many more readers than authors). In future work we anticipate automated testing as a way of reducing
+the author burden by automatically exercising the generated code.
 
-The most important consideration here
+For the more complex cases, where current LLMs are unable to generate a full solution in one step, we envisage
+a more interactive workflow, similar to working with Copilot. For the present paper we are focused on
+exploring the limits of current LLM capabilities and validating the basic idea.
 

### 2025-11-20 paper/reviews/iclr2026/response.md
Onto Bezq.

@@
 a more interactive workflow, similar to working with Copilot. For the present paper we are focused on
-exploring the limits of current LLM capabilities and validating the basic idea.
+exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.
+
+## Reviewer Bezq
+
+_Re. subset of SciGen, could you provide details on the selection logic, sampling criteria, and exact number
+of examples used in evaluation?_
+
+For the submitted version we relied on a small subset of SciGen, with only XX example replacement tasks.
 

### 2025-11-20 paper/reviews/iclr2026/response.md
Reinstate class files.

@@
 
-_Re. subset of SciGen, could you provide details on the selection logic, sampling criteria, and exact number
-of examples used in evaluation?_
+_Can authors report statistics about the manual validation process?_
+
+Unfortunately a user study is out-of-scope for this work; see [...]
+
+_Re. the subset of SciGen used, could you provide details on the selection logic, sampling criteria, and exact
+number of examples used in evaluation?_
+
+[Number of example papers + replacement problems within each of those]
+
+The paper states that experiments were conducted on “a subsample of the SciGen dataset” (line 367), but it remains unclear how this subset was chosen.
+Could the authors provide more details on the selection logic, sampling criteria, and the exact number of examples used in evaluation?
+
+It seems this framework rely on predefined helper routines such as trendWord or growShrink, which encode semantic logic that the model merely invokes rather than learns.
+Moreover, the large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning.
+Could the authors conduct an ablation study removing or varying these helper components to determine how much of the system’s success derives from the LLM’s own reasoning versus predefined components?
 

### 2025-11-20 paper/reviews/iclr2026/response.md
WIP.

@@
 
-_Re. the subset of SciGen used, could you provide details on the selection logic, sampling criteria, and exact
-number of examples used in evaluation?_
+_Re. the subset of SciGen used, could you provide details on the selection logic/sampling criteria and exact
+number of examples used in the evaluation?_
 
 [Number of example papers + replacement problems within each of those]
+
+However we have increased the size of the dataset in two ways:
+- for measuring correctness only, we have added automation scripts to allow use to process the whole of SciGen (400+ examples)
+- for measuring
 

### 2025-11-21 paper/reviews/iclr2026/response.md
WIP.

@@
 
-Unfortunately a user study is out-of-scope for this work; see [...]
+Unfortunately a user study is out of scope for this work. We plan to do that next now that we have a good
+handle on the required LLM competencies.
 
@@
 
-[Number of example papers + replacement problems within each of those]
+[19 of 400 + n replacement problems within each of those]
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Remove line breaks.

@@
 
-We thank our three reviewers for their useful feedback and suggestions for improving the paper. We address
-specific comments of each reviewer below, and then summarise our planned revisions to the paper.
+We thank our three reviewers for their useful feedback and suggestions for revising the paper. We address specific comments of each reviewer below, and then summarise how we plan to improve the paper.
 
@@
 
-_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our
-evaluation we have only considered excerpts from single papers. A multi-paper setup is an interesting use case
-and we will consider for a real-world tool.
+_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our evaluation we have only considered excerpts from single papers.
+A multi-paper setup is an interesting use case and one we will consider for a real-world tool.
 
-_Checks for more complex claims cannot be generated with LLMs. Even for the simple cases, a human has to
-verify the generated links._ Keeping a human in the loop for verification purposes seems important from a
-trust point of view. For the simple cases, where it may seem unfortunate that author verification is still
-needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since
-there are many more readers than authors). In future work we anticipate automated testing as a way of reducing
-the author burden by automatically exercising the generated code.
+_Checks for more complex claims cannot be generated with LLMs.
+Even for the simple cases, a human has to verify the generated links._ Keeping a human in the loop for verification purposes seems important from a trust point of view.
+For the simple cases, where it may seem unfortunate that author verification is still needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since there are many more readers than authors).
+In future work we anticipate providing automated testing as a way of reducing author burden by automatically exercising the generated code.
 
-For the more complex cases, where current LLMs are unable to generate a full solution in one step, we envisage
-a more interactive workflow, similar to working with Copilot. For the present paper we are focused on
-exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.
+For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot.
+For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.
 
@@
 
-Unfortunately a user study is out of scope for this work. We plan to do that next now that we have a good
-handle on the required LLM competencies.
+Unfortunately a user study is out of scope for this work. We plan to do that next now that we have a good handle on the required LLM competencies.
 
-_Re. the subset of SciGen used, could you provide details on the selection logic/sampling criteria and exact
-number of examples used in the evaluation?_
+_Re. the subset of SciGen used, could you provide details on the selection logic/sampling criteria and exact number of examples used in the evaluation?_
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Continue with Bezq.

@@
 
-Unfortunately a user study is out of scope for this work. We plan to do that next now that we have a good handle on the required LLM competencies.
+A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow.
+For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of of the required LLM competencies.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Planned revisions.

@@
 ## Planned revisions to paper
+
+- Clarify that manually annotated subsets of SciGen are randomly chosen, and carry out additional manual annotation to grow subset from 2% to 5%
+- Evaluate SuggestionAgent performance against manual annotations
+- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations

### 2025-11-21 paper/reviews/iclr2026/response.md
Planned revisions.

--- 2025-11-21 paper/reviews/iclr2026/response.md
@@
 
+For the submitted version we were only able to manually annotated a 2% subset of the 400 SciGen issues, with only XX example replacement tasks.
+
 [19 of 400 + n replacement problems within each of those]
@@
 
-For the submitted version we relied on a small subset of SciGen, with only XX example replacement tasks.
+## Planned revisions to paper
 
-## Planned revisions to paper
+- Clarify that manually annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%
+- Evaluate SuggestionAgent performance against manual annotations
+- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations
+
+We will include our manually annotated subset as supplementary material and release it as an open source benchmark.

### 2025-11-21 paper/reviews/iclr2026/response.md
WIP.

--- 2025-11-21 paper/reviews/iclr2026/response.md
@@
 
-[19 of 400 + n replacement problems within each of those]
+For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues (approximately 2% of the total), resulting in XX labelled replacement tasks.
+We have drawn a second (larger) random subsample (5% of the total) for manual annotation and will report on these in the final version of the paper.
+We have also added an automation pipeline to process every SciGen example using the Suggestion Agent.
+By evaluating the performance of the SuggestionAgent against the manual annotations to estimate the noise rate, we are now able to evaluate the InterpretationAgent using the entirety of the SciGen data set, with confidence intervals adjusted to incorporate estimated noise.
 
-However we have increased the size of the dataset in two ways:
-- for measuring correctness only, we have added automation scripts to allow use to process the whole of SciGen (400+ examples)
-- for measuring
+_The framework rely on predefined helper routines such as trendWord or growShrink_. Could the authors conduct an ablation study removing or varying these helper components?
 
-The paper states that experiments were conducted on “a subsample of the SciGen dataset” (line 367), but it remains unclear how this subset was chosen.
-Could the authors provide more details on the selection logic, sampling criteria, and the exact number of examples used in evaluation?
+This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string).
+It is also worth emphasising that a fixed/standardised set of definitions (for a given paper or research community, say) might also be helpful, but your point remains valid.
 
-It seems this framework rely on predefined helper routines such as trendWord or growShrink, which encode semantic logic that the model merely invokes rather than learns.
-Moreover, the large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning.
-Could the authors conduct an ablation study removing or varying these helper components to determine how much of the system’s success derives from the LLM’s own reasoning versus predefined components?
-
-For the submitted version we relied on a small subset of SciGen, with only XX example replacement tasks.
+_Large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning._
 
 ## Planned revisions to paper
+
+- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%
+- Evaluate SuggestionAgent performance against manual annotations
+- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)
+
+We will include our manually annotated subset as supplementary material and release it as an open source benchmark.

### 2025-11-21 paper/reviews/iclr2026/response.md
Reviewer VRkd.

@@
 A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow.
-For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of of the required LLM competencies.
+For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of the required LLM competencies.
 
@@
 For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues (approximately 2% of the total), resulting in XX labelled replacement tasks.
-We have drawn a second (larger) random subsample (5% of the total) for manual annotation and will report on these in the final version of the paper.
-We have also added an automation pipeline to process every SciGen example using the Suggestion Agent.
-By evaluating the performance of the SuggestionAgent against the manual annotations to estimate the noise rate, we are now able to evaluate the InterpretationAgent using the entirety of the SciGen data set, with confidence intervals adjusted to incorporate estimated noise.
+We have drawn a second (larger) random subsample (5% of the total) for manual annotation and will report on this in the final version of the paper.
+We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate SuggestionAgent performance using manual annotations to estimate the noise rate.
+We are now able to evaluate the InterpretationAgent using the entirety of the SciGen data set, with confidence intervals adjusted to incorporate estimated noise.
 
@@
 This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string).
-It is also worth emphasising that a fixed/standardised set of definitions (for a given paper or research community, say) might also be helpful, but your point remains valid.
+It is worth emphasising that a fixed/standardised set of terminological definitions (for a given paper or research community, say) might be helpful (or even important in some domains, such as IPCC Summary for Policymaker reports), but your point remains valid.
 
@@
 
+## Reviewer VRkd
+
+> Many of the niche details, like Table 1, are perhaps unnecessary for the main paper
+
+We would like to keep Table 1 in the body of the paper to be explicit about the specific linguistic/quantitative LLM competencies we evaluate in the paper.
+However it may be possible to streamline the figure a bit.
+We will also make a pass over the paper for other details that might be spurious for our audience here.
+
 ## Planned revisions to paper
 
-- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%
-- Evaluate SuggestionAgent performance against manual annotations
+- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%.
+We will include this data set as supplementary material and release it as an open source benchmark.
+- Evaluate SuggestionAgent performance against manual annotations; estimate noise in SuggestionAgent labelling.
 - Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)
-
-We will include our manually annotated subset as supplementary material and release it as an open source benchmark.

### 2025-11-21 paper/reviews/iclr2026/response.md
Progress on Bezq.

@@
 
-A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow.
+A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow; we recognise there is a risk (as with all AI-generated code) of simply shifting author workload from writing to debugging.
 For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of the required LLM competencies.
@@
 
+In response to additional points:
+
+_No analysis of statistical variance beyond standard deviation._ We are now reporting using box plots to give
+a better sense of the underlying distribution, annotated with the number of samples.
+
 ## Reviewer VRkd

### 2025-11-21 paper/reviews/iclr2026/response.md
Progress on Bezq.

@@
 
-In response to additional points:
+The target string is present in one of our main use cases (where the reader or author is trying to retrofit an interpretation to text already written), so this doesn't qualify as answer leakage per se.
+Counterfactual testing, which is possible for manually annotated solutions, exposes whether some kind of computational reasoning has actually happened.
+We will improve our reporting on this (RQ2).
 
-_No analysis of statistical variance beyond standard deviation._ We are now reporting using box plots to give
-a better sense of the underlying distribution, annotated with the number of samples.
+To respond to your additional points:
+
+_No analysis of statistical variance beyond standard deviation._ We have changed our figures to use box plots for a better sense of the underlying distribution, annotated with the number of samples.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Progress on Bezq.

@@
 
-We would like to keep Table 1 in the body of the paper to be explicit about the specific linguistic/quantitative LLM competencies we evaluate in the paper.
-However it may be possible to streamline the figure a bit.
-We will also make a pass over the paper for other details that might be spurious for our audience here.
+We prefer to keep Table 1 in the body of the paper, to be explicit about the specific linguistic/quantitative LLM competencies evaluated in the paper; however it should be possible to streamline the figure a bit.
+We will also make a pass over the paper for other details that might be spurious for the ICLR audience.
 
@@
 - Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)
+- Expand the treatment of RQ2

### 2025-11-21 paper/reviews/iclr2026/response.md
Progress on VRkd.

@@
 
-_No analysis of statistical variance beyond standard deviation._ We have changed our figures to use box plots for a better sense of the underlying distribution, annotated with the number of samples.
+_No analysis of statistical variance beyond standard deviation._
+
+Our figures now use box plots for a better sense of the underlying distribution, annotated with the number of samples.
+
+_Assess generalization to unseen writing styles or datasets._
+
+In our experiments the LLM was not fine-tuned or prompted with SciGen examples other than as part of the test cases themselves, so all the SciGen examples were acting as out-of-distribution problems.
+
+## Reviewer azjF
+
+_More engineering-oriented, with relatively limited academic research value._
+
+_Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered)._
+
+Doing something new does indeed make comparison to baseline methods difficult.
+However we agree more in-depth ablation studies would be appropriate.
+We are investigating removing the predefined helper routines (see response to reviewer Bezq above); we will consider other ablation possibilities and include those in our reporting on RQ1 and RQ2.
 
 ## Reviewer VRkd
+
+> Isn't this supposed to be an HCI paper? To me, there isn't any technical contribution for an ML venue beyond getting the system implemented.
+> Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too.
+
+The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study.
+This will likelygo to venue like CHI or IUI, as suggested.
+For this paper, we aim to validate key components of such a tool, including:
+
+- competency evaluation for LLMs (which will substantially improve for this paper)
+- proof-of-concept agentic architecture, including support for compiler-in-the-loop and manual validation (both important for this task)
+
+We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper was in need further work, and we plan to substantially address that in the final revision.
+
+### The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some documented NLP tasks (like scientific literature?) to begin with.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Progress on VRkd.

@@
 
-Doing something new does indeed make comparison to baseline methods difficult.
-However we agree more in-depth ablation studies would be appropriate.
-We are investigating removing the predefined helper routines (see response to reviewer Bezq above); we will consider other ablation possibilities and include those in our reporting on RQ1 and RQ2.
+The lack of directly related prior work does indeed make comparison to baseline methods difficult.
+We agree more in-depth ablation studies would be appropriate; we will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other other similar experiments, and include those in our reporting on RQ1 and RQ2.
 
@@
 The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study.
-This will likelygo to venue like CHI or IUI, as suggested.
-For this paper, we aim to validate key components of such a tool, including:
+This will likely go to venue like CHI or IUI, as suggested.
+For this paper, our aim to validate key components of such a tool, including:
 
-- competency evaluation for LLMs (which will substantially improve for this paper)
+- competency evaluation for LLMs (which we will substantially improve for this paper)
 - proof-of-concept agentic architecture, including support for compiler-in-the-loop and manual validation (both important for this task)

### 2025-11-21 paper/reviews/iclr2026/response.md
Do we need to respond to azjF on 'relatively limited academic research value' point?

@@
 
-_More engineering-oriented, with relatively limited academic research value._
+_More engineering-oriented, with relatively limited academic research value._ It is an engineering-oriented paper, however we address important research questions around how such an agent-based tool might be designed, how currently LLM competencies meet the needs of such a design, and how manual and automatic verification can be integrated and assessed.
 
-_Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered)._
-
-The lack of directly related prior work does indeed make comparison to baseline methods difficult.
+_Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered)._ The lack of directly related prior work does indeed make comparison to baseline methods difficult.
 We agree more in-depth ablation studies would be appropriate; we will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other other similar experiments, and include those in our reporting on RQ1 and RQ2.
@@
 
-> Isn't this supposed to be an HCI paper? To me, there isn't any technical contribution for an ML venue beyond getting the system implemented.
-> Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too.
-
-The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study.
-This will likely go to venue like CHI or IUI, as suggested.
+_Isn't this supposed to be an HCI paper?
+To me, there isn't any technical contribution for an ML venue beyond getting the system implemented.
+Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too._ The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study.
+And indeed this would likely go to venue like CHI or IUI.
 For this paper, our aim to validate key components of such a tool, including:
@@
 
-We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper was in need further work, and we plan to substantially address that in the final revision.
+We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper was in need further work.
+We discuss elsewhere how we plan to address that in the final revision.
 
@@
 
-> Many of the niche details, like Table 1, are perhaps unnecessary for the main paper
-
-We prefer to keep Table 1 in the body of the paper, to be explicit about the specific linguistic/quantitative LLM competencies evaluated in the paper; however it should be possible to streamline the figure a bit.
+_Many of the niche details, like Table 1, are perhaps unnecessary for the main paper._ We prefer to keep Table 1 in the body of the paper, to be explicit about the specific linguistic/quantitative LLM competencies evaluated in the paper; however it should be possible to streamline the figure a bit.
 We will also make a pass over the paper for other details that might be spurious for the ICLR audience.

### 2025-11-21 paper/reviews/iclr2026/response.md
Separate comment structure.

@@
-# Response to reviewers
+# General Response to Reviewers
+[Post as initial comment, then one comment per reviewer, as below]
 
-We thank our three reviewers for their useful feedback and suggestions for revising the paper. We address specific comments of each reviewer below, and then summarise how we plan to improve the paper.
+We thank all the reviewers for their time and expertise in evaluating our submission.
+We are grateful for the positive feedback, including the recognition of [..] contribution (Reviewiers ???), the acknowledgment that [..] well-motivated (Reviewers ??), and the appreciation of (Reviewers ??).
+The constructive feedback provided by the reviewers has been invaluable in improving the paper.
+We have provided detailed responses to each reviewer's comments individually and outline below the key results from new experiments conducted in response to multiple reviewer requests.
 
-## Reviewer AyvP
+## Planned Revisions and New Experiments
+
+- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%.
+We will include this data set as supplementary material and release it as an open source benchmark.
+- Evaluate SuggestionAgent performance against manual annotations; estimate noise in SuggestionAgent labelling.
+- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)
+- Expand the treatment of RQ2
+
+# Response to Reviewer AyvP
 
@@
 
-## Reviewer Bezq
+# Response to Reviewer Bezq
 
@@
 
-## Reviewer azjF
+# Response to Reviewer azjF
 
@@
 
-## Reviewer VRkd
+# Response to Reviewer VRkd
 
@@
 
-### The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some documented NLP tasks (like scientific literature?) to begin with.
+_The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some documented NLP tasks (like scientific literature?) to begin with._
 
@@
 We will also make a pass over the paper for other details that might be spurious for the ICLR audience.
-
-## Planned revisions to paper
-
-- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%.
-We will include this data set as supplementary material and release it as an open source benchmark.
-- Evaluate SuggestionAgent performance against manual annotations; estimate noise in SuggestionAgent labelling.
-- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)
-- Expand the treatment of RQ2

### 2025-11-21 src/main/java/authoringassistant/SuggestionAgent.java
[update] loopback for wrong categories generation

@@ -1,6 +1,7 @@
 package authoringassistant;
 
 import authoringassistant.paragraph.Expression;
+import authoringassistant.paragraph.ExpressionCategory;
 import authoringassistant.paragraph.Literal;
 import authoringassistant.paragraph.Paragraph;
 import it.unisa.cluelab.lllm.llm.LLMEvaluatorAgent;
@@ -21,7 +22,7 @@ import java.util.regex.Pattern;
 
 public class SuggestionAgent {
 
-    private static final Pattern REPLACE = Pattern.compile("\\[REPLACE value=\"?(.*?)\"?]");
+    private static final Pattern REPLACE = Pattern.compile("\\[REPLACE value=\"?(.*?)\"? categories=\"?(.*?)\"?]");
     private static final Path SYSTEM_PROMPT_PATH = Path.of("system-prompt", "suggestion-agent", "system-prompt.txt");
     private final LLMEvaluatorAgent<String> llm;
 
@@ -38,12 +39,27 @@ public class SuggestionAgent {
             text = extractText(p);
         }
         PromptList prompts = buildPrompts(text);
-        String result = llm.evaluate(prompts, null);
-        if(result == null) {
-            return p;
+        String result;
+        while(true) {
+            try {
+                result = llm.evaluate(prompts, null);
+                if (result == null) {
+                    return p;
+                }
+                Paragraph paragraph = parseParagraph(result);
+                return new Program(paragraph, p.getDatasets(),p.getImports(),p.getCode(),p.get_loadedDatasets(),p.getTestCaseFileName(),p.getTest_datasets());
+            } catch (IllegalArgumentException ex) {
+                prompts.addUserPrompt("Invalid category! Please enter only one of the following categories:   - COMPARISON\n" +
+                        "  - RANK\n" +
+                        "  - RATIO\n" +
+                        "  - DATA_RETRIEVAL\n" +
+                        "  - AVERAGE\n" +
+                        "  - SUM\n" +
+                        "  - MIN_MAX");
+                continue;
+            }
         }
-        Paragraph paragraph = parseParagraph(result);
-        return new Program(paragraph, p.getDatasets(),p.getImports(),p.getCode(),p.get_loadedDatasets(),p.getTestCaseFileName(),p.getTest_datasets());
+
     }
 
     private Paragraph parseParagraph(String text) {
@@ -59,7 +75,9 @@ public class SuggestionAgent {
             }
             // Expression
[…]

### 2025-11-21 paper/reviews/iclr2026/response.md
Some burble in 'general response'.

@@
+[Post each of the top-level sections below a separate comment, with title as given]
+
 # General Response to Reviewers
-[Post as initial comment, then one comment per reviewer, as below]
 
-We thank all the reviewers for their time and expertise in evaluating our submission.
-We are grateful for the positive feedback, including the recognition of [..] contribution (Reviewiers ???), the acknowledgment that [..] well-motivated (Reviewers ??), and the appreciation of (Reviewers ??).
-The constructive feedback provided by the reviewers has been invaluable in improving the paper.
-We have provided detailed responses to each reviewer's comments individually and outline below the key results from new experiments conducted in response to multiple reviewer requests.
+We thank all reviewers for their time and expertise in evaluating our submission.
+We are grateful for the positive feedback, including the recognition that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), the appreciation that the exploration of a connection between LLM-based code synthesis and data provenance systems is novel (Reviewer Bezq) and that our experimental results, while in need of some further work, show that the approach looks feasible in practice (Reviewers ??).
+The constructive feedback provided by the reviewers has been invaluable in understanding how to improve the evaluation.
+We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper and the new experiments conducted in response to multiple reviewer requests.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Start organising around explicit quotes of reviewer comments.

@@
 
-_Restriction to single papers._ Nothing about our design restricts it to single papers, although for our evaluation we have only considered excerpts from single papers.
-A multi-paper setup is an interesting use case and one we will consider for a real-world tool.
+> - The proposed system still seems to be in early stages with respect to what can be verified, which limits its usefulness in practice.
+First, as far as I understand, it is limited to single papers and does not allow to check claims a paper makes about results presented in another paper.
+This is where a system like the proposed would be most useful
 
-_Checks for more complex claims cannot be generated with LLMs.
-Even for the simple cases, a human has to verify the generated links._ Keeping a human in the loop for verification purposes seems important from a trust point of view.
+Thank you for the suggestion; a multi-paper setup is indeed an interesting use case.
+Although for our evaluation we only considered excerpts from single papers, nothing about our design restricts it to such, so we will consider this for a real-world tool.
+
+> - Second, the claims that can be checked seem to be quite simple (checks for more complex claims cannot be generated with LLMs) and thus easy to check manually.
+Even then, a human has to verify all generated links, which in at least some cases are wrong.
+This begs the question of whether a manual system where only larger and more important claims are annotated would not be more useful in practice.
+
+Keeping a human in the loop for verification purposes seems important from a trust point of view.
 For the simple cases, where it may seem unfortunate that author verification is still needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since there are many more readers than authors).

### 2025-11-21 paper/reviews/iclr2026/response.md
Progress.

@@
 - Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%.
-We will include this data set as supplementary material and release it as an open source benchmark.
+We will include this data set as supplementary material and also plan to release it as an open source benchmark.
 - Evaluate SuggestionAgent performance against manual annotations; estimate noise in SuggestionAgent labelling.
@@
 
-_Can authors report statistics about the manual validation process?_
+> - Although the paper provides useful category-level statistics, it does not analyze statistical variance beyond reporting standard deviation or assess generalization to unseen writing styles or datasets.
 
+We have changed our figures now use box plots to provide a better sense of the underlying distribution, and annotated with number of samples; thanks for the suggestion.
+
+In our experiments the LLM was not fine-tuned or prompted with SciGen examples other than as part of the testing step, so all the SciGen examples were acting as out-of-distribution problems.
+We will make sure this is clear.
+
+> - Can authors report statistics about the manual validation process?
+For examples, the ratio of accepted to rejected edits, average validation time per fragment, or the most common sources of rejection?
+Such data would illustrate how practical and scalable the workflow is for real authors.
+
+Thank you for this point.
 A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow; we recognise there is a risk (as with all AI-generated code) of simply shifting author workload from writing to debugging.
@@
 
-_Re. the subset of SciGen used, could you provide details on the selection logic/sampling criteria and exact number of examples used in the evaluation?_
+> - The paper states that experiments were conducted on “a subsample of the SciGen dataset” (line 367), but it remains unclear how this subset was chosen.
+Could the authors provide more details on the selection logic, sampling criteria, and the exact number of examples used in evaluation?
 
@@
 
-_The framework rely on predefined helper routines such as trendWord or growShrink_. Could the authors conduct an ablation study removing or varying these helper components?
+> - It seems this framework rely on predefined helper routines such as trendWord or growShrink, which encode semantic logic that the model merely invokes rather than learns.
+Moreover, the large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning.
+Could the authors conduct an ablation study removing or varying these helper components to determine how much of the system’s success derives from the LLM’s own reasoning versus predefined components?
 
 This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string).
-It is worth emphasising that a fixed/standardised set of terminological definitions (for a given paper or research community, say) might be helpful (or even important in some domains, such as IPCC Summary for Policymaker reports), but your point remains valid.
+It is worth emphasising that adopting a fixed or standardised set of terminological definitions (for a given paper or research community, say) might be helpful (or even important in some domains, such as IPCC Summary for Policymaker reports), but your point remains valid.
 
@@
 
-The target string is present in one of our main use cases (where the reader or author is trying to retrofit an interpretation to text already written), so this doesn't qualify as answer leakage per se.
-Counterfactual testing, which is possible for manually annotated solutions, exposes whether some kind of computational reasoning has actually happened.
+In one of our main use cases (where the reader or author is trying to retrofit an interpretation to text already written), the target string is present an important part of the query, so this probably doesn't qualify as answer leakage per se.
+Your point about LLM reasoning is a good one though; counterfactual testing, which is possible for any manually annotated solutions, exposes whether some kind of computational reasoning has actually happened.
 We will improve our reporting on this (RQ2).
-
-To respond to your additional points:
-
-_No analysis of statistical variance beyond standard deviation._
-
-Our figures now use box plots for a better sense of the underlying distribution, annotated with the number of samples.
-
-_Assess generalization to unseen writing styles or datasets._
-
-In our experiments the LLM was not fine-tuned or prompted with SciGen examples other than as part of the test cases themselves, so all the SciGen examples were acting as out-of-distribution problems.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Cleanup.

@@
 We thank all reviewers for their time and expertise in evaluating our submission.
-We are grateful for the positive feedback, including the recognition that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), the appreciation that the exploration of a connection between LLM-based code synthesis and data provenance systems is novel (Reviewer Bezq) and that our experimental results, while in need of some further work, show that the approach looks feasible in practice (Reviewers ??).
+We are grateful for the positive feedback, including the recognition that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that the connection between LLM-based code synthesis and data provenance systems is novel (Reviewer Bezq), and that our experimental results, while in need of some further work, indicate that the approach is feasible in practice (Reviewer azjF).
 The constructive feedback provided by the reviewers has been invaluable in understanding how to improve the evaluation.
-We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper and the new experiments conducted in response to multiple reviewer requests.
+We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper, including the new experiments conducted in response to multiple reviewer requests.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Fixing up Bezq.

@@
 
-_Large performance gap between target-value-sharing (74.9%) and no-target (57.1%) suggests potential reliance on implicit answer leakage rather than actual LLM reasoning._
-
-In one of our main use cases (where the reader or author is trying to retrofit an interpretation to text already written), the target string is present an important part of the query, so this probably doesn't qualify as answer leakage per se.
-Your point about LLM reasoning is a good one though; counterfactual testing, which is possible for any manually annotated solutions, exposes whether some kind of computational reasoning has actually happened.
+Re.
+the target sharing point, the target string is an important part of the query for the important use case where the reader or author is trying to retrofit an interpretation to text already written.
+So this probably doesn't qualify as answer leakage per se.
+Your point about determining when LLM reasoning is actually happening a good one though; counterfactual testing, which is possible only given a hand-generated solution, is designed to reveal this.
 We will improve our reporting on this (RQ2).

### 2025-11-21 paper/reviews/iclr2026/response.md
azjF.

@@
 
-_More engineering-oriented, with relatively limited academic research value._ It is an engineering-oriented paper, however we address important research questions around how such an agent-based tool might be designed, how currently LLM competencies meet the needs of such a design, and how manual and automatic verification can be integrated and assessed.
+> - The system presented in this paper is more engineering-oriented, with relatively limited academic research value.
+> - Overall, it can be regarded as a well-executed engineering paper.
 
-_Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered)._ The lack of directly related prior work does indeed make comparison to baseline methods difficult.
-We agree more in-depth ablation studies would be appropriate; we will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other other similar experiments, and include those in our reporting on RQ1 and RQ2.
+We agree that it is an engineering-oriented paper, and thank you for the positive evaluation in that respect.
+However we also address important research questions around how such an agent-based tool might be designed, how current LLM competencies meet the needs of such a design, and how manual and automatic verification can be integrated and assessed.
+
+> - The experimental design lacks a comprehensive justification.
+Without comparisons to other baseline methods, it is difficult to determine whether this system represents the optimal solution (though, as a pioneering work, more in-depth ablation studies could be considered).
+
+The lack of directly related prior work does indeed make comparison to baseline methods difficult.
+We agree more in-depth ablation studies are needed; we will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other other similar experiments, and include those in our reporting on RQ1 and RQ2.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
WIP.

@@
 
-_Isn't this supposed to be an HCI paper?
-To me, there isn't any technical contribution for an ML venue beyond getting the system implemented.
-Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too._ The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study.
+My greatest confusion is: isn't this supposed to be an HCI paper?
+Shouldn't some version of this go to CHI/CSCW/UIST/...?
+To me, there isn't any technical contribution for an ML venue beyond getting the system implemented, but this type of contribution should be very suitable for the HCI venues.
+
+The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study.
 And indeed this would likely go to venue like CHI or IUI.
@@
 
-_The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some documented NLP tasks (like scientific literature?) to begin with._
+> - Since "human" is a really central part in this system, it would be nice to get evaluated by actual human/users too.
 
-_Many of the niche details, like Table 1, are perhaps unnecessary for the main paper._ We prefer to keep Table 1 in the body of the paper, to be explicit about the specific linguistic/quantitative LLM competencies evaluated in the paper; however it should be possible to streamline the figure a bit.
+Agreed, but unfortunately this is not feasible to achieve for this paper, not least because it would require a more mature toolchain to do.
+
+> - The evaluation, from an LLM/NLP standpoint, is perhaps not enough. It would be nice to show if the system could do well on some document NLP tasks (like scientific literature?) to begin with.
+
+[TODO]
+
+> Many of the niche details, like Table 1, are perhaps unnecessary for the main paper, when under consideration at an ML venue.
+
+We prefer to keep Table 1 in the body of the paper, to be explicit about the specific linguistic/quantitative LLM competencies evaluated in the paper; however it should be possible to streamline the figure a bit.
 We will also make a pass over the paper for other details that might be spurious for the ICLR audience.

### 2025-11-21 paper/reviews/iclr2026/response.md
Rough pass.

@@
 
-My greatest confusion is: isn't this supposed to be an HCI paper?
+> - My greatest confusion is: isn't this supposed to be an HCI paper?
 Shouldn't some version of this go to CHI/CSCW/UIST/...?
@@
 
-The natural next step of this work will include a Copilot-like full-featured tool, which will enable a full user study.
-And indeed this would likely go to venue like CHI or IUI.
+The natural next step of this work will be a Copilot-like full-featured tool, which will enable a full user study, and which would likely go to venue like CHI or IUI.
 For this paper, our aim to validate key components of such a tool, including:
@@
 
-We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper was in need further work.
-We discuss elsewhere how we plan to address that in the final revision.
+We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper is in need further work.
+We have discussed elsewhere how we plan to address that in the final revision.
 
@@
 
-Agreed, but unfortunately this is not feasible to achieve for this paper, not least because it would require a more mature toolchain to do.
+Thank you for emphasising the importance of this. We agree, but unfortunately this is not within ascope for this paper, not least because it would require a more mature tool, as discussed above.
 
@@
 
-[TODO]
+There are several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, that are highly relevant and could form part of the capabilityset" required for a tool providing "automated interpretation" of a scientific paper.
+From the computational lingustics literature, data sets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) look relevant, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular looks closely related to some of our natural language tasks.
+We will add some discussion around this to both Related Work and Future Work, and will revisit it as we move towards a more mature tool.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Tweak summary of changes.

@@
 We thank all reviewers for their time and expertise in evaluating our submission.
-We are grateful for the positive feedback, including the recognition that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that the connection between LLM-based code synthesis and data provenance systems is novel (Reviewer Bezq), and that our experimental results, while in need of some further work, indicate that the approach is feasible in practice (Reviewer azjF).
-The constructive feedback provided by the reviewers has been invaluable in understanding how to improve the evaluation.
+We are grateful for the positive feedback, and are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that the connection between LLM-based code synthesis and data provenance systems is novel (Reviewer Bezq), and that our experimental results support the claim that the approach is feasible in practice (Reviewer azjF).
+The constructive feedback provided by the reviewers has been invaluable in understanding how to improve various aspects of the evaluation.
 We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper, including the new experiments conducted in response to multiple reviewer requests.
@@
 We will include this data set as supplementary material and also plan to release it as an open source benchmark.
-- Evaluate SuggestionAgent performance against manual annotations; estimate noise in SuggestionAgent labelling.
-- Extend our correctness evaluation (Fig. 6(b)) to full SciGen dataset, using SuggestionAgent annotations (and adjusted for noise)
+- We will evaluate SuggestionAgent performance against manual annotations and use this to estimate noise in the SuggestionAgent labelling.
+- We will extend the success rate by linguistic category study to the full SciGen dataset, using the SuggestionAgent annotations (with confidence intervals adjusted for noise)
 - Expand the treatment of RQ2

### 2025-11-21 paper/reviews/iclr2026/response.md
Break down by RQ1 and RQ2.

@@
 
-- Clarify manually-annotated subsets of SciGen are randomly chosen; add additional manual annotations to grow subset from 2% to 5%.
-We will include this data set as supplementary material and also plan to release it as an open source benchmark.
-- We will evaluate SuggestionAgent performance against manual annotations and use this to estimate noise in the SuggestionAgent labelling.
-- We will extend the success rate by linguistic category study to the full SciGen dataset, using the SuggestionAgent annotations (with confidence intervals adjusted for noise)
-- Expand the treatment of RQ2
+### Improvements to RQ1 study
+
+We are implementing the following improvements to the RQ1 dataset and evaluation:
+
+- Clarify that manually-annotated subsets of SciGen are generated by randomly selecting from each linguistic category, and add sufficient additional manual annotations to substantially increase the size of the hand-labelled dataset (to 5% of the SciGen issues).
+We will include this dataset as supplementary material and also plan to release it as an open source benchmark.
+- Evaluate SuggestionAgent performance against manual annotations and use this to estimate noise in the SuggestionAgent labelling.
+- Extend the success rate by linguistic category study to the full SciGen dataset, using the SuggestionAgent annotations (with confidence intervals adjusted for noise).
+- Additional ablation study to assess contribution of predefined helpers.
+
+### Improvements to RQ2 study
+
+RQ2 (counterfactual testing) was given only a cursor treatment. We are adding the following improvements:
+- Redo the study of whether LLM solutions which seem otherwise "correct" actually generalise under perturbations of the data, using the improved gold solution dataset
+- Report on counterfactual robustness disaggregated by linguistic category
 
@@
 We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate SuggestionAgent performance using manual annotations to estimate the noise rate.
-We are now able to evaluate the InterpretationAgent using the entirety of the SciGen data set, with confidence intervals adjusted to incorporate estimated noise.
+We are now able to evaluate the InterpretationAgent using the entirety of the SciGen dataset, with confidence intervals adjusted to incorporate estimated noise.
 
@@
 There are several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, that are highly relevant and could form part of the capabilityset" required for a tool providing "automated interpretation" of a scientific paper.
-From the computational lingustics literature, data sets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) look relevant, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular looks closely related to some of our natural language tasks.
+From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) look relevant, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular looks closely related to some of our natural language tasks.
 We will add some discussion around this to both Related Work and Future Work, and will revisit it as we move towards a more mature tool.

### 2025-11-21 paper/sec/authoring-workflow.tex
Fix link to section 4.
Break down by RQ1 and RQ2.

@@
 expression. This is intended to cover the situation where the author has made a claim which is
-\emph{incorrect}, and the data set and surrounding natural language have led the LLM to synthesise an
+\emph{incorrect}, and the dataset and surrounding natural language have led the LLM to synthesise an
 expression which generates a different value from the one specified by the user.
@@
 document configuration. The author can republish the web page hosting the document and interact with the
-proposed revision. As shown in \secref{}, this is an important validation step that can reveal errors in the
-generated expression. If the interactions look reasonable, the author can approve the new document state; this
-is the primary successful path through the workflow and returns the system to the entry state where it is
-waiting for another top-level interaction from the author. Otherwise, the author rejects the proposed change
-and returns to the entry state without any change to the document.
+proposed revision. As shown in \secref{evaluation}, this is an important validation step that can reveal
+errors in the generated expression. If the interactions look reasonable, the author can approve the new
+document state; this is the primary successful path through the workflow and returns the system to the entry
+state where it is waiting for another top-level interaction from the author. Otherwise, the author rejects the
+proposed change and returns to the entry state without any change to the document.
 

### 2025-11-21 paper/sec/introduction.tex
Fix borked refs.
Break down by RQ1 and RQ2.

@@
 
-\paragraph{Contributions.} Our specific contributions are as follows. (\secref{conclusion:future-work}). We
-leave implementing a full Copilot-like authoring plugin for an IDE such as VSCode or Cursor for future work.
+\paragraph{Contributions.} Our specific contributions are as follows. We leave implementing a full
+Copilot-like authoring plugin for an IDE such as VSCode or Cursor for future work (\secref{conclusion}).
 
@@
 \item A proof-of-concept LLM-based tool for iteratively transforming a preexisting opaque document and
-associated data set into a transparent, data-driven counterpart (\secref{authoring-workflow});
+associated dataset into a transparent, data-driven counterpart (\secref{authoring-workflow});
 \item A summary of the natural language idioms we have studied (\secref{nl-idioms}) and an empirical

### 2025-11-21 paper/reviews/iclr2026/response.md
Break down by RQ1 and RQ2.

@@
 
-- Clarify that manually-annotated subsets of SciGen are generated by randomly selecting from each linguistic category, and add sufficient additional manual annotations to substantially increase the size of the hand-labelled dataset (to 5% of the SciGen issues).
-We will include this dataset as supplementary material and also plan to release it as an open source benchmark.
-- Evaluate SuggestionAgent performance against manual annotations and use this to estimate noise in the SuggestionAgent labelling.
-- Extend the success rate by linguistic category study to the full SciGen dataset, using the SuggestionAgent annotations (with confidence intervals adjusted for noise).
+- Clarify that the hand-annotated subsets of SciGen are generated by randomly selecting from each linguistic category.
+- Add sufficient additional manual annotations to substantially increase the size of the hand-labelled dataset (to 5% of the SciGen issues).
+We will include this dataset as supplementary material and will release it as an open source benchmark alonside the final version of the paper.
+- Enhance the SelectionAgent to do an initial labelling by linguistic category. We will evaluate this against our manual annotated subset to estimate the noise in the SuggestionAgent labelling.
+- TRhese machine-generated annotations will allow us to extend the first study (Fig. 6(a)) to the full SciGen dataset (adjusting confidence intervals for noise).
 - Additional ablation study to assess contribution of predefined helpers.
@@
 
-RQ2 (counterfactual testing) was given only a cursor treatment. We are adding the following improvements:
+RQ2 (counterfactual testing) was given only a cursory treatment. We are adding the following improvements:
+
 - Redo the study of whether LLM solutions which seem otherwise "correct" actually generalise under perturbations of the data, using the improved gold solution dataset

### 2025-11-21 paper/reviews/iclr2026/response.md
Further tweaklets.

@@
 
-- Redo the study of whether LLM solutions which seem otherwise "correct" actually generalise under perturbations of the data, using the improved gold solution dataset
-- Report on counterfactual robustness disaggregated by linguistic category
+- Redo the RQ2 study of whether LLM solutions which seem otherwise "correct" actually generalise under perturbations of the data, using the improved gold solution dataset.
+- Report on counterfactual robustness, disaggregated by linguistic category (omitted from submitted version).
 
@@
 Thank you for the suggestion; a multi-paper setup is indeed an interesting use case.
-Although for our evaluation we only considered excerpts from single papers, nothing about our design restricts it to such, so we will consider this for a real-world tool.
+For our evaluation we have only considered excerpts from single papers, but nothing about our design restricts it to such, so we will consider this for a real-world tool.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Further tweaklets.

@@
 We thank all reviewers for their time and expertise in evaluating our submission.
-We are grateful for the positive feedback, and are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that the connection between LLM-based code synthesis and data provenance systems is novel (Reviewer Bezq), and that our experimental results support the claim that the approach is feasible in practice (Reviewer azjF).
-The constructive feedback provided by the reviewers has been invaluable in understanding how to improve various aspects of the evaluation.
-We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper, including the new experiments conducted in response to multiple reviewer requests.
+We are grateful for the positive feedback, and are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that the connection between LLM-based code synthesis and data provenance systems seems to be novel (Reviewer Bezq), and that our experimental results support the claim that the approach is feasible in practice (Reviewer azjF).
+The constructive feedback provided by the reviewers has been invaluable in understanding how to improve various aspects of the empirical evaluation.
+We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper, including the new experiments conducted in response to reviewer requests.
 
@@
 
+- Add sufficient additional manual annotations to substantially increase the size of the hand-labelled dataset (to 5% of the SciGen issues).
+We will include this dataset as supplementary material and will release it as an open source benchmark alongside the final version of the paper.
 - Clarify that the hand-annotated subsets of SciGen are generated by randomly selecting from each linguistic category.
-- Add sufficient additional manual annotations to substantially increase the size of the hand-labelled dataset (to 5% of the SciGen issues).
-We will include this dataset as supplementary material and will release it as an open source benchmark alonside the final version of the paper.
-- Enhance the SelectionAgent to do an initial labelling by linguistic category. We will evaluate this against our manual annotated subset to estimate the noise in the SuggestionAgent labelling.
-- TRhese machine-generated annotations will allow us to extend the first study (Fig. 6(a)) to the full SciGen dataset (adjusting confidence intervals for noise).
+- Enhance the SelectionAgent to do an initial labelling by linguistic category, rather than merely identifying replacement candidates.
+We will evaluate this against our manual annotations to estimate the noise in the SuggestionAgent labelling.
+- These machine-generated labels have allowed us to extend the first study (Fig.
+6(a)) to the full SciGen dataset (adjusting confidence intervals for noise), removing some anomalies (e.g.
+success rates of 0% and 100%) in the original data.
 - Additional ablation study to assess contribution of predefined helpers.
@@
 First, as far as I understand, it is limited to single papers and does not allow to check claims a paper makes about results presented in another paper.
-This is where a system like the proposed would be most useful
+This is where a system like the proposed would be most useful -- while directly linking to evidence in the same paper is useful, manually checking this information is not nearly as laborious as checking something in another paper.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Further tweaklets.

@@
 
-- Redo the RQ2 study of whether LLM solutions which seem otherwise "correct" actually generalise under perturbations of the data, using the improved gold solution dataset.
-- Report on counterfactual robustness, disaggregated by linguistic category (omitted from submitted version).
+- Using the improved gold solution dataset, redo the RQ2 study of whether LLM solutions which seem otherwise "correct" actually generalise under perturbations of the data.
+- Report on counterfactual robustness (omitted from submitted version), namely the mean proportion of responses that are both correct and counterfactually correct, by linguistic category.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Tweaks.

@@
 
-The natural next step of this work will be a Copilot-like full-featured tool, which will enable a full user study, and which would likely go to venue like CHI or IUI.
+The natural next step of this work will be a full-featured Copilot-like authoring (or reading) tool, which will enable a full user study, and which would likely go to venue like CHI, UIST, or IUI.
 For this paper, our aim to validate key components of such a tool, including:
@@
 We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper is in need further work.
-We have discussed elsewhere how we plan to address that in the final revision.
+We have discussed elsewhere how we are addressing that for the final revision.
 
@@
 
-Thank you for emphasising the importance of this. We agree, but unfortunately this is not within ascope for this paper, not least because it would require a more mature tool, as discussed above.
+Thank you for emphasising the importance of this. We agree, but unfortunately this is not within scope for this paper, not least because it would require a more mature tool, as discussed above.
 
@@
 
-There are several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, that are highly relevant and could form part of the capabilityset" required for a tool providing "automated interpretation" of a scientific paper.
+There are several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which are highly relevant and could form part of the "capability set" required for a tool like the one we have in mind.
 From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) look relevant, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular looks closely related to some of our natural language tasks.
-We will add some discussion around this to both Related Work and Future Work, and will revisit it as we move towards a more mature tool.
+We will add some discussion around these to Related Work and consider how some of these challenge problems would become relevant in a more mature tool.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Tweaks.

@@
-[Post each of the top-level sections below a separate comment, with title as given]
+[Post each of the top-level sections below as a separate comment, with title as given]
 
@@
 We thank all reviewers for their time and expertise in evaluating our submission.
-We are grateful for the positive feedback, and are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that the connection between LLM-based code synthesis and data provenance systems seems to be novel (Reviewer Bezq), and that our experimental results support the claim that the approach is feasible in practice (Reviewer azjF).
+We are grateful for the positive feedback, and are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that the connection between LLM-based code synthesis and data provenance systems seems to be novel (Reviewer Bezq), and that our experimental results suggest that the approach is feasible in practice (Reviewer azjF).
 The constructive feedback provided by the reviewers has been invaluable in understanding how to improve various aspects of the empirical evaluation.
@@
 We will include this dataset as supplementary material and will release it as an open source benchmark alongside the final version of the paper.
-- Clarify that the hand-annotated subsets of SciGen are generated by randomly selecting from each linguistic category.
-- Enhance the SelectionAgent to do an initial labelling by linguistic category, rather than merely identifying replacement candidates.
-We will evaluate this against our manual annotations to estimate the noise in the SuggestionAgent labelling.
-- These machine-generated labels have allowed us to extend the first study (Fig.
-6(a)) to the full SciGen dataset (adjusting confidence intervals for noise), removing some anomalies (e.g.
-success rates of 0% and 100%) in the original data.
+- Clarify that the subset chosen for hand-annotation was selected by randomly choosing from each linguistic category.
+- Enhanced the SelectionAgent so that as well as identifying replacement candidates, it also does an initial labelling by linguistic category.
+We will evaluate this against the manual annotations to estimate the noise in the SuggestionAgent labelling.
+- These machine-generated labels have allowed us to extend the first study to the full SciGen dataset (adjusting confidence intervals for noise), removing some anomalies in the original data (e.g.
+success rates of 0% and 100%).
 - Additional ablation study to assess contribution of predefined helpers.

### 2025-11-21 paper/reviews/iclr2026/response.md
Tweaks.

@@
 For the simple cases, where it may seem unfortunate that author verification is still needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since there are many more readers than authors).
-In future work we anticipate providing automated testing as a way of reducing author burden by automatically exercising the generated code.
+In future work we anticipate providing automated testing (e.g.
+automatically generating counterfactual test cases) as a way of reducing author burden by highlighting potential problems automatically.
 
 For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot.
-For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler architecture with less interaction.
+For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler architecture less dependent on interaction.
 
@@
 
-We have changed our figures now use box plots to provide a better sense of the underlying distribution, and annotated with number of samples; thanks for the suggestion.
+We have changed our figures now use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; thanks for the suggestion.
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Tweaks.

@@
 
-We have changed our figures now use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; thanks for the suggestion.
+We have changed our figures now use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; thank you for the suggestion.
 
@@
 Thank you for this point.
-A proper user study with the necessary experimental setup is out of scope for this work, but certainly something we plan to do as the next step, alongside implementing a more complete interactive workflow; we recognise there is a risk (as with all AI-generated code) of simply shifting author workload from writing to debugging.
+A proper user study with the necessary experimental setup is out of scope for this work, but something we plan to do as the next step, alongside implementing a more complete interactive workflow.
+We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.
 For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of the required LLM competencies.
+(It is also worth emphasising that "debugging" by simply interacting with the output is less technically demanding than using a traditional debugger!)
 

### 2025-11-21 paper/reviews/iclr2026/response.md
Tweaks.

@@
 For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues (approximately 2% of the total), resulting in XX labelled replacement tasks.
-We have drawn a second (larger) random subsample (5% of the total) for manual annotation and will report on this in the final version of the paper.
-We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate SuggestionAgent performance using manual annotations to estimate the noise rate.
+We have since selected a second (larger) random subsample (5% of the total) for manual annotation and will report on this in the final version of the paper.
+We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate its performance to estimate the noise rate relative to the manual annotations.
 We are now able to evaluate the InterpretationAgent using the entirety of the SciGen dataset, with confidence intervals adjusted to incorporate estimated noise.
@@
 
-We agree that it is an engineering-oriented paper, and thank you for the positive evaluation in that respect.
+We agree that it is an engineering-oriented paper, and appreciate your positive evaluation in that respect.
 However we also address important research questions around how such an agent-based tool might be designed, how current LLM competencies meet the needs of such a design, and how manual and automatic verification can be integrated and assessed.

### 2025-11-22 paper/reviews/iclr2026/response.md
Tweak.

@@
 We thank all reviewers for their time and expertise in evaluating our submission.
-We are grateful for the positive feedback, and are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that the connection between LLM-based code synthesis and data provenance systems seems to be novel (Reviewer Bezq), and that our experimental results suggest that the approach is feasible in practice (Reviewer azjF).
-The constructive feedback provided by the reviewers has been invaluable in understanding how to improve various aspects of the empirical evaluation.
+We are grateful for the positive feedback; we are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that combining LLM-based code synthesis with data provenance systems seems to be a novel approach to the problem (Reviewer Bezq), and that our experimental results suggest that the approach may be practical (Reviewer azjF).
+The critical feedback provided by the reviewers has also been invaluable in understanding how to improve various aspects of the empirical evaluation.
 We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper, including the new experiments conducted in response to reviewer requests.

### 2025-11-22 paper/reviews/iclr2026/response.md
Tweak.

@@
 We thank all reviewers for their time and expertise in evaluating our submission.
-We are grateful for the positive feedback; we are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that combining LLM-based code synthesis with data provenance systems seems to be a novel approach to the problem (Reviewer Bezq), and that our experimental results suggest that the approach may be practical (Reviewer azjF).
+We are grateful for the positive feedback; we are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that combining LLM-based code synthesis with data provenance systems seems to be a novel approach to the problem (Reviewer Bezq), and that our experimental results suggest that the approach is practical (Reviewer azjF).
 The critical feedback provided by the reviewers has also been invaluable in understanding how to improve various aspects of the empirical evaluation.

### 2025-11-22 paper/reviews/iclr2026/response.md
Tweak.

@@
 We are grateful for the positive feedback; we are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that combining LLM-based code synthesis with data provenance systems seems to be a novel approach to the problem (Reviewer Bezq), and that our experimental results suggest that the approach is practical (Reviewer azjF).
-The critical feedback provided by the reviewers has also been invaluable in understanding how to improve various aspects of the empirical evaluation.
+The critical feedback provided by the reviewers has also been invaluable in understanding how to improve various aspects of the empirical study and analysis.
 We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper, including the new experiments conducted in response to reviewer requests.

### 2025-11-22 paper/reviews/iclr2026/response.md
Tweak.

@@
 The critical feedback provided by the reviewers has also been invaluable in understanding how to improve various aspects of the empirical study and analysis.
-We have provided detailed responses to each reviewer individually and outline below how we plan to further revise the paper, including the new experiments conducted in response to reviewer requests.
+We have provided detailed responses to each reviewer separately; below we outline how we plan to further revise the paper, including new experiments conducted in response to reviewer suggestions.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Tewaks to RQ1 improvements.

@@
 
-- Add sufficient additional manual annotations to substantially increase the size of the hand-labelled dataset (to 5% of the SciGen issues).
-We will include this dataset as supplementary material and will release it as an open source benchmark alongside the final version of the paper.
-- Clarify that the subset chosen for hand-annotation was selected by randomly choosing from each linguistic category.
-- Enhanced the SelectionAgent so that as well as identifying replacement candidates, it also does an initial labelling by linguistic category.
+- We are adding sufficient additional manual annotations to substantially increase the size of the hand-labelled dataset (now covering 5% of SciGen).
+We will include this dataset as supplementary material, and release as an open source benchmark alongside the final version of the paper.
+- We will clarify that the subset chosen for hand-annotation was selected by randomly choosing from each linguistic category.
+- We have enhanced the SelectionAgent so that as well as identifying replacement candidates, it also performs an initial labelling by linguistic category.
 We will evaluate this against the manual annotations to estimate the noise in the SuggestionAgent labelling.

### 2025-11-22 paper/reviews/iclr2026/response.md
Tewaks to RQ1 improvements.

@@
 - We will clarify that the subset chosen for hand-annotation was selected by randomly choosing from each linguistic category.
-- We have enhanced the SelectionAgent so that as well as identifying replacement candidates, it also performs an initial labelling by linguistic category.
-We will evaluate this against the manual annotations to estimate the noise in the SuggestionAgent labelling.
+- The SelectionAgent now also performs an initial labelling by linguistic category, in addition to identifying phrases to be replaced.
+We are evaluating this against the manual annotations to estimate the noise in the automated labelling.
 - These machine-generated labels have allowed us to extend the first study to the full SciGen dataset (adjusting confidence intervals for noise), removing some anomalies in the original data (e.g.

### 2025-11-22 paper/reviews/iclr2026/response.md
Tewaks to RQ1 improvements.

@@
 We are evaluating this against the manual annotations to estimate the noise in the automated labelling.
-- These machine-generated labels have allowed us to extend the first study to the full SciGen dataset (adjusting confidence intervals for noise), removing some anomalies in the original data (e.g.
-success rates of 0% and 100%).
+- Using this labelling, we can now extend the first study to the full SciGen dataset (adjusting confidence intervals for noise), removing some degenerate estimates (e.g.
+success rates of 0% and 100%) attributable to sparsity in the original sample.
 - Additional ablation study to assess contribution of predefined helpers.

### 2025-11-22 paper/reviews/iclr2026/response.md
Tewaks to RQ1 improvements.

--- 2025-11-22 paper/reviews/iclr2026/response.md
@@
 
-- We are adding sufficient additional manual annotations to substantially increase the size of the hand-labelled dataset (now covering 5% of SciGen).
-We will include this dataset as supplementary material, and release as an open source benchmark alongside the final version of the paper.
+- We are adding additional manual annotations to substantially increase the size of the hand-labelled subset (now 5% of SciGen).
 - We will clarify that the subset chosen for hand-annotation was selected by randomly choosing from each linguistic category.
@@
 We are evaluating this against the manual annotations to estimate the noise in the automated labelling.
-- These machine-generated labels have allowed us to extend the first study to the full SciGen dataset (adjusting confidence intervals for noise), removing some anomalies in the original data (e.g.
-success rates of 0% and 100%).
+- Using this method, we can now derive an annotated dataset based on the full SciGen dataset.
+We will include this dataset as supplementary material, and release as an open source benchmark alongside the final version of the paper.
+- We will redo the RQ1 study using the full dataset (adjusting confidence intervals for noise); this should remove some degenerate estimates (e.g.
+success rates of 0% and 100%) attributable to sparsity in the original subset.
 - Additional ablation study to assess contribution of predefined helpers.

### 2025-11-22 paper/reviews/iclr2026/response.md
Tewaks to RQ1 improvements.
Tweaks to RQ2 improvements.

@@
 - We are adding additional manual annotations to substantially increase the size of the hand-labelled subset (now 5% of SciGen).
-- We will clarify that the subset chosen for hand-annotation was selected by randomly choosing from each linguistic category.
+We will clarify that the subset was chosen by randomly choosing from each linguistic category.
 - The SelectionAgent now also performs an initial labelling by linguistic category, in addition to identifying phrases to be replaced.
 We are evaluating this against the manual annotations to estimate the noise in the automated labelling.
-- Using this method, we can now derive an annotated dataset based on the full SciGen dataset.
+- This allows us to derive an annotated dataset based on the full SciGen dataset, which we call SciGen-interpret.
 We will include this dataset as supplementary material, and release as an open source benchmark alongside the final version of the paper.
@@
 success rates of 0% and 100%) attributable to sparsity in the original subset.
-- Additional ablation study to assess contribution of predefined helpers.
+- We will perform an additional ablation study to assess the contribution of predefined helpers to success rates.
 
@@
 
-RQ2 (counterfactual testing) was given only a cursory treatment. We are adding the following improvements:
+RQ2 (counterfactual testing) was given only a cursory treatment in the submitted version. We are implementing the following improvements:
 
-- Using the improved gold solution dataset, redo the RQ2 study of whether LLM solutions which seem otherwise "correct" actually generalise under perturbations of the data.
-- Report on counterfactual robustness (omitted from submitted version), namely the mean proportion of responses that are both correct and counterfactually correct, by linguistic category.
+- Using the improved hand-labelled dataset, redo the RQ2 study of whether LLM solutions which are correct (in the sense of evaluating to the target strong) are also counterfactually correct (i.e.
+generalise under perturbations of the data).
+- Report on counterfactual robustness (omitted from submitted version), namely the mean proportion of responses that are both correct and counterfactually correct, disaggregated by linguistic category.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Burble on AyvP.

@@
 Keeping a human in the loop for verification purposes seems important from a trust point of view.
-For the simple cases, where it may seem unfortunate that author verification is still needed, a small amount of author effort nevertheless translates into a significant _reader_ benefit (since there are many more readers than authors).
-In future work we anticipate providing automated testing (e.g.
-automatically generating counterfactual test cases) as a way of reducing author burden by highlighting potential problems automatically.
+However we agreed that for simple cases, it may seem unfortunate that author verification is still needed.
+It is important to emphasise that a small amount of author effort nevertheless translates into a significant _reader_ benefit, since there are many more readers than authors.
+In future work we anticipate supporting authors better via automated testing (e.g.
+synthesising counterfactual test cases) which will highlight potential problems automatically.
 
-For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot.
-For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler architecture less dependent on interaction.
+For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow based on iterative refinement, similar to working with Copilot, where users rarely accept long completions wholesale, but instead nudge, re-prompt, and constrain via partial code.
+For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler agent-based architecture where user interaction is restricted mainly to validation rather than iterative refinement.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Burble on AyvP.

@@
 Keeping a human in the loop for verification purposes seems important from a trust point of view.
-However we agreed that for simple cases, it may seem unfortunate that author verification is still needed.
+However we agree that for simple cases, it may seem unfortunate that author verification is still needed.
 It is important to emphasise that a small amount of author effort nevertheless translates into a significant _reader_ benefit, since there are many more readers than authors.
-In future work we anticipate supporting authors better via automated testing (e.g.
-synthesising counterfactual test cases) which will highlight potential problems automatically.
+In future work we anticipate supporting authors better via automated testing, for example by synthesising counterfactual test cases which highlight potential problems automatically.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Burble on AyvP.

@@
 
-We have changed our figures now use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; thank you for the suggestion.
+We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; thank you for the suggestion.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Burble on Bezq.

@@
 
-We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; thank you for the suggestion.
+We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; thank you for the comment.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Burble on Bezq.

@@
 For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of the required LLM competencies.
-(It is also worth emphasising that "debugging" by simply interacting with the output is less technically demanding than using a traditional debugger!)
+(It is also worth emphasising that "debugging" by simply interacting with the output is less technically demanding than using a traditional debugger.)
 
@@
 
-For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues (approximately 2% of the total), resulting in XX labelled replacement tasks.
+For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues (approximately 2% of the total), resulting in only 56 labelled replacement tasks for that dataset.
 We have since selected a second (larger) random subsample (5% of the total) for manual annotation and will report on this in the final version of the paper.

### 2025-11-22 paper/reviews/iclr2026/response.md
Burble on Bezq.

@@
 Keeping a human in the loop for verification purposes seems important from a trust point of view.
-However we agree that for simple cases, it may seem unfortunate that author verification is still needed.
-It is important to emphasise that a small amount of author effort nevertheless translates into a significant _reader_ benefit, since there are many more readers than authors.
+Although for simple cases it may seem unfortunate that author verification is still needed, it is important to emphasise that a small amount of author effort nevertheless translates into a significant _reader_ benefit, since there are many more readers than authors.
 In future work we anticipate supporting authors better via automated testing, for example by synthesising counterfactual test cases which highlight potential problems automatically.

### 2025-11-22 paper/reviews/iclr2026/response.md
Burble on Bezq.

@@
 
-For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow based on iterative refinement, similar to working with Copilot, where users rarely accept long completions wholesale, but instead nudge, re-prompt, and constrain via partial code.
+For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot, where users iteratively refine the goal by nudging, re-prompting, and constraining via partial code.
 For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler agent-based architecture where user interaction is restricted mainly to validation rather than iterative refinement.

### 2025-11-22 paper/reviews/iclr2026/response.md
Burble on Bezq.

@@
 Although for simple cases it may seem unfortunate that author verification is still needed, it is important to emphasise that a small amount of author effort nevertheless translates into a significant _reader_ benefit, since there are many more readers than authors.
+It is also worth emphasising that "debugging" by simply interacting with the output is less technically demanding than using a traditional debugger.
 In future work we anticipate supporting authors better via automated testing, for example by synthesising counterfactual test cases which highlight potential problems automatically.
@@
 
-We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; thank you for the comment.
+We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; we appreciate the comment.
 
@@
 A proper user study with the necessary experimental setup is out of scope for this work, but something we plan to do as the next step, alongside implementing a more complete interactive workflow.
-We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.
-For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis and a solid evaluation of the required LLM competencies.
-(It is also worth emphasising that "debugging" by simply interacting with the output is less technically demanding than using a traditional debugger.)
+(We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.) For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Bezq tweaks.

@@
 We have since selected a second (larger) random subsample (5% of the total) for manual annotation and will report on this in the final version of the paper.
-We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate its performance to estimate the noise rate relative to the manual annotations.
-We are now able to evaluate the InterpretationAgent using the entirety of the SciGen dataset, with confidence intervals adjusted to incorporate estimated noise.
+We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate these annotations relative to the manual baseline.
+This means we can now evaluate the core InterpretationAgent on the entirety of the SciGen dataset (with confidence intervals adjusted to incorporate estimated noise).
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Bezq tweaks.

@@
 We have since selected a second (larger) random subsample (5% of the total) for manual annotation and will report on this in the final version of the paper.
-We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate these annotations relative to the manual baseline.
-This means we can now evaluate the core InterpretationAgent on the entirety of the SciGen dataset (with confidence intervals adjusted to incorporate estimated noise).
+We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate those annotations relative to the manual baseline; this means we can now evaluate the core InterpretationAgent on the entirety of the SciGen dataset (with confidence intervals adjusted to incorporate estimated noise).
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Bezq tweaks.

@@
 This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string).
-It is worth emphasising that adopting a fixed or standardised set of terminological definitions (for a given paper or research community, say) might be helpful (or even important in some domains, such as IPCC Summary for Policymaker reports), but your point remains valid.
+It is worth emphasising that adopting a fixed or standardised set of terminological definitions (for a given paper or research community, say) might be helpful or even important in some domains; for example the IPCC Summary for Policymaker report uses terms like _extremely likely_ in a numerically precise way.
+Nevertheless your point remains valid.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Bezq tweaks.

--- 2025-11-22 paper/reviews/iclr2026/response.md
@@
 This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string).
-It is worth emphasising that adopting a fixed or standardised set of terminological definitions (for a given paper or research community, say) might be helpful (or even important in some domains, such as IPCC Summary for Policymaker reports), but your point remains valid.
+It is worth pointing out that fixed or standardised terminological definitions (for a given paper or research community, say) might be helpful or even important in some domains; for example the IPCC Summary for Policymaker report uses terms like _extremely likely_ in a numerically precise way.
+Nevertheless your point remains valid.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Bezq tweaks.

--- 2025-11-22 paper/reviews/iclr2026/response.md
@@
 This is a good suggestion; we will do this and also consider other ablations studies (in addition to excluding the target string).
-It is worth emphasising that adopting a fixed or standardised set of terminological definitions (for a given paper or research community, say) might be helpful (or even important in some domains, such as IPCC Summary for Policymaker reports), but your point remains valid.
+It is worth pointing out that fixed or standardised terminological definitions (for a given paper or research community, say) might be helpful or even important in some domains; for example the IPCC Summary for Policymaker reports use terms like _extremely likely_ in a specific predefined ways.
+Nevertheless your general point remains valid.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
Bezq tweaks.

@@
 Re.
-the target sharing point, the target string is an important part of the query for the important use case where the reader or author is trying to retrofit an interpretation to text already written.
-So this probably doesn't qualify as answer leakage per se.
-Your point about determining when LLM reasoning is actually happening a good one though; counterfactual testing, which is possible only given a hand-generated solution, is designed to reveal this.
+the target-sharing point, the target string is an important part of the query for the key use case where the reader or author is trying to retrofit an interpretation to text already written.
+So this probably does not qualify as answer leakage per se.
+Your point about determining whether LLM reasoning is actually happening is important though; counterfactual testing, which is possible only given a hand-generated solution, is precisely designed to catch this.
 We will improve our reporting on this (RQ2).

### 2025-11-22 paper/reviews/iclr2026/response.md
azjF tweaks.

@@
 The lack of directly related prior work does indeed make comparison to baseline methods difficult.
-We agree more in-depth ablation studies are needed; we will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other other similar experiments, and include those in our reporting on RQ1 and RQ2.
+We agree more in-depth ablation studies are needed; we will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other similar experiments, and include those in our reporting on RQ1 and RQ2.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
VRkd tweaks.

@@
 
-The natural next step of this work will be a full-featured Copilot-like authoring (or reading) tool, which will enable a full user study, and which would likely go to venue like CHI, UIST, or IUI.
+The natural next step of this work will be a full-featured Copilot-like authoring (or reading) tool, which will enable a full user study.
+This will likely go to a venue like CHI, UIST, or IUI.
 For this paper, our aim to validate key components of such a tool, including:
@@
 
-There are several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which are highly relevant and could form part of the "capability set" required for a tool like the one we have in mind.
-From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) look relevant, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular looks closely related to some of our natural language tasks.
+There are several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which could form part of the "capability set" required for a tool like the one we have in mind.
+From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) have some relevance, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular looks closely related to some of our natural language tasks.
 We will add some discussion around these to Related Work and consider how some of these challenge problems would become relevant in a more mature tool.

### 2025-11-22 paper/reviews/iclr2026/response.md
VRkd tweaks.

@@
 There are several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which could form part of the "capability set" required for a tool like the one we have in mind.
-From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) have some relevance, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular looks closely related to some of our natural language tasks.
+From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) have some relevance, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) has problems that look closely related to some of our natural language tasks.
 We will add some discussion around these to Related Work and consider how some of these challenge problems would become relevant in a more mature tool.

### 2025-11-22 paper/reviews/iclr2026/response.md
VRkd tweaks.

@@
 
-Thank you for emphasising the importance of this. We agree, but unfortunately this is not within scope for this paper, not least because it would require a more mature tool, as discussed above.
+We agree, but unfortunately this is not within scope for this paper, not least because it would require a more mature tool, as discussed above.
 
@@
 
-There are several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which could form part of the "capability set" required for a tool like the one we have in mind.
+Thank you for highlighting this.
+There are indeed several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which could form part of the "capability set" required for a tool like the one we have in mind.
 From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) have some relevance, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) has problems that look closely related to some of our natural language tasks.

### 2025-11-22 paper/reviews/iclr2026/response.md
VRkd tweaks.

@@
 
-We agree, but unfortunately this is not within scope for this paper, not least because it would require a more mature tool, as discussed above.
+We agree, but unfortunately this is not within scope for this paper, not least because it would require a more mature tool implementation, as discussed above.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
VRkd tweaks.

@@
 Thank you for highlighting this.
-There are indeed several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which could form part of the "capability set" required for a tool like the one we have in mind.
+There are indeed several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which could form part of the "capability set" required for the mature tool we have in mind.
 From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) have some relevance, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) has problems that look closely related to some of our natural language tasks.

### 2025-11-22 paper/reviews/iclr2026/response.md
VRkd tweaks.

@@
 There are indeed several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which could form part of the "capability set" required for the mature tool we have in mind.
-From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) have some relevance, and [TAT-QA](https://aclanthology.org/2021.acl-long.254/) has problems that look closely related to some of our natural language tasks.
-We will add some discussion around these to Related Work and consider how some of these challenge problems would become relevant in a more mature tool.
+From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) have some relevance; [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular has problems that look closely related to some of our natural language tasks.
+We will add some discussion around these to Related Work and also set out in the conclusion how some of these other problems might become relevant for future versions of the tool.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
VRkd tweaks.

@@
 Thank you for highlighting this.
-There are indeed several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which could form part of the "capability set" required for the mature tool we have in mind.
+There are indeed several well-known NLP problems, beyond argument mining and related techniques already discussed in the paper, which could form part of the "capability set" required for this sort of comprehension assistant.
 From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) have some relevance; [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular has problems that look closely related to some of our natural language tasks.
-We will add some discussion around these to Related Work and also set out in the conclusion how some of these other problems might become relevant for future versions of the tool.
+We will add some discussion around these to Related Work and also set out in the conclusion how some of these other problems might become relevant for a mature version of the tool.
 

### 2025-11-22 paper/reviews/iclr2026/response.md
VRkd tweaks.

@@
 From the computational lingustics literature, datasets like [SciERC](https://aclanthology.org/D18-1360/), [SCIREX](https://aclanthology.org/2020.acl-main.670.pdf) and [Qasper](https://aclanthology.org/2021.naacl-main.365/) (among others) have some relevance; [TAT-QA](https://aclanthology.org/2021.acl-long.254/) in particular has problems that look closely related to some of our natural language tasks.
-We will add some discussion around these to Related Work and also set out in the conclusion how some of these other problems might become relevant for a mature version of the tool.
+We will add some discussion around these to Related Work and also set out in the conclusion how some of these known problems might be relevant to a mature version of the tool.
 

### 2025-11-22 paper/reviews/iclr2026/response.md

@@
 
-We prefer to keep Table 1 in the body of the paper, to be explicit about the specific linguistic/quantitative LLM competencies evaluated in the paper; however it should be possible to streamline the figure a bit.
+We prefer to keep Table 1 in the body of the paper, to be explicit about the specific linguistic/quantitative LLM competencies evaluated in the paper; however it should be possible to streamline the figure a bit, perhaps by moving the example gold solutions to the appendix.
 We will also make a pass over the paper for other details that might be spurious for the ICLR audience.

### 2025-11-24 script/authoring-assistant.sh
fix - missing £

@@ -6,5 +6,5 @@ if [[ $OSTYPE == 'darwin'* ]]; then
   export JAVA_HOME="/Library/Java/JavaVirtualMachines/jdk-22.0.2.jdk/Contents/Home"
 fi
 java --version
-command="java --enable-preview -jar target/authoringAssistant-0.1-jar-with-dependencies.jar"
+command="java --enable-preview -jar target/authoringAssistant-0.1-jar-with-dependencies.jar $1"
 output=$(eval "$command")

### 2025-11-24 src/main/java/authoringassistant/Main.java
fix indentation

@@ -64,7 +64,7 @@ public class Main {
 
         for (Program program : programs) {
             Program p = sa.generateTemplateProgram(program);
-            String json = p.toJsonProgram().toString();
+            String json = p.toJsonProgram().toString(2);
 
             String fileName = STR."\{Path.of(p.getTestCaseFileName()).getFileName()}.json";
             Path outputPath = Paths.get(outputFolder, fileName);

### 2025-11-24 paper/reviews/iclr2026/response.md
Tweaks to AyvP response.

@@
 
-Keeping a human in the loop for verification purposes seems important from a trust point of view.
-Although for simple cases it may seem unfortunate that author verification is still needed, it is important to emphasise that a small amount of author effort nevertheless translates into a significant _reader_ benefit, since there are many more readers than authors.
-It is also worth emphasising that "debugging" by simply interacting with the output is less technically demanding than using a traditional debugger.
-In future work we anticipate supporting authors better via automated testing, for example by synthesising counterfactual test cases which highlight potential problems automatically.
+Although we cannot remove the need for author verification entirely, we do at least make it much easier than it would be using a traditional debugger, since the author can interact directly with the output.
+Keeping a human in the loop for verification purposes is also likely to be important from a trust point of view.
 

### 2025-11-24 paper/reviews/iclr2026/response.md
Tweak to Bezq response.

@@
 
-For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues (approximately 2% of the total), resulting in only 56 labelled replacement tasks for that dataset.
-We have since selected a second (larger) random subsample (5% of the total) for manual annotation and will report on this in the final version of the paper.
+For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues, resulting in only 56 labelled replacement tasks for that dataset.
+We have since selected a substantially larger subsample (also randomly chosen) for manual annotation, and will use this for the final version of the paper.
 We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate those annotations relative to the manual baseline; this means we can now evaluate the core InterpretationAgent on the entirety of the SciGen dataset (with confidence intervals adjusted to incorporate estimated noise).

### 2025-11-24 paper/reviews/iclr2026/response.md
Tweak to Bezq response.

@@
 For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues, resulting in only 56 labelled replacement tasks for that dataset.
-We have since selected a substantially larger subsample (also randomly chosen) for manual annotation, and will use this for the final version of the paper.
-We have also added an automation pipeline to process every SciGen example using the Suggestion Agent, and will evaluate those annotations relative to the manual baseline; this means we can now evaluate the core InterpretationAgent on the entirety of the SciGen dataset (with confidence intervals adjusted to incorporate estimated noise).
+We have since selected a substantially larger subsample (also randomly chosen) for manual annotation, which we will use for the final version of the paper.
+
+In addition, we have also added an automation pipeline to process every SciGen example using the SuggestionAgent.
+This means we can now evaluate the core InterpretationAgent on the *entirety* of the SciGen dataset (with confidence intervals adjusted to incorporate noise estimated by evaluating SuggestionAgent annotations relative to the manual baseline).
 

### 2025-11-24 paper/reviews/iclr2026/response.md
Tweak to Bezq response.

@@
 
-For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues, resulting in only 56 labelled replacement tasks for that dataset.
+For the submitted version we were only able to manually annotate 9 randomly-chosen SciGen issues, resulting in 56 labelled replacement tasks for that dataset.
 We have since selected a substantially larger subsample (also randomly chosen) for manual annotation, which we will use for the final version of the paper.

### 2025-11-24 paper/reviews/iclr2026/response.md
Typo.

@@
 
-- Using the improved hand-labelled dataset, redo the RQ2 study of whether LLM solutions which are correct (in the sense of evaluating to the target strong) are also counterfactually correct (i.e.
+- Using the improved hand-labelled dataset, redo the RQ2 study of whether LLM solutions which are "correct" (in the sense of evaluating to the target string) are also counterfactually correct (i.e.
 generalise under perturbations of the data).

### 2025-11-24 paper/reviews/iclr2026/response.md
Tweak to Bezq response.

@@
 
-Although we cannot remove the need for author verification entirely, we do at least make it much easier than it would be using a traditional debugger, since the author can interact directly with the output.
+Although we cannot remove the need for author verification entirely, we can at least make it much easier than it would be using a traditional debugger, since the author is now able to interact directly with the output.
 Keeping a human in the loop for verification purposes is also likely to be important from a trust point of view.
@@
 For the more complex cases, where current LLMs are unable to generate a full solution in one step, our current design should be able to form the basis of a more interactive workflow, similar to working with Copilot, where users iteratively refine the goal by nudging, re-prompting, and constraining via partial code.
-For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler agent-based architecture where user interaction is restricted mainly to validation rather than iterative refinement.
+For the present paper we are focused on exploring the limits of current LLM capabilities and validating a simpler agent-based architecture, where user interaction is restricted mainly to validation rather than iterative refinement.
 
@@
 Re.
-the target-sharing point, the target string is an important part of the query for the key use case where the reader or author is trying to retrofit an interpretation to text already written.
-So this probably does not qualify as answer leakage per se.
-Your point about determining whether LLM reasoning is actually happening is important though; counterfactual testing, which is possible only given a hand-generated solution, is precisely designed to catch this.
+the target-sharing point, the target string is an essential part of the problem specification, so not part of the answer.
+(The key use case for us is a reader or author trying to retrofit an interpretation to text already written.) Your point about determining whether LLM reasoning is actually happening is important though; counterfactual testing, which is possible only given a hand-generated solution, is precisely designed to catch this.
 We will improve our reporting on this (RQ2).

### 2025-11-24 paper/reviews/iclr2026/response.md
Tweak to azjF response.

@@
 We agree that it is an engineering-oriented paper, and appreciate your positive evaluation in that respect.
-However we also address important research questions around how such an agent-based tool might be designed, how current LLM competencies meet the needs of such a design, and how manual and automatic verification can be integrated and assessed.
+However we also address important research questions around how an agent-based tool for retrofitting computational interpretations to scholarly articles might be designed, how current LLM competencies meet the needs of such a design, and how manual and automatic verification can be integrated and assessed.
 

### 2025-11-24 paper/reviews/iclr2026/response.md
Drop 'compiler in the loop'.

@@
 - competency evaluation for LLMs (which we will substantially improve for this paper)
-- proof-of-concept agentic architecture, including support for compiler-in-the-loop and manual validation (both important for this task)
+- proof-of-concept agentic architecture, with both automatic and manual validation (both important for this task)
 

### 2025-11-24 paper/reviews/iclr2026/response.md
Tweak to VRkd response.
Further tweaklets.

@@
-[Post each of the top-level sections below as a separate comment, with title as given]
+[Post each top-level section below as separate comment, with title as given]
 
@@
 We thank all reviewers for their time and expertise in evaluating our submission.
-We are grateful for the positive feedback; we are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that combining LLM-based code synthesis with data provenance systems seems to be a novel approach to the problem (Reviewer Bezq), and that our experimental results suggest that the approach is practical (Reviewer azjF).
-The critical feedback provided by the reviewers has also been invaluable in understanding how to improve various aspects of the empirical study and analysis.
+We are grateful for the positive feedback; we are glad to hear that the proposed system addresses an important research direction (Reviewers AyvP, Bezq and VRkd), that combining LLM-based code synthesis with data provenance techniques seems to be a novel approach to the problem (Reviewer Bezq), and that our experimental results suggest the approach is practical (Reviewer azjF).
+The critical feedback provided by reviewers has also been invaluable in understanding how to improve various aspects of the empirical study and analysis.
 We have provided detailed responses to each reviewer separately; below we outline how we plan to further revise the paper, including new experiments conducted in response to reviewer suggestions.
@@
 
-- We are adding additional manual annotations to substantially increase the size of the hand-labelled subset (now 5% of SciGen).
+- We are adding additional manual annotations to substantially increase the size of the hand-labelled subset.
 We will clarify that the subset was chosen by randomly choosing from each linguistic category.
@@
 We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper is in need further work.
-We have discussed elsewhere how we are addressing that for the final revision.
+We are substantially improving that section as summarised in "Planned Revisions and New Experiments" above.
 

### 2025-11-24 paper/reviews/iclr2026/response.md

@@
 - We will redo the RQ1 study using the full dataset (adjusting confidence intervals for noise); this should remove some degenerate estimates (e.g.
-success rates of 0% and 100%) attributable to sparsity in the original subset.
+success rates of 0% and 100%) attributable to sparsity in the original subsample.
 - We will perform an additional ablation study to assess the contribution of predefined helpers to success rates.
@@
 
-We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; we appreciate the comment.
+We have changed our figures to use box plots to provide a better sense of the underlying distribution, additionally annotated with number of samples; we hope this addresses your point.
 
@@
 
-In addition, we have also added an automation pipeline to process every SciGen example using the SuggestionAgent.
-This means we can now evaluate the core InterpretationAgent on the *entirety* of the SciGen dataset (with confidence intervals adjusted to incorporate noise estimated by evaluating SuggestionAgent annotations relative to the manual baseline).
+In addition, we have added an automation pipeline to process every SciGen example using the SuggestionAgent.
+This means we can now evaluate the core InterpretationAgent on the *entirety* of the SciGen dataset (with confidence intervals adjusted to incorporate noise, estimated by evaluating SuggestionAgent annotations relative to the manual baseline).
 
@@
 The lack of directly related prior work does indeed make comparison to baseline methods difficult.
-We agree more in-depth ablation studies are needed; we will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other similar experiments, and include those in our reporting on RQ1 and RQ2.
+We agree more in-depth ablation studies are needed.
+We will investigate removing the predefined helper routines (see response to reviewer Bezq above), along with other similar experiments, and include those in our reporting on RQ1 and RQ2.
 
@@
 This will likely go to a venue like CHI, UIST, or IUI.
-For this paper, our aim to validate key components of such a tool, including:
+For the present paper, our aim to validate key components of such a tool, including:
 
@@
 We believe that this is of interest to the ML community, at least from NLP and tool architecture points of view, although we concede that the evaluation in the submitted version of the paper is in need further work.
-We are substantially improving that section as summarised in "Planned Revisions and New Experiments" above.
+We are making improvements to that section as summarised in "Planned Revisions and New Experiments" above.
 

### 2025-11-24 src/main/java/authoringassistant/SuggestionAgent.java
[new] check output from llm

@@ -40,25 +40,28 @@ public class SuggestionAgent {
         }
         PromptList prompts = buildPrompts(text);
         String result;
-        while(true) {
+        int attempts = 0;
+        while(attempts < Settings.getLimit()) {
+            attempts++;
             try {
                 result = llm.evaluate(prompts, null);
                 if (result == null) {
                     return p;
                 }
                 Paragraph paragraph = parseParagraph(result);
+                if(!p.getParagraph().equals(paragraph)) {
+                    prompts.addUserPrompt("Your response contains extra text outside the annotated paragraph. Please provide ONLY the original text with [REPLACE value=\"...\" categories=\"...\"] annotations inserted inline. Do not add any explanations, comments, markdown formatting, or other additional content.");
+                    continue;
+                }
                 return new Program(paragraph, p.getDatasets(),p.getImports(),p.getCode(),p.get_loadedDatasets(),p.getTestCaseFileName(),p.getTest_datasets());
             } catch (IllegalArgumentException ex) {
-                prompts.addUserPrompt("Invalid category! Please enter only one of the following categories:   - COMPARISON\n" +
-                        "  - RANK\n" +
-                        "  - RATIO\n" +
-                        "  - DATA_RETRIEVAL\n" +
-                        "  - AVERAGE\n" +
-                        "  - SUM\n" +
-                        "  - MIN_MAX");
+                prompts.addUserPrompt("Invalid category! Please use only the following categories: COMPARISON, RANK, RATIO, DATA_RETRIEVAL, DIFFERENCE, AVERAGE, SUM, MIN_MAX. Return ONLY the annotated paragraph with [REPLACE ...] tags, without any additional comments or explanations.");
                 continue;
             }
         }
+        
+        // Limite di tentativi raggiunto, restituisci il programma originale
+        return p;
 
     }
 

### 2025-11-25 script/chart-generation.py
[new] boxplot chart

@@ -15,6 +15,7 @@ def get_latest_csv(folder_path):
         exit(1)
 
     latest_file = max(csv_files, key=os.path.getmtime)
+    print(f"Using latest CSV file: {latest_file}")
     return latest_file
 
 def generate_success_rate_test_case_plot(df, plot):
@@ -81,6 +82,73 @@ def generate_aggregated_plot(df, plot):
     plt.savefig(f"{fig_dir}/success_rate_by_category.png")
     plt.close()
 
+def generate_aggregated_boxplot(df, plot):
+    df['expression-type'] = df['expression-type'].str.strip('[]').str.split(',')
+    df_exploded = df.explode('expression-type')
+    df_exploded['expression-type'] = df_exploded['expression-type'].str.strip()
+
+    summary = (
+        df_exploded.groupby(["runId", "expression-type", "target-value"])
+        .agg(success_rate=("success", "mean"))
+        .reset_index()
+    )
+
+    category_counts = (
+        df_exploded.groupby("expression-type")
+        .apply(lambda x: x[['test-case', 'target-value']].drop_duplicates().shape[0])
+        .to_dict()
+    )
+
+    # labels for  target-value
+    plt.figure(figsize=(8,6))
+    label_map = {
+        1: "Present",
+        0: "Absent"
+    }
+    summary["target_label"] = summary["target-value"].map(label_map)
+
+    ax = sns.boxplot(
+        data=summary,
+        x="expression-type",
+        y="success_rate",
+        hue="target_label",
+        palette="Set2",
+        order=sorted(summary['expression-type'].unique())
+    )
+    ax.legend(title="Target value", loc='lower right', bbox_to_anchor=(1.0, -0.25))
+
+    categories = sorted(summary['expression-type'].unique())
+    target_values = sorted(summary['target-value'].unique())
+    
+    for i, category in enumerate(categories):
+        for j, target_val in enumerate(target_values):
+            count = df_exploded[
+                (df_exploded['expression-type'] == category) & 
+                (df_exploded['target-value'] == target_val)
+            ].shape[0] // df_exploded['runId'].nunique()
+            
+            if count > 0:
+                x_offset = -0.2 if j == 0 else 0.2
+                mask = (summary['expression-type'] == category) & (summary['target-value'] == target_val)
+                if mask.sum() > 0:
[…]

### 2025-11-25 paper/reviews/iclr2026/response.md
Update response.md (future user study)

@@
 A proper user study with the necessary experimental setup is out of scope for this work, but something we plan to do as the next step, alongside implementing a more complete interactive workflow.
-(We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.) For this paper, our main focus is on a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.
+We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.
+To illustrate how we plan to conduct our future user study, we briefly describe its basic design:
+- Participants: scientific authors with experience in writing/reviewing manuscripts and knowledge of Fluid.
+- Tasks: participants will validate the interpretations generated by the system and, if necessary, correct them.
+The time taken for validation and the time for possible corrections will be measured.
+A comparison will be made with the baseline methodology in which participants will work without LLM support.
+- Metrics: acceptance rate for the suggestions, average time for validation and corrections, and participant workload and satisfaction will be measured using standard questionnaires.
+For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.
 
@@
 
-We agree, but unfortunately this is not within scope for this paper, not least because it would require a more mature tool implementation, as discussed above.
+We agree that evaluating the tool with real users is important.
+However, this is beyond the scope of this work, as it would require a fully mature authoring tool.
+As discussed in more detail above (in the response to Reviewer Bezq), we plan to conduct a user study as future work, once the tool is finalised.
 

### 2025-11-25 src/main/java/authoringassistant/Main.java
[update] move logs in a new folder, and update version library for pretty-printing

@@ -74,7 +74,7 @@ public class Main {
     }
 
     private static void writeLog(ArrayList<Pair<Program, QueryResult>> results, String agent, int learningContextSize) throws IOException {
-        try (PrintWriter out = new PrintWriter(new FileOutputStream(STR."\{Settings.getLogFolder()}/log_\{System.currentTimeMillis()}.csv"))) {
+        try (PrintWriter out = new PrintWriter(new FileOutputStream("results/results.csv"))) {
             String[] headers = {
                     "runId", "test-case", "llm-agent", "temperature", "num-token", "is-negative", "in-context-learning-size",
                     "attempts", "result", "target-value", "expression-type", "generated-expression", "expected-value", "expected-expression", "parseErrors", "counterfactualFails", "nullExpressions", "onlyLiteralExpressions", "duration(ms)"

### 2025-11-25 script/chart-generation.py
[update] update folder for charts

@@ -194,7 +194,7 @@ def generate_success_rate_by_category_count(df, plot):
     plt.close()
 
 def generate_charts():
-    latest_csv = get_latest_csv("logs")
+    latest_csv = get_latest_csv("results")
     print(f"Using latest CSV file: {latest_csv}")
     df = pd.read_csv(latest_csv, delimiter=';', quotechar='"', encoding='utf-8')
     df["success"] = df["generated-expression"].notna().astype(int)

### 2025-11-25 paper/reviews/iclr2026/response.md
Streamline slightly.

@@
 We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.
-To illustrate how we plan to conduct our future user study, we briefly describe its basic design:
-- Participants: scientific authors with experience in writing/reviewing manuscripts and knowledge of Fluid.
+
+We anticipate a study with the following basic design:
+- Participants: scientific authors with experience in writing/reviewing manuscripts and coding experience
 - Tasks: participants will validate the interpretations generated by the system and, if necessary, correct them.
-The time taken for validation and the time for possible corrections will be measured.
-A comparison will be made with the baseline methodology in which participants will work without LLM support.
+Time taken for validation and any corrections will be measured, and a comparison made with a baseline workflow in which participants work without LLM support.
 - Metrics: acceptance rate for the suggestions, average time for validation and corrections, and participant workload and satisfaction will be measured using standard questionnaires.
+
 For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.

### 2025-11-25 paper/reviews/iclr2026/response.md
Tweak to VRkd.

@@
 We agree that evaluating the tool with real users is important.
-However, this is beyond the scope of this work, as it would require a fully mature authoring tool.
-As discussed in more detail above (in the response to Reviewer Bezq), we plan to conduct a user study as future work, once the tool is finalised.
+However, this is beyond the scope of the present work, as it would require a mature authoring tool.
+As discussed in more detail above (in response to Reviewer Bezq), we plan to conduct a user study as future work once the tool is finalised.
 

### 2025-11-25 paper/reviews/iclr2026/response.md
Some reorg of Bezq.

@@
 
-Thank you for this point.
-A proper user study with the necessary experimental setup is out of scope for this work, but something we plan to do as the next step, alongside implementing a more complete interactive workflow.
-We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.
-
-We anticipate a study with the following basic design:
+Thank you for this point. A proper user study with the necessary experimental setup is out of scope for this work, but we anticipate a study with the following basic design:
 - Participants: scientific authors with experience in writing/reviewing manuscripts and coding experience
@@
 
-For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.
+We will clarify in Future Work that this is something we plan to do as the next step, alongside implementing a more complete interactive workflow.
+(We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.) For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.
 

### 2025-11-25 src/main/java/authoringassistant/Main.java
[update] folder structure

@@ -74,7 +74,8 @@ public class Main {
     }
 
     private static void writeLog(ArrayList<Pair<Program, QueryResult>> results, String agent, int learningContextSize) throws IOException {
-        try (PrintWriter out = new PrintWriter(new FileOutputStream("results/results.csv"))) {
+        Files.createDirectories(Path.of(STR."results/\{Path.of(Settings.getTestCaseFolder()).getFileName()}/"));
+        try (PrintWriter out = new PrintWriter(new FileOutputStream(STR."results/\{Path.of(Settings.getTestCaseFolder()).getFileName()}/results.csv"))) {
             String[] headers = {
                     "runId", "test-case", "llm-agent", "temperature", "num-token", "is-negative", "in-context-learning-size",
                     "attempts", "result", "target-value", "expression-type", "generated-expression", "expected-value", "expected-expression", "parseErrors", "counterfactualFails", "nullExpressions", "onlyLiteralExpressions", "duration(ms)"

### 2025-11-25 paper/reviews/iclr2026/response.md

@@
 - Participants: scientific authors with experience in writing/reviewing manuscripts and coding experience
-- Tasks: participants will validate the interpretations generated by the system and, if necessary, correct them.
-Time taken for validation and any corrections will be measured, and a comparison made with a baseline workflow in which participants work without LLM support.
+- Tasks: participants will validate the interpretations generated by the system and, if necessary, correct them; time taken for validation and any corrections will be measured, compared to a baseline workflow in which participants work without LLM support
 - Metrics: acceptance rate for the suggestions, average time for validation and corrections, and participant workload and satisfaction will be measured using standard questionnaires.
@@
 We will clarify in Future Work that this is something we plan to do as the next step, alongside implementing a more complete interactive workflow.
-(We certainly recognise that there is a risk, as with all AI-generated code, of simply shifting author workload from writing to debugging/validation.) For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.
+(We certainly recognise that there is a risk, as with AI-generated code in general, of simply shifting author workload from writing to debugging/validation.) For this paper, our main focus remains on developing a basic agent-based architecture for interpretation synthesis, plus a solid evaluation of the required LLM competencies.
 

### 2025-11-25 script/generate-scigen-manual.sh
generating-scigen-manual-script

@@ -1,0 +1,24 @@
+#!/bin/bash
+set -xe
+
+RAW_SCIGEN_DIR="./testCases/scigen-SuggestionAgent-raw/"
+TARGET_DIR="./testCases/scigen-manual-2/"
+RANDOM_SEED=2025 
+TARGET_SIZE=20
+
+mkdir -p "$TARGET_DIR"
+
+# Find all json files, shuffle them and take up to target size
+find "$RAW_SCIGEN_DIR" -maxdepth 1 -name "*.json" | \
+awk -v RANDOM_SEED="$RANDOM_SEED" 'BEGIN {srand(RANDOM_SEED)} {print rand() "\t" $0}' | \
+sort -k1,1n | \
+cut -f2- | \
+head -n "$TARGET_SIZE" | \
+while read -r json_path; do
+    
+    # Copy json then .fld file
+    cp "$json_path" "$TARGET_DIR/"
+    fld_path="${json_path%.json}.fld"
+    cp "$fld_path" "$TARGET_DIR/"
+    
+done

### 2025-11-25 script/chart-generation.py
[update] generate charts for each subfolder in result

@@ -6,7 +6,31 @@ import glob
 import os
 import json
 
-fig_dir = "paper/fig"
+def count_problems_per_category(df):
+    """
+    Count the number of unique problems per linguistic category.
+    
+    Args:
+        df: DataFrame with columns 'expression-type', 'test-case', 'target-value'
+    
+    Returns:
+        Series with category names as index and problem counts as values
+    """
+    # Explode expression-type categories (format: "[Category1,Category2]")
+    df_exploded = df.copy()
+    # Convert to string and handle NaN values
+    df_exploded['expression-type'] = df_exploded['expression-type'].astype(str)
+    df_exploded['expression-type'] = df_exploded['expression-type'].str.strip('[]').str.split(',')
+    df_exploded = df_exploded.explode('expression-type')
+    df_exploded['expression-type'] = df_exploded['expression-type'].str.strip()
+    
+    # Count unique (test-case, target-value) pairs per category
+    # This avoids double-counting the same problem across multiple runs
+    category_counts = df_exploded.groupby("expression-type").apply(
+        lambda x: x[['test-case', 'target-value']].drop_duplicates().shape[0]
+    )
+    
+    return category_counts
 
 def get_latest_csv(folder_path):
     csv_files = glob.glob(os.path.join(folder_path, "*.csv"))
@@ -18,7 +42,7 @@ def get_latest_csv(folder_path):
     print(f"Using latest CSV file: {latest_file}")
     return latest_file
 
-def generate_success_rate_test_case_plot(df, plot):
+def generate_success_rate_test_case_plot(df, plot, fig_dir):
     summary = (
         df.groupby("test-case-short")
         .agg(success_rate=("success", "mean"), avg_attempts=("attempts", "mean"), num_queries=("test-case-short", "count"))
@@ -35,9 +59,9 @@ def generate_success_rate_test_case_plot(df, plot):
     plt.savefig(f"{fig_dir}/success_rate_by_test_case.png")
     plt.close()
 
-def generate_aggregated_plot(df, plot):
+def generate_aggregated_plot(df, plot, fig_dir):
     # esplodo le categorie
-    df['expression-type'] = df['expression-type'].str.strip('[]').str.split(',')
+    df['expression-type'] = df['expression-type'].astype(str).str.strip('[]').str.split(',')
     df_exploded = df.explode('expression-type')
     df_exploded['expression-type'] = df_exploded['expression-type'].str.strip()
 
[…]

### 2025-11-25 script/chart-generation.py
[update] fix aggregation on problems_per_category

@@ -7,15 +7,6 @@ import os
 import json
 
 def count_problems_per_category(df):
-    """
-    Count the number of unique problems per linguistic category.
-    
-    Args:
-        df: DataFrame with columns 'expression-type', 'test-case', 'target-value'
-    
-    Returns:
-        Series with category names as index and problem counts as values
-    """
     # Explode expression-type categories (format: "[Category1,Category2]")
     df_exploded = df.copy()
     # Convert to string and handle NaN values
@@ -24,10 +15,10 @@ def count_problems_per_category(df):
     df_exploded = df_exploded.explode('expression-type')
     df_exploded['expression-type'] = df_exploded['expression-type'].str.strip()
     
-    # Count unique (test-case, target-value) pairs per category
-    # This avoids double-counting the same problem across multiple runs
+    # Count unique (expected-expression, target-value) pairs per category
+    # This counts the actual expressions, not the test case files
     category_counts = df_exploded.groupby("expression-type").apply(
-        lambda x: x[['test-case', 'target-value']].drop_duplicates().shape[0]
+        lambda x: x[['expected-expression', 'target-value']].drop_duplicates().shape[0]
     )
     
     return category_counts

### 2025-11-25 src/main/java/authoringassistant/paragraph/ExpressionCategory.java
[fix] category

@@ -23,7 +23,7 @@ public enum ExpressionCategory {
 
     public static ExpressionCategory of(String category) {
         for (ExpressionCategory cat : values()) {
-            if (cat.name().equalsIgnoreCase(category)) {
+            if (cat.getName().equalsIgnoreCase(category)) {
                 return cat;
             }
         }

### 2025-11-25 src/main/java/authoringassistant/paragraph/ExpressionCategory.java
[fix] category

@@ -8,7 +8,7 @@ public enum ExpressionCategory {
     DIFFERENCE("Difference"),
     AVERAGE("Average"),
     SUM("Sum"),
-    MIN_MAX("Min/Max"),
+    MIN_MAX("Min Max"),
     GENERALISED_QUANTIFIER("Generalised Quantifier");
 
 

### 2025-11-25 src/main/java/authoringassistant/paragraph/ExpressionCategory.java
[fix] category

@@ -8,7 +8,7 @@ public enum ExpressionCategory {
     DIFFERENCE("Difference"),
     AVERAGE("Average"),
     SUM("Sum"),
-    MIN_MAX("Min Max"),
+    MIN_MAX("Min/Max"),
     GENERALISED_QUANTIFIER("Generalised Quantifier");
 
 

### 2025-11-25 script/chart-generation.py
[fix] remove negative cases from charts

@@ -222,6 +222,8 @@ def process_csv_file(csv_file):
     print(f"{'='*60}")
     
     df = pd.read_csv(csv_file, delimiter=';', quotechar='"', encoding='utf-8')
+    # Filter out rows where is-negative=true
+    df = df[df["is-negative"].astype(str).str.lower() != "true"]
     df["success"] = df["generated-expression"].notna().astype(int)
     df["target-value"] = df["target-value"].astype(int)
     df["attempts"] = pd.to_numeric(df["attempts"], errors="coerce")

### 2025-11-25 script/chart-generation.py
fix label

@@ -98,6 +98,9 @@ def generate_aggregated_plot(df, plot, fig_dir):
     plt.close()
 
 def generate_aggregated_boxplot(df, plot, fig_dir):
+    # Use the reusable function to count problems per category BEFORE exploding df
+    category_counts = count_problems_per_category(df)
+    
     df['expression-type'] = df['expression-type'].astype(str).str.strip('[]').str.split(',')
     df_exploded = df.explode('expression-type')
     df_exploded['expression-type'] = df_exploded['expression-type'].str.strip()
@@ -107,11 +110,7 @@ def generate_aggregated_boxplot(df, plot, fig_dir):
         .agg(success_rate=("success", "mean"))
         .reset_index()
     )
-
-    # Use the reusable function to count problems per category
-    category_counts = count_problems_per_category(df)
-    category_counts_dict = category_counts.to_dict()
-
+    
     # labels for  target-value
     plt.figure(figsize=(8,6))
     label_map = {
@@ -130,28 +129,31 @@ def generate_aggregated_boxplot(df, plot, fig_dir):
     )
     ax.legend(title="Target value", loc='lower right', bbox_to_anchor=(1.0, -0.25))
 
+    # Use categories from summary for iteration
     categories = sorted(summary['expression-type'].unique())
     target_values = sorted(summary['target-value'].unique())
     
     for i, category in enumerate(categories):
-        # Get problem count for this category
-        n_problems = category_counts_dict.get(category, 0)
-        n_runs = df_exploded['runId'].nunique()
-        count = n_problems // n_runs if n_runs > 0 else n_problems
+        # Get problem count for this category from category_counts
+        # category_counts already contains the total unique problems (not per run)
+        if category in category_counts.index:
+            count = int(category_counts[category])
+        else:
+            count = 0
         
         for j, target_val in enumerate(target_values):
-            if count > 0:
-                x_offset = -0.2 if j == 0 else 0.2
-                mask = (summary['expression-type'] == category) & (summary['target-value'] == target_val)
-                if mask.sum() > 0:
-                    median_y = summary[mask]['success_rate'].median()
-                else:
-                    median_y = 0.5  
-                
-                ax.text(i + x_offset, median_y, f'n:{count}', 
-                        ha='center', va='center', 
-                        fontsize=8, color='black', weight='bold',
-                        bbox=dict(boxstyle='round,pad=0.2', facecolor='white', alpha=0.8))
[…]

### 2025-11-28 paper/sec/evaluation.tex
[paper] update references

--- 2025-09-25 paper/sec/evaluation.tex
@@
         \centering
-        \includegraphics[width=\linewidth]{fig/success_rate_by_category}
+        \includegraphics[width=\linewidth]{fig/scigen-manual-old/success_rate_by_category_boxplot}
         \caption{Success rate by Linguistic Category}
@@
         \centering
-        \includegraphics[width=\linewidth]{fig/success_rate_by_complexity}
+        \includegraphics[width=\linewidth]{fig/scigen-manual-old/success_rate_by_complexity}
         \caption{Success rate by Complexity}

### 2025-12-04 paper/sec/evaluation.tex
Update fig.

--- 2025-09-25 paper/sec/evaluation.tex
@@
         \centering
-        \includegraphics[width=\linewidth]{fig/success_rate_by_category}
+        \includegraphics[width=\linewidth]{fig/scigen-SuggestionAgent-downsampled/success_rate_by_category}
         \caption{Success rate by Linguistic Category}
@@
         \centering
-        \includegraphics[width=\linewidth]{fig/success_rate_by_complexity}
+        \includegraphics[width=\linewidth]{fig/scigen-SuggestionAgent-downsampled/success_rate_by_complexity}
         \caption{Success rate by Complexity}
